---
title: "Product Advancement Final Report"
output: html_document
params:
  toDownload: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = TRUE
)

library(DT)
library(magrittr)
library(shiny)
library(ggplot2)

load("resultProdAdv.RData")
final_table <- final_table_export
```

```{r, include=FALSE}
printTable <- function(DT, pageLength = 25,
                       numericColNames = NULL, numRound = 3,
                       scrollXOpt = TRUE, filt = "top", ...) {

  oldnames <- colnames(DT)

  DT <- data.frame(lapply(DT, function(x) {
    if (is.numeric(x)) round(x, numRound) else x
  }))

  colnames(DT) <- oldnames

  table <- DT::datatable(
    data = DT,
    filter = filt,
    options = list(
      dom = "l<<t>pB>",
      buttons = c("copy", "csv", "excel", "print"),
      pageLength = pageLength,
      searchHighlight = TRUE,
      lengthMenu = c(10, 25, 50, 100),
      scrollX = scrollXOpt
    ),
    extensions = "Buttons",
    rownames = FALSE,
    ...
  )

  if (length(numericColNames) > 0) {
    table <- table %>%
      DT::formatRound(columns = numericColNames, digits = numRound)
  }

  table
}

```

### About this analysis


**Inputs used by the Pre-Advancement Module:**

- **MTA predictions (BLUPs):** Multi-trial analysis best linear unbiased predictions across environments, providing the overall genetic merit estimates for each candidate.
- **STA predictions:** Single-trial analysis per-environment estimates used for spatial performance visualisation and lollipop plots.
- **Pedigree or genomic relatedness data (optional):** Used to compute family groupings and relatedness plots for diversity monitoring.
- **Environmental covariates (optional):** Climate variables (temperature, rainfall, humidity indices) used for environment clustering and TPE (Target Population of Environments) visualisations.

**Outputs produced:**

- A **named final selection stamp** containing each designation with its selection status (**SELECTED**, **NOT SELECTED**, **REVISE**, or **CHECK**), index values, and individual trait values.
- The final decision table consolidates initial selection (index-based), table-level overrides, and plot-level manual overrides using the hierarchy: plot > table > initial.

**Sharing with stakeholders:** The output RData file from this analysis can be shared with other stakeholders for use in the **Advancement Meeting Dashboard** module, enabling collaborative review and decision-making across teams.

---

```{r summary-section, results='asis'}
# Compute summary statistics
if (!is.null(final_table) && nrow(final_table) > 0 && "final_decision" %in% colnames(final_table)) {
  n_evaluated <- nrow(final_table)
  n_selected <- sum(final_table$final_decision == "SELECTED", na.rm = TRUE)
  n_not_selected <- sum(final_table$final_decision == "NOT SELECTED", na.rm = TRUE)
  n_revise <- sum(final_table$final_decision == "REVISE", na.rm = TRUE)
  n_checks <- sum(final_table$final_decision == "CHECK", na.rm = TRUE)
  
  analysis_label <- if (exists("analysis_name") && !is.null(analysis_name) && nzchar(analysis_name)) {
    analysis_name
  } else {
    "Unnamed analysis"
  }
  
  cat(paste0(
    "### Summary\n\n",
    "**Analysis name:** ", analysis_label, "\n\n",
    "- **Candidates evaluated:** ", n_evaluated, "\n",
    "- **Selected:** ", n_selected, "\n",
    "- **Not selected:** ", n_not_selected, "\n",
    "- **Revise:** ", n_revise, "\n",
    "- **Checks:** ", n_checks, "\n\n"
  ))
} else {
  cat("**No final selection stamp exists.** Complete the Final Review workflow to see results here.\n\n")
}
```

---

###  Final decision table

The table below shows all candidates with their trait values, index score, and final selection status. The `final_decision` column reflects the consolidated decision after applying the hierarchy: plot-level overrides take precedence over table-level overrides, which take precedence over the initial index-based selection. Rows are color-coded by decision status (blue = selected, orange = not selected, pink = check, yellow = revise).

```{r, include=FALSE}
htmltools::tagList(printTable(mtcars))

# Remove trait_decision columns (keep trait values only)
trait_decision_cols <- grep("_trait_decision$", colnames(final_table), value = TRUE)
final_table <- final_table[, setdiff(colnames(final_table), trait_decision_cols), drop = FALSE]

# Define decision columns in desired order: initial, table, plot, final
main_decision_cols <- intersect(
  c("initial_decision", "table_decision", "plot_decision", "final_decision"),
  colnames(final_table)
)

# Everything else (e.g. designation, numeric traits, index_value)
other_cols <- setdiff(colnames(final_table), main_decision_cols)

# Reorder: trait values first, then decision columns
final_table <- final_table[, c(
  other_cols,
  main_decision_cols
), drop = FALSE]

```


```{r}
dt <- printTable(
  final_table,
  pageLength = 25,
  scrollXOpt = TRUE,
  numericColNames = names(final_table)[vapply(final_table, is.numeric, logical(1))]
)

decision_cols <- grep(
  "_decision$",
  colnames(final_table),
  value = TRUE
)

decision_cols <- intersect(
  decision_cols,
  colnames(final_table)
)

for (col in decision_cols) {
  dt <- DT::formatStyle(
    dt,
    columns = col,
    backgroundColor = DT::styleEqual(
      c("SELECTED", "NOT SELECTED", "CHECK","REVISE"),
      c("#D6EAF8", "#F5C4A5", "#F5DCE9","#FFF3CD")
    ),
    fontWeight = "bold"
  )
}

# Colour the trait/index cells by each row's final decision, matching the
# gradient-filled Decision Table on the Output tab.
if ("final_decision" %in% colnames(final_table)) {
  value_cols <- setdiff(
    names(final_table)[vapply(final_table, is.numeric, logical(1))],
    decision_cols
  )
  if (length(value_cols) > 0) {
    dt <- DT::formatStyle(
      dt,
      columns = value_cols,
      valueColumns = "final_decision",
      backgroundColor = DT::styleEqual(
        c("SELECTED", "NOT SELECTED", "CHECK", "REVISE"),
        c("#EAF4FB", "#FBEEE4", "#F7E9F1", "#FDF6E0")
      )
    )
  }
}

htmltools::tagList(dt)

```

---

```{r set-download-eval, include=FALSE}
# The remaining sections (Population Statistics, TPP Breakdown, Global Options
# Summary) are rendered by Shiny outputs on the live dashboard. They are only
# included in the standalone downloaded HTML (params$toDownload == TRUE).
is_download <- tryCatch(isTRUE(params$toDownload), error = function(e) FALSE)
```

```{r population-statistics-header, results='asis', eval=is_download}
cat("\n\n### Population Statistics\n\n")
cat("For every metric - the selection index and each trait used in the selection - the table below contrasts the **selected** set against **all candidates** (checks excluded) and against the **checks** on their own. Rows are colour-coded to match the distribution plot: blue = selected, orange = all candidates, purple = checks.\n\n")
```

```{r population-statistics, results='asis', eval=is_download}
have_stats <- exists("review_long") && !is.null(review_long) &&
  is.data.frame(review_long) && nrow(review_long) > 0

if (!have_stats) {
  cat("_No trait distribution data available for this selection._\n\n")
}
```

```{r population-statistics-tables, results='asis', eval=is_download}
# Rendered as plain styled HTML rather than DT: these are small fixed 3-row
# summaries, and DT's JavaScript width calculation detaches the header from the
# body when several tables are emitted into a static document.
if (have_stats) {
  group_colors <- if (exists("REVIEW_GROUP_COLORS")) {
    REVIEW_GROUP_COLORS
  } else {
    c("Selected" = "#85C1E9", "All candidates" = "#E8A87C", "Checks" = "#C39BD3")
  }

  # Wash the group colours toward white so the text stays readable
  fade <- function(hex, opacity = 0.35) {
    r <- strtoi(substr(hex, 2, 3), 16L)
    g <- strtoi(substr(hex, 4, 5), 16L)
    b <- strtoi(substr(hex, 6, 7), 16L)
    sprintf("#%02X%02X%02X",
            as.integer(r * opacity + 255 * (1 - opacity)),
            as.integer(g * opacity + 255 * (1 - opacity)),
            as.integer(b * opacity + 255 * (1 - opacity)))
  }
  faded <- vapply(group_colors, fade, character(1))
  group_levels <- names(group_colors)
  stat_cols <- c("Min", "Max", "Mean", "Median", "SD")

  fmt <- function(x) if (is.na(x)) "&ndash;" else formatC(round(x, 3), format = "f", digits = 3)

  stats_for <- function(sub) {
    do.call(rbind, lapply(group_levels, function(g) {
      v <- sub$value[sub$group == g]
      v <- v[is.finite(v)]
      if (length(v) == 0) {
        data.frame(Group = g, Min = NA_real_, Max = NA_real_, Mean = NA_real_,
                   Median = NA_real_, SD = NA_real_, stringsAsFactors = FALSE)
      } else {
        data.frame(Group = g, Min = min(v), Max = max(v), Mean = mean(v),
                   Median = stats::median(v), SD = stats::sd(v),
                   stringsAsFactors = FALSE)
      }
    }))
  }

  for (m in unique(review_long$metric)) {
    sub <- review_long[review_long$metric == m, , drop = FALSE]
    label <- if (identical(m, "index_value")) "Index Value" else m
    sdf <- stats_for(sub)

    header <- paste0(
      "<tr>",
      "<th style='text-align:left; padding:7px 10px; border-bottom:2px solid #2C3E50;'>Group</th>",
      paste0("<th style='text-align:right; padding:7px 10px; border-bottom:2px solid #2C3E50;'>",
             stat_cols, "</th>", collapse = ""),
      "</tr>"
    )

    body <- paste0(vapply(seq_len(nrow(sdf)), function(i) {
      bg <- faded[[sdf$Group[i]]]
      paste0(
        "<tr style='background-color:", bg, ";'>",
        "<td style='text-align:left; padding:6px 10px; border-bottom:1px solid #fff;'>",
        sdf$Group[i], "</td>",
        paste0("<td style='text-align:right; padding:6px 10px; border-bottom:1px solid #fff;'>",
               vapply(stat_cols, function(cl) fmt(sdf[[cl]][i]), character(1)),
               "</td>", collapse = ""),
        "</tr>"
      )
    }, character(1)), collapse = "")

    cat(paste0(
      "<div style='margin: 0 0 22px 0;'>",
      "<div style='font-weight:700; font-size:14px; color:#2C3E50;",
      " padding:0 0 5px 2px; border-bottom:2px solid #2C3E50; margin-bottom:0;'>",
      label, "</div>",
      "<table style='width:100%; border-collapse:collapse; font-size:13px;",
      " table-layout:fixed; margin-top:0;'>",
      "<thead>", header, "</thead><tbody>", body, "</tbody></table>",
      "</div>\n"
    ))
  }
}
```

```{r plot-intro-text, results="asis", eval=is_download}
cat("The plot below shows, for each metric, how the selected set is distributed against the full candidate set. The dashed vertical line marks the mean of the checks.\n\n")
```

```{r population-statistics-height, include=FALSE, eval=is_download}
# Scale the figure with the number of facet rows (2 panels per row)
stats_fig_height <- 4
if (have_stats) {
  n_panels <- length(unique(review_long$metric))
  stats_fig_height <- max(3.5, ceiling(n_panels / 2) * 2.6)
}
```

```{r population-statistics-plot, fig.width=11, fig.height=if(exists("stats_fig_height")) stats_fig_height else 4, eval=is_download}
if (have_stats) {
  plot_df <- review_long[review_long$group %in% c("All candidates", "Selected"), , drop = FALSE]

  if (nrow(plot_df) > 0) {
    metric_order <- unique(review_long$metric)
    label_order <- vapply(metric_order, function(m) {
      if (identical(m, "index_value")) "Index Value" else m
    }, character(1))

    plot_df$label <- factor(plot_df$label, levels = unname(label_order))
    plot_df$group <- factor(plot_df$group, levels = c("All candidates", "Selected"))

    check_df <- review_long[review_long$group == "Checks", , drop = FALSE]
    check_means <- NULL
    if (nrow(check_df) > 0) {
      check_means <- stats::aggregate(value ~ label, data = check_df,
                                      FUN = mean, na.rm = TRUE)
      check_means$label <- factor(check_means$label, levels = unname(label_order))
    }

    group_colors <- if (exists("REVIEW_GROUP_COLORS")) {
      REVIEW_GROUP_COLORS
    } else {
      c("Selected" = "#85C1E9", "All candidates" = "#E8A87C", "Checks" = "#C39BD3")
    }

    p <- ggplot2::ggplot(plot_df, ggplot2::aes(x = value, fill = group)) +
      ggplot2::geom_histogram(position = "identity", alpha = 0.55,
                              bins = 30, colour = NA) +
      ggplot2::facet_wrap(~ label, scales = "free", ncol = 2) +
      ggplot2::scale_fill_manual(
        values = c(
          "All candidates" = unname(group_colors[["All candidates"]]),
          "Selected"       = unname(group_colors[["Selected"]])
        ),
        name = NULL
      ) +
      ggplot2::labs(x = "Predicted value", y = "Count") +
      ggplot2::theme_minimal(base_size = 12) +
      ggplot2::theme(legend.position = "top",
                     strip.text = ggplot2::element_text(face = "bold"))

    if (!is.null(check_means) && nrow(check_means) > 0) {
      p <- p + ggplot2::geom_vline(
        data = check_means,
        ggplot2::aes(xintercept = value),
        colour = unname(group_colors[["Checks"]]),
        linetype = "dashed", linewidth = 0.9
      )
    }

    # Mean lines for Selected and All candidates
    sel_means <- stats::aggregate(value ~ label, data = plot_df[plot_df$group == "Selected", , drop = FALSE], FUN = mean, na.rm = TRUE)
    all_means <- stats::aggregate(value ~ label, data = plot_df[plot_df$group == "All candidates", , drop = FALSE], FUN = mean, na.rm = TRUE)
    if (nrow(sel_means) > 0) {
      sel_means$label <- factor(sel_means$label, levels = levels(plot_df$label))
      p <- p + ggplot2::geom_vline(
        data = sel_means,
        ggplot2::aes(xintercept = value),
        colour = unname(group_colors[["Selected"]]),
        linetype = "dashed", linewidth = 0.7
      )
    }
    if (nrow(all_means) > 0) {
      all_means$label <- factor(all_means$label, levels = levels(plot_df$label))
      p <- p + ggplot2::geom_vline(
        data = all_means,
        ggplot2::aes(xintercept = value),
        colour = unname(group_colors[["All candidates"]]),
        linetype = "dashed", linewidth = 0.7
      )
    }

    print(p)
  }
}
```

```{r hr1, results="asis", eval=is_download}
cat("\n\n---\n\n")
```

```{r tpp-header, results="asis", eval=is_download}
cat("### TPP Breakdown\n\n")
```

```{r tpp-breakdown, results='asis', eval=is_download}
have_bd <- exists("breakdown_detail") && !is.null(breakdown_detail) &&
  !is.null(breakdown_detail$table) && is.data.frame(breakdown_detail$table) &&
  nrow(breakdown_detail$table) > 0

if (!have_bd) {
  cat("_No TPP information available for this selection._\n\n")
} else {
  cat(paste0(
    "This table compares what the Target Product Profile asks for against what was ",
    "actually evaluated on your phenotypic data. The `_tpp` columns reproduce the TPP ",
    "verbatim; the `_data` columns show what was resolved against your data - relative ",
    "specifications are converted to the equivalent absolute threshold using the mean of ",
    "the checks you mapped. `Sel_pct_meeting_criteria` is the share of the **selected** ",
    "individuals meeting the resolved threshold and `All_pct_meeting_criteria` the share ",
    "across all candidates, so comparing the two shows how much selection enriched each ",
    "criterion. Traits with no phenotypic mapping read \"not evaluated\" and traits you ",
    "dropped read \"trait not used\".\n\n"
  ))
}
```

```{r tpp-breakdown-table, eval=is_download}
if (have_bd) {
  # Colour-code the two Target_proximity columns. Defined inline because the
  # knit environment has no access to the package's internal helpers.
  .prox_bg <- c(
    "Comfortably meeting"   = "#D5F5E3",
    "Moderately meeting"    = "#E8F8F5",
    "Barely meeting"        = "#FCF3CF",
    "Barely failing"        = "#FDEBD0",
    "Moderately failing"    = "#FADBD8",
    "Substantially failing" = "#F5B7B1"
  )
  .prox_fg <- c(
    "Comfortably meeting"   = "#186A3B",
    "Moderately meeting"    = "#148F77",
    "Barely meeting"        = "#7D6608",
    "Barely failing"        = "#935116",
    "Moderately failing"    = "#943126",
    "Substantially failing" = "#78281F"
  )

  .bd <- printTable(breakdown_detail$table, pageLength = 20,
                    scrollXOpt = TRUE, filt = "none")
  for (.cl in intersect(c("Target_proximity_sel", "Target_proximity_all"),
                        colnames(breakdown_detail$table))) {
    .bd <- DT::formatStyle(
      .bd,
      columns = .cl,
      backgroundColor = DT::styleEqual(names(.prox_bg), unname(.prox_bg)),
      color = DT::styleEqual(names(.prox_fg), unname(.prox_fg)),
      fontWeight = DT::styleEqual(names(.prox_bg), rep("600", length(.prox_bg)))
    )
  }

  htmltools::tagList(.bd)
}
```

```{r tpp-breakdown-notes, results='asis', eval=is_download}
if (have_bd && length(breakdown_detail$env_notes) > 0) {
  cat("**Traits evaluated on a subset of environments**\n\n")
  for (n in breakdown_detail$env_notes) cat(paste0("- ", n, "\n"))
  cat("\n")
}
```

```{r hr2, results="asis", eval=is_download}
cat("\n\n---\n\n")
```

```{r opts-header, results="asis", eval=is_download}
cat("### Global Options Summary\n\n")
```

```{r global-options, results='asis', eval=is_download}
have_opts <- exists("options_table") && !is.null(options_table) &&
  is.data.frame(options_table) && nrow(options_table) > 0

if (!have_opts) {
  cat("_No options configured for this selection._\n\n")
} else {
  cat(paste0(
    "This table records the **actual selection criteria applied** when the selection ran ",
    "- not what the TPP asks for. For each trait it shows the index weight it carried, ",
    "the threshold value and direction taken from its trait rule card, and the reference ",
    "check(s) used where the rule was defined relative to a check. Traits with a weight of ",
    "0 did not contribute to the index, and traits showing \"no threshold\" were not ",
    "filtered on a cut-off.\n\n"
  ))
}
```

```{r global-options-table, eval=is_download}
if (have_opts) {
  htmltools::tagList(
    printTable(options_table, pageLength = 25, scrollXOpt = TRUE, filt = "none")
  )
}
```
