---
title: "Quality Assurance Phenotypic Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
  toDownload: FALSE
---

```{r setup, include=FALSE}

# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
options(knitr.kable.NA = '')

# loading necessary R packages ####
## data manipulation
library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric

## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(grafify)
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                         numericColNames = NULL, numRound = 3, 
                         scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,...) {
    
    DT <- data.frame(lapply(X = DT, 
                            FUN = function(x) {
                              if(is.numeric(x)){
                                round(x, numRound)
                              } else {
                                x
                              }
                            }))
    
    table <- DT::datatable(data = DT, 
                           colnames = colNames,
                           filter = "top", 
                           options = list(autoWidth = autoWidthOpt,
                                          dom = 'l<<t>pB>', 
                                          buttons = c('copy', 'csv', 'excel', 'print'),
                                          pageLength = pageLength,
                                          searchHighlight = TRUE,
                                          lengthMenu = c(7, 14, 21, 28, 35),
                                          scrollX = scrollXOpt),
                           extensions = 'Buttons',
                           rownames = FALSE,
                           ...)
    if (length(numericColNames) > 0){
      table <- table %>% DT::formatRound(columns = numericColNames,
                                     digits = numRound)
    }
    
    table
  }
```

```{r }

suppressWarnings(tryCatch({
  if(file.exists("./outputs/result.RData")){
    load("./outputs/resultQaPheno.RData")
  }else{
    load("resultQaPheno.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

```


### Table of modifications
<p>&nbsp;</p>

The following table shows the modifications that have been saved in your analysis object and that will be applied to the raw phenotypic data when this specific time stamp (ID) is selected/used in other analytical modules (e.g., single trial analysis module).

```{r }

if (is.null(result)){
  # invisible(NULL)
} else {
  modifications <- result$modifications$pheno
  idQa <- result$status[which(result$status$module %in% c("qaRaw","qaMb","qaFilter","qaDesign","qaConsist")),"analysisId"];
  idQa <- idQa[length(idQa)]
  idQaType <- result$status[result$status$analysisId == idQa, "module"]
  modifications <- modifications[which( modifications$analysisId == idQa & modifications$module == idQaType),]
  traitsTable <- unique(modifications[which( modifications$analysisId == idQa),"trait"])
  
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("traitQa"), 
                       label = "Trait:", 
                       choices = traitsTable, 
    )
  }
  
}

```


```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  numericalColumns <- c("value","row") 
  
  if("params" %in% ls()){ # we will download the document
    printTable(modifications, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(modifications),
           numericColNames = numericalColumns,
           numRound = 4)
  }else{
    
    DT::renderDT(printTable(modifications[modifications$trait==input$traitQa,],
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(modifications[modifications$trait==input$traitQa,]),
                        numericColNames = numericalColumns,
                        numRound = 4), server = FALSE)
  }
}

```



<p>&nbsp;</p>

### Boxplot of modifications

The following boxplot allows you to visualize which records were tagged as outliers and will be ignored in posterior analyses if this QA time stamp is selected in the posterior analyses.

<p>&nbsp;</p>

```{r }

if (is.null(result)){
  # invisible(NULL)
} else {
  # modifications <- result$modifications$pheno
  # idQa <- result$status[which(result$status$module %in% c("qaRaw","qaMb","qaFilter","qaDesign","qaConsist")),"analysisId"];
  # idQa <- idQa[length(idQa)]
  # idQaType <- result$status[result$status$analysisId == idQa, "module"]
  traitsBoxplot <- unique(modifications[which( modifications$analysisId == idQa),"trait"])
  if(idQaType == "qaMb"){traitsBoxplot <- paste0(traitsBoxplot,"-residual")}
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("traitQaBox"), 
                       label = "Trait:", 
                       choices = traitsBoxplot, 
    )
  }
  
}

```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  # object <- result
  mydata <- result$data$pheno
  ### change column names for mapping
  paramsPheno <- result$metadata$pheno
  paramsPheno <- paramsPheno[which(paramsPheno$parameter != "trait"),]
  colnames(mydata) <- cgiarBase::replaceValues(colnames(mydata), 
                                               Search = paramsPheno$value, 
                                               Replace = paramsPheno$parameter )
  ###
  mydata$rowindex <- 1:nrow(mydata)
  mydata[, "environment"] <- as.factor(mydata[, "environment"])
  mydata[, "designation"] <- as.factor(mydata[, "designation"])
  mo <-result$modifications$pheno
  analysisIds <- unique(mo$analysisId)
  analysisIds <- analysisIds[length(analysisIds)]
  mo <- mo[which(mo$analysisId == analysisIds),]
  if(idQaType == "qaMb"){
    mo$trait <- paste0(mo$trait,"-residual")
  }
  
  if("params" %in% ls()){ # we will download the document
    
    cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
    
    for(iTrait in traitsBoxplot){
      cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
      
      mydata$color <- "valid"
      if(nrow(mo) > 0){
        mydata$color[which(mydata$rowindex %in% unique(mo$row[which(mo$trait == iTrait)]))]="tagged"
      }
      mydata$color <- as.factor(mydata$color)
      mydata$predictedValue <- mydata[,iTrait]
      
      p <- ggplot(mydata, aes(x=as.factor(environment), y=as.numeric(predictedValue))) +
        geom_boxplot(fill='#A4A4A4', color="black", notch = TRUE, outliers = FALSE)+
        theme_classic()+
        geom_jitter(aes(colour = color), alpha = 0.6) +
        xlab("Environment") + ylab("Trait value") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
        scale_color_manual(values = c(valid = "#66C2A5", tagged = "#FC8D62")) # specifying colors names avoids having valid points in orange in absence of potential outliers. With only colour = color, valid points are in orange in that case.
      print(p)
    }
    
  }else{
    
    shiny::renderPlot({
      
      mydata$color <- "valid"
      if(nrow(mo) > 0){
        mydata$color[which(mydata$rowindex %in% unique(mo$row[which(mo$trait == input$traitQaBox)]))]="tagged"
      }
      mydata$color <- as.factor(mydata$color)
      mydata$predictedValue <- mydata[,input$traitQaBox]
      
      ggplot(mydata, aes(x=as.factor(environment), y=as.numeric(predictedValue))) +
        geom_boxplot(fill='#A4A4A4', color="black", notch = TRUE, outliers = FALSE)+
        theme_classic()+
        geom_jitter(aes(colour = color), alpha = 0.6) +
        xlab("Environment") + ylab("Trait value") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
        scale_color_manual(values = c(valid = "#66C2A5", tagged = "#FC8D62"))
      
    })
    
  }
  
}
```



### References on methods used

Tukey, J. W. (1977). Exploratory Data Analysis. Section 2C.

McGill, R., Tukey, J. W. and Larsen, W. A. (1978). Variations of box plots. The American Statistician, 32, 12–16. doi:10.2307/2683468.

Velleman, P. F. and Hoaglin, D. C. (1981). Applications, Basics and Computing of Exploratory Data Analysis. Duxbury Press.


<p>&nbsp;</p>





