---
title: "Selection Index Report"
author: ""
date: "December 2023"
output: html_document
params:
 toDownload: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
```

```{r printfxn, include=FALSE}
# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, ...) {
  
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  
  table <- DT::datatable(data = DT, 
                         filter = "top", 
                         options = list(autoWidth = TRUE, 
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'excel'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35)),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r plot_init, include = FALSE}
# Init Step to make sure that the dependencies are loaded
# was done to enable proper printing of the tab header and the graphs within the for loop below using print(htmltools::tagList(ggplotly(...)))
htmltools::tagList(printTable(mtcars))
htmltools::tagList(plotly::ggplotly(ggplot2::ggplot()))

# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}

```

### Across-environment trait table

The following table allows you to inspect the across-environment predictions for the different entries (rows) and different traits (columns).

<p>&nbsp;</p>

```{r, results='asis'}
suppressWarnings(tryCatch({
  if(file.exists("./outputs/resultIndex.RData")){
    load("./outputs/resultIndex.RData")
  }else{
    load("resultIndex.RData")
  }
}, error = function(e) {
  NULL
}))

if (is.null(result)){
  # invisible(NULL)
} else {
  mydata = result$predictions
  idIndex <- result$status[which(result$status$module == "indexD"),"analysisId"];
  idIndex <- idIndex[length(idIndex)]
  idMet <-result$modeling[which(result$modeling$parameter %in% "analysisId" & result$modeling$analysisId == idIndex),"value"]
  mydata <- mydata[which(mydata$analysisId %in% c(idMet,idIndex)),]
  wide <- stats::reshape(mydata[,c(c("designation"),"trait",c("predictedValue"))], direction = "wide", idvar = c("designation"),
                         timevar = "trait", v.names = c("predictedValue"), sep= "_")
  colnames(wide) <- gsub("predictedValue_","",colnames(wide))
  numeric.output <- colnames(wide)[-c(1)]
  
  if("params" %in% ls()){ # we will download the document
    printTable(wide)
  }else{
    DT::renderDT(DT::formatRound(DT::datatable(wide, extensions = 'Buttons',
                                               options = list(dom = 'Blfrtip',scrollX = TRUE,buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                                                              lengthMenu = list(c(5,20,50,-1), c(5,20,50,'All')))
    ), numeric.output)
    )
  }
  
}
```

<p>&nbsp;</p>

### Radar Plot

This plot allows the user to visualize the distance between the population means and the target values.

<p>&nbsp;</p>

```{r, results='asis'}
library(dplyr)
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    
    # cat(paste0("#### B  {.tabset .tabset-pills}    \n\n  "))
    # cat("\n\n#####", "Radar", "{.tabset .tabset-pills}       \n\n")
    
    dtIdxD <- result$predictions
    idIndex <- result$status[which(result$status$module == "indexD"),"analysisId"];
    idIndex <- idIndex[length(idIndex)]
    idMet <-result$modeling[which(result$modeling$parameter %in% "analysisId" & result$modeling$analysisId == idIndex),"value"]
    mydata <- dtIdxD[which(dtIdxD$analysisId == idMet),setdiff(colnames(dtIdxD),c("module","analysisId"))]
    # subset to only traits being part of the index
    desirev <-result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"value"]
    uTraits <- result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"trait"]
    uTraits <- gsub("_scaled","",uTraits)
    dtIdxD <- dtIdxD[which(dtIdxD$trait %in% uTraits),]
    ##
    environmentPredictionsRadar2="across"
    traitFilterPredictionsRadar2=unique(dtIdxD[which(dtIdxD$analysisId == idMet),"trait"])
    ## ensure product profile means come sorted
    meanGroupPredictionsRadar= result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),c("trait","value")]
    meanGroupPredictionsRadar <- meanGroupPredictionsRadar[with(meanGroupPredictionsRadar, order(as.numeric(as.factor(trait)))), ]
    meanGroupPredictionsRadar <- meanGroupPredictionsRadar[,"value"]
    
    fontSizeRadar=12; r0Radar=NULL; neRadar=NULL; plotSdRadar=FALSE
    
    mydata = mydata[which( (mydata$environment %in% environmentPredictionsRadar2) &
                             (mydata$trait %in% traitFilterPredictionsRadar2)
    ),]
    
    if(!is.null(environmentPredictionsRadar2) & !is.null(traitFilterPredictionsRadar2)){
      mm <- stats::aggregate(predictedValue~trait, data=mydata, FUN=mean, na.rm=TRUE)
      mmsd <- stats::aggregate(predictedValue~trait, data=mydata, FUN=stats::sd, na.rm=TRUE)
      mm2 <- stats::aggregate(predictedValue~trait, data=mydata, FUN=min, na.rm=TRUE)
      mm3 <- stats::aggregate(predictedValue~trait, data=mydata, FUN=max, na.rm=TRUE)
      namesA <- mm[,1]
      meanA <- mm[,2]/(mm3[,2])
      
      # meanA <- rep(0.75,nrow(mm))
      fig <-  plotly::plot_ly(
        type = 'scatterpolar',
        fill = 'toself'
      )
      fig <- fig %>%
        plotly::add_trace(
          r = meanA, #c(39, 28, 8, 7, 28, 39),
          theta =paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
          name = 'Breeding population means',
          text=round(mm[,2],3)
        )
      ## add product profile means
      if(!is.null(meanGroupPredictionsRadar)){
        meanB <- as.numeric(unlist(strsplit(meanGroupPredictionsRadar,",")))
        if(length(meanB) == length(mm[,2])){
          meanB <- meanB + mm[,2]
          # r2 = (meanB*0.75)/mm[,2]
          r2 = meanB/(mm3[,2])
          dToSearch <- meanB #- mm[,2]
          fig <- fig %>%
            plotly::add_trace(
              r = r2, # c(1.5, 10, 39, 31, 15, 1.5),
              theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
              name = "Desired population", #'Group B'
              # mode="text",
              text=round(dToSearch,2)
              # textfont = list(color = '#000000', size = fontSizeRadar)
            )
        }
      }
      # add SD polygon
      if(plotSdRadar){
        meanC <- (mm[,2]+mmsd[,2])/(mm3[,2])
        fig <- fig %>%
          plotly::add_trace(
            r = meanC, # c(1.5, 10, 39, 31, 15, 1.5),
            theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
            name = "Mean + Standard Deviation from pop", #'Group B'
            text=round(mm[,2]+mmsd[,2],3)
          )
        meanD <- (mm[,2]-mmsd[,2])/(mm3[,2])
        fig <- fig %>%
          plotly::add_trace(
            r = meanD, # c(1.5, 10, 39, 31, 15, 1.5),
            theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
            name = "Mean - Standard Deviation from pop", #'Group B'
            text=round(mm[,2]-mmsd[,2],3)
          )
      }
      # add selection limits polygon
      ## add product profile means'
      if(!is.null(r0Radar)){
        meanR <- as.numeric(unlist(strsplit(r0Radar,",")))/100
        if(length(meanR) == length(mm[,2])){
          meanR <- mm[,2] + ((meanR*mm[,2])*neRadar*2) # selection limits 2*Ne*R0
          r3 = (meanR*0.75)/mm[,2]
          dToSearchR <- meanR #- mm[,2]
          fig <- fig %>%
            plotly::add_trace(
              r = r3, # c(1.5, 10, 39, 31, 15, 1.5),
              theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
              name = "Selection limits", #'Group B'
              # mode="text",
              text=round(dToSearchR,2)
              # textfont = list(color = '#000000', size = fontSizeRadar)
            )
        }
      }
    }
    print(htmltools::tagList(plotly::ggplotly(fig)))
    
  }else{
    
    plotly::renderPlotly({
      dtIdxD <- result$predictions
      idIndex <- result$status[which(result$status$module == "indexD"),"analysisId"];
      idIndex <- idIndex[length(idIndex)]
      idMet <-result$modeling[which(result$modeling$parameter %in% "analysisId" & result$modeling$analysisId == idIndex),"value"]
      mydata <- dtIdxD[which(dtIdxD$analysisId == idMet),setdiff(colnames(dtIdxD),c("module","analysisId"))]
      # subset to only traits being part of the index
      desirev <-result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"value"]
      uTraits <- result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"trait"]
      uTraits <- gsub("_scaled","",uTraits)
      dtIdxD <- dtIdxD[which(dtIdxD$trait %in% uTraits),]
      ##
      environmentPredictionsRadar2="across"
      traitFilterPredictionsRadar2=unique(dtIdxD[which(dtIdxD$analysisId == idMet),"trait"])
      ## ensure product profile means come sorted
      meanGroupPredictionsRadar= result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),c("trait","value")]
      meanGroupPredictionsRadar <- meanGroupPredictionsRadar[with(meanGroupPredictionsRadar, order(as.numeric(as.factor(trait)))), ]
      meanGroupPredictionsRadar <- meanGroupPredictionsRadar[,"value"]
      
      fontSizeRadar=12; r0Radar=NULL; neRadar=NULL; plotSdRadar=FALSE
      
      mydata = mydata[which( (mydata$environment %in% environmentPredictionsRadar2) &
                               (mydata$trait %in% traitFilterPredictionsRadar2)
      ),]
      
      if(!is.null(environmentPredictionsRadar2) & !is.null(traitFilterPredictionsRadar2)){
        mm <- stats::aggregate(predictedValue~trait, data=mydata, FUN=mean, na.rm=TRUE)
        mmsd <- stats::aggregate(predictedValue~trait, data=mydata, FUN=stats::sd, na.rm=TRUE)
        mm2 <- stats::aggregate(predictedValue~trait, data=mydata, FUN=min, na.rm=TRUE)
        mm3 <- stats::aggregate(predictedValue~trait, data=mydata, FUN=max, na.rm=TRUE)
        namesA <- mm[,1]
        meanA <- mm[,2]/(mm3[,2])
        
        # meanA <- rep(0.75,nrow(mm))
        fig <-  plotly::plot_ly(
          type = 'scatterpolar',
          fill = 'toself'
        )
        fig <- fig %>%
          plotly::add_trace(
            r = meanA, #c(39, 28, 8, 7, 28, 39),
            theta =paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
            name = 'Breeding population means',
            text=round(mm[,2],3)
          )
        ## add product profile means
        if(!is.null(meanGroupPredictionsRadar)){
          meanB <- as.numeric(unlist(strsplit(meanGroupPredictionsRadar,",")))
          if(length(meanB) == length(mm[,2])){
            meanB <- meanB + mm[,2]
            # r2 = (meanB*0.75)/mm[,2]
            r2 = meanB/(mm3[,2])
            dToSearch <- meanB #- mm[,2]
            fig <- fig %>%
              plotly::add_trace(
                r = r2, # c(1.5, 10, 39, 31, 15, 1.5),
                theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
                name = "Desired population", #'Group B'
                # mode="text",
                text=round(dToSearch,2)
                # textfont = list(color = '#000000', size = fontSizeRadar)
              )
          }
        }
        # add SD polygon
        if(plotSdRadar){
          meanC <- (mm[,2]+mmsd[,2])/(mm3[,2])
          fig <- fig %>%
            plotly::add_trace(
              r = meanC, # c(1.5, 10, 39, 31, 15, 1.5),
              theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
              name = "Mean + Standard Deviation from pop", #'Group B'
              text=round(mm[,2]+mmsd[,2],3)
            )
          meanD <- (mm[,2]-mmsd[,2])/(mm3[,2])
          fig <- fig %>%
            plotly::add_trace(
              r = meanD, # c(1.5, 10, 39, 31, 15, 1.5),
              theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
              name = "Mean - Standard Deviation from pop", #'Group B'
              text=round(mm[,2]-mmsd[,2],3)
            )
        }
        # add selection limits polygon
        ## add product profile means'
        if(!is.null(r0Radar)){
          meanR <- as.numeric(unlist(strsplit(r0Radar,",")))/100
          if(length(meanR) == length(mm[,2])){
            meanR <- mm[,2] + ((meanR*mm[,2])*neRadar*2) # selection limits 2*Ne*R0
            r3 = (meanR*0.75)/mm[,2]
            dToSearchR <- meanR #- mm[,2]
            fig <- fig %>%
              plotly::add_trace(
                r = r3, # c(1.5, 10, 39, 31, 15, 1.5),
                theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
                name = "Selection limits", #'Group B'
                # mode="text",
                text=round(dToSearchR,2)
                # textfont = list(color = '#000000', size = fontSizeRadar)
              )
          }
        }
      }
      fig
    }) # enf of render plotly
    
  }
  
}

```


### Expected Response Plot

This plot allows the user to visualize the expected response in the new generation given the individuals selected (best 20 percent) using the selection index.

<p>&nbsp;</p>

```{r}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    proportionTrait <- c(0.05,0.10,0.15,0.20)
  }else{
    shiny::numericInput(ns("proportionTrait"), label = "Selected proportion", value = 0.1, min=0.001,max=1, step=0.05)
  }
  
}
```

```{r, results='asis'}
library(dplyr)
if (is.null(result)){
  # invisible(NULL)
} else {
  
  dtIdxD <- result$predictions
  idIndex <- result$status[which(result$status$module == "indexD"),"analysisId"];
  idIndex <- idIndex[length(idIndex)]
  idMet <-result$modeling[which(result$modeling$parameter %in% "analysisId" & result$modeling$analysisId == idIndex),"value"]
  mydata <- dtIdxD[which(dtIdxD$analysisId == idMet),setdiff(colnames(dtIdxD),c("module","analysisId"))]
  desirev <-result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"value"]
  uTraits <- result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"trait"]
  uTraits <- gsub("_scaled","",uTraits)
  # uTraits <- unique(dtIdxD[which(dtIdxD$analysisId == idMet),"trait"])
  
  object=result
  environmentPredictionsRadar2="across"
  traitFilterPredictionsRadar2=uTraits
  meanGroupPredictionsRadar=desirev
  analysisId=idMet
  trait=uTraits
  desirev=desirev
  scaled=as.logical(result$modeling[which(result$modeling$analysisId == idIndex & result$modeling$parameter == "scaled"),"value"])
  
  mydata <- object$predictions
  mydata <- mydata[which(mydata$analysisId == analysisId),]
  
  mydata = mydata[which( (mydata$environment %in% environmentPredictionsRadar2) &
                           (mydata$trait %in% traitFilterPredictionsRadar2)
  ),]
  mydata$selected <- "no"
  ## pop means
  mm <- stats::aggregate(predictedValue~trait, data=mydata, FUN=mean, na.rm=TRUE)
  
  
  if("params" %in% ls()){ # we will download the document
    
    cat(paste0("#### Proportion selected  {.tabset .tabset-pills}    \n\n  "))
    for(iProportionTrait in proportionTrait){
      cat("\n\n#####", iProportionTrait, "{.tabset .tabset-pills}       \n\n")
      
      if(length(as.numeric(unlist(strsplit(meanGroupPredictionsRadar,",")))) == nrow(mm)){
        ## add product profile means'
        mm$threshold <- as.numeric(unlist(strsplit(meanGroupPredictionsRadar,","))) + mm[,2]
        mm$desire <- mm$threshold - mm$predictedValue
        # index calculation
        resInd <- cgiarPipeline::indexDesire(
          phenoDTfile= object, # input data structure
          analysisId=analysisId, # analysis to be picked from predictions database
          trait= trait, # traits to include in the index
          desirev = as.numeric(unlist(strsplit(desirev,","))), # vector of desired values
          scaled=scaled, # whether predicted values should be scaled or not
          verbose=FALSE # sh
        )
        myIndex <- resInd$predictions
        mtas <- resInd$status[which(resInd$status$module == "indexD"),"analysisId"]; mtaId <- mtas[length(mtas)]
        myIndex <- myIndex[which(myIndex$analysisId == mtaId),]
        myIndex <- myIndex[with(myIndex, order(-predictedValue)), ]
        selected <- myIndex[1:round(nrow(myIndex)*iProportionTrait),"designation"]
        mydata$selected[which(mydata$designation %in% selected)] <- "yes"
        # update mm
        mm2 <- stats::aggregate(predictedValue~trait, data=mydata[which(mydata$selected == "yes"),], FUN=mean, na.rm=TRUE)
      }
      
      if(length(as.numeric(unlist(strsplit(meanGroupPredictionsRadar,",")))) == nrow(mm)){
        p <- ggplot2::ggplot(mydata, ggplot2::aes(x=predictedValue, color=selected, fill = selected)) +
          ggplot2::geom_histogram() +
          ggplot2::ylab("Frequency") +
          ggplot2::geom_rug(sides="t", length = ggplot2::unit(0.3, "cm")) +
          ggplot2::facet_wrap(~trait, ncol=3, scales = "free") +
          ggplot2::geom_vline(data = mm, ggplot2::aes(xintercept = predictedValue), linetype="dotdash", col="red") +
          ggplot2::geom_vline(data = mm2, ggplot2::aes(xintercept = predictedValue), linetype="dotdash", col="cadetblue")
      }else{
        p <- ggplot2::ggplot(mydata, ggplot2::aes(x=predictedValue, color=selected, fill = selected)) +
          ggplot2::geom_histogram() +
          ggplot2::ylab("Frequency") +
          ggplot2::geom_rug(sides="t", length = ggplot2::unit(0.3, "cm")) +
          ggplot2::facet_wrap(~trait, ncol=3, scales = "free") +
          ggplot2::geom_vline(data = mm, ggplot2::aes(xintercept = predictedValue), linetype="dotdash", col="red")
      }
      print(htmltools::tagList(plotly::ggplotly(p)))
    }
    
  }else{
    
    shiny::renderPlot({ # plotly::renderPlotly({
      ## if user provides threshold values let them know the selection differentials
      if(length(as.numeric(unlist(strsplit(meanGroupPredictionsRadar,",")))) == nrow(mm)){
        ## add product profile means'
        mm$threshold <- as.numeric(unlist(strsplit(meanGroupPredictionsRadar,","))) + mm[,2]
        mm$desire <- mm$threshold - mm$predictedValue
        # index calculation
        resInd <- cgiarPipeline::indexDesire(
          phenoDTfile= object, # input data structure
          analysisId=analysisId, # analysis to be picked from predictions database
          trait= trait, # traits to include in the index
          desirev = as.numeric(unlist(strsplit(desirev,","))), # vector of desired values
          scaled=scaled, # whether predicted values should be scaled or not
          verbose=FALSE # sh
        )
        myIndex <- resInd$predictions
        mtas <- resInd$status[which(resInd$status$module == "indexD"),"analysisId"]; mtaId <- mtas[length(mtas)]
        myIndex <- myIndex[which(myIndex$analysisId == mtaId),]
        myIndex <- myIndex[with(myIndex, order(-predictedValue)), ]
        selected <- myIndex[1:round(nrow(myIndex)*input$proportionTrait),"designation"]
        mydata$selected[which(mydata$designation %in% selected)] <- "yes"
        # update mm
        mm2 <- stats::aggregate(predictedValue~trait, data=mydata[which(mydata$selected == "yes"),], FUN=mean, na.rm=TRUE)
      }
      
      if(length(as.numeric(unlist(strsplit(meanGroupPredictionsRadar,",")))) == nrow(mm)){
        p <- ggplot2::ggplot(mydata, ggplot2::aes(x=predictedValue, color=selected, fill = selected)) +
          ggplot2::geom_histogram() +
          ggplot2::ylab("Frequency") +
          ggplot2::geom_rug(sides="t", length = ggplot2::unit(0.3, "cm")) +
          ggplot2::facet_wrap(~trait, ncol=3, scales = "free") +
          ggplot2::geom_vline(data = mm, ggplot2::aes(xintercept = predictedValue), linetype="dotdash", col="red") +
          ggplot2::geom_vline(data = mm2, ggplot2::aes(xintercept = predictedValue), linetype="dotdash", col="cadetblue")
      }else{
        p <- ggplot2::ggplot(mydata, ggplot2::aes(x=predictedValue, color=selected, fill = selected)) +
          ggplot2::geom_histogram() +
          ggplot2::ylab("Frequency") +
          ggplot2::geom_rug(sides="t", length = ggplot2::unit(0.3, "cm")) +
          ggplot2::facet_wrap(~trait, ncol=3, scales = "free") +
          ggplot2::geom_vline(data = mm, ggplot2::aes(xintercept = predictedValue), linetype="dotdash", col="red")
      }
      # fig2 <-  plotly::ggplotly(p)
      # fig2
      p
    })
    
  }
  
  
  
}

```

### References of methods used

Pesek, J., & Baker, R. J. (1969). Desired improvement in relation to selection indices. Canadian journal of plant science, 49(6), 803-804.

Ceron-Rojas, J. J., & Crossa, J. (2018). Linear selection indices in modern plant breeding (p. 256). Springer Nature.

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

<p>&nbsp;</p>

