---
title: "Population Structure Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
  toDownload: FALSE
---
  
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
options(knitr.kable.NA = '')
# loading necessary R packages ####
## data manipulation
# library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric
## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}
# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, ...) {
  
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  
  table <- DT::datatable(data = DT, 
                         filter = "top", 
                         options = list(autoWidth = TRUE, 
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'excel'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35)),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r, include=FALSE}
# Init Step to make sure that the dependencies are loaded
htmltools::tagList(printTable(mtcars))
htmltools::tagList(ggplotly(ggplot()))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
```

### Summary diversity

The following table allows you to inspect the principal statistics for the diversity analysis.

<p>&nbsp;</p>

```{r, results='asis'}
suppressWarnings(tryCatch({
  if(file.exists(normalizePath("./data/resultPopStr.RData"))){
    load(normalizePath("./data/resultPopStr.RData"))   
  }else{
    load("resultPopStr.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

    seedatosum<-as.data.frame(result[["PopStr"]][["SummaryDiversityAnalysis"]])
    colnames(seedatosum)=c("Parameter","Value")
    seedatosum[,2]=as.numeric(seedatosum[,2])
if("params" %in% ls()){ # we will download the document
    DT::formatRound(DT::datatable(
      seedatosum,
      rownames = FALSE,
      class = 'cell-border stripe',
      extensions = 'Buttons',
      options = list(dom = 'Blfrtip',scrollX = TRUE,buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                     lengthMenu = list(c(10,20,50,-1), c(10,20,50,'All')))
    ),"Value",5)
  }else{
    DT::renderDT({DT::formatRound(DT::datatable(
      seedatosum,
      rownames = FALSE,
      class = 'cell-border stripe',
      extensions = 'Buttons',
      options = list(dom = 'Blfrtip',scrollX = TRUE,buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                     lengthMenu = list(c(10,20,50,-1), c(10,20,50,'All')))
    ),"Value",5)
    }, server = FALSE)
  }
  
```

<p>&nbsp;</p>

### Percent of missing values
  
  This plot allows the user to visualize the distribution of percent of missing values for genotypes and for markers.

<p>&nbsp;</p>
```{r, results='asis'}
suppressWarnings(tryCatch({
  if(file.exists(normalizePath("./data/resultPopStr.RData"))){
    load(normalizePath("./data/resultPopStr.RData"))
  }else{
    load("resultPopStr.RData")	
  }
}, error = function(e) {
  shinyjs::hide()
}))

      NAmark=result[["PopStr"]][["CalculusPerMarker"]][["%NA"]]
      NAgen=result[["PopStr"]][["CalculusPerGenotype"]][["%NA"]]

if("params" %in% ls()){ # we will download the document
      fig <- plotly::plot_ly(y = ~NAmark, type = "box", name="missing values in markers")
      fig <- fig %>% plotly::add_trace(y = ~NAgen, name="missing values in genotypes")
      fig <- fig %>% plotly::layout(yaxis = list(title = 'Percent'),height = "550px",width = "750px")
      fig
}else{
	plotly::renderPlotly({
      fig <- plotly::plot_ly(y = ~NAmark, type = "box", name="missing values in markers")
      fig <- fig %>% plotly::add_trace(y = ~NAgen, name="missing values in genotypes")
      fig <- fig %>% plotly::layout(yaxis = list(title = 'Percent'),height = "550px",width = "750px")
      fig
	})
}
```

<p>&nbsp;</p>

<p>&nbsp;</p>

### AMOVA

 Is a powerful tool that can help support hypotheses of population structure due to clonal reproduction or isolation without making assumptions about Hardy-Weinberg equilibrium. We have shown that we can reject the Ho
 of random mating between the two populations and have strong evidence that these populations are significantly differentiated at all stratifications.

<p>&nbsp;</p>

```{r, results='asis'}
suppressWarnings(tryCatch({
  if(file.exists(normalizePath("./data/resultPopStr.RData"))){
    load(normalizePath("./data/resultPopStr.RData"))   
  }else{
    load("resultPopStr.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

    seeamova<-as.data.frame(result[["PopStr"]][["AMOVA"]])
    colnames(seeamova)=c("source","Df","SumOfSqs","MS","F","Pr.F","Sigma" ,"PercVar","Phi")
	if("params" %in% ls()){ # we will download the document
		DT::formatRound(DT::datatable(
      seeamova,
      rownames = FALSE,
      class = 'cell-border stripe',
      extensions = 'Buttons',
      options = list(dom = 'Blfrtip',scrollX = TRUE,buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                     lengthMenu = list(c(10,20,50,-1), c(10,20,50,'All')))
    ),c("SumOfSqs","MS","F","Pr.F","Sigma" ,"PercVar","Phi"),5)
	}else{
		DT::renderDT({DT::formatRound(DT::datatable(
      seeamova,
      rownames = FALSE,
      class = 'cell-border stripe',
      extensions = 'Buttons',
      options = list(dom = 'Blfrtip',scrollX = TRUE,buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                     lengthMenu = list(c(10,20,50,-1), c(10,20,50,'All')))
    ),c("SumOfSqs","MS","F","Pr.F","Sigma" ,"PercVar","Phi"),5)
    }, server = FALSE)
	}
    

```

<p>&nbsp;</p>

### Multidimensional scaling plot
  
  This plot allows the user to visualize the closeness or distance between genotypes to check if the population structure was as expected.

<p>&nbsp;</p>
```{r, results='asis'}
suppressWarnings(tryCatch({
  if(file.exists(normalizePath("./data/resultPopStr.RData"))){
    load(normalizePath("./data/resultPopStr.RData"))
	load(normalizePath("./data/parmDendMDS.RData"))
  }else{
    load("resultPopStr.RData")
	load("parmDendMDS.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

      uno=as.data.frame(result[["PopStr"]][["Plots"]][[2]])
	if(length(names(uno))>5){catv<-names(uno)[6]}else{catv <-"GroupClust"}
	eti="Gen"

	txlab <- paste0('Factor 1 (',result[["PopStr"]][["Plots"]][[1]][1],'%)')
	tylab <- paste0('Factor 2 (',result[["PopStr"]][["Plots"]][[1]][2],'%)')
	tzlab <- paste0('Factor 3 (',result[["PopStr"]][["Plots"]][[1]][3],'%)')

	txlab2=txlab
	if(plist[[1]][[1]]=="Factor2") txlab2=tylab
	if(plist[[1]][[1]]=="Factor3") txlab2=tzlab
	tylab2=txlab
	if(plist[[1]][[2]]=="Factor2") tylab2=tylab
	if(plist[[1]][[2]]=="Factor3") tylab2=tzlab

if("params" %in% ls()){ # we will download the document
	p<-plotly::plot_ly(data=uno,x=uno[,plist[[1]][[1]]],y=uno[,plist[[1]][[2]]],color=uno[,catv],
                   type="scatter",mode="markers",colors = plist[[1]][[3]],xaxis=F, yaxis=F,
                   text=uno[,eti],marker=list(size=plist[[1]][[4]]))
	#color de fondo del grafico
	p<-p %>% plotly::layout(plot_bgcolor=plist[[1]][[5]],autosize=F, height = "550px",width = "750px")
	#titulo y etiquetas ejes
	p<-p %>% plotly::layout(title=plist[[1]][[6]],titlefont=list(size=plist[[1]][[7]],color=plist[[1]][[8]]), xaxis = list(title = txlab2, titlefont=list(size=plist[[1]][[9]],color=plist[[1]][[10]])),
                        yaxis = list(title = tylab2,titlefont=list(size=plist[[1]][[9]],color=plist[[1]][[10]])))
	p
}else{
	plotly::renderPlotly({
		p<-plotly::plot_ly(data=uno,x=uno[,plist[[1]][[1]]],y=uno[,plist[[1]][[2]]],color=uno[,catv],
                   type="scatter",mode="markers",colors = plist[[1]][[3]],xaxis=F, yaxis=F,
                   text=uno[,eti],marker=list(size=plist[[1]][[4]]))
	#color de fondo del grafico
	p<-p %>% plotly::layout(plot_bgcolor=plist[[1]][[5]],autosize=F, height = "550px",width = "750px")
	#titulo y etiquetas ejes
	p<-p %>% plotly::layout(title=plist[[1]][[6]],titlefont=list(size=plist[[1]][[7]],color=plist[[1]][[8]]), xaxis = list(title = txlab2, titlefont=list(size=plist[[1]][[9]],color=plist[[1]][[10]])),
                        yaxis = list(title = tylab2,titlefont=list(size=plist[[1]][[9]],color=plist[[1]][[10]])))
	p
})
}
```

<p>&nbsp;</p>

<p>&nbsp;</p>

### Dendogram
  
  This plot allows the user to visualize in one way the groupings of genotypes.

<p>&nbsp;</p>
```{r, results='asis',fig.width=14, fig.height=14}
suppressWarnings(tryCatch({
  if(file.exists(normalizePath("./data/resultPopStr.RData"))){
    load(normalizePath("./data/resultPopStr.RData"))
	load(normalizePath("./data/parmDendMDS.RData"))
  }else{
    load("resultPopStr.RData")
	load("parmDendMDS.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

      data=as.data.frame(result[["PopStr"]][["Plots"]][[2]])
      if(length(names(data))>5){catv<-names(data)[6]}else{catv <-"GroupClust"}
	  txlab <- paste0('Factor 1 (',result[["PopStr"]][["Plots"]][[1]][1],'%)')
	  tylab <- paste0('Factor 2 (',result[["PopStr"]][["Plots"]][[1]][2],'%)')
	  tzlab <- paste0('Factor 3 (',result[["PopStr"]][["Plots"]][[1]][3],'%)')
	  
      info<- data[,c("Gen",catv)]
      info<- cbind(ID=info$Gen,info)
      names(info)=c("ID","Gen","Group")
      tree=result[["PopStr"]][["Plots"]][[3]]
	  
if("params" %in% ls()){ # we will download the document
	if (plist[[2]][[1]]=="rectangular"){
			plot(tree, type = "phylogram", cex = plist[[2]][[2]], label.offset = plist[[2]][[3]], show.tip.label = TRUE, edge.color = "black", edge.width =plist[[2]][[4]], edge.lty = 1,tip.color = plist[[2]][[5]][info$Group])
			legend(plist[[2]][[6]], legend=levels(info$Group), fill=plist[[2]][[5]],box.lty=0)
	}else{
			plot(tree, type = "fan", cex = plist[[2]][[2]], label.offset = plist[[2]][[3]], show.tip.label = TRUE, edge.color = "black", edge.width =plist[[2]][[4]], edge.lty = 1,tip.color = plist[[2]][[5]][info$Group])
			legend(plist[[2]][[6]], legend=levels(info$Group), fill=plist[[2]][[5]],box.lty=0)
	}
}else{
	if (plist[[2]][[1]]=="rectangular"){
		shiny::renderPlot({
			plot(tree, type = "phylogram", cex = plist[[2]][[2]], label.offset = plist[[2]][[3]], show.tip.label = TRUE, edge.color = "black", edge.width =plist[[2]][[4]], edge.lty = 1,tip.color = plist[[2]][[5]][info$Group])
			legend(plist[[2]][[6]], legend=levels(info$Group), fill=plist[[2]][[5]],box.lty=0)
		},width = 950,height = 850)
	}else{
		shiny::renderPlot({
			plot(tree, type = "fan", cex = plist[[2]][[2]], label.offset = plist[[2]][[3]], show.tip.label = TRUE, edge.color = "black", edge.width =plist[[2]][[4]], edge.lty = 1,tip.color = plist[[2]][[5]][info$Group])
			legend(plist[[2]][[6]], legend=levels(info$Group), fill=plist[[2]][[5]],box.lty=0)
		},width = 950,height = 850)
	}
}
```


