---
title: "Optimal Cross Selection Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
 toDownload: FALSE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
```

```{r printfxn, include=FALSE}
# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                         numericColNames = NULL, numRound = 3, 
                         scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,...) {
    oldnames <- colnames(DT)
    DT <- data.frame(lapply(X = DT, 
                            FUN = function(x) {
                              if(is.numeric(x)){
                                round(x, numRound)
                              } else {
                                x
                              }
                            }))
    colnames(DT) <- oldnames
    table <- DT::datatable(data = DT, 
                           colnames = colNames,
                           filter = "top", 
                           options = list(autoWidth = autoWidthOpt,
                                          dom = 'l<<t>pB>', 
                                          buttons = c('copy', 'csv', 'excel', 'print'),
                                          pageLength = pageLength,
                                          searchHighlight = TRUE,
                                          lengthMenu = c(7, 14, 21, 28, 35),
                                          scrollX = scrollXOpt),
                           extensions = 'Buttons',
                           rownames = FALSE,
                           ...)
    if (length(numericColNames) > 0){
      table <- table %>% DT::formatRound(columns = numericColNames,
                                     digits = numRound)
    }
    
    table
  }
```

```{r plot_init, include = FALSE}
# Init Step to make sure that the dependencies are loaded
# was done to enable proper printing of the tab header and the graphs within the for loop below using print(htmltools::tagList(ggplotly(...)))
htmltools::tagList(plotly::ggplotly(ggplot2::ggplot()))
htmltools::tagList(printTable(mtcars))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}

```

### Objectives of Optimal Cross Selection

The objective of this dashboard is to help scientist to understand the following points:

1. Optimal crosses for the trade off between performance and genetic variance specified

2. Expected performance of crosses by parent

3. Performance of parents selected to go into the crossing block

4. Number of crosses proposed per parent selected to go into the crossing block

5. Difference in mean and variance observed for different degree runs

Understanding these data features should allow the scientist to identify which crosses should be performed in the crossing block. It should also allow the scientist to assess the relationship between the degree selected (aggressiveness) and the number of parents selected and times that each parent shows up.  

Keep in mind that the suggested crosses may over-represent some parents depending on the degrees selected. Some users tend to ask if they can constraint the number of times a parent should be used. Although this is possible, this implies that the user is looking for a lower sub-optimal solution since the specification of that constraint implies that the user is looking for keeping even more variance (higher degrees). 

### Cross performance table

The following table provides an overview of the performance of the predicted/proposed crosses across different traits. There should be as many crosses as requested in the interface. If you requested different number of crosses or degrees, the treatment column allows you to visualize the results from different runs. Remember that zero degree implies more interest in gain and 90 degrees implies more interest in genetic variance.

<p>&nbsp;</p>

```{r, results='asis'}
suppressWarnings(tryCatch({
  if(file.exists("./outputs/resultOcs.RData")){
    load("./outputs/resultOcs.RData")
  }else{
    load("resultOcs.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

if (is.null(result)){
  # invisible(NULL)
} else {
  mydata = result$predictions
  idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
  idOcs <- idOcs[length(idOcs)]
  ocsTrait <- unique(result$modeling[which(result$modeling$analysisId == idOcs & result$modeling$parameter == "ocsFormula"),"trait"]);
  mydata <- mydata[which(mydata$analysisId %in% c(idOcs)),]
  
  treatments <- unique(mydata$environment)
  wideList <- list()
  for(iTreatment in treatments){
    wideList[[iTreatment]] <- stats::reshape(mydata[which(mydata$environment == iTreatment),c(c("designation"),"trait",c("predictedValue"))], 
                                             direction = "wide", idvar = c("designation"),
                                             timevar = "trait", v.names = c("predictedValue"), sep= "_")
    wideList[[iTreatment]]$treatment <- iTreatment
    colnames(wideList[[iTreatment]]) <- gsub("predictedValue_","",colnames(wideList[[iTreatment]]))
  }
  wide <- do.call(rbind, wideList)
  rownames(wide) <- NULL
  myTypes <- unlist(lapply(wide, class))
  numeric.output0 <- names(myTypes[which( myTypes %in% c("numeric","integer") )])
  
  # cat("The trait used for OCS was: ",ocsTrait)
  
  if("params" %in% ls()){ # we will download the document
    printTable(wide, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(wide),
           numericColNames = numeric.output0,
           numRound = 4)
  }else{
    
    DT::renderDT(printTable(wide,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(wide),
                        numericColNames = numeric.output0,
                        numRound = 4), server = FALSE)
  }
  
}
```

<p>&nbsp;</p>

### Cross performance per parent

The following boxplot allows you to see the performance of the different crosses predicted/proposed (dots in the boxplot) grouped by the different parents selected to go into the crossing block (x-axis).

<p>&nbsp;</p>
```{r}
if (is.null(result)){
  # invisible(NULL)
} else {
  predictions <- result$predictions
  idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
  idOcs <- idOcs[length(idOcs)]
  myTraits <- predictions[which(predictions$analysisId == idOcs),"trait"]
  
  if("params" %in% ls()){  # if user will download the report
    shinyjs::hide()
  }else{ # if user will just visualize the report in the website
    
    shiny::selectInput(ns("traitFilterPredictions2D2"), 
                       label = "Trait:", 
                       multiple = TRUE,
                       selected = unique(myTraits),
                       choices = unique(myTraits)
    )
    
    
  }
  
}
```

```{r, results='asis'}
library(dplyr)
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){  # if user will download the report
    
    cat(paste0("###  {.tabset .tabset-pills}    \n\n  "))
    for(iTrait in unique(myTraits)){
      cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
      mydata = result$predictions
      idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
      idOcs <- idOcs[length(idOcs)]
      mydata <- mydata[which(mydata$analysisId %in% c(idOcs)),]
      ## get the required traits
      mydata <- mydata[which(mydata$trait %in% c(iTrait)),]
      ##
      mydataF <- mydataM <- mydata
      mydataF <- mydataF[,-which(colnames(mydataF)=="mother")]
      colnames(mydataF)[which(colnames(mydataF)=="father")] <- "parent"
      mydataM <- mydataM[,-which(colnames(mydataM)=="father")]
      colnames(mydataM)[which(colnames(mydataM)=="mother")] <- "parent"
      mydata2 <- rbind(mydataM,mydataF[,colnames(mydataM)])
      
      p <- ggplot2::ggplot(mydata2, ggplot2::aes(x=reorder(parent, -predictedValue), y=predictedValue, label=designation) ) + 
        ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0), ggplot2::aes(colour=parent), alpha=0.9) +
        ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=parent)) +
        ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
                       axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
        ggplot2::labs(x="", y="Trait value") #+ ggplot2::facet_grid(trait~environment, scales = "free")
      print(htmltools::tagList(plotly::ggplotly(p)))
      
    }
    
  }else{ # if user will just visualize the report in the website
    plotly::renderPlotly({
      mydata = result$predictions
      idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
      idOcs <- idOcs[length(idOcs)]
      mydata <- mydata[which(mydata$analysisId %in% c(idOcs)),]
      ## get the required traits
      mydata <- mydata[which(mydata$trait %in% c(input$traitFilterPredictions2D2)),]
      ##
      mydataF <- mydataM <- mydata
      mydataF <- mydataF[,-which(colnames(mydataF)=="mother")]
      colnames(mydataF)[which(colnames(mydataF)=="father")] <- "parent"
      mydataM <- mydataM[,-which(colnames(mydataM)=="father")]
      colnames(mydataM)[which(colnames(mydataM)=="mother")] <- "parent"
      mydata2 <- rbind(mydataM,mydataF[,colnames(mydataM)])
      
      
      p <- ggplot2::ggplot(mydata2, ggplot2::aes(x=reorder(parent, -predictedValue), y=predictedValue, label=designation) ) + 
        ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0), ggplot2::aes(colour=parent), alpha=0.9) +
        ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=parent)) +
        ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
                       axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
        ggplot2::labs(x="", y="Trait value") + ggplot2::facet_grid(trait~environment, scales = "free")# +
      # ggplot2::guides(color=ggplot2::guide_legend(""))
      
      fig <-  plotly::ggplotly(p)#, height = get_h()*2);
      fig
    })
  }
  
}
```

### Parent performance table

This table shows the performance of the parents selected by the OCS algorithm. The parental means are extracted from the multi-trial analysis results.

<p>&nbsp;</p>

```{r}

if (is.null(result)){
  # invisible(NULL)
} else {
  mydata = result$predictions
  idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
  idOcs <- idOcs[length(idOcs)]
  # get parents selected in crosses
  mydataOcs <- mydata[which(mydata$analysisId %in% c(idOcs)),]
  parentsSelected <- unique(c(mydataOcs$mother, mydataOcs$father))
  # what was the input of ocs? index or mta?
  idInputOcs <- result$modeling[which(result$modeling$analysisId == idOcs & result$modeling$trait == "inputObject"), "value"]
  moduleInputOcs <- result$status[which(result$status$analysisId == idInputOcs),"module"];
  if(moduleInputOcs != "mta"){ # we keep looking for mta since we had an index
    idInputOcs <- result$modeling[which(result$modeling$analysisId == idInputOcs & result$modeling$trait == "inputObject"), "value"]
    moduleInputOcs <- result$status[which(result$status$analysisId == idInputOcs),"module"];
  }else{} # we stop
  
  mydata <- mydata[which(mydata$analysisId %in% c(idInputOcs)),]
  wideParents <- stats::reshape(mydata[,c(c("designation"),"trait",c("predictedValue"))], direction = "wide", idvar = c("designation"),
                                timevar = "trait", v.names = c("predictedValue"), sep= "_")
  colnames(wideParents) <- gsub("predictedValue_","",colnames(wideParents))
  numeric.output <- colnames(wideParents)[-c(1)]
  wideParents <- wideParents[which(wideParents$designation %in% parentsSelected),]
  rownames(wideParents) <- NULL
  
  if("params" %in% ls()){  # if user will download the report
    printTable(wideParents, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(wideParents),
           numericColNames = numeric.output,
           numRound = 4)
  }else{ # if user will just examinate the report on the website
    DT::renderDT(printTable(wideParents,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(wideParents),
                        numericColNames = numeric.output,
                        numRound = 4), server = FALSE)
  }
  
}
```

<p>&nbsp;</p>

### Number of crosses per parent

The following barplot allows you to assess the number of times (y-axis) that a given parent (x-axis) is suggested to be used in the crossing block for certain OCS treatment (number of crosses by degrees specifications).

<p>&nbsp;</p>

```{r}
if (is.null(result)){
  # invisible(NULL)
} else {
  predictions <- result$predictions
  idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
  idOcs <- idOcs[length(idOcs)]
  myEnvs <- predictions[which(predictions$analysisId == idOcs),"environment"]
  if("params" %in% ls()){  # if user will download the report
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("environ"), 
                       label = "Treatment:", 
                       multiple = TRUE,
                       selected = unique(myEnvs),
                       choices = unique(myEnvs)
    )
  }
}
```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){  # if user will download the report
    
    cat(paste0("###  {.tabset .tabset-pills}    \n\n  "))
    counter <- 1
    for(iEnv in unique(myEnvs)){
      cat("\n\n####", iEnv, "{.tabset .tabset-pills}       \n\n")
      mydata = result$predictions
      idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
      idOcs <- idOcs[length(idOcs)]
      ocsTrait <- result$modeling[which(result$modeling$analysisId == idOcs & result$modeling$parameter == "ocsFormula"),"trait"];
      mydata <- mydata[which(mydata$analysisId %in% c(idOcs[length(idOcs)])),]
      mydata <- mydata[which(mydata$trait == ocsTrait),]
      mydata <- mydata[which(mydata$environment %in% iEnv),]
      
      treatments <- unique(mydata$environment)
      wideList <- list()
      for(iTreatment in treatments){
        overlayCounts <- with(mydata[which(mydata$environment == iTreatment),],enhancer::overlay(mother,father))
        countsCol <- data.frame(times=apply(overlayCounts,2,sum))
        countsCol$parent <- as.factor(rownames(countsCol))
        countsCol$treatment <- iTreatment
        wideList[[iTreatment]] <- countsCol
      }
      countsCol <- do.call(rbind, wideList)
      
      p2 <- ggplot2::ggplot(data=countsCol, ggplot2::aes(x=parent, y=times)) +
        ggplot2::geom_bar(stat="identity", fill="steelblue")+
        ggplot2::geom_text(ggplot2::aes(label=times), vjust=6, size=3.5)+
        ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45)) + ggplot2::facet_grid(~treatment) + 
        ggplot2::labs(x="Parent", y="Times used in a cross") 
      print(htmltools::tagList(plotly::ggplotly(p2)))
      cat("\n")
    }
    cat("\n")
  }else{
    
    plotly::renderPlotly({
      
      mydata = result$predictions
      idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
      idOcs <- idOcs[length(idOcs)]
      ocsTrait <- result$modeling[which(result$modeling$analysisId == idOcs & result$modeling$parameter == "ocsFormula"),"trait"];
      mydata <- mydata[which(mydata$analysisId %in% c(idOcs[length(idOcs)])),]
      mydata <- mydata[which(mydata$trait == ocsTrait),]
      mydata <- mydata[which(mydata$environment %in% input$environ),]
      
      treatments <- unique(mydata$environment)
      wideList <- list()
      for(iTreatment in treatments){
        overlayCounts <- with(mydata[which(mydata$environment == iTreatment),],enhancer::overlay(mother,father))
        countsCol <- data.frame(times=apply(overlayCounts,2,sum))
        countsCol$parent <- as.factor(rownames(countsCol))
        countsCol$treatment <- iTreatment
        wideList[[iTreatment]] <- countsCol
      }
      countsCol <- do.call(rbind, wideList)
      
      p2 <- ggplot2::ggplot(data=countsCol, ggplot2::aes(x=parent, y=times)) +
        ggplot2::geom_bar(stat="identity", fill="steelblue")+
        ggplot2::geom_text(ggplot2::aes(label=times), vjust=6, size=3.5)+
        ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45)) + ggplot2::facet_grid(~treatment) + 
        ggplot2::labs(x="Parent", y="Times used in a cross") 
      fig2 <-  plotly::ggplotly(p2)#, height = get_h()); 
      fig2
      
    })
    
  }
  
}
```

<p>&nbsp;</p>

### Treatment comparison

The following table summarizes the trait means for the different OCS runs (number of crosses and degrees treatments) in case you want to compare what would have been the resulting means and variances under different degrees or number of crosses. 

<p>&nbsp;</p>

```{r}

if (is.null(result)){
  # invisible(NULL)
} else {
  
  idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
  idOcs <- idOcs[length(idOcs)]
  metrics <- result$metrics[which(result$metrics$analysisId == idOcs),]
  metrics <- metrics[which(metrics$parameter == "meanValue"),]
  metrics <- metrics[with(metrics, order(trait)), c("trait","environment","parameter","value","stdError")]
  colnames(metrics) <- cgiarBase::replaceValues(colnames(metrics), Search = "environment", Replace = "treatment")
  rownames(metrics) <- NULL
  numeric.output2 <- c("value","stdError")
  
  if("params" %in% ls()){  # if user will download the report
    printTable(metrics, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(metrics),
           numericColNames = numeric.output2,
           numRound = 4)
  }else{ # if user will just examinate the report on the website
    DT::renderDT(printTable(metrics,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(metrics),
                        numericColNames = numeric.output2,
                        numRound = 4), server = FALSE)
  }
  
}
```

<p>&nbsp;</p>

### References of methods used

Kinghorn, B. (1999). 19. Mate Selection for the tactical implementation of breeding programs. Proceedings of the Advancement of Animal Breeding and Genetics, 13, 130-133.

Woolliams, J. A., Berg, P., Dagnachew, B. S., & Meuwissen, T. H. E. (2015). Genetic contributions and their optimization. Journal of Animal Breeding and Genetics, 132(2), 89-99.

https://alphagenes.roslin.ed.ac.uk/wp/wp-content/uploads/2019/05/01_OptimalContributionSelection.pdf?x44213

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

https://github.com/gaynorr/QuantGenResources

<p>&nbsp;</p>




