---
title: "Advancement Meeting Report"
output: html_document
params:
  toDownload: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = TRUE
)

library(DT)
library(magrittr)
library(shiny)
library(ggplot2)

load("resultAdvMeeting.RData")
```

```{r, include=FALSE}
printTable <- function(DT, pageLength = 25,
                       numericColNames = NULL, numRound = 3,
                       scrollXOpt = TRUE, filt = "top", ...) {

  oldnames <- colnames(DT)

  DT <- data.frame(lapply(DT, function(x) {
    if (is.numeric(x)) round(x, numRound) else x
  }))

  colnames(DT) <- oldnames

  table <- DT::datatable(
    data = DT,
    filter = filt,
    options = list(
      dom = "l<<t>pB>",
      buttons = c("copy", "csv", "excel", "print"),
      pageLength = pageLength,
      searchHighlight = TRUE,
      lengthMenu = c(10, 25, 50, 100),
      scrollX = scrollXOpt
    ),
    extensions = "Buttons",
    rownames = FALSE,
    ...
  )

  if (length(numericColNames) > 0) {
    table <- table %>%
      DT::formatRound(columns = numericColNames, digits = numRound)
  }

  table
}
```

### About this analysis

**Advancement Meeting Dashboard** enables multi-stakeholder consensus on product advancement decisions. Each breeder independently completes the Pre-advancement workflow and saves their output. A meeting facilitator merges all stakeholder files and uses this module to compare decisions, discuss disagreements, and record consensus outcomes.

**Inputs used:**

- **Multiple Final Selection Stamps:** One per stakeholder, each containing per-designation selection decisions (SELECTED, NOT SELECTED, REVISE, CHECK).
- **Shared MTA predictions (BLUPs):** The underlying multi-trial analysis providing trait performance data for all candidates.

**Outputs produced:**

- A **Meeting Stamp** recording the consensus decision for every designation, combining auto-assigned unanimous decisions with manually resolved controversial candidates.

---

```{r summary-section, results='asis'}
# Expected variables from .RData:
# comparison_matrix, meeting_decisions_df, mta_stamp, participants, meeting_name,
# trait_directions, predictions_data

if (!exists("meeting_name") || is.null(meeting_name)) meeting_name <- "Unnamed meeting"
if (!exists("participants") || is.null(participants)) participants <- character(0)
if (!exists("comparison_matrix") || is.null(comparison_matrix)) comparison_matrix <- data.frame()
if (!exists("meeting_decisions_df") || is.null(meeting_decisions_df)) meeting_decisions_df <- data.frame()

n_participants <- length(participants)
participant_names <- paste(participants, collapse = ", ")
n_total <- nrow(meeting_decisions_df)

n_selected <- if (n_total > 0 && "decision" %in% colnames(meeting_decisions_df)) {
  sum(meeting_decisions_df$decision == "SELECTED", na.rm = TRUE)
} else { 0 }

n_not_selected <- if (n_total > 0 && "decision" %in% colnames(meeting_decisions_df)) {
  sum(meeting_decisions_df$decision == "NOT SELECTED", na.rm = TRUE)
} else { 0 }

# Discussed and resolved = controversial candidates that received a joint decision
# These are candidates whose controversy_score is strictly between 0 and 1
# (or all REVISE) that were manually resolved
if (n_total > 0 && "controversy_score" %in% colnames(meeting_decisions_df)) {
  n_discussed <- sum(
    meeting_decisions_df$controversy_score > 0 &
    meeting_decisions_df$controversy_score < 1,
    na.rm = TRUE
  )
} else {
  n_discussed <- 0
}

cat(paste0(
  "### Summary\n\n",
  "**Meeting name:** ", meeting_name, "\n\n",
  "- **Number of participants:** ", n_participants, "\n",
  "- **Participants:** ", participant_names, "\n",
  "- **Total candidates evaluated:** ", n_total, "\n",
  "- **Consensus SELECTED:** ", n_selected, "\n",
  "- **Consensus NOT SELECTED:** ", n_not_selected, "\n",
  "- **Discussed and resolved:** ", n_discussed, "\n\n"
))
```

---

### Final Meeting Decisions Table

The table below shows all candidates with their meeting consensus decision, controversy score, and each stakeholder's individual decision. Cells are color-coded by decision status: blue = SELECTED, orange = NOT SELECTED, yellow = REVISE, pink = CHECK.

```{r, include=FALSE}
htmltools::tagList(printTable(mtcars))
```

```{r decisions-table}
# STATUS_COLORS: SELECTED = "#0072B2", NOT SELECTED = "#D55E00",
#                REVISE = "#F9A825", CHECK = "#C2185B"
STATUS_COLORS <- c(
  "SELECTED"     = "#0072B2",
  "NOT SELECTED" = "#D55E00",
  "REVISE"       = "#F9A825",
  "CHECK"        = "#C2185B"
)

# Light background versions for table cells
STATUS_BG_COLORS <- c(
  "SELECTED"     = "#D6EAF8",
  "NOT SELECTED" = "#F5C4A5",
  "REVISE"       = "#FFF3CD",
  "CHECK"        = "#F5DCE9"
)

if (nrow(comparison_matrix) > 0 && nrow(meeting_decisions_df) > 0) {

  # Build display table: designation, meeting_decision, controversy_score, stakeholder columns
  stakeholder_cols <- setdiff(
    colnames(comparison_matrix),
    c("designation", "controversy_score")
  )

  display_df <- data.frame(
    designation = comparison_matrix$designation,
    stringsAsFactors = FALSE
  )

  # Add meeting decision
  display_df$meeting_decision <- meeting_decisions_df$decision[
    match(display_df$designation, meeting_decisions_df$designation)
  ]

  # Add controversy score
  if ("controversy_score" %in% colnames(comparison_matrix)) {
    display_df$controversy_score <- comparison_matrix$controversy_score
  }

  # Add stakeholder columns
  for (sc in stakeholder_cols) {
    display_df[[sc]] <- comparison_matrix[[sc]]
  }

  # Sort by controversy (most controversial first: closest to 0.5)
  if ("controversy_score" %in% colnames(display_df)) {
    display_df <- display_df[
      order(abs(display_df$controversy_score - 0.5)), , drop = FALSE
    ]
  }

  # Render DT with color coding
  dt <- printTable(
    display_df,
    pageLength = 25,
    scrollXOpt = TRUE,
    numericColNames = "controversy_score"
  )

  # Apply color coding to decision columns
  decision_cols <- c("meeting_decision", stakeholder_cols)
  decision_cols <- intersect(decision_cols, colnames(display_df))

  for (col in decision_cols) {
    dt <- DT::formatStyle(
      dt,
      columns = col,
      backgroundColor = DT::styleEqual(
        c("SELECTED", "NOT SELECTED", "CHECK", "REVISE"),
        c(STATUS_BG_COLORS[["SELECTED"]], STATUS_BG_COLORS[["NOT SELECTED"]],
          STATUS_BG_COLORS[["CHECK"]], STATUS_BG_COLORS[["REVISE"]])
      ),
      fontWeight = "bold"
    )
  }

  htmltools::tagList(dt)
} else {
  cat("No meeting decisions data available.\n")
}
```

---

### Agreement Heatmap

This heatmap shows the comparison matrix of stakeholder decisions. Rows represent designations sorted by controversy level (most disagreement at top), columns represent stakeholders. Cell colors indicate each stakeholder's decision.

```{r agreement-heatmap, fig.height=10, fig.width=10}
if (nrow(comparison_matrix) > 0) {

  stakeholder_cols_hm <- setdiff(
    colnames(comparison_matrix),
    c("designation", "controversy_score")
  )

  if (length(stakeholder_cols_hm) > 0) {

    # Sort by controversy (most controversial at top = closest to 0.5)
    hm_df <- comparison_matrix
    if ("controversy_score" %in% colnames(hm_df)) {
      hm_df <- hm_df[order(abs(hm_df$controversy_score - 0.5)), , drop = FALSE]
    }

    # Convert to long format for ggplot
    hm_long <- do.call(rbind, lapply(stakeholder_cols_hm, function(st) {
      data.frame(
        designation = hm_df$designation,
        stakeholder = st,
        decision = hm_df[[st]],
        stringsAsFactors = FALSE
      )
    }))

    # Create ordered factor for designation (preserve controversy sort order)
    hm_long$designation <- factor(
      hm_long$designation,
      levels = rev(hm_df$designation)
    )

    # Ensure decision is a factor with known levels
    hm_long$decision <- factor(
      hm_long$decision,
      levels = c("SELECTED", "NOT SELECTED", "REVISE", "CHECK")
    )

    p_hm <- ggplot(hm_long, aes(x = stakeholder, y = designation, fill = decision)) +
      geom_tile(color = "white", linewidth = 0.5) +
      scale_fill_manual(
        values = STATUS_COLORS,
        name = "Decision",
        drop = FALSE
      ) +
      theme_minimal(base_size = 12) +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = 11),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size = 11),
        legend.title = element_text(size = 12, face = "bold"),
        panel.grid = element_blank()
      ) +
      labs(
        title = "Agreement Heatmap (sorted by controversy)",
        subtitle = "Most controversial candidates at top"
      )

    print(p_hm)
  } else {
    cat("No stakeholder columns found in comparison matrix.\n")
  }
} else {
  cat("No comparison matrix data available for heatmap.\n")
}
```

---

### Trait Distribution per Status

This plot compares the distribution of trait values across the meeting consensus decision groups (SELECTED, NOT SELECTED, CHECK). Each trait is shown as a horizontal section. Within each section, three bars summarize each group's performance: the thin line spans the full range (min to max), the thick rectangle covers the interquartile range (Q1 to Q3), the diamond marks the mean, and the vertical line marks the median. An arrow next to each trait name indicates the breeding direction (→ = higher is better, ← = lower is better).

```{r trait-distribution-plot, fig.height=14, fig.width=12}
# STATUS colors for trait distribution plot (3 groups only)
status_colors_plot <- c(
  "SELECTED"     = "#0072B2",
  "NOT SELECTED" = "#D55E00",
  "CHECK"        = "#C2185B"
)

# Use trait_directions from the RData if available
if (!exists("trait_directions")) trait_directions <- list()
if (!exists("predictions_data") || is.null(predictions_data)) {
  predictions_data <- data.frame()
}

# Build trait data from predictions_data (long format: designation, trait, predictedValue)
# and meeting_decisions_df (designation, decision)

if (nrow(predictions_data) > 0 && nrow(meeting_decisions_df) > 0 &&
    "designation" %in% colnames(predictions_data) &&
    "trait" %in% colnames(predictions_data) &&
    "predictedValue" %in% colnames(predictions_data)) {

  # Merge decisions onto predictions
  pred_df <- merge(
    predictions_data[, c("designation", "trait", "predictedValue"), drop = FALSE],
    meeting_decisions_df[, c("designation", "decision"), drop = FALSE],
    by = "designation",
    all.x = FALSE
  )

  # Map REVISE to NOT SELECTED for the distribution plot (only 3 groups)
  pred_df$plot_status <- pred_df$decision
  pred_df$plot_status[pred_df$plot_status == "REVISE"] <- "NOT SELECTED"
  pred_df <- pred_df[pred_df$plot_status %in% c("SELECTED", "NOT SELECTED", "CHECK"), , drop = FALSE]
  pred_df <- pred_df[!is.na(pred_df$predictedValue), , drop = FALSE]

  numeric_trait_cols <- unique(pred_df$trait)

  if (length(numeric_trait_cols) > 0) {

    # Compute summary stats per trait x status
    summary_df <- do.call(rbind, lapply(numeric_trait_cols, function(tr) {
      tr_df <- pred_df[pred_df$trait == tr, , drop = FALSE]
      do.call(rbind, lapply(c("SELECTED", "NOT SELECTED", "CHECK"), function(st) {
        vals <- tr_df$predictedValue[tr_df$plot_status == st]
        if (length(vals) == 0) return(NULL)
        data.frame(
          trait = tr,
          status = st,
          xmin = min(vals, na.rm = TRUE),
          q1 = quantile(vals, 0.25, na.rm = TRUE),
          median = median(vals, na.rm = TRUE),
          mean_val = mean(vals, na.rm = TRUE),
          q3 = quantile(vals, 0.75, na.rm = TRUE),
          xmax = max(vals, na.rm = TRUE),
          n = length(vals),
          stringsAsFactors = FALSE
        )
      }))
    }))
    rownames(summary_df) <- NULL

    # Build direction labels
    dir_labels <- vapply(numeric_trait_cols, function(tr) {
      d <- trait_directions[[tr]]
      if (is.null(d) || identical(d, "Higher is better")) {
        "\u2192 better"
      } else {
        "\u2190 better"
      }
    }, character(1))

    # Create ordered trait factor with direction annotation
    trait_labels <- paste0(numeric_trait_cols, "  (", dir_labels, ")")
    summary_df$trait_label <- factor(
      trait_labels[match(summary_df$trait, numeric_trait_cols)],
      levels = rev(trait_labels)
    )

    # Status factor for consistent ordering
    summary_df$status <- factor(
      summary_df$status,
      levels = c("CHECK", "NOT SELECTED", "SELECTED")
    )

    # Create y positions
    summary_df$y_pos <- as.numeric(summary_df$trait_label) +
      as.numeric(summary_df$status) * 0.25 - 0.5

    # Trait distribution plot
    p <- ggplot(summary_df) +
      # Thin line: full range (min to max)
      geom_segment(
        aes(x = xmin, xend = xmax, y = y_pos, yend = y_pos, color = status),
        linewidth = 0.4, alpha = 0.6
      ) +
      # Thick bar: IQR (Q1 to Q3)
      geom_rect(
        aes(xmin = q1, xmax = q3, ymin = y_pos - 0.08, ymax = y_pos + 0.08, fill = status),
        alpha = 0.7, color = NA
      ) +
      # Mean marker (diamond)
      geom_point(
        aes(x = mean_val, y = y_pos, color = status),
        shape = 18, size = 3
      ) +
      # Median marker (vertical tick)
      geom_segment(
        aes(x = median, xend = median, y = y_pos - 0.08, yend = y_pos + 0.08, color = status),
        linewidth = 1
      ) +

      # Colors
      scale_color_manual(values = status_colors_plot, name = "Group", drop = FALSE) +
      scale_fill_manual(values = status_colors_plot, name = "Group", drop = FALSE) +
      # Facet by trait (free x scales since traits have different units)
      facet_grid(trait_label ~ ., scales = "free", switch = "y", space = "fixed") +
      # Theme
      theme_minimal(base_size = 14) +
      theme(
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.title.x = element_text(size = 13),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        strip.text.y.left = element_text(
          angle = 0, hjust = 1, face = "bold", size = 12
        ),
        strip.placement = "outside",
        legend.position = "bottom",
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 13, face = "bold"),
        panel.spacing = unit(0.5, "lines"),
        plot.caption = element_text(size = 10, color = "#555555")
      ) +
      labs(
        x = "Trait value",
        caption = paste0(
          "Thick bar = IQR (Q1-Q3) | Diamond = mean | ",
          "Vertical line = median | Thin line = range (min-max)"
        )
      )

    print(p)
  } else {
    cat("No numeric trait data available for trait distribution plot.\n")
  }
} else {
  cat("Trait distribution plot requires predictions_data and meeting_decisions_df.\n")
}
```
