---
title: "Marker Assisted Selection Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
  toDownload: FALSE
---

```{r setup, include=FALSE}

# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
options(knitr.kable.NA = '')

# loading necessary R packages ####
## data manipulation
library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric

## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(grafify)
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, 
                       scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,...) {
  oldnames <- colnames(DT)
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  colnames(DT) <- oldnames
  table <- DT::datatable(data = DT, 
                         colnames = colNames,
                         filter = "top", 
                         options = list(autoWidth = autoWidthOpt,
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'csv', 'excel', 'print'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35),
                                        scrollX = scrollXOpt),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% DT::formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r, include=FALSE}
# Init Step to make sure that the dependencies are loaded
htmltools::tagList(printTable(mtcars))
htmltools::tagList(ggplotly(ggplot()))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
```

```{r }

suppressWarnings(tryCatch({
  if(file.exists("./outputs/result.RData")){
    load("./outputs/result.RData")
  }else{
    load("result.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

```


### Table of MAS predictions
<p>&nbsp;</p>

The following table shows the total merit that has been calculated based on the weights assigned for each QTL.

```{r }

if (is.null(result)){
  # invisible(NULL)
} else {
  predictions <- result$predictions
  idQa <- result$status[which(result$status$module %in% c("mas")),"analysisId"];
  idQa <- idQa[length(idQa)]
  predictions <- predictions[which( predictions$analysisId == idQa),]
  # traitsTable <- unique(predictions[,"trait"])
  
  
  # if("params" %in% ls()){ # we will download the document
  #   shinyjs::hide()
  # }else{
  #   shiny::selectInput(ns("traitQa"), 
  #                      label = "Trait:", 
  #                      choices = traitsTable, 
  #   )
  # }
  
}

```


```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  numericalColumns <- c("predictedValue","stdError") 
  
  if("params" %in% ls()){ # we will download the document
    printTable(predictions, autoWidthOpt = TRUE, scrollXOpt = FALSE,
           colNames = colnames(predictions),
           numericColNames = numericalColumns,
           numRound = 4)
  }else{
    DT::renderDT(printTable(predictions,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(predictions),
                        numericColNames = numericalColumns,
                        numRound = 4), server = FALSE)
  }
  
}

```

### Graph of merit
<p>&nbsp;</p>

The following graph shows the total merit that has been calculated based on the weights assigned for each QTL.

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  p <- ggplot(predictions, aes(x=designation, y=predictedValue, text=designation)) + 
    geom_point(size=2, alpha = 0.6 ) + ylab("Merit") + xlab("Designation") +
    theme(axis.text.x=element_blank())
  
  if("params" %in% ls()){ # we will download the document
     print(htmltools::tagList(plotly::ggplotly(p)))
  }else{
    
    plotly::renderPlotly({
      plotly::ggplotly(p)
    })
    
  }
  
}

```

### Graph of QTL frequencies
<p>&nbsp;</p>


### References on methods used

Tukey, J. W. (1977). Exploratory Data Analysis. Section 2C.

McGill, R., Tukey, J. W. and Larsen, W. A. (1978). Variations of box plots. The American Statistician, 32, 12–16. doi:10.2307/2683468.

Velleman, P. F. and Hoaglin, D. C. (1981). Applications, Basics and Computing of Exploratory Data Analysis. Duxbury Press.


<p>&nbsp;</p>





