---
title: "MTA Report"
author: ""
date: "2023-11-03"
output: html_document
params:
 toDownload: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
font.plot_ly <- list(family="Courier New, monospace", size=9)
```

```{r printfxn, include=FALSE}
# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, ...) {
  
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  
  table <- DT::datatable(data = DT, 
                         filter = "top", 
                         options = list(autoWidth = TRUE, 
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'excel'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35)),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r plot_init, include = FALSE}
# Init Step to make sure that the dependencies are loaded
# was done to enable proper printing of the tab header and the graphs within the for loop below using print(htmltools::tagList(ggplotly(...)))
library(ggplot2)
htmltools::tagList(plotly::ggplotly(ggplot2::ggplot()))
htmltools::tagList(printTable(mtcars))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}

```

```{r }

suppressWarnings(tryCatch({
  if(file.exists("./outputs/result.RData")){
    load("./outputs/resultMta.RData")
  }else{
    load("resultMta.RData")
  }
}, error = function(e) {
  # NULL
}))

if (is.null(result)){
  # invisible(NULL)
} else {
  # result$status <- result$status[1:6,]
  idMta <- result$status[which(result$status$module == "mta"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred <- result$predictions
  pred <- pred[pred$analysisId == idMta,]
  traits <- unique(result$modeling[which(result$modeling$analysisId == idMta  & result$modeling$environment == "across"),"trait"]) # unique(pred$trait)
}
```

### Distribution by environment

The following boxplot allows to see the distribution of predicted values (y-axis) in the different environments for each **trait**.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    
  }else{
    selectInput(ns("traitMta"), 
                label = "", 
                choices = traits
    )
  }
  
}

```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"]
  idSta <- idSta[length(idSta)]
  pred.sta <- result$predictions
  pred.sta <- pred.sta[pred.sta$analysisId == idSta,]
  pred.sta.boxplot <- subset(pred.sta, select = c(trait,designation,environment,predictedValue))
  
  if("params" %in% ls()){ # we will download the document
    
    cat(paste0("### Distribution by environment  {.tabset .tabset-pills}    \n\n  "))
    
    for(iTrait in traits){
      cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
      
      temp.pred.box <- pred.sta.boxplot[pred.sta.boxplot$trait== iTrait,]
      
      p <- plotly::plot_ly( data = temp.pred.box, y = ~predictedValue, x = ~environment,
                            type = "box", color = ~environment
      ) %>%  plotly::layout(font=font.plot_ly, xaxis=list( title=""), yaxis=list( title="Trait value")  )
      
      print(htmltools::tagList(plotly::ggplotly(p)))
      
    }
    
  }else{
    
    
    plotly::renderPlotly({
      temp.pred.box <- pred.sta.boxplot[pred.sta.boxplot$trait==input$traitMta,]
      p <- plotly::plot_ly( data = temp.pred.box, y = ~predictedValue,
                            x = ~environment,  type = "box", color = ~environment
      ) %>%  plotly::layout( font=font.plot_ly,
                             xaxis=list( title=""), yaxis=list( title="Trait value") )
    })
    
    # temp.pred.box <- pred.sta.boxplot[pred.sta.boxplot$trait==input$traitMta,]
    # p <- temp.pred.box %>% 
    #   ggplot() +
    #   aes(x=environment, y=predictedValue, fill = environment) +
    #   geom_boxplot() +
    #   labs(x = "", y = "Trait value") +
    #   scale_fill_discrete(name = "") +
    #   theme_minimal() +
    #   theme(axis.text.x = element_blank(),
    #         axis.ticks.x = element_blank(),
    #         text = element_text(family = "Courier New, monospace", size=9)) 
    # p
    
  }
}
```

### Distribution across environment

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  result$status <- result$status#[1:6,]
  idMta <- result$status[which(result$status$module == "mta"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred <- result$predictions
  pred <- pred[pred$analysisId == idMta,]
  traits <- unique(pred$trait)
  
  if("params" %in% ls()){ # we will download the document
    
  }else{
    selectInput(ns("traitMta2"),
                label = "",
                choices = traits
    )
  }
}
```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    
    # cat(paste0("### Distribution across environments  {.tabset .tabset-pills}    \n\n  "))
    cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
    
    for(iTrait in traits){
      cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
      
      temp.pred.box <- pred[pred$trait== iTrait,]
      
      p <- ggplot2::ggplot(
        data = temp.pred.box, 
        mapping = ggplot2::aes(
          x = reorder(designation, -predictedValue),
          y = predictedValue,
          col = entryType,
          text = paste(
            paste0('<br>', "designation: ", designation),
            paste0('<br>', "entry type: ", entryType),
            paste0('<br>', "predicted value: ", round(predictedValue, 4)),
            paste0('<br>', "standard error: ", round(stdError, 4)),
            # paste0("<br>95% CI: (", round(predictedValue - CI, 2) , ", ", 
            #        round(predictedValue + CI, 2), ")")))) +
            paste0("<br>predVal +- stdError: (", round(predictedValue - stdError, 2) , ", ", 
                   round(predictedValue + stdError, 4), ")")))) +
        ggplot2::theme_classic() +  ggplot2::geom_point() +
        ggplot2::geom_errorbar(ggplot2::aes(ymax = predictedValue + stdError,
                                            ymin = predictedValue - stdError), position = ggplot2::position_dodge(0.5),
                               width = 0.10) +
        ggplot2::labs(title = "across", x = "designation", y = "predicted value", col = "entry type") +
        ggplot2::theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7))
      if(length(unique(temp.pred.box$designation)) > 200){ # go fixed format
        print(p)
      }else{ # go dynamic
        print(htmltools::tagList(plotly::ggplotly(p)))
      }
    }
    
  }else{
    
    if(length(unique(pred$designation)) > 200){ # go fixed format

      shiny::renderPlot({

        temp.pred.box <- pred[pred$trait== input$traitMta2,]
        p <- ggplot2::ggplot(
          data = temp.pred.box,
          mapping = ggplot2::aes(
            x = reorder(designation, -predictedValue),
            y = predictedValue,
            col = entryType,
            text = paste(
              paste0('<br>', "designation: ", designation),
              paste0('<br>', "entry type: ", entryType),
              paste0('<br>', "predicted value: ", round(predictedValue, 4)),
              paste0('<br>', "standard error: ", round(stdError, 4)),
              # paste0("<br>95% CI: (", round(predictedValue - CI, 2) , ", ",
              #        round(predictedValue + CI, 2), ")")))) +
              paste0("<br>predVal +- stdError: (", round(predictedValue - stdError, 2) , ", ",
                     round(predictedValue + stdError, 4), ")")))) +
          ggplot2::theme_classic() +  ggplot2::geom_point() +
          ggplot2::geom_errorbar(ggplot2::aes(ymax = predictedValue + stdError,
                                              ymin = predictedValue - stdError), position = ggplot2::position_dodge(0.5),
                                 width = 0.10) +
          ggplot2::labs(title = "across", x = "designation", y = "predicted value", col = "entry type") +
          ggplot2::theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7))
        p

      })
    }else{ # go dynamic
      plotly::renderPlotly({

        temp.pred.box <- pred[pred$trait== input$traitMta2,]
        p <- ggplot2::ggplot(
          data = temp.pred.box,
          mapping = ggplot2::aes(
            x = reorder(designation, -predictedValue),
            y = predictedValue,
            col = entryType,
            text = paste(
              paste0('<br>', "designation: ", designation),
              paste0('<br>', "entry type: ", entryType),
              paste0('<br>', "predicted value: ", round(predictedValue, 4)),
              paste0('<br>', "standard error: ", round(stdError, 4)),
              # paste0("<br>95% CI: (", round(predictedValue - CI, 2) , ", ",
              #        round(predictedValue + CI, 2), ")")))) +
              paste0("<br>predVal +- stdError: (", round(predictedValue - stdError, 2) , ", ",
                     round(predictedValue + stdError, 4), ")")))) +
          ggplot2::theme_classic() +  ggplot2::geom_point() +
          ggplot2::geom_errorbar(ggplot2::aes(ymax = predictedValue + stdError,
                                              ymin = predictedValue - stdError), position = ggplot2::position_dodge(0.5),
                                 width = 0.10) +
          ggplot2::labs(title = "across", x = "designation", y = "predicted value", col = "entry type") +
          ggplot2::theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7))
        plotly::ggplotly(p)
      })
    }
    
  }
}
```


<p>&nbsp;</p>

### Genetic correlations between environments

The following graph allows to see the genetic correlations of the different environment for each **trait**.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  idMta <- result$status[which(result$status$module == "mta"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred <- result$predictions
  pred <- pred[pred$analysisId == idMta,]
  traits <- unique(result$modeling[which(result$modeling$analysisId == idMta  & result$modeling$environment == "across"),"trait"])# unique(pred$trait)
  if("params" %in% ls()){ # we will download the document
    
  }else{
    selectInput(ns("traitPredictionsCorrelation"),
                label = "",
                choices = traits
    )
  }
  
}

```


```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"]
  idSta <- idSta[length(idSta)]
  pred.sta <- result$predictions
  pred.sta <- pred.sta[pred.sta$analysisId == idSta,]
  traits <- unique(pred.sta$trait)
  predictions.gcorrE <- subset(pred.sta, select = c(trait,designation,environment,predictedValue))
  envs <- unique(predictions.gcorrE[,"environment"])
  addTextInPlots <- ifelse(length(envs) > 40, FALSE, TRUE)
  if("params" %in% ls()){ # we will download the document
    
    # cat(paste0("### Correlation  {.tabset .tabset-pills}    \n\n  "))
    cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
    
    for(iTrait in traits){
      cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
      
      p.cor.pred.env <- corPlotPredictions(predictions.gcorrE, iTrait, unitOfCorrelation="designation", correlatedAcross="environment", valueForCorrelation ="predictedValue",checkboxCluster=TRUE, checkboxText=addTextInPlots, checkboxAxis= addTextInPlots)
      print(htmltools::tagList(plotly::ggplotly(p.cor.pred.env)))
      
    }
    
  }else{
    
    plotly::renderPlotly({
      p.cor.pred.env <- corPlotPredictions(predictions.gcorrE, input$traitPredictionsCorrelation, unitOfCorrelation="designation", correlatedAcross="environment",
                                           valueForCorrelation="predictedValue",checkboxCluster=TRUE, checkboxText=addTextInPlots,
                                           checkboxAxis= addTextInPlots)
      plotly::layout(p.cor.pred.env,font=font.plot_ly)
    })
    
    
  }
  
}
```

<p>&nbsp;</p>

## Genetic correlations between traits

The following graph allows to see the genetic correlations among traits using across environment estimates.

<p>&nbsp;</p>

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  idMta <- result$status[which(result$status$module == "mta"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred.mta <- result$predictions
  pred.mta <- pred.mta[pred.mta$analysisId == idMta,]
  predictions.gcorrT <- subset(pred.mta, select = c(trait,designation,environment,predictedValue))
  nTraits <- unique(predictions.gcorrE[,"trait"])
  addTextInPlots <- ifelse(length(nTraits) > 40, FALSE, TRUE)
  if(length(nTraits) > 1){
    if("params" %in% ls()){ # we will download the document
      
      p.cor.pred.traits <- corPlotPredictions(predictions.gcorrT, traitPredictionsCorrelation=NULL, unitOfCorrelation="designation", correlatedAcross="trait",
                                              valueForCorrelation="predictedValue",checkboxCluster=TRUE, checkboxText=addTextInPlots, 
                                              checkboxAxis= addTextInPlots)
      print(htmltools::tagList(plotly::ggplotly(p.cor.pred.traits)))
      
    }else{
      
      plotly::renderPlotly({
        p.cor.pred.traits <- corPlotPredictions(predictions.gcorrT, traitPredictionsCorrelation=NULL, unitOfCorrelation="designation", correlatedAcross="trait",
                                                valueForCorrelation="predictedValue",checkboxCluster=TRUE, checkboxText=addTextInPlots, 
                                                checkboxAxis= addTextInPlots)
        plotly::layout(p.cor.pred.traits,font=font.plot_ly)
      })
      
    }
  }else{
    cat("Only one trait fitted. Skipping correlation plot.")
  }
  
}
```

<p>&nbsp;</p>

## Biplot by trait

The following graph allows to see the performance of the genotypes over the enevironments for each **trait**.

<p>&nbsp;</p>

```{r}
if (is.null(result)){
  # invisible(NULL)
} else {
  idMta <- result$status[which(result$status$module == "mta"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred.mta <- result$predictions
  pred.mta <- pred.mta[pred.mta$analysisId == idMta,]
  traits <- unique(result$modeling[which(result$modeling$analysisId == idMta  & result$modeling$environment == "across"),"trait"]) # unique(pred.mta$trait)
  
  if("params" %in% ls()){ # we will download the document
    
  }else{
    selectInput(ns("traitBiplot"),
                label = "",
                choices = traits
    )
  }
}

```

```{r, results='asis'}

if (is.null(result)){
  # invisible(NULL)
} else {
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"]
  idSta <- idSta[length(idSta)]
  pred.sta <- result$predictions
  pred.sta <- pred.sta[pred.sta$analysisId == idSta,]
  
  if("params" %in% ls()){ # we will download the document
    
    # cat(paste0("### Correlation  {.tabset .tabset-pills}    \n\n  "))
    cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
    
    for(iTrait in traits){
      cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
      
      pred.mta.biplot <- reshape(pred.sta[which(pred.sta$trait == iTrait),c("designation","environment","predictedValue")], 
                                 direction = "wide", idvar = "designation", timevar = "environment")
      
      rownames(pred.mta.biplot) <- pred.mta.biplot[,"designation"]
      pred.mta.biplot <- pred.mta.biplot[,-1]
      names(pred.mta.biplot) <- gsub("predictedValue.","",names(pred.mta.biplot))
      
      # keep only 
      missing <- apply(pred.mta.biplot,1, function(x){length(which(is.na(x)))/length(x)})
      pred.mta.biplot <- pred.mta.biplot[which(missing < 0.8),]
      missing <- apply(pred.mta.biplot,2, function(x){length(which(is.na(x)))/length(x)})
      pred.mta.biplot <- pred.mta.biplot[,which(missing < 0.8)]
      pred.mta.biplot <- apply(pred.mta.biplot,2, sommer::imputev)
      # pred.mta.biplot <- na.omit(pred.mta.biplot[,2:min(c(ncol(pred.mta.biplot),5))])
      prin_comp <- prcomp(pred.mta.biplot, rank = 2)
      components <- prin_comp[["x"]]
      components <- data.frame(components)
      components$PC2 <- -components$PC2
      
      explained_variance <- summary(prin_comp)[["sdev"]]
      explained_variance <- explained_variance[1:2]
      comp <- prin_comp[["rotation"]]
      comp[,'PC2'] <- - comp[,'PC2']
      loadings <- comp
      for (i in seq(explained_variance) ){
        loadings[,i] <- comp[,i] * explained_variance[i]
      }
      p.mta.biplot <- plotly::plot_ly (type = 'scatter', mode = 'markers')
      p.mta.biplot <- plotly::add_trace(p.mta.biplot, 
                                        x=components$PC1,
                                        y=components$PC2,
                                        text=rownames(components),
                                        marker=list(color="blue"),
                                        hoverinfo='text',
                                        showlegend=FALSE
      )
      for (i in 1:ncol(pred.mta.biplot)){
        p.mta.biplot <- p.mta.biplot %>%
          plotly::add_segments(x = 0, xend = loadings[i, 1], y = 0, yend = loadings[i, 2], line = list(color = 'red'),inherit = FALSE, showlegend = FALSE)
        if(ncol(pred.mta.biplot) < 20){
          p.mta.biplot <- p.mta.biplot %>%
            plotly::add_annotations(x=loadings[i, 1], y=loadings[i, 2], ax = 0, ay = 0,text = rownames(loadings)[i], xanchor = 'center', yanchor= 'bottom')
        }
      }
      print(htmltools::tagList(plotly::ggplotly(p.mta.biplot)))
      
    }
    
  }else{
    
    plotly::renderPlotly({
      pred.mta.biplot <- reshape(pred.sta[which(pred.sta$trait == input$traitBiplot),c("designation","environment","predictedValue")], 
                                 direction = "wide", idvar = "designation", timevar = "environment")
      
      rownames(pred.mta.biplot) <- pred.mta.biplot[,"designation"]
      pred.mta.biplot <- pred.mta.biplot[,-1]
      names(pred.mta.biplot) <- gsub("predictedValue.","",names(pred.mta.biplot))
      
      # keep only 
      missing <- apply(pred.mta.biplot,1, function(x){length(which(is.na(x)))/length(x)})
      pred.mta.biplot <- pred.mta.biplot[which(missing < 0.8),]
      missing <- apply(pred.mta.biplot,2, function(x){length(which(is.na(x)))/length(x)})
      pred.mta.biplot <- pred.mta.biplot[,which(missing < 0.8)]
      pred.mta.biplot <- apply(pred.mta.biplot,2, sommer::imputev)
      # pred.mta.biplot <- na.omit(pred.mta.biplot[,2:min(c(ncol(pred.mta.biplot),5))])
      
      prin_comp <- prcomp(pred.mta.biplot, rank = 2)
      components <- prin_comp[["x"]]
      components <- data.frame(components)
      components$PC2 <- -components$PC2
      
      explained_variance <- summary(prin_comp)[["sdev"]]
      explained_variance <- explained_variance[1:2]
      comp <- prin_comp[["rotation"]]
      comp[,'PC2'] <- - comp[,'PC2']
      loadings <- comp
      for (i in seq(explained_variance) ){
        loadings[,i] <- comp[,i] * explained_variance[i]
      }
      
      p.mta.biplot <- plotly::plot_ly (type = 'scatter', mode = 'markers')
      p.mta.biplot <- plotly::add_trace(p.mta.biplot, 
                                        x=components$PC1,
                                        y=components$PC2,
                                        text=rownames(components),
                                        marker=list(color="blue"),
                                        hoverinfo='text',
                                        showlegend=FALSE
      )
      
      
      for (i in 1:ncol(pred.mta.biplot)){
        p.mta.biplot <- p.mta.biplot %>%
          plotly::add_segments(x = 0, xend = loadings[i, 1], y = 0, yend = loadings[i, 2], line = list(color = 'red'),inherit = FALSE, showlegend = FALSE)
        if(ncol(pred.mta.biplot) < 20){
          p.mta.biplot <- p.mta.biplot %>%
            plotly::add_annotations(x=loadings[i, 1], y=loadings[i, 2], ax = 0, ay = 0,text = rownames(loadings)[i], xanchor = 'center', yanchor= 'bottom')
        }
      }
      plotly::layout(p.mta.biplot,font=font.plot_ly)
    })
    
  }
  
}

```


<p>&nbsp;</p>

### Predictions 

The following table allows to check the trait predictions in wide format together with the QTL profile in case those are available.

<p>&nbsp;</p>

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  idMta <- result$status[which(result$status$module == "mta"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred.mta <- result$predictions
  pred.mta <- pred.mta[pred.mta$analysisId == idMta,]
  traits <- unique(pred.mta$trait)
  pred.mta <- subset(pred.mta, select = c(trait,designation,environment,predictedValue))
  pred.mta.wide <- reshape(
    data=pred.mta, 
    timevar = c("trait"), 
    idvar = c("environment","designation"), 
    direction="wide"
  )
  names(pred.mta.wide) <- c("designation","environment",traits)
  
  if(!is.null(result$data$qtl)){ # there's QTL information
    qtls <- result$data$qtl
    colnames(qtls) <- cgiarBase::replaceValues(Source = colnames(qtls), Search = result$metadata$qtl$value , Replace = result$metadata$qtl$parameter)
    pred.mta.wide <- merge(pred.mta.wide, qtls, by="designation", all.x = TRUE)
  }
  
  if("params" %in% ls()){ # we will download the document
    printTable(pred.mta.wide)
  }else{
    DT::renderDT({DT::formatRound(DT::datatable(
      pred.mta.wide,
      extensions = 'Buttons',
      rownames = FALSE,
      class = 'cell-border',
      options = list(
        dom = 'Blfrtip',
        scrollY = "400px",
        scrollX = "400px",
        buttons = c('copy','csv','excel','pdf','print'),
        paging = F)
    ),traits)
    })
  }
  
}

```




