---
title: "Predicted Genetic Gain Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
  toDownload: FALSE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
library(dplyr)
library(ggplot2)
```

```{r plot_init, include = FALSE}
# Init Step to make sure that the dependencies are loaded
# was done to enable proper printing of the tab header and the graphs within the for loop below using print(htmltools::tagList(ggplotly(...)))
htmltools::tagList(plotly::ggplotly(ggplot2::ggplot()))

# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}

```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, 
                       scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,...) {
  oldnames <- colnames(DT)
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  colnames(DT) <- oldnames
  table <- DT::datatable(data = DT, 
                         colnames = colNames,
                         filter = "top", 
                         options = list(autoWidth = autoWidthOpt,
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'csv', 'excel', 'print'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35),
                                        scrollX = scrollXOpt),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% DT::formatRound(columns = numericColNames,
                                       digits = numRound)
  }
  
  table
}
```

### Summary metrics

The following table allows you to review the different metrics calculated for the genetic gain analysis. You can filter the metrics using the search bar.

```{r, results='asis'}
if (is.null(result)){
  
} else {
  
  pggId <- result$status$analysisId[length(result$status$analysisId)]
  mod <- result$modeling %>% filter(analysisId==pggId)
  predsId <- mod[mod$analysisId == pggId & mod$trait == "inputObject","value"]
  pred <- result$predictions %>% filter(analysisId == pggId) # before we use the predsId but now we store the designation predictions in the rgg functions
  met2 <- result$metrics %>% filter(analysisId==pggId)
  
  numeric.output <- c("value", "stdError")
  if("params" %in% ls()){ # we will download the document
    # cat("\n\n## Table       \n\n")
    printTable(met2, autoWidthOpt = TRUE, scrollXOpt = FALSE,
               colNames = colnames(met2),
               numericColNames = numeric.output,
               numRound = 4)
    # cat("\n")
  }else{
    h2("Table")
    
    DT::renderDT(printTable(met2,
                            autoWidthOpt = FALSE, scrollXOpt = TRUE,
                            colNames = colnames(met2),
                            numericColNames = numeric.output,
                            numRound = 4), server = FALSE)
  }
  
}

```

### Predicted genetic gain

The following density plot allows you to see the expected change in the next generation compared to the current generation given the current genetic gain parameters (accuracy, intensity, genetic variance) obtained in a particular MET for the selected traits.

<p>&nbsp;</p>

```{r }

suppressWarnings(tryCatch({
  if(file.exists("./outputs/resultPgg.RData")){
    load("./outputs/resultPgg.RData")
  }else{
    load("resultPgg.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

if (is.null(result)){
  
} else {
  modeling <- result$modeling
  idPgg <- result$status[which(result$status$module == "pgg"),"analysisId"];
  idPgg <- idPgg[length(idPgg)]
  modeling <- setdiff( unique(modeling[which(modeling$analysisId == idPgg),"trait"]), "inputObject")
  
  metrics <- result$metrics
  idPgg <- result$status[which(result$status$module == "pgg"),"analysisId"];
  idPgg <- idPgg[length(idPgg)]
  current.metrics <- metrics[which(metrics$analysisId %in% c(idPgg)),]
  all.vals <- list()

for(iTrait in unique(modeling)){

  mu0 <- current.metrics[current.metrics$parameter == "meanG" &  current.metrics$trait == iTrait, "value"]
  sd0 <- current.metrics[current.metrics$parameter == "sigmaG" & current.metrics$trait == iTrait, "value"]
  R <- current.metrics[current.metrics$parameter == "PGG" &  current.metrics$trait == iTrait, "value"]
  m1 <- mu0 + R
  envs <- unique(current.metrics$environment[current.metrics$trait == iTrait])
  vals0 <- as.data.frame(apply(data.frame(mu0, sd0), 1,function(x) rnorm(1000, mean = x[1], sd = x[2])))
  vals1 <- as.data.frame(apply(data.frame(m1, sd0), 1,function(x) rnorm(1000, mean = x[1], sd = x[2])))
  vals0$Subject <- vals1$Subject <- paste0("a", 1:nrow(vals0))
  vals0l <- reshape(vals0, idvar = "Subject",
                    varying = list(1:length(mu0)),
                    v.names = "conc",
                    direction = "long",
                    times = envs)

  vals0l$treat <- "original"

  vals1l <- reshape(vals1, idvar = "Subject",
                    varying = list(1:length(mu0)),
                    v.names = "conc",
                    direction = "long",
                    times = envs)

  vals1l$treat <- "selected"

  vals <- rbind(vals0l, vals1l)
  vals$trait <- iTrait

  all.vals[[iTrait]] <- vals
}

vals <- do.call(rbind, all.vals)

    if("params" %in% ls()){
      
    cat(paste0("### Selection differentials and expected response  {.tabset .tabset-pills}    \n\n  "))  
      p<-ggplot(vals, aes(x = conc, fill = treat)) +  geom_density(alpha = 0.4) +
         facet_wrap(~trait, scales = "free") +  xlab("Trait value")
      print(p)
      
  }else{
    plotly::renderPlotly({
      
      p<-ggplot(vals, aes(x = conc, fill = treat)) +  geom_density(alpha = 0.4) +
         facet_wrap(~trait, scales = "free") +  xlab("Trait value")
      plotly::ggplotly(p)
      
    })
  }
}
```


### References of methods used

Lush, J. L. (2013). Animal breeding plans. Read Books Ltd.

Mrode, R. A. (2014). Linear models for the prediction of animal breeding values. Cabi.

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

<p>&nbsp;</p>


