---
title: "Selection Index Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
 toDownload: FALSE
---

```{r setup, include=FALSE}
# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
options(knitr.kable.NA = '')
# loading necessary R packages ####
## data manipulation
# library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric
## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, 
                       scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,...) {
  oldnames <- colnames(DT)
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  colnames(DT) <- oldnames
  table <- DT::datatable(data = DT, 
                         colnames = colNames,
                         filter = "top", 
                         options = list(autoWidth = autoWidthOpt,
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'csv', 'excel', 'print'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35),
                                        scrollX = scrollXOpt),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% DT::formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r, include=FALSE}
# Init Step to make sure that the dependencies are loaded
htmltools::tagList(printTable(mtcars))
htmltools::tagList(ggplotly(ggplot()))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
```

```{r}
suppressWarnings(tryCatch({
  if(file.exists("./outputs/result.RData")){
    load("./outputs/resultIndex.RData")
  }else{
    load("resultIndex.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

if (is.null(result)){
  # invisible(NULL)
} else {
  mydata <- result$predictions
  idIndex <- result$status[which(result$status$module == "indexB"),"analysisId"];
  idIndex <- idIndex[length(idIndex)]
  idMta <-result$modeling[which(result$modeling$parameter %in% "analysisId" & result$modeling$analysisId == idIndex),"value"]
  mydata <- mydata[which(mydata$analysisId %in% c(idMta,idIndex)),]
  # envsUsed <- unique(result$modeling[which(result$modeling$analysisId %in% idIndex),"environment"])
  # mydata <- mydata[which(mydata$environment %in% envsUsed),]
  mTrait <- unique(result$modeling[which(result$modeling$analysisId %in% idIndex),"trait"])
  mTrait <- gsub("_scaled","",mTrait)
  uTrait <- c(mTrait[which(mTrait %in% unique(mydata$trait))],"baseIndex")
  mydata <- mydata[which(mydata$trait %in% uTrait),]
  wide <- stats::reshape(mydata[,c(c("designation"),"trait",c("predictedValue", "entryType"))], direction = "wide", idvar = c("designation"),
                         timevar = "trait", v.names = c("predictedValue"), sep= "_")
  colnames(wide) <- gsub("predictedValue_","",colnames(wide))
  numeric.output <- colnames(wide)[-c(1:2)]
  wide <- wide[order(wide$baseIndex,decreasing=TRUE),]
  wide$Rank <- c(1:sum(!is.na(wide$baseIndex)),rep(NA,sum(is.na(wide$baseIndex))))
  
}
```

### Objectives of a Selection Index

The objective of this dashboard is to help scientist to understand the following points:

1. Individual across environment predictions for each trait (input) 

2. Individual sensitivity values from the Finlay-Wilkinson model (Finlay & Wilkinson, 1963) (input)

3. Weighted sum of BLUP values with user-provided economic weights (output)

### Across-environment trait table

The following table allows you to inspect the across-environment predictions for the different entries (rows) and different traits (columns) that were used as input to calculate the selection index.

<p>&nbsp;</p>

```{r, results='asis'}

if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    printTable(wide, autoWidthOpt = TRUE, scrollXOpt = FALSE,
           colNames = colnames(wide),
           numericColNames = numeric.output,
           numRound = 4)
  }else{
    
    DT::renderDT(printTable(wide,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(wide),
                        numericColNames = numeric.output,
                        numRound = 4), server = FALSE)
  }
  
}
```

<p>&nbsp;</p>

### Modeling table

The following table aims to keep record of the corresponding weights for this run.

```{r, results='asis'}

if (is.null(result)){
  # invisible(NULL)
} else {
  
  modelingTable <- result$modeling
  modelingTable <- modelingTable[which(modelingTable$analysisId %in% c(idIndex)),]
  
  if("params" %in% ls()){ # we will download the document
    printTable(modelingTable, autoWidthOpt = TRUE, scrollXOpt = FALSE,
           colNames = colnames(modelingTable))
  }else{
    
    DT::renderDT(printTable(modelingTable,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(modelingTable)), server = FALSE)
  }
  
}
```

### References of methods used

Brim, C. A., Johnson, H. W., & Cockerham, C. C. (1959). Multiple selection criteria in soybeans 1. Agronomy Journal, 51(1), 42-46.

Ceron-Rojas, J. J., & Crossa, J. (2018). Linear selection indices in modern plant breeding (p. 256). Springer Nature.

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

<p>&nbsp;</p>

