---
title: "Weather Data Extraction Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
  toDownload: FALSE
---

```{r setup, include=FALSE}

# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
options(knitr.kable.NA = '')

# loading necessary R packages ####
## data manipulation
library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric

## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(grafify)
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                         numericColNames = NULL, numRound = 3, 
                         scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,...) {
    
    DT <- data.frame(lapply(X = DT, 
                            FUN = function(x) {
                              if(is.numeric(x)){
                                round(x, numRound)
                              } else {
                                x
                              }
                            }))
    
    table <- DT::datatable(data = DT, 
                           colnames = colNames,
                           filter = "top", 
                           options = list(autoWidth = autoWidthOpt,
                                          dom = 'l<<t>pB>', 
                                          buttons = c('copy', 'csv', 'excel', 'print'),
                                          pageLength = pageLength,
                                          searchHighlight = TRUE,
                                          lengthMenu = c(7, 14, 21, 28, 35),
                                          scrollX = scrollXOpt),
                           extensions = 'Buttons',
                           rownames = FALSE,
                           ...)
    if (length(numericColNames) > 0){
      table <- table %>% DT::formatRound(columns = numericColNames,
                                     digits = numRound)
    }
    
    table
  }
```

```{r }

suppressWarnings(tryCatch({
  if(file.exists("./outputs/result.RData")){
    load("./outputs/resultWeather.RData")
  }else{
    load("resultWeather.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

```


### Weather data table
<p>&nbsp;</p>

The following table shows the modifications saved to be applied for the raw phenotypic data when is used in one of the analytical modules.


```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  numericalColumns1 <- c("RH2M","T2M","PRECTOTCORR") 
  
  if("params" %in% ls()){ # we will download the document
    printTable(result$data$weather, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(result$data$weather),
           numericColNames = numericalColumns1,
           numRound = 4)
  }else{
    DT::renderDT(printTable(result$data$weather,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(result$data$weather),
                        numericColNames = numericalColumns1,
                        numRound = 4), server = FALSE)
  }
  
}

```


### Weather data summary
<p>&nbsp;</p>

The following table shows the modifications saved to be applied for the raw phenotypic data when is used in one of the analytical modules.


```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  numericalColumns2 <- c("value") 
  
  if("params" %in% ls()){ # we will download the document
    printTable(result$metadata$weather, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(result$metadata$weather),
           numericColNames = numericalColumns2,
           numRound = 4)
  }else{
    DT::renderDT(printTable(result$metadata$weather,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(result$metadata$weather),
                        numericColNames = numericalColumns2,
                        numRound = 4), server = FALSE)
    
  }
  
}

```

<p>&nbsp;</p>

### Coordinates map

The following map allows you to assess the location where trials are planted.

<p>&nbsp;</p>

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  xx <- result$metadata$weather
  xx <- xx[which(xx$trait %in% c("latitude","longitude")),c("environment","trait","value")]
  xx <- reshape(xx, direction = "wide", idvar = "environment",
                timevar = "trait", v.names = "value", sep= "_")
  colnames(xx) <- cgiarBase::replaceValues(colnames(xx), Search = c("value_latitude","value_longitude") , Replace = c("latitude","longitude") )
  
  if("params" %in% ls()){ # we will download the document
    
    world <- ggplot2::map_data("world")
    p <- ggplot2::ggplot() + # + 
      ggplot2::theme(legend.position="bottom", legend.text = ggplot2::element_text(size=4), legend.title = ggplot2::element_text(size=0)) +
      ggplot2::geom_map(
        data = world, map = world,
        ggplot2::aes(long, lat, map_id = region),
        color = "black", fill = "lightgray", linewidth=0.1
      )  + ggplot2::ylab(" Latitude") + ggplot2::xlab("Longitude") +
      ggplot2::geom_point(
        data = xx,
        ggplot2::aes(longitude, latitude, color = environment),
        alpha = 0.7
      ) 
    print(p)
    
  }else{
    
    plotly::renderPlotly({
      
      fig <- xx
      fig <- fig %>%
        plotly::plot_ly(
          lat = ~latitude,
          lon = ~longitude,
          type = "scattermapbox",
          hovertext = ~environment, #us_cities[,"City"],
          marker = list(color = "fuchsia"))
      fig <- fig %>%
        plotly::layout(
          mapbox = list(
            style = 'open-street-map',
            zoom =1,
            center = list(lon = 0, lat = 0)
          )
        )
      # }
      fig
      
    })
    
  }
  
}
```

### References of methods used

Sparks A (2023). nasapower: NASA-POWER Data from R. doi:10.5281/zenodo.1040727 <https://doi.org/10.5281/zenodo.1040727>, R package version 4.0.11

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

<p>&nbsp;</p>




