---
title: "Quality Assurance Genotype Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
  toDownload: FALSE
---

```{r setup, include=FALSE}

# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
echo = FALSE,
cache = FALSE,
warning = FALSE,
message = FALSE,
comment = NA,
out.width = "100%",
error = TRUE)
options(knitr.kable.NA = '')

# loading necessary R packages ####
## data manipulation
library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric

## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(grafify)
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                         numericColNames = NULL, numRound = 3, 
                         scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,...) {
    
    DT <- data.frame(lapply(X = DT, 
                            FUN = function(x) {
                              if(is.numeric(x)){
                                round(x, numRound)
                              } else {
                                x
                              }
                            }))
    
    table <- DT::datatable(data = DT, 
                           colnames = colNames,
                           filter = "top", 
                           options = list(autoWidth = autoWidthOpt,
                                          dom = 'l<<t>pB>', 
                                          buttons = c('copy', 'csv', 'excel', 'print'),
                                          pageLength = pageLength,
                                          searchHighlight = TRUE,
                                          lengthMenu = c(7, 14, 21, 28, 35),
                                          scrollX = scrollXOpt),
                           extensions = 'Buttons',
                           rownames = FALSE,
                           ...)
    if (length(numericColNames) > 0){
      table <- table %>% DT::formatRound(columns = numericColNames,
                                     digits = numRound)
    }
    
    table
  }
```

```{r }

suppressWarnings(tryCatch({
  if(file.exists("./outputs/result.RData")){
    load("./outputs/resultQaPheno.RData")
  }else{
    load("resultQaPheno.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

```


### Table of modifications
<p>&nbsp;</p>

The following table shows the modifications that have been saved in your analysis object and that will be applied to the raw genotype data when this specific time stamp (ID) is selected/used in other analytical modules (e.g., multi trial analysis module).

```{r }

if (is.null(result)){
  # invisible(NULL)
} else {
  modifications <- result$modifications$geno
  idQa <- result$status[which(result$status$module %in% c("qaGeno")),"analysisId"];
  idQa <- idQa[length(idQa)]
  idQaType <- result$status[result$status$analysisId == idQa, "module"]
  modifications <- modifications[which( modifications$analysisId == idQa & modifications$module == idQaType),]
  traitsTable <- unique(modifications[which( modifications$analysisId == idQa),"reason"])
  
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("traitQa"), 
                       label = "Reason for QA:", 
                       choices = traitsTable, 
    )
  }
  
}

```


```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  numericalColumns <- c("value","row") 
  
  if("params" %in% ls()){ # we will download the document
    printTable(modifications, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(modifications),
           numericColNames = numericalColumns,
           numRound = 4)
  }else{
    DT::renderDT(printTable(modifications[modifications$reason==input$traitQa,],
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(modifications[modifications$reason==input$traitQa,]),
                        numericColNames = numericalColumns,
                        numRound = 4), server = FALSE)
  }
  
}

```


<p>&nbsp;</p>

### References on methods used

Tukey, J. W. (1977). Exploratory Data Analysis. Section 2C.

Velleman, P. F. and Hoaglin, D. C. (1981). Applications, Basics and Computing of Exploratory Data Analysis. Duxbury Press.


<p>&nbsp;</p>



