---
title: "Realized Genetic Gain Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
 toDownload: FALSE
---

```{r setup, include=FALSE}

# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
options(knitr.kable.NA = '')

# loading necessary R packages ####
## data manipulation
library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric

## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(grafify)
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                         numericColNames = NULL, numRound = 3, 
                         scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,...) {
    oldnames <- colnames(DT)
    DT <- data.frame(lapply(X = DT, 
                            FUN = function(x) {
                              if(is.numeric(x)){
                                round(x, numRound)
                              } else {
                                x
                              }
                            }))
    colnames(DT) <- oldnames
    table <- DT::datatable(data = DT, 
                           colnames = colNames,
                           filter = "top", 
                           options = list(autoWidth = autoWidthOpt,
                                          dom = 'l<<t>pB>', 
                                          buttons = c('copy', 'csv', 'excel', 'print'),
                                          pageLength = pageLength,
                                          searchHighlight = TRUE,
                                          lengthMenu = c(7, 14, 21, 28, 35),
                                          scrollX = scrollXOpt),
                           extensions = 'Buttons',
                           rownames = FALSE,
                           ...)
    if (length(numericColNames) > 0){
      table <- table %>% DT::formatRound(columns = numericColNames,
                                     digits = numRound)
    }
    
    table
  }
```

<!-- END of setup chunk -->

```{r }
suppressWarnings(tryCatch({
  if(file.exists("./outputs/resultRgg.RData")){
    load("./outputs/resultRgg.RData")
  }else{
    load("resultRgg.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

if (is.null(result)){
  # invisible(NULL)
} else {
  
  rggId <- result$status$analysisId[length(result$status$analysisId)]
  mod <- result$modeling %>% filter(analysisId==rggId)
  # predsId <- mod[mod$analysisId == rggId & mod$trait == "inputObject","value"]
  pred <- result$predictions %>% filter(analysisId == rggId) # before we use the predsId but now we store the designation predictions in the rgg functions
  met <- result$metrics %>% filter(analysisId==rggId)
  traits <- unique(met$trait)
  ped <- result$data$pedigree
  paramsPed <- result$metadata$pedigree
  colnames(ped) <- cgiarBase::replaceValues(colnames(ped), Search = paramsPed$value, Replace = paramsPed$parameter )
  ped <- unique(ped[,c("designation","yearOfOrigin")])
  # merge yearOfOrigin
  myResult <- merge(pred,ped,by="designation")
  myResult <- myResult[which(!is.na(myResult$yearOfOrigin)),]
}
```

```{r, include=FALSE}
# Init Step to make sure that the dependencies are loaded
htmltools::tagList(printTable(mtcars))
htmltools::tagList(ggplotly(ggplot()))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
```

### Summary metrics

The following table allows you to review the different metrics calculated for the genetic gain analysis. You can filter the metrics using the search bar.

```{r, results='asis'}

modeling = result$modeling[result$modeling$module == "rgg",]
lessThanFive = modeling[modeling$parameter == "lessThanFiveYears","value"]
if(any(lessThanFive) == "TRUE"){
  cat("> **WARNING:** Less than five years were used for genetic gain analysis. Interpret results with caution.\n\n")
}

```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  numeric.output <- c("value", "stdError")
  if("params" %in% ls()){ # we will download the document
    # cat("\n\n## Table       \n\n")
    printTable(met, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(met),
           numericColNames = numeric.output,
           numRound = 4)
    # cat("\n")
  }else{
    h2("Table")
    
    DT::renderDT(printTable(met,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(met),
                        numericColNames = numeric.output,
                        numRound = 4), server = FALSE)
  }
  
}

```

### Summary metrics (wide-format)

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  wide1 <- reshape(met[,c("trait","parameter","value")], direction = "wide", idvar = "trait",
                   timevar = "parameter", v.names = "value", sep= "_")
  colnames(wide1) <- gsub("value_","",colnames(wide1))
  wide2 <- reshape(met[which(met$parameter %in% c("gg%","ggInter","ggSlope") ),c("trait","parameter","stdError")], direction = "wide", idvar = "trait",
                   timevar = "parameter", v.names = "stdError", sep= "_")
  wide <- merge(wide1, wide2, by="trait")
  
  numeric.output2 <- which( colnames(wide) %in% setdiff(colnames(wide), "trait") )
  if("params" %in% ls()){ # we will download the document
    # cat("\n\n## Table       \n\n")
    printTable(wide, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(wide),
           numericColNames = numeric.output2,
           numRound = 4)
    # cat("\n")
  }else{
    # h2("Table")
    DT::renderDT(printTable(wide,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(wide),
                        numericColNames = numeric.output2,
                        numRound = 4), server = FALSE)
  }
  
}

```

### Regression plot

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    selectInput(ns("traitSta3"), 
                label = "Trait:", 
                multiple = FALSE,
                choices = unique(traits), 
    )
  }
  
}

```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    cat("\n\n### {.tabset .tabset-pills}       \n\n")
    
    
    for (j in 1:length(traits)){
      cat("\n\n####", traits[j], "{.tabset .tabset-pills}       \n\n")
      
      # subset by trait
      myPreds <- myResult %>% filter(trait == traits[j])
      myMetrics <- met %>% filter(trait == traits[j])
      myLabel <- paste0("Y = ", round(myMetrics[myMetrics$parameter == "ggInter","value"],4), " + ",
                        round(myMetrics[myMetrics$parameter == "ggSlope","value"],4), " x Year, ", 
                        round(myMetrics[myMetrics$parameter == "gg%(average.year)","value"],2),"% Gain")
      p <- ggplot(myPreds, aes(x=yearOfOrigin, y=predictedValue  ) ) +
        geom_smooth(method = "lm", formula = y ~ x) + 
        theme_classic()+
        geom_jitter(alpha = 0.6, width = 0.3) +
        labs(title=myLabel,x="Year of origin", y = "Trait performance") +
        theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
      print(htmltools::tagList(plotly::ggplotly(p)))
      
      # cat("\n")
    }
    
    cat("\n")
  }else{ # regular rendering
    plotly::renderPlotly({
      myPreds <- myResult %>% filter(trait == input$traitSta3)
      myMetrics <- met %>% filter(trait == input$traitSta3)
      myLabel <- paste0("Y = ", round(myMetrics[myMetrics$parameter == "ggInter","value"],4), " + ",
                        round(myMetrics[myMetrics$parameter == "ggSlope","value"],4), " x Year, ", 
                        round(myMetrics[myMetrics$parameter == "gg%(average.year)","value"],2),"% Gain")
      p <- ggplot(myPreds, aes(x=yearOfOrigin, y=predictedValue  ) ) +
        geom_smooth(method = "lm", formula = y ~ x) + 
        theme_classic()+
        geom_jitter(alpha = 0.6, width = 0.3) +
        labs(title=myLabel,x="Year of origin", y = "Trait performance") +
        theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
      fig <-  plotly::ggplotly(p)#, height = get_h()*2);
      fig
    })
  }
  
}

```


### References on methods used

Mackay, I., Horwell, A., Garner, J., White, J., McKee, J., & Philpott, H. (2011). Reanalyses of the historical series of UK variety trials to quantify the contributions of genetic and environmental factors to trends and variability in yield over time. Theoretical and Applied Genetics, 122, 225-238.

Laidig, F., Piepho, H. P., Drobek, T., & Meyer, U. (2014). Genetic and non-genetic long-term trends of 12 different crops in German official variety performance trials and on-farm yield trends. Theoretical and Applied Genetics, 127, 2599-2617.

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

<p>&nbsp;</p>

