---
title: "Predicted Genetic Gain Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
  toDownload: FALSE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
library(dplyr)
library(ggplot2)
```

```{r plot_init, include = FALSE}
# Init Step to make sure that the dependencies are loaded
# was done to enable proper printing of the tab header and the graphs within the for loop below using print(htmltools::tagList(ggplotly(...)))
htmltools::tagList(plotly::ggplotly(ggplot2::ggplot()))

# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}

```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, 
                       scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,...) {
  oldnames <- colnames(DT)
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  colnames(DT) <- oldnames
  table <- DT::datatable(data = DT, 
                         colnames = colNames,
                         filter = "top", 
                         options = list(autoWidth = autoWidthOpt,
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'csv', 'excel', 'print'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35),
                                        scrollX = scrollXOpt),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% DT::formatRound(columns = numericColNames,
                                       digits = numRound)
  }
  
  table
}
```

### Summary metrics

The following table allows you to review the different metrics calculated for the genetic gain analysis. You can filter the metrics using the search bar.

```{r, results='asis'}
if (is.null(result)){
  
} else {
  
  pggId <- result$status$analysisId[length(result$status$analysisId)]
  mod <- result$modeling %>% filter(analysisId==pggId)
  predsId <- mod[mod$analysisId == pggId & mod$trait == "inputObject","value"]
  pred <- result$predictions %>% filter(analysisId == pggId) # before we use the predsId but now we store the designation predictions in the rgg functions
  met2 <- result$metrics %>% filter(analysisId==pggId)
  
  numeric.output <- c("value", "stdError")
  if("params" %in% ls()){ # we will download the document
    # cat("\n\n## Table       \n\n")
    printTable(met2, autoWidthOpt = TRUE, scrollXOpt = FALSE,
               colNames = colnames(met2),
               numericColNames = numeric.output,
               numRound = 4)
    # cat("\n")
  }else{
    h2("Table")
    
    DT::renderDT(printTable(met2,
                            autoWidthOpt = FALSE, scrollXOpt = TRUE,
                            colNames = colnames(met2),
                            numericColNames = numeric.output,
                            numRound = 4), server = FALSE)
  }
  
}

```

### Predicted genetic gain

The following density plot allows you to see the expected change in the next generation compared to the current generation given the current genetic gain parameters (accuracy, intensity, genetic variance) obtained in a particular MET for the selected traits.

<p>&nbsp;</p>

```{r }

suppressWarnings(tryCatch({
  if(file.exists("./outputs/resultPgg.RData")){
    load("./outputs/resultPgg.RData")
  }else{
    load("resultPgg.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

if (is.null(result)){
  
} else {
  modeling <- result$modeling
  idPgg <- result$status[which(result$status$module == "pgg"),"analysisId"];
  idPgg <- idPgg[length(idPgg)]
  modeling <- setdiff( unique(modeling[which(modeling$analysisId == idPgg),"trait"]), "inputObject")
  if("params" %in% ls()){ # input <- list(traitFilterPredictions2D2=modeling)
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("traitFilterPredictions2D2"),
                       label = "Trait:",
                       choices = modeling,
    )
  }
}

```

```{r, results='asis'}
if (is.null(result)){
  
} else {
  
  metrics <- result$metrics
  idPgg <- result$status[which(result$status$module == "pgg"),"analysisId"];
  idPgg <- idPgg[length(idPgg)]
  current.metrics <- metrics[which(metrics$analysisId %in% c(idPgg)),]
  
  if("params" %in% ls()){
    cat(paste0("### Selection differentials and expected response  {.tabset .tabset-pills}    \n\n  "))
    # l <- htmltools::tagList()
    # counter <- 1
    for(iTrait in unique(modeling)){
      cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
      
      mu0 <-current.metrics[which(current.metrics$parameter == "meanG" & current.metrics$trait == iTrait ),"value"];
      sd0 <-current.metrics[which(current.metrics$parameter == "sigmaG" & current.metrics$trait == iTrait ),"value"];
      R <-current.metrics[which(current.metrics$parameter == "R" & current.metrics$trait == iTrait ),"value"];
      m1 <- mu0 + R
      envs <- unique(current.metrics$environment[which(current.metrics$trait == iTrait)])
      
      vals0 <- as.data.frame(apply(data.frame(mu0,sd0),1,function(x){rnorm(1000, mean=x[1], sd = x[2] )}))
      vals1 <- as.data.frame(apply(data.frame(m1,sd0),1,function(x){rnorm(1000, mean=x[1], sd = x[2] )}))
      vals0$Subject <- vals1$Subject <- paste0("a",1:nrow(vals0))
      
      vals0l <- reshape(vals0, idvar = "Subject", varying = list(1:(length(mu0))),
                        v.names = "conc", direction = "long", times = envs)
      vals0l$treat <- "original"
      vals1l <- reshape(vals1, idvar = "Subject", varying = list(1:(length(mu0))),
                        v.names = "conc", direction = "long", times = envs)
      vals1l$treat <- "selected"
      vals <- rbind(vals0l,vals1l)
      
      p<-ggplot2::ggplot(vals, ggplot2::aes(x=conc, fill=treat)) +
        ggplot2::geom_density(alpha=0.4) +  ggplot2::facet_wrap(~time) +
        ggplot2::xlab("Trait value")
      
      print(p)
      
    }
  }else{
    plotly::renderPlotly({
      
      mu0 <-current.metrics[which(current.metrics$parameter == "meanG" & current.metrics$trait == input$traitFilterPredictions2D2 ),"value"];
      sd0 <-current.metrics[which(current.metrics$parameter == "sigmaG" & current.metrics$trait == input$traitFilterPredictions2D2 ),"value"];
      R <-current.metrics[which(current.metrics$parameter == "R" & current.metrics$trait == input$traitFilterPredictions2D2 ),"value"];
      m1 <- mu0 + R
      envs <- unique(current.metrics$environment[which(current.metrics$trait == input$traitFilterPredictions2D2)])
      
      vals0 <- as.data.frame(apply(data.frame(mu0,sd0),1,function(x){rnorm(1000, mean=x[1], sd = x[2] )}))
      vals1 <- as.data.frame(apply(data.frame(m1,sd0),1,function(x){rnorm(1000, mean=x[1], sd = x[2] )}))
      vals0$Subject <- vals1$Subject <- paste0("a",1:nrow(vals0))
      
      vals0l <- reshape(vals0, idvar = "Subject", varying = list(1:(length(mu0))),
                        v.names = "conc", direction = "long", times = envs)
      vals0l$treat <- "original"
      vals1l <- reshape(vals1, idvar = "Subject", varying = list(1:(length(mu0))),
                        v.names = "conc", direction = "long", times = envs)
      vals1l$treat <- "selected"
      vals <- rbind(vals0l,vals1l)
      
      p<-ggplot2::ggplot(vals, ggplot2::aes(x=conc, fill=treat)) +
        ggplot2::geom_density(alpha=0.4) +  ggplot2::facet_wrap(~time) + 
        ggplot2::xlab("Trait value")
      
      plotly::ggplotly(p)
      
    })
  }
}
```

### Barplot for metrics 

The following barplot allows you to compare the parameters with respect to the classification variables.

```{r}
if (is.null(result)){
  # invisible(NULL)
} else {
  # analysis id
  idPgg <- result$status[which(result$status$module == "pgg"),"analysisId"];
  idPgg <- idPgg[length(idPgg)]
  # tables
  modelingPgg <- result$modeling[which(result$modeling$analysisId == idPgg),]
  metricsPgg <- result$metrics[which(result$metrics$analysisId == idPgg),]
  # choices
  traitsPgg <- setdiff(unique(metricsPgg$trait),c("inputObject"))
  paramPgg <- setdiff( unique(metricsPgg$parameter), "min.G" )
  classPgg <- unique(modelingPgg[which(modelingPgg$parameter == "classifier"),"value"])
  if("params" %in% ls()){  # if user will download the report
    shinyjs::hide()
  }else{
    
    inputPanel(
      selectInput(ns("paramMetricsBy"),
                  label = "Trait:",
                  # selected = "environment",
                  choices = traitsPgg
      ),
      shiny::selectInput(ns("paramMetrics"), 
                         label = "Parameter to filter:", 
                         multiple = FALSE,
                         # selected = unique(parameters),
                         choices = paramPgg
      ),
      
    )
    
    
  }
}
```
<p>&nbsp;</p>

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){  # if user will download the report
    
    cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
    counter <- 1
    
    for(iView in traitsPgg ){
      cat("\n\n####", iView, "{.tabset .tabset-pills}       \n\n")
      # for(iTrait in staTraits){
      for(iParam in paramPgg ){
        cat("\n\n#####", iParam, "{.tabset .tabset-pills}       \n\n")
        
        metricsPgg2 <- metricsPgg[which(metricsPgg$trait == iView ),]
        metricsPgg3 <- metricsPgg2[which(metricsPgg2$parameter == iParam ),]
        
        if(nrow(metricsPgg3) > 0){
          
          p<-ggplot(metricsPgg3, aes(x=environment, y=value)) +
            geom_bar(stat="identity")+theme_minimal()
          
          print(htmltools::tagList(plotly::ggplotly(p)))
        }
        
      }
      cat("\n")
      
    }
    
    cat("\n")
    
  }else{ # display in website
    
    plotly::renderPlotly({
      
      metricsPgg2 <- metricsPgg[which(metricsPgg$trait == input$paramMetricsBy ),]
      metricsPgg3 <- metricsPgg2[which(metricsPgg2$parameter == input$paramMetrics ),]
      
      p<-ggplot(metricsPgg3, aes(x=environment, y=value)) +
        geom_bar(stat="identity")+theme_minimal()
      p
      
      plotly::ggplotly(p)
    })
    
  }
  
}
```

### References of methods used

Lush, J. L. (2013). Animal breeding plans. Read Books Ltd.

Mrode, R. A. (2014). Linear models for the prediction of animal breeding values. Cabi.

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

<p>&nbsp;</p>


