---
title: "Selection Index Report"
author: ""
date: "December 2023"
output: html_document
params:
 toDownload: FALSE
---

```{r setup, include=FALSE}
# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
options(knitr.kable.NA = '')
# loading necessary R packages ####
## data manipulation
# library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric
## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, ...) {
  
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  
  table <- DT::datatable(data = DT, 
                         filter = "top", 
                         options = list(autoWidth = TRUE, 
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'excel'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35)),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r, include=FALSE}
# Init Step to make sure that the dependencies are loaded
htmltools::tagList(printTable(mtcars))
htmltools::tagList(ggplotly(ggplot()))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
```

```{r}
suppressWarnings(tryCatch({
  if(file.exists("./outputs/resultIndex.RData")){
    load("./outputs/resultIndex.RData")
  }else{
    load("resultIndex.RData")
  }
}, error = function(e) {
 shinyjs::hide()
}))

if (is.null(result)){
  # invisible(NULL)
} else {
  mydata = result$predictions
  idIndex <- result$status[which(result$status$module == "indexD"),"analysisId"];
  idIndex <- idIndex[length(idIndex)]
  idMta <-result$modeling[which(result$modeling$parameter %in% "analysisId" & result$modeling$analysisId == idIndex),"value"]
  mydata <- mydata[which(mydata$analysisId %in% c(idMta,idIndex)),]
  wide <- stats::reshape(mydata[,c(c("designation"),"trait",c("predictedValue"))], direction = "wide", idvar = c("designation"),
                         timevar = "trait", v.names = c("predictedValue"), sep= "_")
  colnames(wide) <- gsub("predictedValue_","",colnames(wide))
  numeric.output <- colnames(wide)[-c(1)]
  
}
```

### Across-environment trait table

The following table allows you to inspect the across-environment predictions for the different entries (rows) and different traits (columns).

<p>&nbsp;</p>

```{r, results='asis'}

if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    printTable(wide)
  }else{
    DT::renderDT(DT::formatRound(DT::datatable(wide, extensions = 'Buttons',
                                               options = list(dom = 'Blfrtip',scrollX = TRUE,buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                                                              lengthMenu = list(c(5,20,50,-1), c(5,20,50,'All')))
    ), numeric.output)
    )
  }
  
}
```

<p>&nbsp;</p>

### Radar Plot

This plot allows the user to visualize the distance between the population means and the target values.

<p>&nbsp;</p>

```{r, results='asis'}
library(dplyr)
if (is.null(result)){
  # invisible(NULL)
} else {
  
  dtIdxD <- result$predictions
  idIndex <- result$status[which(result$status$module == "indexD"),"analysisId"];
  idIndex <- idIndex[length(idIndex)]
  # idIndex <- modelingAbi[modelingAbi$parameter == "indexD", "value"]
  # idMta <- modelingAbi[modelingAbi$parameter == "mta", "value"] 
  idMta <- result$modeling[which(result$modeling$parameter %in% "analysisId" & result$modeling$analysisId == idIndex),"value"]
  mydata <- dtIdxD[which(dtIdxD$analysisId == idMta),setdiff(colnames(dtIdxD),c("module","analysisId"))]
  # subset to only traits being part of the index
  desirev <-result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"value"]
  uTraits <- result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"trait"]
  uTraits <- gsub("_scaled","",uTraits)
  dtIdxD <- dtIdxD[which(dtIdxD$trait %in% uTraits),]
  ##
  environmentPredictionsRadar2="across"
  traitFilterPredictionsRadar2=unique(dtIdxD[which(dtIdxD$analysisId == idMta),"trait"])
  ## ensure product profile means come sorted
  meanGroupPredictionsRadar= result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),c("trait","value")]
  meanGroupPredictionsRadar <- meanGroupPredictionsRadar[with(meanGroupPredictionsRadar, order(as.numeric(as.factor(trait)))), ]
  meanGroupPredictionsRadar <- meanGroupPredictionsRadar[,"value"]
  
  fontSizeRadar=12; r0Radar=NULL; neRadar=NULL; plotSdRadar=FALSE
  
  mydata = mydata[which( (mydata$environment %in% environmentPredictionsRadar2) &
                           (mydata$trait %in% traitFilterPredictionsRadar2)
  ),]
  
  if(!is.null(environmentPredictionsRadar2) & !is.null(traitFilterPredictionsRadar2)){
    mm <- stats::aggregate(predictedValue~trait, data=mydata, FUN=mean, na.rm=TRUE)
    mmsd <- stats::aggregate(predictedValue~trait, data=mydata, FUN=stats::sd, na.rm=TRUE)
    mm2 <- stats::aggregate(predictedValue~trait, data=mydata, FUN=min, na.rm=TRUE)
    mm3 <- stats::aggregate(predictedValue~trait, data=mydata, FUN=max, na.rm=TRUE)
    namesA <- mm[,1]
    meanA <- mm[,2]/(mm3[,2])
    
    # meanA <- rep(0.75,nrow(mm))
    fig <-  plotly::plot_ly(
      type = 'scatterpolar',
      fill = 'toself'
    )
    fig <- fig %>%
      plotly::add_trace(
        r = meanA, #c(39, 28, 8, 7, 28, 39),
        theta =paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
        name = 'Breeding population means',
        text=round(mm[,2],3)
      )
    ## add product profile means
    if(!is.null(meanGroupPredictionsRadar)){
      meanB <- as.numeric(unlist(strsplit(meanGroupPredictionsRadar,",")))
      if(length(meanB) == length(mm[,2])){
        meanB <- meanB + mm[,2]
        # r2 = (meanB*0.75)/mm[,2]
        r2 = meanB/(mm3[,2])
        dToSearch <- meanB #- mm[,2]
        fig <- fig %>%
          plotly::add_trace(
            r = r2, # c(1.5, 10, 39, 31, 15, 1.5),
            theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
            name = "Desired population", #'Group B'
            # mode="text",
            text=round(dToSearch,2)
            # textfont = list(color = '#000000', size = fontSizeRadar)
          )
      }
    }
    # add SD polygon
    if(plotSdRadar){
      meanC <- (mm[,2]+mmsd[,2])/(mm3[,2])
      fig <- fig %>%
        plotly::add_trace(
          r = meanC, # c(1.5, 10, 39, 31, 15, 1.5),
          theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
          name = "Mean + Standard Deviation from pop", #'Group B'
          text=round(mm[,2]+mmsd[,2],3)
        )
      meanD <- (mm[,2]-mmsd[,2])/(mm3[,2])
      fig <- fig %>%
        plotly::add_trace(
          r = meanD, # c(1.5, 10, 39, 31, 15, 1.5),
          theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
          name = "Mean - Standard Deviation from pop", #'Group B'
          text=round(mm[,2]-mmsd[,2],3)
        )
    }
    # add selection limits polygon
    ## add product profile means'
    if(!is.null(r0Radar)){
      meanR <- as.numeric(unlist(strsplit(r0Radar,",")))/100
      if(length(meanR) == length(mm[,2])){
        meanR <- mm[,2] + ((meanR*mm[,2])*neRadar*2) # selection limits 2*Ne*R0
        r3 = (meanR*0.75)/mm[,2]
        dToSearchR <- meanR #- mm[,2]
        fig <- fig %>%
          plotly::add_trace(
            r = r3, # c(1.5, 10, 39, 31, 15, 1.5),
            theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
            name = "Selection limits", #'Group B'
            # mode="text",
            text=round(dToSearchR,2)
            # textfont = list(color = '#000000', size = fontSizeRadar)
          )
      }
    }
  }
  
  if("params" %in% ls()){ # we will download the document
    
    
    print(htmltools::tagList(plotly::ggplotly(fig)))
    
  }else{
    
    plotly::renderPlotly({
      fig
    }) # enf of render plotly
    
  }
  
}

```


### Expected Response Plot

This plot allows the user to visualize the expected response in the new generation given the individuals selected (best 20 percent) using the selection index.

<p>&nbsp;</p>

```{r}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    proportionTrait <- c(0.05,0.10,0.15,0.20)
  }else{
    shiny::numericInput(ns("proportionTrait"), label = "Selected proportion", value = 0.1, min=0.001,max=1, step=0.05)
  }
  
}
```

```{r, results='asis'}
library(dplyr)
if (is.null(result)){
  # invisible(NULL)
} else {
  
  dtIdxD <- result$predictions
  idIndex <- result$status[which(result$status$module == "indexD"),"analysisId"];
  idIndex <- idIndex[length(idIndex)]
  idMta <-result$modeling[which(result$modeling$parameter %in% "analysisId" & result$modeling$analysisId == idIndex),"value"]
  mydataMta <- dtIdxD[which(dtIdxD$analysisId == idMta),setdiff(colnames(dtIdxD),c("module","analysisId"))]
  mydataIndex <- dtIdxD[which(dtIdxD$analysisId == idIndex),setdiff(colnames(dtIdxD),c("module","analysisId"))]
  mydata <- rbind(mydataMta,mydataIndex)
  desirev <-result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"value"]
  uTraits <- unique(mydata$trait)#result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"trait"]
  uTraits <- gsub("_scaled","",uTraits)
  mydata$selected <- "no"
  
  
  if("params" %in% ls()){ # we will download the document
    
    cat(paste0("#### Proportion selected  {.tabset .tabset-pills}    \n\n  "))
    for(iProportionTrait in proportionTrait){
      cat("\n\n#####", iProportionTrait, "{.tabset .tabset-pills}       \n\n")
      
      for(iTrait in uTraits){ # iTrait = uTraits[1]
        prov <- mydata[mydata$trait == "desireIndex", ]
        prov <- prov[with(prov, order(-predictedValue)), ]
        selected <- prov[1:round(nrow(prov)*iProportionTrait),"designation"]
        mydata[which( (mydata$trait == iTrait) & (mydata$designation %in% selected) ), "selected"] <- "yes"
      }
      p <- ggplot2::ggplot(mydata, ggplot2::aes(x=predictedValue, color=selected, fill = selected)) +
        ggplot2::geom_histogram() +
        ggplot2::ylab("Frequency") +
        ggplot2::geom_rug(sides="t", length = ggplot2::unit(0.3, "cm")) +
        ggplot2::facet_wrap(~trait, ncol=3, scales = "free")
      print(htmltools::tagList(plotly::ggplotly(p)))
    }
    
  }else{
    
    # shiny::renderPlot({ # 
    plotly::renderPlotly({
      ## if user provides threshold values let them know the selection differentials
      for(iTrait in uTraits){ # iTrait = uTraits[1]
        prov <- mydata[mydata$trait == "desireIndex", ]
        prov <- prov[with(prov, order(-predictedValue)), ]
        selected <- prov[1:round(nrow(prov)*input$proportionTrait),"designation"]
        mydata[which( (mydata$trait == iTrait) & (mydata$designation %in% selected) ), "selected"] <- "yes"
      }
      p <- ggplot2::ggplot(mydata, ggplot2::aes(x=predictedValue, color=selected, fill = selected)) +
        ggplot2::geom_histogram() +
        ggplot2::ylab("Frequency") +
        ggplot2::geom_rug(sides="t", length = ggplot2::unit(0.3, "cm")) +
        ggplot2::facet_wrap(~trait, ncol=3, scales = "free")
      
      # p
      plotly::ggplotly(p)
    })
    
  }
  
  
  
}

```

### Index versus Trait Plot

This plot allows the user to visualize the relationship between the selection index with other traits. This can be specially useful for selecting materials with high total merit and stability for yield.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  
  
  mtaTraits <- setdiff( colnames(wide), c("designation","desireIndex"))
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("traitMtaScatter"), 
                       label = "Trait:", 
                       multiple = FALSE,
                       choices = mtaTraits, 
    )
  }
  
}

```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    
    cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
    
    for(iTrait in mtaTraits){ # iTrait="Yield_Mg_ha"
      cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
      
      wide$trait1 <- wide[,iTrait]
      p <- ggplot(wide, aes(x=trait1, y=desireIndex, text=designation)) + 
        geom_point(size=2, alpha = 0.6) + xlab(iTrait) + ylab("Selection Index") 
      
      if(nrow(wide) > 1000){ # go fixed format
        print(p)
      }else{ # go dynamic
        print(htmltools::tagList(plotly::ggplotly(p)))
      }
    }
    
  }else{ # render in screen
    
    plotly::renderPlotly({
      wide$trait1 <- wide[,input$traitMtaScatter]
      p <- ggplot(wide, aes(x=trait1, y=desireIndex, text=designation)) + 
        geom_point(size=2, alpha = 0.6) + xlab(input$traitMtaScatter) + ylab("Selection Index") 
      
      plotly::ggplotly(p)
    })
    
  }
  
}
```

### References of methods used

Pesek, J., & Baker, R. J. (1969). Desired improvement in relation to selection indices. Canadian journal of plant science, 49(6), 803-804.

Ceron-Rojas, J. J., & Crossa, J. (2018). Linear selection indices in modern plant breeding (p. 256). Springer Nature.

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

<p>&nbsp;</p>

