---
title: "Multi-Trial Analysis Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
  toDownload: FALSE
  # modelUsed: NULL
---

```{r setup, include=FALSE}
# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
# ,fig.width=12, fig.height=8)
options(knitr.kable.NA = '')

# loading necessary R packages ####
## data manipulation
# library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric
## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(data.table)
library(shiny)
library(maps)
```

```{r printfxn, include=FALSE}
# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, 
                       scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,
                       fixedCol = 0, filt = "top",...) {
  oldnames <- colnames(DT)
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  colnames(DT) <- oldnames
  table <- DT::datatable(data = DT, 
                         colnames = colNames,
                         filter = filt,
                         options = list(autoWidth = autoWidthOpt,
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'csv', 'excel', 'print'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35),
                                        scrollX = scrollXOpt,
                                        fixedColumns = list(leftColumns = fixedCol)
                                        ),
                         extensions = c('Buttons', 'FixedColumns'),
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% DT::formatRound(columns = numericColNames,
                                       digits = numRound)
  }
  
  table
}
```

```{r, include=FALSE}
# Init Step to make sure that the dependencies are loaded
htmltools::tagList(printTable(mtcars))
htmltools::tagList(ggplotly(ggplot()))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
```

```{r }

suppressWarnings(tryCatch({
  if(file.exists("./outputs/result.RData")){
    load("./outputs/resultMtaLMMsolver.RData")
  }else{
    load("resultMtaLMMsolver.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

if (is.null(result)){
  # invisible(NULL)
} else {
  # result$status <- result$status[1:6,]
  idMta <- result$status[which(result$status$module == "mtaLmms"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred <- result$predictions
  pred <- pred[pred$analysisId == idMta,]
  traits <- setdiff( unique(result$modeling[which(result$modeling$analysisId == idMta  & result$modeling$environment == "across"),"trait"]), "inputObject") 
  metrics <- result$metrics
  modeling <- result$modeling
  modelingFieldsIncludedInMta <- modeling[modeling$analysisId == idMta & modeling$parameter == "includedInMta", ]
  # predictions in all formats to be used later
  predSplit <- split(pred, pred[,"effectType"])
  predSplitSplit <- lapply(predSplit, function(x){split(x, x[,"trait"])})
  predSplitSplit2 <- lapply(predSplit, function(x){split(x, x[,"environment"])})
  predSplitSplitWide <- lapply(predSplitSplit, function(y){lapply(y,function(x){
    tableForm <- table(x$environment)
    ninds <- tableForm; nEnvs <- length(tableForm)
    x <- x[with(x, order(environment, designation)), ]
    
    x$idsmock <- gsub(paste(unique(modelingFieldsIncludedInMta$environment), collapse = "|"), "", x$designation)
    x$idsmock <- gsub("^:|:$", "", x$idsmock)
    
    wide0 <- reshape(x[,c("environment","predictedValue","idsmock")], direction = "wide", idvar = "idsmock",
                     timevar = "environment", v.names = "predictedValue", sep= "_")
    rownames(wide0) <- wide0$idsmock; wide0 <- wide0[,-1,drop=FALSE]
    wide0 <- wide0[,apply(wide0,2,enhancer::propMissing) != 1, drop=FALSE]
    colnames(wide0) <- gsub("predictedValue_","",colnames(wide0))
    return(wide0)
  }
  )})
  predSplitSplitWide2 <- lapply(predSplitSplit2, function(y){ # for each environment do
    lapply(y,function(x){
      x <- x[with(x, order(trait, designation)), ]
      
      x$idsmock <- gsub("^:|:$", "", 
                      gsub(paste(unique(modelingFieldsIncludedInMta$environment), collapse = "|"), "", x$designation))
      
      wide0 <- reshape(x[,c("trait","predictedValue","idsmock")], direction = "wide", idvar = "idsmock",
                       timevar = "trait", v.names = "predictedValue", sep= "_")
      rownamesWide0 <- wide0[,1]
      wide0 <- wide0[,-1,drop=FALSE]; 
      wide0 <- apply(wide0,2,enhancer::imputev)
      if(is.matrix(wide0)){
        rownames(wide0) <- rownamesWide0
        colnames(wide0) <- gsub("predictedValue_","",colnames(wide0))
      }
      return(wide0)
    })
  })
  effectTypes <- rev(names(predSplitSplit))
  envirTypes <- unique(pred$environment)
  # check names
  entryTypePresent <- table((unique(pred[,c("designation","entryType")]))$entryType)
  checksFound <- grep("^(?!.*#).*(check|chck|bnchmark|benchmark|chek|chec).*", names(entryTypePresent),ignore.case = TRUE,perl = TRUE,value = FALSE)
  labelChecks <- names(entryTypePresent)[checksFound]
  
  nameOfChecks <- pred$designation[(pred$entryType %in% labelChecks) & (pred$effectType == "designation")]

  idSta <- modeling[which(modeling$trait == "inputObject" & modeling$analysisId == idMta),"value"]
  envs2 <- unique(result$data$pheno[,result$metadata$pheno[which(result$metadata$pheno$parameter == "environment"), "value"]])
  traits2 <- unique(result$metadata$pheno[which(result$metadata$pheno$parameter == "trait"), "value"])
    gxeList <- list(
  "gxeCS" = c("environment_designation", "designation_environment", 
              "environment:designation", "designation:environment"),
  "gxeDMD" = c(paste0(envs2,"_designation"), paste0("designation_",envs2),
               paste0(envs2,":designation"), paste0("designation:",envs2)),
  "gxeFW"  = c(paste0("value_",traits2,"envIndex_designation"),
               paste0("designation_value_",traits2,"envIndex"),
               paste0("value_",traits2,"envIndex:designation"),
               paste0("designation:value_",traits2,"envIndex"))
)
  
  gxeModelNum <- 0
  for(i in 1:3){
    pat <- paste(gxeList[[i]], collapse = "|")  # union of all patterns for this i
    if (any(grepl(pat, effectTypes))) {
      gxeModelNum <- i
      break
      }
  }
  
  if(gxeModelNum == 2){
    effectTypes2 <- setdiff(effectTypes, c(effectTypes[which(effectTypes %in% gxeList[[2]])],
                                           effectTypes[which(effectTypes %in% envs2)]))
    if(any(grepl(":",effectTypes[which(effectTypes %in% gxeList[[2]])]))){
      effectTypes2 <- c(effectTypes2, "environment:designation")
    } else{
      effectTypes2 <- c(effectTypes2, "environment_designation")
    }
  } else if(gxeModelNum == 3){
    effectTypes2 <- setdiff(effectTypes, effectTypes[which(effectTypes %in% gxeList[[3]])])
  } else{
    effectTypes2 <- effectTypes
  }
  
}
```

### Objectives of Multi-Trial Analysis

The objective of this dashboard is to help scientist to understand the following points:

1. Overall number of entries and effect types included in the multi trial analysis (input)

2. High-level view of the phenotypic adjusted means included in the analysis (input)

3. Number of entries connecting the different environments per trait (input)

4. Phenotypic correlation between environments for the traits present (input)

5. Across environment KPIs per trait such as reliability and variance component ratios (output) 

6. Individual across environment predictions for each trait (output) 

7. Individual sensitivity and stability values across environments (output)

8. Percent check comparison against the different benchmark varieties present in the dataset (output)

9. Genetic correlation between the traits (output)

Understanding these data features should allow the scientist to identify which traits express more genotype by environment interactions and how they should be selected. It should also allow the scientist to assess the correlation between traits and how the product profile should be addressed to maximize genetic gains while developing the needed varieties at the same time. Materials with the highest performance and sensitivity to the environment could be potential nominations for advancement although we recommend to don't select new parents or products until the results from the multi-trial analysis are analyzed with a selection index.  

### Metrics: Table of parameters

The following table aims to make a high-level assessment of the different types of entries included in the analysis across environments.

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  metricsMta <- result$metrics[which(metrics$analysisId == idMta),]
  metricsMta <- metricsMta[, setdiff(colnames(metricsMta),c("module","analysisId"))]
  
  if("params" %in% ls()){ # we will download the document
    printTable(metricsMta, autoWidthOpt = TRUE, scrollXOpt = FALSE,
               colNames = colnames(metricsMta),
               numericColNames = c("value","stdError"),
               numRound = 4)
  }else{
    DT::renderDT(printTable(metricsMta,
                            autoWidthOpt = FALSE, scrollXOpt = TRUE,
                            colNames = colnames(metricsMta),
                            numericColNames = c("value","stdError"),
                            numRound = 4), server = FALSE)
    
  }
  
}

```

### Metadata: Map of trials planted

The following map allows you to assess the location where trials are planted.

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if(!is.null(result$data$weather)){ # if weather data was extracted
    if(nrow(result$data$weather)>0){
    yy <- result$metadata$weather
    yy <- yy[yy$parameter != "trait",]
    xx <- cgiarPipeline::summaryWeather(result, wide=TRUE)
    xx$environment <- rownames(xx)
    metaWeather <- result$metadata$weather
    latitude <- unique(metaWeather[which(metaWeather$parameter %in% c("latitude") ),"value"])
    longitude <- unique(metaWeather[which(metaWeather$parameter %in% c("longitude") ),"value"])
    
    if(length(c(latitude,longitude)) == 2){
      colnames(xx) <- cgiarBase::replaceValues(colnames(xx), Search = paste0("value_",c(latitude,longitude),"mean") , Replace = c("latitude","longitude"))
      if("params" %in% ls()){ # we will download the document
        
        world <- ggplot2::map_data("world")
        p <- ggplot2::ggplot() + # + 
          ggplot2::theme(legend.position="bottom", legend.text = ggplot2::element_text(size=4), legend.title = ggplot2::element_text(size=0)) +
          ggplot2::geom_map(
            data = world, map = world,
            ggplot2::aes(long, lat, map_id = region),
            color = "black", fill = "lightgray", linewidth=0.1
          )  + ggplot2::ylab(" Latitude") + ggplot2::xlab("Longitude") +
          ggplot2::geom_point(
            data = xx,
            ggplot2::aes(longitude, latitude, color = environment),
            alpha = 0.7
          ) 
        print(p)
        # plotly::ggplotly(p)
        
      }else{
        
        plotly::renderPlotly({
          
          fig <- xx
          fig <- fig %>%
            plotly::plot_ly(
              lat = ~latitude,
              lon = ~longitude,
              type = "scattermapbox",
              hovertext = ~environment, #us_cities[,"City"],
              marker = list(color = "fuchsia"))
          fig <- fig %>%
            plotly::layout(
              mapbox = list(
                style = 'open-street-map',
                zoom =1,
                center = list(lon = 0, lat = 0)
              )
            )
          # }
          fig
          
        })
        
      }
    }
    }
  }else{
    cat("No coordinates available. Skipping planting map.")
  }
  
}
```

### Predictions: By environment merit distribution

The following boxplot allows you to inspect the distribution of adjusted means (y-axis) from the different environments for each trait that were used as input for the analysis. The environments are sorted by environmental mean to understand the slope of the regression fitted in the Finlay-Wilkinson model. It is recommended that you have at least 6 environments to fit the random regressions over the environmental indices or other weather and soil variables

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{ # input <- list(boxplotMtaPredsTrait="YLDTONHA",boxplotMtaPredsType="environment_designation")
    inputPanel(
      selectInput(ns("boxplotMtaPredsType"),
                  label = "Effect type",
                  choices = setdiff(effectTypes2, c("environment", "(Intercept)"))
      ),
      column(width=12),
      selectInput(ns("boxplotMtaPredsTrait"),
                  label = "Trait",
                  choices = traits
      ),
    )
  }
  
}

```

<div style = "width:auto; height:auto; overflow:auto">

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    
    cat(paste0("###  {.tabset .tabset-pills}    \n\n  "))
    for(iType in setdiff(effectTypes2, c("environment", "(Intercept)"))){
      cat("\n\n####", iType, "{.tabset .tabset-pills}       \n\n")
      for(iTrait in traits){
        cat("\n\n#####", iTrait, "{.tabset .tabset-pills}       \n\n")
        if(gxeModelNum == 2){
          if(iType %in% c("environment_designation", "environment:designation")){
            level1 <- predSplitSplit[effectTypes[which(effectTypes %in% gxeList[[2]])]]
            level2 <- lapply(level1, function(x){x[iTrait]})
            level3 <- NULL
            for(i in 1:length(level2)){
              level3 <- rbind(level3,level2[[i]][[1]])
            }
            temp.pred.box <- level3
            # temp.pred.box$environment <- substr(temp.pred.box$environment, 4, length(temp.pred.box$environment))
          } else{
            temp.pred.box <- predSplitSplit[[iType]][[iTrait]]
          }
        } else{
          temp.pred.box <- predSplitSplit[[iType]][[iTrait]]
        }
        # temp.pred.box <- temp.pred.box[which(temp.pred.box$environment %in% modelingFieldsIncludedInMta[modelingFieldsIncludedInMta$trait == iTrait,"environment"]),]
        if(!is.null(temp.pred.box)){
          if(nrow(temp.pred.box) > 0){
            p <- temp.pred.box %>%
              ggplot() +
              aes(x=reorder(environment,-predictedValue), y=predictedValue, fill = environment) +
              geom_boxplot() +
              labs(x = "", y = "Trait value") +
              scale_fill_discrete(name = "") +
              theme(legend.position = "none")
            # guides(fill=guide_legend(nrow=3,byrow=TRUE))
            
            if(iType %in% c("environment_designation", "designation_environment",
                            "environment:designation", "designation:environment")){
              print(htmltools::tagList(plotly::ggplotly(p + labs(x = "Environment")) %>% layout(xaxis = list(tickangle =-45),autosize = T, height = 800)))
            } else{
              print(htmltools::tagList(plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank()))))
            }
          }else{
            print(htmltools::tagList(plotly::ggplotly(ggplot2::ggplot() + ggplot2::ggtitle("Combination not available"))))
          }# end of nrow > 0
        }else{
          print(htmltools::tagList(plotly::ggplotly(ggplot2::ggplot() + ggplot2::ggtitle("Combination not available"))))
        }
      }# end of for each trait
    } # enf of for each type
    
  } else{ # display in website
    plotly::renderPlotly({
      if(gxeModelNum == 2){
        if(input$boxplotMtaPredsType %in% c("environment_designation", "environment:designation")){
          level1 <- predSplitSplit[effectTypes[which(effectTypes %in% gxeList[[2]])]]
          level2 <- lapply(level1, function(x){x[input$boxplotMtaPredsTrait]})
          level3 <- NULL
          for(i in 1:length(level2)){
            level3 <- rbind(level3,level2[[i]][[1]])
          }
          temp.pred.box <- level3
          # temp.pred.box$environment <- substr(temp.pred.box$environment, 4, length(temp.pred.box$environment))
        } else{
          temp.pred.box <- predSplitSplit[[input$boxplotMtaPredsType]][[input$boxplotMtaPredsTrait]]
        }
      } else{
        temp.pred.box <- predSplitSplit[[input$boxplotMtaPredsType]][[input$boxplotMtaPredsTrait]]
      }
      
      # temp.pred.box <- temp.pred.box[which(temp.pred.box$environment %in% modelingFieldsIncludedInMta[modelingFieldsIncludedInMta$trait == input$boxplotMtaPredsTrait,"environment"]),]
      if(!is.null(temp.pred.box)){
        if(nrow(temp.pred.box) > 0){
          p <- temp.pred.box %>%
            ggplot() +
            aes(x=reorder(environment,-predictedValue), y=predictedValue, fill = environment) +
            geom_boxplot() +
            labs(x = "", y = "Trait value") +
            scale_fill_discrete(name = "") +
            theme(legend.position = "none")
          # guides(fill=guide_legend(nrow=3,byrow=TRUE))
          
          if(input$boxplotMtaPredsType %in% c("environment_designation", "designation_environment",
                                              "environment:designation", "designation:environment")){
            plotly::ggplotly(p + labs(x = "Environment")) %>% layout(xaxis = list(tickangle =-45),autosize = T, height = 800)
          } else{
            plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank()))
          }
        }else{
          plotly::ggplotly(ggplot2::ggplot() + ggplot2::ggtitle("Combination not available"))
        }
      }else{
        plotly::ggplotly(ggplot2::ggplot() + ggplot2::ggtitle("Combination not available"))
      }
      
    })
    
  }
  
}
```
</div>

### Predictions: Connectivity between environments

The following heatmap and histogram allows you to assess the connectivity (gennotypes in common) between different environments for a given trait. Our recommendation is that each pair of environments should at least have 30 genotypes in common in order to estimate genetic correlations and at least 2-3 entries in common to adjust across environment means. We classify as low everything below 30, intermediate connectivity between 30 to 60, and high everything above 60.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{ # input<-list(traitMtaConnect="YLDTONHA")
    inputPanel(shiny::selectInput(ns("traitMtaConnect"), 
                                  label = "Trait", 
                                  choices = traits
    ),
    shiny::checkboxInput(ns("freqConnect"),
                         label = "Show frequency distribution",
                         value = TRUE
    )
    )
    
  }
  
}

```

<div style = "width:auto; height:auto; overflow:auto">

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  mydataSta <- result$predictions[which(result$predictions$analysisId %in% idSta),] # extract predictions
  # mydataSta[,"environment"] <- gsub("[[:punct:]]", "", mydataSta[,"environment"] )
  
  if("params" %in% ls()){ # we will download the document
    
    cat(paste0("###  {.tabset .tabset-pills}    \n\n  "))
    
    for(iTrait in traits){
      cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
      
      goodFields <- modelingFieldsIncludedInMta[modelingFieldsIncludedInMta$value == "TRUE" & modelingFieldsIncludedInMta$trait == iTrait, "environment"]
      if(length(goodFields) > 0){
        
        mydataSta2 <- mydataSta[which(mydataSta$trait == iTrait & mydataSta$environment %in% goodFields), ]
        splitAggregate <- with(mydataSta2,  split(mydataSta2[,"designation"],mydataSta2[,"environment"]) ) # split by environment
        splitAggregate <- lapply(splitAggregate,unique); nag <- length(splitAggregate) # get unique individual names
        nagm <- matrix(0,nag,nag); rownames(nagm) <- colnames(nagm) <- names(splitAggregate) # prefilled matrix
        for(i in 1:length(splitAggregate)){ # fill the matrix of intersection of individuals between pair of environments
          for(j in 1:i){
            nagm[i,j] <- length(intersect(splitAggregate[[i]],splitAggregate[[j]]))
          }
        }
        nagm[upper.tri(nagm)] <- t(nagm)[upper.tri(nagm)] # fill the upper triangular
        
        mydataSta4 <- cgiarBase::matToTab(nagm) # matrix to a dataframe for plot
        maxVal <- max(nagm, na.rm = TRUE) # get the maximum value found in the matrix of connectivity
        minVal <- min(nagm, na.rm = TRUE)
        midVal <- (maxVal - minVal)/2
        p <- ggplot2::ggplot(data = mydataSta4, ggplot2::aes(Var2, Var1, fill = Freq))+
          ggplot2::geom_tile(color = "white")+
          ggplot2::scale_fill_gradient2(low = "firebrick", high = "#038542", mid = "gold",
                                        midpoint = midVal, limit = c(0,maxVal+20), space = "Lab",
                                        name="Connectivity") +
          ggplot2::theme_minimal()+
          ggplot2::ylab("") + ggplot2::xlab("") +
          ggplot2::coord_fixed() 
        
        # if(nrow(nagm) < 8){ # if user wants to add text to the cells
        #   p <- p + ggplot2::geom_text(ggplot2::aes(label = Freq), color = "white", size = 4)
        # }
        # Basic histogram
        mydataSta4$Group <- ifelse(mydataSta4$Freq < 31, "Low", ifelse(mydataSta4$Freq < 80, "Medium", "High"))
        group.colors <- c(Low="firebrick",Medium="gold", High="#038542")
        ub <- max(mydataSta4$Freq, na.rm = TRUE)
        lb <- min(mydataSta4$Freq, na.rm = TRUE)
        mb <- (ub - lb)/2
        # p2 <- ggplot(mydataSta4[which(mydataSta4$Var1 != mydataSta4$Var2),], aes(x=Freq, fill=Group)) + 
        #   geom_histogram() + ggplot2::scale_x_continuous(limits = c(0,ub)) +
        #   scale_fill_manual(values=group.colors) + ggplot2::xlab("Frequency")
        
        p2 <- ggplot() + geom_histogram(data = mydataSta4[which(mydataSta4$Var1 != mydataSta4$Var2),], aes(x = Freq, fill = ..x..), binwidth=10) +
          ggplot2::scale_x_continuous(limits = c(0,ub+20)) +
          scale_fill_gradient2(low="firebrick",mid="gold", high="#038542", midpoint=mb,
                               limit = c(0,ub+20), space = "Lab",
                               name="Connectivity") +
          ggplot2::xlab("Value") + ggplot2::ylab("")
        
        
        for(iFreq in c("with frequency distribution", "without frequency distribution")){
          cat("\n\n#####", iFreq, "{.tabset .tabset-pills}       \n\n")
          if(iFreq == "with frequency distribution"){
            # combine plots
            print(htmltools::tagList( plotly::subplot(plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank()) + geom_text(aes(label=Freq), size = 2)) , plotly::ggplotly(p2), nrows=1, shareY = FALSE, shareX = FALSE ) %>% layout(autosize = T, width = 1600)))
          } else{
            print(htmltools::tagList(plotly::ggplotly(p + geom_text(aes(label=Freq), size = 3)) %>% layout(xaxis = list(tickangle =-45),autosize = T, height = 800)))
          }
          
        }
        
      } else{
        print(htmltools::tagList(plotly::ggplotly(ggplot2::ggplot() + ggplot2::ggtitle("Combination not available"))))
      }
      
    }
    
  } else{ # display in website
    
    # shiny::renderPlot({ # static
    plotly::renderPlotly({ # dynamic
      goodFields <- modelingFieldsIncludedInMta[modelingFieldsIncludedInMta$value == "TRUE" & modelingFieldsIncludedInMta$trait == input$traitMtaConnect, "environment"]
      if(length(goodFields) > 0){
        mydataSta2 <- mydataSta[which(mydataSta$trait == input$traitMtaConnect & mydataSta$environment %in% goodFields), ]
        splitAggregate <- with(mydataSta2,  split(mydataSta2[,"designation"],mydataSta2[,"environment"]) ) # split by environment
        splitAggregate <- lapply(splitAggregate,unique); nag <- length(splitAggregate) # get unique individual names
        nagm <- matrix(0,nag,nag); rownames(nagm) <- colnames(nagm) <- names(splitAggregate) # prefilled matrix
        for(i in 1:length(splitAggregate)){ # fill the matrix of intersection of individuals between pair of environments
          for(j in 1:i){
            nagm[i,j] <- length(intersect(splitAggregate[[i]],splitAggregate[[j]]))
          }
        }
        nagm[upper.tri(nagm)] <- t(nagm)[upper.tri(nagm)] # fill the upper triangular
        
        mydataSta4 <- cgiarBase::matToTab(nagm) # matrix to a dataframe for plot
        maxVal <- max(nagm, na.rm = TRUE) # get the maximum value found in the matrix of connectivity
        minVal <- min(nagm, na.rm = TRUE)
        midVal <- (maxVal - minVal)/2
        p <- ggplot2::ggplot(data = mydataSta4, ggplot2::aes(Var2, Var1, fill = Freq))+
          ggplot2::geom_tile(color = "white")+
          ggplot2::scale_fill_gradient2(low = "firebrick", high = "#038542", mid = "gold",
                                        midpoint = midVal, limit = c(0,maxVal+20), space = "Lab",
                                        name="Connectivity") +
          ggplot2::theme_minimal()+
          ggplot2::ylab("") + ggplot2::xlab("") +
          ggplot2::coord_fixed()
        # ggplot2::coord_fixed() + 
        # ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank() )
        # if(nrow(nagm) < 8){ # if user wants to add text to the cells
        #   p <- p + ggplot2::geom_text(ggplot2::aes(label = Freq), color = "white", size = 4)
        # }
        # Basic histogram
        mydataSta4$Group <- ifelse(mydataSta4$Freq < 31, "Low", ifelse(mydataSta4$Freq < 80, "Medium", "High"))
        group.colors <- c(Low="firebrick",Medium="gold", High="#038542")
        ub <- max(mydataSta4$Freq, na.rm = TRUE)
        lb <- min(mydataSta4$Freq, na.rm = TRUE)
        mb <- (ub-lb)/2
        # p2 <- ggplot(mydataSta4[which(mydataSta4$Var1 != mydataSta4$Var2),], aes(x=Freq, fill=Group)) + 
        #   geom_histogram() + ggplot2::scale_x_continuous(limits = c(0,ub)) +
        #   scale_fill_manual(values=group.colors) + ggplot2::xlab("Frequency")
        
        p2 <- ggplot() + geom_histogram(data = mydataSta4[which(mydataSta4$Var1 != mydataSta4$Var2),], aes(x = Freq, fill = ..x..), binwidth=10) +
          ggplot2::scale_x_continuous(limits = c(0,ub+20)) +
          scale_fill_gradient2(low="firebrick",mid="gold", high="#038542", midpoint=mb,
                               limit = c(0,ub+20), space = "Lab",
                               name="Connectivity") +
          ggplot2::xlab("Value") + ggplot2::ylab("")
        
        if(input$freqConnect == TRUE){
          # combine plots
          ply1 <- plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank()) + geom_text(aes(label=Freq), size = 3)) 
          ply2 <- plotly::ggplotly(p2)
          plotly::subplot(ply1, ply2, nrows=1, shareY = FALSE, shareX = FALSE ) %>% layout(autosize = T, width = 1600)
        } else{
          plotly::ggplotly(p  + geom_text(aes(label=Freq))) %>% layout(xaxis = list(tickangle =-45),autosize = T, height = 800)
        }
      } else{
        plotly::ggplotly(ggplot2::ggplot() + ggplot2::ggtitle("Combination not available"))
      }
      
      
    })
    
  }
  
}
```

</div>

### Predictions: Correlations between environments

The following heatmap and histogram allows to assess the genetic correlations among the different environment one trait at a time. If the user modeled GxE, this matrix is calculated as the correlation between the environment-specific estimates for individuals. We recommend that only environments that are on-average positively correlated with the main cluster (i.e., rG>0) are included in the multi-trial analysis to guarantee decent rates of genetic gain and adapted products in the TPE.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{ # input <- list(traitPredictionsCorrelation=traits[1],corrplotMtaPredsType=effectTypes[1])
    # if(length(setdiff(effectTypes, c("environment", "designation", "(Intercept)"))) > 0){
    if(any(c("environment_designation", "designation_environment",
             "environment:designation", "designation:environment") %in% effectTypes2)){
      inputPanel(
        # selectInput(ns("corrplotMtaPredsType"),
        #             label = "Effect type",
        #             choices = setdiff(effectTypes, c("environment", "designation", "(Intercept)")), multiple = TRUE, selected = setdiff(effectTypes, c("environment", "designation", "(Intercept)"))[1]
        # ),
        # column(width=12),
        selectInput(ns("traitPredictionsCorrelation"),
                    label = "Trait",
                    choices = traits
        ),
        column(width=12),
        shiny::checkboxInput(ns("freqCor"),
                             label = "Show frequency distribution",
                             value = TRUE
        )
      )
    }
  }
}

```

<div style = "width:auto; height:auto; overflow:auto">

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  # if(length(setdiff(effectTypes, c("environment", "designation", "(Intercept)"))) > 0){
  if(gxeModelNum == 2){
    if("params" %in% ls()){ # we will download the document
      
      cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
      
      for(iTrait2 in traits){
        
        cat("\n\n####", iTrait2, "{.tabset .tabset-pills}       \n\n")
        
        level1 <- predSplitSplitWide[effectTypes[which(effectTypes %in% gxeList[[2]])]]
        level2 <- lapply(level1, function(x){x[iTrait2]})
        level3 <- NULL
        
        for(i in 1:length(level2)){
          level2[[i]][[1]]$designation <- gsub(paste0(colnames(level2[[i]][[1]]),":"),"",
                                          gsub(paste0(":",colnames(level2[[i]][[1]])),"",
                                               row.names(level2[[i]][[1]])))
          if(i == 1){
            level3 <- level2[[i]][[1]]
          } else{
            level3 <- merge(level3, level2[[i]][[1]], by = "designation", all = TRUE)
          }
        }
        
        wide2 <- level3
        row.names(wide2) <-wide2$designation
        wide2 <- wide2[,-1]
        # colnames(wide2) <- substring(colnames(wide2),4,nchar(colnames(wide2)))
        
        if(!is.null(wide2)){
          if(nrow(wide2) > 1){
            corr <- round(stats::cor(wide2, use="pairwise.complete.obs"),2)
            if(nrow(corr) > 1){
              if( length(which(is.na(corr))) == 0){
                ord <- hclust( dist(corr, method = "euclidean"), method = "ward.D" )$order
                corr <- corr[ord,ord]
              }
            }else{
              corr[1,1]=1
            }
            mydata4 <- cgiarBase::matToTab(corr)
            p <- ggplot2::ggplot(data = mydata4, ggplot2::aes(Var2, Var1, fill = Freq)) +
              ggplot2::geom_tile(color = "white") +
              ggplot2::scale_fill_gradient2(low = "#E46726", high = "#038542", mid = "white",
                                            midpoint = 0, limit = c(-1,1), space = "Lab",
                                            name="Pearson\nCorrelation") +
              ggplot2::theme_minimal() +
              ggplot2::ylab("") + 
              ggplot2::xlab("") +
              ggplot2::coord_fixed()
            
            # Basic histogram
            mydata4$Group <- ifelse(mydata4$Freq < 0, "Low", ifelse(mydata4$Freq < 0.2, "Medium", "High"))
            group.colors <- c(Low="#E46726",Medium="white", High="#038542")
            p2 <- ggplot() + 
              geom_histogram(data = mydata4, aes(x = Freq, fill = ..x..), binwidth=0.1 ) + 
              ggplot2::scale_x_continuous(limits = c(-1,1)) +
              scale_fill_gradient2(low='#E46726', mid='white', high='#038542', midpoint=0, 
                                   limit = c(-1,1), space = "Lab", 
                                   name="Pearson\nCorrelation") + 
              ggplot2::xlab("Value") + 
              ggplot2::ylab("")
          } else{
            p <- ggplot2::ggplot() + ggplot2::ggtitle("Not enough rows")
            p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Not enough rows")
          }
        } else{
          p <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
          p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
        }
        
        for(iFreq in c("with frequency distribution", "without frequency distribution")){
          cat("\n\n#####", iFreq, "{.tabset .tabset-pills}       \n\n")
          if(iFreq == "with frequency distribution"){
            # combine plots
            if(!is.null(wide2)){
              if(nrow(wide2) > 1){
                print(htmltools::tagList( plotly::subplot(plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank()) + geom_text(aes(label=Freq), size = 2)) , plotly::ggplotly(p2), nrows=1, shareY = FALSE, shareX = FALSE ) %>% layout(autosize = T, width = 1600)))
              } else{
                print(htmltools::tagList( plotly::subplot(plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank())) , plotly::ggplotly(p2), nrows=1, shareY = FALSE, shareX = FALSE ) ))
              }
            }else{
              print(htmltools::tagList( plotly::subplot(plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank())) , plotly::ggplotly(p2), nrows=1, shareY = FALSE, shareX = FALSE ) ))   
            }
            
          } else{
            if(!is.null(wide2)){
              if(nrow(wide2) > 1){
                print(htmltools::tagList(plotly::ggplotly(p + geom_text(aes(label=Freq), size = 3)) %>% layout(xaxis = list(tickangle =-45),autosize = T, height = 800)))
              }else{
                print(htmltools::tagList(plotly::ggplotly(p)))
              }
            }else{
              print(htmltools::tagList(plotly::ggplotly(p)))
            }
          }
        }
        
      }
    }else{
      plotly::renderPlotly({
        
        level1 <- predSplitSplitWide[effectTypes[which(effectTypes %in% gxeList[[2]])]]
        level2 <- lapply(level1, function(x){x[input$traitPredictionsCorrelation]})
        level3 <- NULL
        for(i in 1:length(level2)){
          level2[[i]][[1]]$designation <- gsub(paste0(colnames(level2[[i]][[1]]),":"),"",
                                          gsub(paste0(":",colnames(level2[[i]][[1]])),"", row.names(level2[[i]][[1]])))
          if(i == 1){
            level3 <- level2[[i]][[1]]
          } else{
            level3 <- merge(level3, level2[[i]][[1]], by = "designation", all = TRUE)
          }
        }
        row.names(level3) <-level3$designation
        level3 <- level3[,-1]
        
        if(length(level3)>0){
          N <- sort(unlist(lapply(level3,length)), decreasing = TRUE)
          wide2 <- do.call(cbind,level3[ names(N)[which(N == N[1])] ])
          # colnames(wide2) <- substring(colnames(wide2),4,nchar(colnames(wide2)))
          corr <- round(stats::cor(wide2, use="pairwise.complete.obs"),2)
          if(nrow(corr) > 1){
            if( length(which(is.na(corr))) == 0){
              ord <- hclust( dist(corr, method = "euclidean"), method = "ward.D" )$order
              corr <- corr[ord,ord]
            }
          }else{
            corr[1,1]=1
          }
          
          mydata4 <- cgiarBase::matToTab(corr)
          p <- ggplot2::ggplot(data = mydata4, ggplot2::aes(Var2, Var1, fill = Freq))+
            ggplot2::geom_tile(color = "white")+
            ggplot2::scale_fill_gradient2(low = "#E46726", high = "#038542", mid = "white",
                                          midpoint = 0, limit = c(-1,1), space = "Lab",
                                          name="Pearson\nCorrelation") +
            ggplot2::theme_minimal()+
            ggplot2::ylab("") + 
            ggplot2::xlab("") +
            ggplot2::coord_fixed()
          
          # Basic histogram
          mydata4$Group <- ifelse(mydata4$Freq < 0, "Low", ifelse(mydata4$Freq < 0.2, "Medium", "High"))
          group.colors <- c(Low="#E46726",Medium="white", High="#038542")
          
          p2 <- ggplot() + geom_histogram(data = mydata4, 
                                          aes(x = Freq, fill = ..x..), binwidth=0.1 ) + 
            ggplot2::scale_x_continuous(limits = c(-1,1)) +
            scale_fill_gradient2(low='#E46726', mid='white', high='#038542', midpoint=0, 
                                 limit = c(-1,1), space = "Lab", 
                                 name="Pearson\nCorrelation") + 
            ggplot2::xlab("Correlation Value") + 
            ggplot2::ylab("")
          
        }else{
          p <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
          p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
        }
        
        if(input$freqCor == TRUE){
          # combine plots
          if(length(level3)>0){
            ply1 <- plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank()) + geom_text(aes(label=Freq), size = 3)) 
            ply2 <- plotly::ggplotly(p2)
            plotly::subplot(ply1, ply2, nrows=1, shareY = FALSE, shareX = FALSE ) %>% layout(autosize = T, width = 1600)
          } else{
            ply1 <- plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank()))
            ply2 <- plotly::ggplotly(p2)
            plotly::subplot(ply1, ply2, nrows=1, shareY = FALSE, shareX = FALSE )
          }
        } else{
          if(length(level3)>0){
            plotly::ggplotly(p  + geom_text(aes(label=Freq))) %>% layout(xaxis = list(tickangle =-45),autosize = T, height = 800)
          } else{
            plotly::ggplotly(p)
          }
        }
      })
    }
  } else if(gxeModelNum == 1){
    
    if("params" %in% ls()){ # we will download the document
      
      cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
      
      # for(iType2 in setdiff(effectTypes, c("environment", "designation", "(Intercept)"))){
      #   cat("\n\n####", iType2, "{.tabset .tabset-pills}       \n\n")
      
      for(iTrait2 in traits){
        cat("\n\n####", iTrait2, "{.tabset .tabset-pills}       \n\n")
        
        # wide2 <- predSplitSplitWide[[iType2]][[iTrait2]] 
        wide2 <- predSplitSplitWide[[which(names(predSplitSplitWide) %in% gxeList[[1]])]][[iTrait2]]
        
        if(!is.null(wide2)){
          if(nrow(wide2) > 1){
            corr <- round(stats::cor(wide2, use="pairwise.complete.obs"),2)
            if(nrow(corr) > 1){
              if( length(which(is.na(corr))) == 0){
                ord <- hclust( dist(corr, method = "euclidean"), method = "ward.D" )$order
                corr <- corr[ord,ord]
              }
            }else{
              corr[1,1]=1
            }
            mydata4 <- cgiarBase::matToTab(corr)
            p <- ggplot2::ggplot(data = mydata4, ggplot2::aes(Var2, Var1, fill = Freq))+
              ggplot2::geom_tile(color = "white")+
              ggplot2::scale_fill_gradient2(low = "#E46726", high = "#038542", mid = "white",
                                            midpoint = 0, limit = c(-1,1), space = "Lab",
                                            name="Pearson\nCorrelation") +
              ggplot2::theme_minimal()+
              ggplot2::ylab("") + ggplot2::xlab("") +
              ggplot2::coord_fixed()
            # ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black"), 
            #                axis.text.x = ggplot2::element_text(angle = 45, hjust=1, vjust = 1))
            # ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank() ) 
            # if(nrow(corr) < 8){ # if user wants to fill cell values
            # p <- p + ggplot2::geom_text(ggplot2::aes(label = round(Freq,2) ), color = "black", size = 4)
            # }
            # Basic histogram
            mydata4$Group <- ifelse(mydata4$Freq < 0, "Low", ifelse(mydata4$Freq < 0.2, "Medium", "High"))
            group.colors <- c(Low="#E46726",Medium="white", High="#038542")
            p2 <- ggplot() + geom_histogram(data = mydata4, aes(x = Freq, fill = ..x..), binwidth=0.1 ) + 
              ggplot2::scale_x_continuous(limits = c(-1,1)) +
              scale_fill_gradient2(low='#E46726', mid='white', high='#038542', midpoint=0, 
                                   limit = c(-1,1), space = "Lab", 
                                   name="Pearson\nCorrelation") + 
              ggplot2::xlab("Value") + ggplot2::ylab("")
            # p2 <- ggplot(mydata4, aes(x=Freq, fill=Group)) + 
            #   geom_histogram() + ggplot2::scale_x_continuous(limits = c(-1,1)) +
            #   scale_fill_manual(values=group.colors) + ggplot2::xlab("Frequency")
          }else{
            p <- ggplot2::ggplot() + ggplot2::ggtitle("Not enough rows")
            p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Not enough rows")
          }
        }else{
          p <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
          p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
        }
        
        for(iFreq in c("with frequency distribution", "without frequency distribution")){
          cat("\n\n#####", iFreq, "{.tabset .tabset-pills}       \n\n")
          if(iFreq == "with frequency distribution"){
            # combine plots
            if(!is.null(wide2)){
              if(nrow(wide2) > 1){
                print(htmltools::tagList( plotly::subplot(plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank()) + geom_text(aes(label=Freq), size = 2)) , plotly::ggplotly(p2), nrows=1, shareY = FALSE, shareX = FALSE ) %>% layout(autosize = T, width = 1600)))
              } else{
                print(htmltools::tagList( plotly::subplot(plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank())) , plotly::ggplotly(p2), nrows=1, shareY = FALSE, shareX = FALSE ) ))
              }
            }else{
              print(htmltools::tagList( plotly::subplot(plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank())) , plotly::ggplotly(p2), nrows=1, shareY = FALSE, shareX = FALSE ) ))   
            }
            
          } else{
            if(!is.null(wide2)){
              if(nrow(wide2) > 1){
                print(htmltools::tagList(plotly::ggplotly(p + geom_text(aes(label=Freq), size = 3)) %>% layout(xaxis = list(tickangle =-45),autosize = T, height = 800)))
              }else{
                print(htmltools::tagList(plotly::ggplotly(p)))
              }
            }else{
              print(htmltools::tagList(plotly::ggplotly(p)))
            }
            
          }
          
        }
        # combine plots
        # print(htmltools::tagList( plotly::subplot(plotly::ggplotly(p), plotly::ggplotly(p2), nrows=1, shareY = TRUE, shareX = TRUE ) ) )
      }
      # }
      
    }else{
      
      plotly::renderPlotly({
        # shiny::renderPlot({
        # if(length(input$corrplotMtaPredsType) > 0){
        # level1 <- predSplitSplitWide[input$corrplotMtaPredsType]
        
        level1 <- predSplitSplitWide[which(names(predSplitSplitWide) %in% gxeList[[1]])]
        level2 <- lapply(level1, function(x){x[input$traitPredictionsCorrelation]})
        level3 <- lapply(rapply(level2, enquote, how="unlist"), eval)
        
        if(length(level3)>0){
          N <- sort(unlist(lapply(level3,length)), decreasing = TRUE)
          wide2 <- do.call(cbind,level3[ names(N)[which(N == N[1])] ])
          colnames(wide2) <- gsub(paste0(effectTypes[which(effectTypes %in% gxeList[[1]])],".",input$traitPredictionsCorrelation,"."),"",colnames(wide2))
          corr <- round(stats::cor(wide2, use="pairwise.complete.obs"),2)
          if(nrow(corr) > 1){
            if( length(which(is.na(corr))) == 0){
              ord <- hclust( dist(corr, method = "euclidean"), method = "ward.D" )$order
              corr <- corr[ord,ord]
            }
          }else{
            corr[1,1]=1
          }
          mydata4 <- cgiarBase::matToTab(corr)
          p <- ggplot2::ggplot(data = mydata4, ggplot2::aes(Var2, Var1, fill = Freq))+
            ggplot2::geom_tile(color = "white")+
            ggplot2::scale_fill_gradient2(low = "#E46726", high = "#038542", mid = "white",
                                          midpoint = 0, limit = c(-1,1), space = "Lab",
                                          name="Pearson\nCorrelation") +
            ggplot2::theme_minimal()+
            ggplot2::ylab("") + ggplot2::xlab("") +
            ggplot2::coord_fixed()
          # ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black"), 
          #                axis.text.x = ggplot2::element_text(angle = 45, hjust=1, vjust = 1) )
          # ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank() ) 
          # if(nrow(corr) < 8){ # if user wants to fill cell values
          # p <- p + ggplot2::geom_text(ggplot2::aes(label = round(Freq,2) ), color = "black", size = 4)
          # }
          # Basic histogram
          mydata4$Group <- ifelse(mydata4$Freq < 0, "Low", ifelse(mydata4$Freq < 0.2, "Medium", "High"))
          group.colors <- c(Low="#E46726",Medium="white", High="#038542")
          
          p2 <- ggplot() + geom_histogram(data = mydata4, 
                                          aes(x = Freq, fill = ..x..), binwidth=0.1 ) + 
            ggplot2::scale_x_continuous(limits = c(-1,1)) +
            scale_fill_gradient2(low='#E46726', mid='white', high='#038542', midpoint=0, 
                                 limit = c(-1,1), space = "Lab", 
                                 name="Pearson\nCorrelation") + 
            ggplot2::xlab("Correlation Value") + ggplot2::ylab("")
          
          # p2 <- ggplot(mydata4, aes(x=Freq, fill=Group)) + 
          #   geom_histogram() + ggplot2::scale_x_continuous(limits = c(-1,1)) +
          #   scale_fill_manual(values=group.colors) + ggplot2::xlab("Frequency")
          # combine plots
          
        }else{
          p <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
          p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
        }
        
        if(input$freqCor == TRUE){
          # combine plots
          if(length(level3)>0){
            ply1 <- plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank()) + geom_text(aes(label=Freq), size = 3)) 
            ply2 <- plotly::ggplotly(p2)
            plotly::subplot(ply1, ply2, nrows=1, shareY = FALSE, shareX = FALSE ) %>% layout(autosize = T, width = 1600)
          } else{
            ply1 <- plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank()))
            ply2 <- plotly::ggplotly(p2)
            plotly::subplot(ply1, ply2, nrows=1, shareY = FALSE, shareX = FALSE )
          }
        } else{
          if(length(level3)>0){
            plotly::ggplotly(p  + geom_text(aes(label=Freq))) %>% layout(xaxis = list(tickangle =-45),autosize = T, height = 800)
          } else{
            plotly::ggplotly(p)
          }
          
        }
        # ply1 <- plotly::ggplotly(p)
        # ply2 <- plotly::ggplotly(p2)
        # plotly::subplot(ply1, ply2, nrows=1, shareY = TRUE, shareX = TRUE )
        # p # static
        # plotly::ggplotly(p) # dynamic
        # }
        
      })
      
    }
  } else{
    cat("No designation x environment interaction present. Skipping correlation plot.")
  }
}
```

</div>

### Predictions: Biplot by trait

The following graph allows to see the clustering of different genotypes in the TPE for each trait. This can help you identify clusters of environments and better define the TPE.

<p>&nbsp;</p>

```{r}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    if(any(c("environment_designation", "designation_environment",
             "environment:designation", "designation:environment") %in% effectTypes2)){
      inputPanel(
        selectInput(ns("traitBiplot"),
                    label = "Trait:",
                    choices = traits
        ),
      )
    }
  }
}
```

<div style = "width:auto; height:auto; overflow:auto">

```{r, results='asis'}

if (is.null(result)){
  # invisible(NULL)
} else {
  # if(length(setdiff(effectTypes, c("environment", "designation", "(Intercept)"))) > 0){
  if(gxeModelNum == 2){
    if("params" %in% ls()){ # we will download the document
      cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
      
      for(iTrait3 in traits){
        cat("\n\n####", iTrait3, "{.tabset .tabset-pills}       \n\n")
        
        level1 <- predSplitSplitWide[effectTypes[which(effectTypes %in% gxeList[[2]])]]
        level2 <- lapply(level1, function(x){x[iTrait3]})
        level3 <- NULL
        for(i in 1:length(level2)){
          level2[[i]][[1]]$designation <- gsub(paste0(colnames(level2[[i]][[1]]),":"),"",
                                               gsub(paste0(":",colnames(level2[[i]][[1]])),"",
                                                    row.names(level2[[i]][[1]])))
          if(i == 1){
            level3 <- level2[[i]][[1]]
          } else{
            level3 <- merge(level3, level2[[i]][[1]], by = "designation", all = TRUE)
          }
        }
        row.names(level3) <-level3$designation
        level3 <- level3[,-1]
        
        if(length(level3)>0){
          N <- sort(unlist(lapply(level3,length)), decreasing = TRUE)
          pred.mta.biplot <- do.call(cbind,level3[names(N)[which(N == N[1])]])
          rownames(pred.mta.biplot) <- rownames(level2[[ which(N == N[1])[1] ]][[1]] )
          for(i in 1:ncol(pred.mta.biplot)){
            rownames(pred.mta.biplot) <- gsub(paste0(colnames(pred.mta.biplot)[i],":"),"",
                                              gsub(paste0(":",colnames(pred.mta.biplot)[i]),"",
                                                   rownames(pred.mta.biplot)))
          }
          # colnames(pred.mta.biplot) <- substring(colnames(pred.mta.biplot),4,nchar(colnames(pred.mta.biplot)))
          pred.mta.biplot <- pred.mta.biplot[,apply(pred.mta.biplot,2,enhancer::propMissing) == 0]
          if( all(c( ncol(pred.mta.biplot) >1, is.matrix(pred.mta.biplot)  ) ) ){
            prin_comp <- prcomp(pred.mta.biplot, rank = 2)
            components <- prin_comp[["x"]]
            components <- data.frame(components)
            components$PC2 <- -components$PC2
            
            explained_variance <- summary(prin_comp)[["sdev"]]
            explained_variance <- explained_variance[1:2]
            comp <- prin_comp[["rotation"]]
            comp[,'PC2'] <- - comp[,'PC2']
            loadings <- comp
            for (i in seq(explained_variance) ){
              loadings[,i] <- comp[,i] * explained_variance[i]
            }
            p.mta.biplot <- plotly::plot_ly (type = 'scatter', mode = 'markers')
            p.mta.biplot <- plotly::add_trace(p.mta.biplot, 
                                              x=components$PC1,
                                              y=components$PC2,
                                              text=rownames(components),
                                              marker=list(color="blue"),
                                              hoverinfo='text',
                                              showlegend=FALSE
            )
            for (i in 1:ncol(pred.mta.biplot)){
              p.mta.biplot <- p.mta.biplot %>%
                plotly::add_segments(x = 0, xend = loadings[i, 1], y = 0, yend = loadings[i, 2], line = list(color = 'red'),inherit = FALSE, showlegend = FALSE)
              if(ncol(pred.mta.biplot) < 20){
                p.mta.biplot <- p.mta.biplot %>%
                  plotly::add_annotations(x=loadings[i, 1], y=loadings[i, 2], ax = 0, ay = 0,text = rownames(loadings)[i], xanchor = 'center', yanchor= 'bottom')
              }
            }
          }else{
            p.mta.biplot <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available or not enough environments")
          }
        }else{ # end of if length > 0
          p.mta.biplot <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available or not enough environments")
        }
        print(htmltools::tagList(plotly::ggplotly(p.mta.biplot)))
      }# end of for each trait
      
    }else{
      
      plotly::renderPlotly({
        level1 <- predSplitSplitWide[effectTypes[which(effectTypes %in% gxeList[[2]])]]
        level2 <- lapply(level1, function(x){x[input$traitBiplot]})
        level3 <- NULL
        for(i in 1:length(level2)){
          level2[[i]][[1]]$designation <- gsub(paste0(colnames(level2[[i]][[1]]),":"),"",
                                               gsub(paste0(":",colnames(level2[[i]][[1]])),"",
                                                    row.names(level2[[i]][[1]])))
          if(i == 1){
            level3 <- level2[[i]][[1]]
          } else{
            level3 <- merge(level3, level2[[i]][[1]], by = "designation", all = TRUE)
          }
        }
        row.names(level3) <-level3$designation
        level3 <- level3[,-1]
        
        if(length(level3) > 0){
          N <- sort(unlist(lapply(level3,length)), decreasing = TRUE)
          pred.mta.biplot <- do.call(cbind,level3[names(N)[which(N == N[1])]])
          rownames(pred.mta.biplot) <- rownames(level2[[ which(N == N[1])[1] ]][[1]] )
          for(i in 1:ncol(pred.mta.biplot)){
            rownames(pred.mta.biplot) <- gsub(paste0(colnames(pred.mta.biplot)[i],":"),"",
                                              gsub(paste0(":",colnames(pred.mta.biplot)[i]),"",
                                                   rownames(pred.mta.biplot)))
          }
          # colnames(pred.mta.biplot) <- substring(colnames(pred.mta.biplot),4,nchar(colnames(pred.mta.biplot)))
          pred.mta.biplot <- pred.mta.biplot[,apply(pred.mta.biplot,2,enhancer::propMissing) == 0]
          if( all(c( ncol(pred.mta.biplot) >1, is.matrix(pred.mta.biplot) ) ) ){
            prin_comp <- prcomp(pred.mta.biplot, rank = 2)
            components <- prin_comp[["x"]]
            components <- data.frame(components)
            components$PC2 <- -components$PC2
            
            explained_variance <- summary(prin_comp)[["sdev"]]
            explained_variance <- explained_variance[1:2]
            comp <- prin_comp[["rotation"]]
            comp[,'PC2'] <- - comp[,'PC2']
            loadings <- comp
            for (i in seq(explained_variance) ){
              loadings[,i] <- comp[,i] * explained_variance[i]
            }
            p.mta.biplot <- plotly::plot_ly (type = 'scatter', mode = 'markers')
            p.mta.biplot <- plotly::add_trace(p.mta.biplot, 
                                              x=components$PC1,
                                              y=components$PC2,
                                              text=rownames(components),
                                              marker=list(color="blue"),
                                              hoverinfo='text',
                                              showlegend=FALSE
            )
            for (i in 1:ncol(pred.mta.biplot)){
              p.mta.biplot <- p.mta.biplot %>%
                plotly::add_segments(x = 0, xend = loadings[i, 1], y = 0, yend = loadings[i, 2], line = list(color = 'red'),inherit = FALSE, showlegend = FALSE)
              if(ncol(pred.mta.biplot) < 20){
                p.mta.biplot <- p.mta.biplot %>%
                  plotly::add_annotations(x=loadings[i, 1], y=loadings[i, 2], ax = 0, ay = 0,text = rownames(loadings)[i], xanchor = 'center', yanchor= 'bottom')
              }
            }
          }else{
            px <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available or not enough environments")
            p.mta.biplot <- plotly::ggplotly(px) 
          } # end of if cols > 1
        }else{
          px <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available or not enough environments")
          p.mta.biplot <- plotly::ggplotly(px) 
        }
        plotly::layout(p.mta.biplot) # ,font=font.plot_ly
      }) # end of plotly
      
    }
    
  } else if(gxeModelNum == 1){
    if(any(effectTypes %in% gxeList[[1]])){
      
      if("params" %in% ls()){ # we will download the document
        
        cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
        
        # for(iType3 in setdiff(effectTypes, c("environment", "designation", "(Intercept)"))){
        #   cat("\n\n####", iType3, "{.tabset .tabset-pills}       \n\n")
        
        for(iTrait3 in traits){
          cat("\n\n####", iTrait3, "{.tabset .tabset-pills}       \n\n")
          
          level1 <- predSplitSplitWide[effectTypes[which(effectTypes %in% gxeList[[1]])]]
          level2 <- lapply(level1, function(x){x[iTrait3]})
          level3 <- lapply(rapply(level2, enquote, how="unlist"), eval)
          
          if(length(level3)>0){
            N <- sort(unlist(lapply(level3,length)), decreasing = TRUE)
            pred.mta.biplot <- do.call(cbind,level3[names(N)[which(N == N[1])]])
            colnames(pred.mta.biplot) <- gsub(paste0("environment_designation.",iTrait3,"."),"",
                                              gsub(paste0("designation_environment.",iTrait3,"."),"",
                                                   gsub(paste0("environment:designation.",iTrait3,"."),"",
                                                        gsub(paste0("designation:environment.",iTrait3,"."),"",
                                                             colnames(pred.mta.biplot))))) 
            rownames(pred.mta.biplot) <- rownames(level2[[ which(N == N[1])[1] ]][[1]] )
            for(i in 1:ncol(pred.mta.biplot)){
              rownames(pred.mta.biplot) <- gsub(paste0(colnames(pred.mta.biplot)[i],":"),"",
                                                gsub(paste0(":",colnames(pred.mta.biplot)[i]),"",
                                                     rownames(pred.mta.biplot)))
            }
            # 
            # for (j in unique(result$data$pheno$environment)){
            #   tempA <- gsub(j, "",rownames(level2[[ which(N == N[1])[1] ]][[1]] ))
            #   a <- c(tempA,a)
            # }
            pred.mta.biplot <- pred.mta.biplot[,apply(pred.mta.biplot,2,enhancer::propMissing) == 0]
            if( all(c( ncol(pred.mta.biplot) >1, is.matrix(pred.mta.biplot)  ) ) ){
              prin_comp <- prcomp(pred.mta.biplot, rank = 2)
              components <- prin_comp[["x"]]
              components <- data.frame(components)
              components$PC2 <- -components$PC2
              
              explained_variance <- summary(prin_comp)[["sdev"]]
              explained_variance <- explained_variance[1:2]
              comp <- prin_comp[["rotation"]]
              comp[,'PC2'] <- - comp[,'PC2']
              loadings <- comp
              for (i in seq(explained_variance) ){
                loadings[,i] <- comp[,i] * explained_variance[i]
              }
              p.mta.biplot <- plotly::plot_ly (type = 'scatter', mode = 'markers')
              p.mta.biplot <- plotly::add_trace(p.mta.biplot, 
                                                x=components$PC1,
                                                y=components$PC2,
                                                text=rownames(components),
                                                marker=list(color="blue"),
                                                hoverinfo='text',
                                                showlegend=FALSE
              )
              for (i in 1:ncol(pred.mta.biplot)){
                p.mta.biplot <- p.mta.biplot %>%
                  plotly::add_segments(x = 0, xend = loadings[i, 1], y = 0, yend = loadings[i, 2], line = list(color = 'red'),inherit = FALSE, showlegend = FALSE)
                if(ncol(pred.mta.biplot) < 20){
                  p.mta.biplot <- p.mta.biplot %>%
                    plotly::add_annotations(x=loadings[i, 1], y=loadings[i, 2], ax = 0, ay = 0,text = rownames(loadings)[i], xanchor = 'center', yanchor= 'bottom')
                }
              }
            }else{
              p.mta.biplot <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available or not enough environments")
            }
          }else{ # end of if length > 0
            p.mta.biplot <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available or not enough environments")
          }
          print(htmltools::tagList(plotly::ggplotly(p.mta.biplot)))
        }# end of for each trait
        
        # }
        
      }else{
        
        plotly::renderPlotly({
          # if(length(input$biplotMtaPredsType) > 0){
          
          # level1 <- predSplitSplitWide[input$biplotMtaPredsType]
          
          level1 <- predSplitSplitWide[effectTypes[which(effectTypes %in% gxeList[[1]])]]
          level2 <- lapply(level1, function(x){x[input$traitBiplot]})
          level3 <- lapply(rapply(level2, enquote, how="unlist"), eval)
          
          if(length(level3) > 0){
            N <- sort(unlist(lapply(level3,length)), decreasing = TRUE)
            pred.mta.biplot <- do.call(cbind,level3[names(N)[which(N == N[1])]])
            rownames(pred.mta.biplot) <- rownames(level2[[ which(N == N[1])[1] ]][[1]] )
            colnames(pred.mta.biplot) <- gsub(paste0("environment_designation.",input$traitBiplot,"."),"",
                                              gsub(paste0("designation_environment.",input$traitBiplot,"."),"",
                                                   gsub(paste0("environment:designation.",input$traitBiplot,"."),"",
                                                        gsub(paste0("designation:environment.",input$traitBiplot,"."),"",
                                                             colnames(pred.mta.biplot))))) 
            rownames(pred.mta.biplot) <- rownames(level2[[ which(N == N[1])[1] ]][[1]] )
            for(i in 1:ncol(pred.mta.biplot)){
              rownames(pred.mta.biplot) <- gsub(paste0(colnames(pred.mta.biplot)[i],":"),"",
                                                gsub(paste0(":",colnames(pred.mta.biplot)[i]),"",
                                                     rownames(pred.mta.biplot)))
            }
            pred.mta.biplot <- pred.mta.biplot[,apply(pred.mta.biplot,2,enhancer::propMissing) == 0]
            if( all(c( ncol(pred.mta.biplot) >1, is.matrix(pred.mta.biplot) ) ) ){
              prin_comp <- prcomp(pred.mta.biplot, rank = 2)
              components <- prin_comp[["x"]]
              components <- data.frame(components)
              components$PC2 <- -components$PC2
              
              explained_variance <- summary(prin_comp)[["sdev"]]
              explained_variance <- explained_variance[1:2]
              comp <- prin_comp[["rotation"]]
              comp[,'PC2'] <- - comp[,'PC2']
              loadings <- comp
              for (i in seq(explained_variance) ){
                loadings[,i] <- comp[,i] * explained_variance[i]
              }
              p.mta.biplot <- plotly::plot_ly (type = 'scatter', mode = 'markers')
              p.mta.biplot <- plotly::add_trace(p.mta.biplot, 
                                                x=components$PC1,
                                                y=components$PC2,
                                                text=rownames(components),
                                                marker=list(color="blue"),
                                                hoverinfo='text',
                                                showlegend=FALSE
              )
              for (i in 1:ncol(pred.mta.biplot)){
                p.mta.biplot <- p.mta.biplot %>%
                  plotly::add_segments(x = 0, xend = loadings[i, 1], y = 0, yend = loadings[i, 2], line = list(color = 'red'),inherit = FALSE, showlegend = FALSE)
                if(ncol(pred.mta.biplot) < 20){
                  p.mta.biplot <- p.mta.biplot %>%
                    plotly::add_annotations(x=loadings[i, 1], y=loadings[i, 2], ax = 0, ay = 0,text = rownames(loadings)[i], xanchor = 'center', yanchor= 'bottom')
                }
              }
            }else{
              px <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available or not enough environments")
              p.mta.biplot <- plotly::ggplotly(px) 
            } # end of if cols > 1
          }else{
            px <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available or not enough environments")
            p.mta.biplot <- plotly::ggplotly(px) 
          }
          # }else{
          #   px <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available or not enough environments")
          #   p.mta.biplot <- plotly::ggplotly(px) 
          # } # end of if biplotMtaPredsType has length > 0
          plotly::layout(p.mta.biplot) # ,font=font.plot_ly
        }) # end of plotly
      }
    } else{
      cat("No designation x environment interaction present. Skipping biplot.")
    }
  } else{
    cat("No designation x environment interaction present. Skipping biplot.")
  }
}
```

</div>

### Predictions: Merit estimates of top entries

In the left-side plot you can observe the comparison between the top 100 entries from each effect type category for the different traits. If a category has less than a 100 entries all individuals are displayed. This should allow you to identify the entries that could potentially become parents or nominated for advanced stages of evaluation. We would recommend you to wait until a selection index is calculated. In the right-side plot you can see a boxplot of the entire distribution of values for each effectType category so you know until which trait-values the rest of the entries that are not plotted in the left-side plot reach. 


```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{ # input <- list(scatterMtaPredsType=effectTypes[1], scatterMtaPredsTrait=traits[1], scatterMtaPredsEnvir=envirTypes[1])
    inputPanel(
      selectInput(ns("scatterMtaPredsType"),
                  label = "Effect type:",
                  choices = setdiff(effectTypes2, c("environment", "(Intercept)"))
                  # effectTypes[effectTypes %in% c("environment_designation", "designation")]
                  # choices = effectTypes, multiple = TRUE,
                  # selected=effectTypes[1]
      ),
      selectInput(ns("scatterMtaPredsTrait"),
                  label = "Trait",
                  choices = traits,
                  selected=traits[1]
      ),
      selectInput(ns("scatterMtaPredsEnvir"),
                  label = "Environment",
                  choices = envirTypes,
                  selected=envirTypes[1]
      ),
      numericInput(ns("scatterMtaPredsTop"),
                   label = "Number of Entries",
                   value = 100,
                   min = 1
      ),
    )
  }
}
```

<div style = "width:auto; height:auto; overflow:auto">

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    
    cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
    
    for(iType4 in setdiff(effectTypes2, c("environment", "(Intercept)"))){
      cat("\n\n####", iType4, "{.tabset .tabset-pills}       \n\n")
      for(iTrait4 in traits){
        cat("\n\n#####", iTrait4, "{.tabset .tabset-pills}       \n\n")
        if(!(iType4 %in% c("environment_designation", "designation_environment",
                           "environment:designation", "designation:environment"))){
          envirTypes2 <- c("(Intercept)")
        } else{
          envirTypes2 <- setdiff(envirTypes,c("(Intercept)"))
        }
        for(iEnv4 in envirTypes2){
          cat("\n\n######", iEnv4, "{.tabset .tabset-pills}       \n\n")
          if(gxeModelNum == 2){
            if(iType4 %in% c("environment_designation","environment:designation")){
              level1 <- predSplitSplit[effectTypes[which(effectTypes %in% gxeList[[2]])]]
              level2 <- lapply(level1, function(x){x[iTrait4]})
            } else{
              level1 <- predSplitSplit[iType4]
              level2 <- lapply(level1, function(x){x[iTrait4]})
            }
          } else{
            level1 <- predSplitSplit[iType4]
            level2 <- lapply(level1, function(x){x[iTrait4]})
          }
          level3 <- list()
          for(o in 1:length(level2)){level3[[o]] <- level2[[o]][[1]]}
          temp.pred.box.top <- do.call(rbind,level3)
          temp.pred.box.top <- temp.pred.box.top[which(temp.pred.box.top$environment %in% iEnv4),]
          if(!is.null(temp.pred.box.top)){
            if(nrow(temp.pred.box.top) > 0){
              ub <- max(temp.pred.box.top$predictedValue + temp.pred.box.top$stdError, na.rm = TRUE)
              lb <- min(temp.pred.box.top$predictedValue - temp.pred.box.top$stdError, na.rm = TRUE)
              
              p1 <- ggplot2::ggplot(temp.pred.box.top, ggplot2::aes(x=reorder(entryType, -predictedValue), y=predictedValue) ) +
                ggplot2::geom_boxplot(alpha = 0.5, show.legend = TRUE, ggplot2::aes(colour=factor(entryType))) +
                ggplot2::theme(axis.text.x = ggplot2::element_blank(), legend.position = "bottom", legend.title = element_blank() ) + # element_text(angle = 45, vjust = 0.5, size = 7, face = "bold")
                ggplot2::labs(x="", y="Trait value", color="Effect type") +  
                ggplot2::guides(colour = guide_legend(nrow = ceiling(length(unique(temp.pred.box.top$entryType))/3) )) +  ggplot2::scale_y_continuous(limits = c(lb,ub) ) 
              
              # plot 2
              temp.pred.box.top.list <- split(temp.pred.box.top, temp.pred.box.top$effectType)
              temp.pred.box.top.list <- lapply(temp.pred.box.top.list, function(x){
                x.c <- x[which(grepl("check|chck|bnchmark|benchmark|chek|chec",
                                    x$entryType, ignore.case = TRUE)),]
                x <- x[which(!grepl("check|chck|bnchmark|benchmark|chek|chec",
                                    x$entryType, ignore.case = TRUE)),]
                x <- x[with(x, order(-predictedValue)), ]
                x <- x[1:min(c(nrow(x), 100)),]
                x <- rbind(x, x.c)
                x <- x[order(-x$predictedValue), ]
                return(x)
              })
              temp.pred.box.top2 <- as.data.frame(do.call(rbind, temp.pred.box.top.list))
              temp.pred.box.top2$entryTypeSimple <- LETTERS[as.numeric(as.factor(temp.pred.box.top2$entryType))]
              
              p2 <- ggplot2::ggplot(
                data = temp.pred.box.top2,
                mapping = ggplot2::aes(
                  x = reorder(designation, -predictedValue),
                  y = predictedValue,
                  col = entryType,
                  text = paste(
                    paste0('<br>', "designation: ", designation),
                    paste0('<br>', "entry type: ", entryType),
                    paste0('<br>', "predicted value: ", round(predictedValue, 4)),
                    paste0('<br>', "standard error: ", round(stdError, 4)),
                    # paste0("<br>95% CI: (", round(predictedValue - CI, 2) , ", ",
                    #        round(predictedValue + CI, 2), ")")))) +
                    paste0("<br>predVal +- stdError: (", round(predictedValue - stdError, 2) , ", ",
                           round(predictedValue + stdError, 4), ")")))) +
                ggplot2::theme_classic() +  ggplot2::geom_point() +
                ggplot2::geom_errorbar(ggplot2::aes(ymax = predictedValue + stdError,
                                                    ymin = predictedValue - stdError), position = ggplot2::position_dodge(0.5),
                                       width = 0.10) +
                ggplot2::labs(title = "", x = "Designation", y = "Predicted value", color = "Entry type") +
                ggplot2::theme(axis.text.x = ggplot2::element_blank() , legend.position = "none", legend.title = element_blank() ) +   # element_text(angle = 45, vjust = 0.5, size = 7, face = "bold")
                ggplot2::guides(colour = guide_legend(nrow = ceiling(length(unique(temp.pred.box.top2$entryType))/3) )) + 
                ggplot2::scale_y_continuous(limits = c(lb,ub) ) 
            }else{
              p1 <- p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
            }
          }else{
            p1 <- p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
          }
          ply1 <- plotly::ggplotly(p1)
          ply2 <- plotly::ggplotly(p2)
          print(htmltools::tagList( plotly::subplot(plotly::ggplotly(p2), plotly::ggplotly(p1), shareX = FALSE, nrows=2, heights = c(0.75,0.25) ) ) )
        } # enf of nrow > 0
        
      }
    }
    
  }else{
    
    plotly::renderPlotly({
      
      if(length(input$scatterMtaPredsType) > 0){
        
        if(gxeModelNum == 2){
          if(input$scatterMtaPredsType %in% c("environment_designation","environment:designation")){
            level1 <- predSplitSplit[effectTypes[which(effectTypes %in% gxeList[[2]])]]
            level2 <- lapply(level1, function(x){x[input$scatterMtaPredsTrait]})
          } else{
            level1 <- predSplitSplit[input$scatterMtaPredsType]
            level2 <- lapply(level1, function(x){x[input$scatterMtaPredsTrait]})
          }
        } else{
          level1 <- predSplitSplit[input$scatterMtaPredsType]
          level2 <- lapply(level1, function(x){x[input$scatterMtaPredsTrait]})
        }
        
        level3 <- list()
        for(o in 1:length(level2)){level3[[o]] <- level2[[o]][[1]]}
        temp.pred.box.top <- do.call(rbind,level3)
        temp.pred.box.top <- temp.pred.box.top[which(temp.pred.box.top$environment %in% input$scatterMtaPredsEnvir),]
        
        if(!is.null(temp.pred.box.top)){
          if(nrow(temp.pred.box.top) > 0){
            ub <- max(temp.pred.box.top$predictedValue + temp.pred.box.top$stdError, na.rm = TRUE)
            lb <- min(temp.pred.box.top$predictedValue - temp.pred.box.top$stdError, na.rm = TRUE)
            
            p1 <- ggplot2::ggplot(temp.pred.box.top, ggplot2::aes(x=reorder(entryType, -predictedValue), y=predictedValue) ) +
              ggplot2::geom_boxplot(alpha = 0.5, show.legend = TRUE, ggplot2::aes(colour=factor(entryType))) +
              ggplot2::theme(axis.text.x = ggplot2::element_blank(), legend.position = "bottom", legend.title = element_blank() ) + # element_text(angle = 45, vjust = 0.5, size = 7, face = "bold")
              ggplot2::labs(x="", y="Trait value", color="Entry type") +  
              ggplot2::guides(colour = guide_legend(nrow = ceiling(length(unique(temp.pred.box.top$entryType))/3) )) +  ggplot2::scale_y_continuous(limits = c(lb,ub) ) 
            
            # plot 2
            temp.pred.box.top.list <- split(temp.pred.box.top, temp.pred.box.top$effectType)
            temp.pred.box.top.list <- lapply(temp.pred.box.top.list, function(x){
              # x.c <- x[which(grepl("check", tolower(x$entryType))),]
              # x <- x[which(!grepl("check", tolower(x$entryType))),]
              x.c <- x[which(grepl("check|chck|bnchmark|benchmark|chek|chec",
                                  x$entryType, ignore.case = TRUE)),]
              x <- x[which(!grepl("check|chck|bnchmark|benchmark|chek|chec",
                                 x$entryType, ignore.case = TRUE)),]
              x <- x[with(x, order(-predictedValue)), ]
              x <- x[1:min(c(nrow(x), ifelse(!is.na(input$scatterMtaPredsTop), input$scatterMtaPredsTop,100))),]
              x <- rbind(x, x.c)
              x <- x[order(-x$predictedValue), ]
              return(x)
            })
            temp.pred.box.top2 <- as.data.frame(do.call(rbind, temp.pred.box.top.list))
            temp.pred.box.top2$entryTypeSimple <- LETTERS[as.numeric(as.factor(temp.pred.box.top2$entryType))]
            
            p2 <- ggplot2::ggplot(
              data = temp.pred.box.top2,
              mapping = ggplot2::aes(
                x = reorder(designation, -predictedValue),
                y = predictedValue,
                col = entryType,
                text = paste(
                  paste0('<br>', "designation: ", designation),
                  paste0('<br>', "entry type: ", entryType),
                  paste0('<br>', "predicted value: ", round(predictedValue, 4)),
                  paste0('<br>', "standard error: ", round(stdError, 4)),
                  # paste0("<br>95% CI: (", round(predictedValue - CI, 2) , ", ",
                  #        round(predictedValue + CI, 2), ")")))) +
                  paste0("<br>predVal +- stdError: (", round(predictedValue - stdError, 2) , ", ",
                         round(predictedValue + stdError, 4), ")")))) +
              ggplot2::theme_classic() +  ggplot2::geom_point() +
              ggplot2::geom_errorbar(ggplot2::aes(ymax = predictedValue + stdError,
                                                  ymin = predictedValue - stdError), position = ggplot2::position_dodge(0.5),
                                     width = 0.10) +
              ggplot2::labs(title = "", x = "Designation", y = "Predicted value", color = "Effect type") +
              ggplot2::theme(axis.text.x = ggplot2::element_blank() , legend.position = "none", legend.title = element_blank() ) +   # element_text(angle = 45, vjust = 0.5, size = 7, face = "bold")
              ggplot2::guides(colour = guide_legend(nrow = ceiling(length(unique(temp.pred.box.top2$entryType))/3) )) + 
              ggplot2::scale_y_continuous(limits = c(lb,ub) ) 
            
            
          }else{
            p1 <- p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
          } # enf of nrow > 0
        }else{
          p1 <- p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
        }
        ply1 <- plotly::ggplotly(p1)
        ply2 <- plotly::ggplotly(p2)
        plotly::subplot(ply2, ply1, nrows=2, shareX = FALSE,  heights = c(0.75,0.25))
        
      } # end of length(input$scatterMtaPredsType) > 0
      
    })
    
  }
}
```

</div>

### Predictions: Table of estimates 

The following table allows you to inspect the trait predictions in wide format together with the QTL profile (in case those are available) to understand the type of data that would be used to calculate a selection index (e.g., desire index).

<p>&nbsp;</p>

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  pred.mta <- subset(pred, select = c(trait,designation,environment,effectType,entryType,predictedValue))
  pred.mta.wide <- reshape(
    data=pred.mta, 
    timevar = c("trait"), 
    idvar = c("environment","designation","effectType","entryType"), 
    direction="wide"
  )
  names(pred.mta.wide) <- gsub("predictedValue.","",colnames(pred.mta.wide)) 
  
  # to show only 1 row per designation effectType (concatenated entryType)
  pred.mta.wide1 <- pred.mta.wide[which(pred.mta.wide$effectType != "designation"),]
  pred.mta.wide2 <- pred.mta.wide[which(pred.mta.wide$effectType == "designation"),]
  
  coalesce_by_column <- function(x) {
  args <- as.list(x)
    
  # Drop NULLs and length-0 elements 
  keep <- vapply(args, function(z) !(is.null(z) || length(z) == 0L), logical(1))
  args <- args[keep]
    
  #NULL/empty define as NA 
  if (length(args) == 0L) return(NA_real_)
    
  #Coalesce values
  dplyr::coalesce(!!!args)
  }
  
  pred.mta.wide2_trait <- pred.mta.wide2 %>% select(c("designation", all_of(traits))) %>%
    group_by(designation) %>% 
    summarise(across(traits, coalesce_by_column)) 
  
  pred.mta.wide2_meta <- pred.mta.wide2 %>% select(!all_of(traits)) %>%
    arrange(designation, entryType) %>% 
    group_by(designation) %>% 
    mutate(entryType2 = paste0(entryType, collapse = "#")) %>%
    filter(duplicated(designation) == FALSE)
  
  pred.mta.wide2 <- merge(pred.mta.wide2_meta, pred.mta.wide2_trait, by = "designation")
  pred.mta.wide2 <- subset(pred.mta.wide2, select = -c(entryType))
  pred.mta.wide2 <- rename(pred.mta.wide2, entryType = entryType2)
  
  pred.mta.wide <- rbind(pred.mta.wide1, pred.mta.wide2)

  # to deduct intercept from the envIndex_designation effectType to get slope for FW model
  # if(params$modelUsed == "fw_model"){
  #   for (i in 1:length(unique(pred.mta$trait))){
  #     pred.mta.wide[grepl("envIndex_designation",pred.mta.wide$effectType),unique(pred.mta$trait)[i]] <-
  #       pred.mta.wide[grepl("envIndex_designation",pred.mta.wide$effectType),unique(pred.mta$trait)[i]] - pred.mta.wide[pred.mta.wide$effectType == "(Intercept)",unique(pred.mta$trait)[i]]
  #   }
  # }
    
  if(!is.null(result$data$qtl)  ){ # there's QTL information
    if(nrow(result$data$qtl)>0){
      qtls <- result$data$qtl
      byMerge <- result$metadata$qtl$value[which(result$metadata$qtl$parameter == "designation")]
      pred.mta.wide <- merge(pred.mta.wide, qtls, by.x="designation", by.y=byMerge, all.x = TRUE)
    }
  }
  
  if("params" %in% ls()){ # we will download the document
    printTable(pred.mta.wide, autoWidthOpt = TRUE, scrollXOpt = FALSE,
               colNames = colnames(pred.mta.wide),
               numericColNames = traits,
               numRound = 4)
  }else{
    DT::renderDT(printTable(pred.mta.wide,
                            autoWidthOpt = FALSE, scrollXOpt = TRUE,
                            colNames = colnames(pred.mta.wide),
                            numericColNames = traits,
                            numRound = 4), server = FALSE)
  }
  
}

```

### Predictions: Correlations between traits

The following heatmap and histogram allows to see the genetic correlations among traits calculated using across environment estimates of merit for the different traits. This can be used to understand the implications of selecting for a set of traits to achieve a product profile and make neccesary adjustment to the selection strategy.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{ # input <- list(corrplotTraitMtaPredsType=effectTypes[1],corrplotTraitMtaPredsEnv=envirTypes[1])
    if(length(traits) > 1){
      inputPanel(
        # selectInput(ns("corrplotTraitMtaPredsType"),
        #             label = "Effect type",
        #             choices = setdiff(effectTypes,c("(Intercept")), multiple = FALSE,
        # ),
        # column(width=12),
        # selectInput(ns("corrplotTraitMtaPredsEnv"),
        #             label = "Environment", 
        #             choices = envirTypes
        # ),
        shiny::checkboxInput(ns("freqCor"),
                             label = "Show frequency distribution",
                             value = TRUE
        )
      )
    }
  }
}

```

<div style = "width:auto; height:auto; overflow:auto">

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if(length(traits) > 1){
    # if(nrow(corr) > 1){
    if("params" %in% ls()){ # we will download the document
      
      cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
      
      # for(iType5 in setdiff(effectTypes, c("(Intercept)"))){
      #   cat("\n\n####", iType5, "{.tabset .tabset-pills}       \n\n")
      
      #       if(iType5 %in% c("environment", "designation")){
      #     envirTypes3 <- c("(Intercept)")
      # } else{
      #     envirTypes3 <- setdiff(envirTypes,c("(Intercept)"))
      #   }
      
      # for(iEnv5 in envirTypes3){
      #   cat("\n\n#####", iEnv5, "{.tabset .tabset-pills}       \n\n")
      
      # wideCorTrait2 <- predSplitSplitWide2[[iType5]][[iEnv5]]
      wideCorTrait2 <- predSplitSplitWide2[["designation"]][["(Intercept)"]]
      if(!is.null(wideCorTrait2)){
        
        if(is.matrix(wideCorTrait2)){
          wideCorTrait2 <- apply(wideCorTrait2,2,enhancer::imputev)
          corrX <- round(stats::cor(wideCorTrait2, use="pairwise.complete.obs"),2)
          mydata4CorTraits <- cgiarBase::matToTab(corrX)
          p <- ggplot2::ggplot(data = mydata4CorTraits, ggplot2::aes(Var2, Var1, fill = Freq))+
            ggplot2::geom_tile(color = "white")+
            ggplot2::scale_fill_gradient2(low = "#E46726", high = "#038542", mid = "white",
                                          midpoint = 0, limit = c(-1,1), space = "Lab",
                                          name="Pearson\nCorrelation") +
            ggplot2::theme_minimal()+
            ggplot2::ylab("") + ggplot2::xlab("") +
            ggplot2::coord_fixed()
          # ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black"), 
          #                axis.text.x = ggplot2::element_text(angle = 45, hjust=1, vjust = 1))
          # if(nrow(corrX) < 8){ # if user wants to fill cell values
          #   p <- p + ggplot2::geom_text(ggplot2::aes(label = round(Freq,2) ), color = "black", size = 4)
          # }
          
          # Basic histogram
          mydata4CorTraits$Group <- ifelse(mydata4CorTraits$Freq < 0, "Low", ifelse(mydata4CorTraits$Freq < 0.2, "Medium", "High"))
          group.colors <- c(Low="#E46726",Medium="white", High="#038542")
          p2 <- ggplot() + geom_histogram(data = mydata4CorTraits[which(mydata4CorTraits$Var1 != mydata4CorTraits$Var2),], 
                                          aes(x = Freq, fill = ..x..), binwidth=0.1 ) + 
            ggplot2::scale_x_continuous(limits = c(-1,1)) +
            scale_fill_gradient2(low='#E46726', mid='white', high='#038542', midpoint=0, 
                                 limit = c(-1,1), space = "Lab", 
                                 name="Pearson\nCorrelation") + 
            ggplot2::xlab("Correlation Value") + ggplot2::ylab("")
          # plotly::ggplotly(p2)
        }else{
          p <- p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
        } # end of nrow > 1
        
      }else{
        p <- p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
      }# end of each env
      
      for(iFreq in c("with frequency distribution", "without frequency distribution")){
        cat("\n\n####", iFreq, "{.tabset .tabset-pills}       \n\n")
        if(iFreq == "with frequency distribution"){
          # combine plots
          if(!is.null(wideCorTrait2) & is.matrix(wideCorTrait2)){
            print(htmltools::tagList( plotly::subplot(plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank()) + geom_text(aes(label=Freq), size = 2)) , plotly::ggplotly(p2), nrows=1, shareY = FALSE, shareX = FALSE ) %>% layout(autosize = T, width = 1600)))
          }else{
            print(htmltools::tagList( plotly::subplot(plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank())) , plotly::ggplotly(p2), nrows=1, shareY = FALSE, shareX = FALSE ) ))
          }
        } else{
          if(!is.null(wideCorTrait2) & is.matrix(wideCorTrait2)){
            print(htmltools::tagList(plotly::ggplotly(p + geom_text(aes(label=Freq), size = 3)) %>% layout(xaxis = list(tickangle =-45),autosize = T, height = 800)))
          }else{
            print(htmltools::tagList(plotly::ggplotly(p)))
          }
        }
      }
      
      # }
      
      # }
      
    }else{
      
      
      plotly::renderPlotly({
        # shiny::renderPlot({
        # wideCorTrait2 <- predSplitSplitWide2[[input$corrplotTraitMtaPredsType]][[input$corrplotTraitMtaPredsEnv]]
        wideCorTrait2 <- predSplitSplitWide2[["designation"]][["(Intercept)"]]
        if(!is.null(wideCorTrait2)){
          if(is.matrix(wideCorTrait2)){
            wideCorTrait2 <- apply(wideCorTrait2,2,enhancer::imputev)
            corrX <- round(stats::cor(wideCorTrait2, use="pairwise.complete.obs"),2)
            mydata4CorTraits <- cgiarBase::matToTab(corrX)
            
            p <- ggplot2::ggplot(data = mydata4CorTraits, ggplot2::aes(Var2, Var1, fill = Freq))+
              ggplot2::geom_tile(color = "white")+
              ggplot2::scale_fill_gradient2(low = "#E46726", high = "#038542", mid = "white",
                                            midpoint = 0, limit = c(-1,1), space = "Lab",
                                            name="Pearson\nCorrelation") +
              ggplot2::theme_minimal()+
              ggplot2::ylab("") + ggplot2::xlab("") +
              ggplot2::coord_fixed() + 
              ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black"), 
                             axis.text.x = ggplot2::element_text(angle = 45, hjust=1, vjust = 1))
            # if(nrow(corrX) < 8){ # if user wants to fill cell values
            #   p <- p + ggplot2::geom_text(ggplot2::aes(label = round(Freq,2) ), color = "black", size = 4)
            # }
            
            # Basic histogram
            mydata4CorTraits$Group <- ifelse(mydata4CorTraits$Freq < 0, "Low", ifelse(mydata4CorTraits$Freq < 0.2, "Medium", "High"))
            group.colors <- c(Low="#E46726",Medium="white", High="#038542")
            
            p2 <- ggplot() + geom_histogram(data = mydata4CorTraits[which(mydata4CorTraits$Var1 != mydata4CorTraits$Var2),], 
                                            aes(x = Freq, fill = ..x..), binwidth=0.1 ) + 
              ggplot2::scale_x_continuous(limits = c(-1,1)) +
              scale_fill_gradient2(low='#E46726', mid='white', high='#038542', midpoint=0, 
                                   limit = c(-1,1), space = "Lab", 
                                   name="Pearson\nCorrelation") + 
              ggplot2::xlab("Correlation Value") + ggplot2::ylab("")
          }else{
            p <- p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
          }
        }else{
          p <- p2 <- ggplot2::ggplot() + ggplot2::ggtitle("Combination not available")
        }
        # combine plots
        if(input$freqCor == TRUE){
          # combine plots
          if(!is.null(wideCorTrait2) & is.matrix(wideCorTrait2)){
            ply1 <- plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank()) + geom_text(aes(label=Freq), size = 3)) 
            ply2 <- plotly::ggplotly(p2)
            plotly::subplot(ply1, ply2, nrows=1, shareY = FALSE, shareX = FALSE ) %>% layout(autosize = T, width = 1600)
          } else{
            ply1 <- plotly::ggplotly(p + ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank())) 
            ply2 <- plotly::ggplotly(p2)
            plotly::subplot(ply1, ply2, nrows=1, shareY = FALSE, shareX = FALSE )
          }
          
        } else{
          if(!is.null(wideCorTrait2) & is.matrix(wideCorTrait2)){
            plotly::ggplotly(p  + geom_text(aes(label=Freq))) %>% layout(xaxis = list(tickangle =-45),autosize = T, height = 800)
          } else{
            plotly::ggplotly(p)
          }
          
        }
        # ply1 <- plotly::ggplotly(p)
        # ply2 <- plotly::ggplotly(p2)
        # plotly::subplot(ply1, ply2, nrows=1, shareY = TRUE, shareX = TRUE)
        # print(p)
      })
      
    }
  } else{
    cat("Only 1 trait selected. Skipping correlation between traits.")
  }
  
}
```

</div>

### Percent check comparison to top 30 entries

The following plots allow the user to compare the top 30 entries against the different checks/benchmarks present in the dataset for each trait. The table below that shows the wide-table of all possible comparisons for all traits.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  result$status <- result$status#[1:6,]
  idMta <- result$status[which(result$status$module == "mtaLmms"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred <- result$predictions
  pred <- pred[pred$analysisId == idMta,]

  traitChecks <- unique(pred$trait)

  entryTypePresent <- table((unique(pred[which(pred$effectType == "designation"),c("designation","entryType")]))$entryType)
  check1 <- which(entryTypePresent < 200)
  check2 <- grep("check|chck|bnchmark|benchmark|chek|chec", names(entryTypePresent), ignore.case = TRUE, value=FALSE)
  checksFound <- intersect(check1,check2)
  if(length(checksFound) > 0){
    labelChecks <- names(entryTypePresent)[checksFound]
    nameOfChecks <- unique(pred[ which(pred$entryType %in% labelChecks & pred$effectType == "designation") , "designation"])

    if("params" %in% ls()){ # we will download the document
      shinyjs::hide()
    }else{
      inputPanel(
        selectInput(ns("checkMta"),
                    label = "Check to compare to:",
                    choices = nameOfChecks
        ),
        selectInput(ns("checkMtaTrait"),
                    label = "View trait:",
                    choices = traitChecks
        ),
      )
    }
  }else{
    shinyjs::hide()
  }

}
```

<div style = "width:auto; height:auto; overflow:auto">

```{r, results='asis', fig.height=20}
if (is.null(result)){
  # invisible(NULL)
} else {

  if(length(checksFound) > 0){

    if("params" %in% ls()){ # we will download the document

      cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))

      for(iCheck in nameOfChecks){

        cat("\n\n####", iCheck, "{.tabset .tabset-pills}       \n\n")

        for(iTrait in traitChecks){

          cat("\n\n#####", iTrait, "{.tabset .tabset-pills}       \n\n")

          # predDta <- pred[which(pred$effectType == "designation"),]
          # predDtaTrait <- predDta$trait
          predListBench <- split(pred[which(pred$effectType == "designation"),], pred[which(pred$effectType == "designation"),]$trait)
          predListBench2 <- lapply(predListBench, function(x){ # x <- predListBench[[1]]
            checkSelected <- which(x$designation == iCheck)
            if(length(checkSelected) > 0){
              x$predictedValue2 <- ( (x$predictedValue / x$predictedValue[checkSelected]) - 1 ) * 100
              x$UB <- ( ( (x$predictedValue+x$stdError) / x$predictedValue[checkSelected]) - 1 ) * 100
              x$LB <- ( ( (x$predictedValue-x$stdError) / x$predictedValue[checkSelected]) - 1 ) * 100
              x <- x[with(x, order(-predictedValue2)), ]
              x_test <- x[which(!x$designation %in% nameOfChecks),]
              x_test <- x_test[c(1:min(c(30, nrow(x_test)))),]
              x <- x[c(which(x$designation %in% x_test$designation),which(x$designation %in% nameOfChecks)),]
              x <- x[with(x, order(-predictedValue2)), ]
              return(x)
            }
          })
          temp.pred.box.bench <- do.call(rbind, predListBench2)
          temp.pred.box.bench$entryTypeSimple <- LETTERS[as.numeric(as.factor(temp.pred.box.bench$entryType))]
          temp.pred.box.bench <- temp.pred.box.bench[which(temp.pred.box.bench$trait == iTrait),]
          if(nrow(temp.pred.box.bench)>0){
            pb <- ggplot2::ggplot(
            data = temp.pred.box.bench,
            mapping = ggplot2::aes(
              x = reorder(designation, -predictedValue2),
              y = predictedValue2,
              col = entryType,
              text = paste(
                paste0('<br>', "designation: ", designation),
                paste0('<br>', "entry type: ", entryType),
                paste0('<br>', "predicted value: ", round(predictedValue2, 4)),
                # paste0('<br>', "standard error: ", round(stdError2, 4)),
                paste0("<br>predVal +- stdError: (", round(LB, 2) , ", ",
                       round(UB, 4), ")")))) +
            ggplot2::theme_classic() +  ggplot2::geom_point() + #facet_grid(~trait, scales = "free_x")+
            ggplot2::geom_errorbar(ggplot2::aes(ymax = UB,# predictedValue2 + stdError2,
                                                ymin = LB#predictedValue2 - stdError2
            ), position = ggplot2::position_dodge(0.5),
            width = 0.10) +
            ggplot2::labs(title = "Percentage over benchmark", x = "designation", y = "% over check entry", col = "Entry type") + ggplot2::geom_hline(yintercept=0, linetype='dotted', col = 'grey') +
            ggplot2::theme(axis.text.x = ggplot2::element_blank(), legend.position = "bottom") + ggplot2::guides(fill=guide_legend(nrow=3,byrow=TRUE)) #+ ggplot2::facet_wrap(~trait, scales = "free", ncol=2)
             
            print(htmltools::tagList(plotly::ggplotly(pb)))
            
            # tablepb <- temp.pred.box.bench[which(temp.pred.box.bench$predictedValue2 > 0),]
            # 
            # printTable(tablepb, autoWidthOpt = TRUE, scrollXOpt = FALSE,
            #            colNames = colnames(tablepb),
            #            numericColNames = c("predictedValue", "stdError", "reliability", 
            #                                "predictedValue2", "UB", "LB"),
            #            numRound = 4)
            # 
            # # cowplot::plot_grid(pb, tableGrob(tablepb), ncol = 1, rel_widths = c(1,2))
            # 
            # print(htmltools::tagList(plotly::ggplotly(pb), 
            #                          printTable(tablepb, autoWidthOpt = TRUE, scrollXOpt = FALSE,
            #                                     colNames = colnames(tablepb),
            #                                     numericColNames = c("predictedValue", "stdError",
            #                                                         "reliability", 
            #                                                         "predictedValue2", "UB", "LB"),
            #                                     numRound = 4)))
        
          }else{
            print(htmltools::tagList(plotly::ggplotly(ggplot2::ggplot() + ggplot2::ggtitle("Combination not available"))))
          }
        }

      }

    }else{

      plotly::renderPlotly({
        predListBench <- split(pred[which(pred$effectType == "designation"),], pred[which(pred$effectType == "designation"),]$trait)
        predListBench2 <- lapply(predListBench, function(x){ # x <- predListBench[[1]]
          checkSelected <- which(x$designation == input$checkMta)
          # checkSelected <- which(x$designation == "IRRI 156")
          if(length(checkSelected) > 0){
            x$predictedValue2 <- ( (x$predictedValue / x$predictedValue[checkSelected]) - 1 ) * 100
            x$UB <- ( ( (x$predictedValue+x$stdError) / x$predictedValue[checkSelected]) - 1 ) * 100
            x$LB <- ( ( (x$predictedValue-x$stdError) / x$predictedValue[checkSelected]) - 1 ) * 100
            x <- x[with(x, order(-predictedValue2)), ]
            x_test <- x[which(!x$designation %in% nameOfChecks),]
            x_test <- x_test[c(1:min(c(30, nrow(x_test)))),]
            x <- x[c(which(x$designation %in% x_test$designation),which(x$designation %in% nameOfChecks)),]
            # x <- x[unique(c(1:min(c(30, nrow(x))),which(x$designation == input$checkMta))),]
            return(x)
          }
        })
        temp.pred.box.bench <- do.call(rbind, predListBench2)
        temp.pred.box.bench$entryTypeSimple <- LETTERS[as.numeric(as.factor(temp.pred.box.bench$entryType))]
        temp.pred.box.bench <- temp.pred.box.bench[which(temp.pred.box.bench$trait == input$checkMtaTrait),]
        if(nrow(temp.pred.box.bench)>0){
          pb <- ggplot2::ggplot(
            data = temp.pred.box.bench,
            mapping = ggplot2::aes(
              x = reorder(designation, -predictedValue2),
              y = predictedValue2,
              col = entryType,
              text = paste(
                paste0('<br>', "designation: ", designation),
                paste0('<br>', "entry type: ", entryType),
                paste0('<br>', "predicted value: ", round(predictedValue2, 4)),
                # paste0('<br>', "standard error: ", round(stdError2, 4)),
                paste0("<br>predVal +- stdError: (", round(LB, 2) , ", ",
                       round(UB, 4), ")")))) +
            ggplot2::theme_classic() +  ggplot2::geom_point() + #facet_grid(~trait, scales = "free_x")+
            ggplot2::geom_errorbar(ggplot2::aes(ymax = UB,# predictedValue2 + stdError2,
                                                ymin = LB#predictedValue2 - stdError2
            ), position = ggplot2::position_dodge(0.5),
            width = 0.10) +
            ggplot2::labs(title = "Percentage over benchmark", x = "designation", y = "% over check entry", col = "Entry type") + ggplot2::geom_hline(yintercept=0, linetype='dotted', col = 'grey') +
            ggplot2::theme(axis.text.x = ggplot2::element_blank(), legend.position = "bottom") + ggplot2::guides(fill=guide_legend(nrow=3,byrow=TRUE)) # + ggplot2::facet_wrap(~trait, scales = "free", ncol=2)
          plotly::ggplotly(pb)
        } else{
          plotly::ggplotly(ggplot2::ggplot() + ggplot2::ggtitle("Combination not available"))
        }
      })

    }

  } else{
    cat("No checks found. Skipping percent check comparison.")
  }

}
```
</div>

```{r, results='asis', fig.height=20}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if(length(checksFound) > 0){
    
    predListBenchForTable <- split(pred[which(pred$effectType == "designation"),], pred[which(pred$effectType == "designation"),]$trait)# make a list by trait
    
    for(iTrait in 1:length(predListBenchForTable)){
      x <- predListBenchForTable[[iTrait]]
      byCheck <- list()
      for(iCheck in nameOfChecks){ # iCheck = nameOfChecks[1]
        checkSelected <- which(x$designation == iCheck)
        if(length(checkSelected) > 0){
          x$predictedValue2 <- ( (x$predictedValue / x$predictedValue[checkSelected]) - 1 ) * 100
          x$UB <- ( ( (x$predictedValue+x$stdError) / x$predictedValue[checkSelected]) - 1 ) * 100
          x$LB <- ( ( (x$predictedValue-x$stdError) / x$predictedValue[checkSelected]) - 1 ) * 100
          x <- x[with(x, order(designation)), ]
          # x <- x[unique(c(1:min(c(30, nrow(x))),which(x$designation == iCheck ))),]
          x$benchmark <- iCheck
          byCheck[[iCheck]] <- x
        }
      }
      long <- do.call(rbind, byCheck)
      wideBench <- reshape(long[, c("benchmark", "designation","predictedValue2")], direction = "wide", idvar = "designation",
                           timevar = c("benchmark"), v.names = c("predictedValue2"), sep= "_")
      colnames(wideBench) <- paste0(names(predListBenchForTable)[iTrait], gsub("predictedValue2","",colnames(wideBench)))
      colnames(wideBench)[1] <- "designation"
      if(iTrait == 1){
        resBench <- wideBench
      }else{
        resBench <- merge(resBench, wideBench, by="designation")
      }
    }
    numCols <- colnames(resBench)[-c(1)]
    
    if("params" %in% ls()){ # we will download the document
      printTable(resBench, colNames = colnames(resBench), fixedCol = 1, filt = "none")
    }else{
      DT::renderDT({DT::formatRound(DT::datatable(
        resBench,
        extensions = c('Buttons', 'FixedColumns'),
        rownames = FALSE,
        class = 'cell-border',
        filter = "none",
        options = list(
          dom = 'Blfrtip',
          scrollY = "400px",
          scrollX = "400px",
          buttons = c('copy','csv','excel','pdf','print'),
          paging = F,
          fixedColumns = list(leftColumns = 1))
      ),numCols)
      })
    }
    
  }
  
}
```

### Modeling parameters

This section aims to provide the modeling table for the analysis in order to keep track of which environments were used in the analysis, what was the final model used for each trait and other potentially important parameters for future reference.

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  modelingMta <- result$modeling[which(modeling$analysisId == idMta),]
  
  if("params" %in% ls()){ # we will download the document #params$toDownload == TRUE
    printTable(modelingMta, autoWidthOpt = TRUE, scrollXOpt = FALSE,
               colNames = colnames(modelingMta))
  }else{
    DT::renderDT(printTable(modelingMta,
                            autoWidthOpt = FALSE, scrollXOpt = TRUE,
                            colNames = colnames(modelingMta)), server = FALSE)
  }
  
}

```


### References of methods used

Finlay, K. W., & Wilkinson, G. N. (1963). The analysis of adaptation in a plant-breeding programme. Australian journal of agricultural research, 14(6), 742-754.

Henderson Jr, C. R. (1982). Analysis of covariance in the mixed model: higher-level, nonhomogeneous, and random regressions. Biometrics, 623-640.

Odegard, J., Indahl, U., Stranden, I., & Meuwissen, T. H. (2018). Large-scale genomic prediction using singular value decomposition of the genotype matrix. Genetics Selection Evolution, 50(1), 1-12.

Xiang, T., Christensen, O. F., & Srensen, A. C. (2016). Genomic evaluation by including dominance effects and inbreeding depression for purebred and crossbred performance with a unified model. Genetics Selection Evolution, 48(1), 84

Batista, L. G., Mello, V. H., Souza, A. P., & Margarido, G. R. A. (2022). Genomic prediction with allele dosage information in highly polyploid species. Theoretical and Applied Genetics, 135(2), 723739.

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

Boer M, van Rossum B (2022). LMMsolver: Linear Mixed Model Solver. R package version 1.0.4.9000.

Covarrubias-Pazaran G. (2024). lme4breeding: enabling genetic evaluation in the era of genomic data. bioRxiv, 2024-05.

Covarrubias-Pazaran G. (2016). Genome assisted prediction of quantitative traits using the R package sommer. PLoS ONE 11(6):1-15.


