---
title: "Genome wide association study Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
  toDownload: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
# font.plot_ly <- list(family="Courier New, monospace", size=9)
```

```{r printfxn, include=FALSE}
# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, ...) {
  
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  
  table <- DT::datatable(data = DT, 
                         filter = "top", 
                         extensions = 'Buttons',
                         options = list(dom = 'Blfrtip',scrollX = TRUE,buttons = c('copy', 'csv', 'excel', 'print'),
                                        lengthMenu = list(c(10,20,50,-1), c(10,20,50,'All'))),
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r plot_init, include = FALSE}
# Init Step to make sure that the dependencies are loaded
htmltools::tagList(plotly::ggplotly(ggplot2::ggplot()))
#htmltools::tagList(printTable(mtcars))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
```

```{r }
suppressWarnings(tryCatch({
  if(file.exists("./outputs/result.RData")){
    load("./outputs/resultGwas.RData")
  }else{
    load("resultGwas.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

if (is.null(result)){
  # invisible(NULL)
} else {
  # result$status <- result$status[1:6,]
  idGwas <- result$status[which(result$status$module == "gwas"),"analysisId"]
  idGwas <- idGwas[length(idGwas)]
  pred <- result$predictions # predictions
  # pred <- pred[pred$analysisId == idGwas,]
  traits <- unique(result$modeling[which(result$modeling$analysisId == idGwas  & result$modeling$environment != "general"),"trait"])
  modeling <- result$modeling
  # modelingFieldsIncludedInMta <- modeling[modeling$analysisId == idMta & modeling$parameter == "includedInMta", ]
}
```


### Objectives of Genome Wide Association Study Analysis

The objective of this dashboard is to help scientist to understand the following points:

1. Overall number of entries and entry types and environments included in the single or multi trial analysis (input)

2. Overall number of markers considered after filtering genetic markers (input)

3. Marker effect in standard format of Manhattan plot and associated QQ-plot for each trait by environment (output)

4. Distribution of phenotype by genotype for top 5 significant markers for each trait by environment combination (output)

5. Individual SNP effect predictions for each trait by environment (output)

Understanding these data features should allow the scientist to identify which markers express a significant statistical association with trait by environment combinations. It should also allow the scientist to assess the Distribution of phenotype by genotype for trait by environment combinations.

### Number of individuals and markers per trait and environment

The following table aims to make a high-level assessment of the different types of entries and markers included in the analysis by trait and environment combinations.

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  idGwas <- result$status[which(result$status$module == "gwas"),"analysisId"]
  idGwas <- idGwas[length(idGwas)]
  # predictions
  pred.gwas <- result$predictions
  pred.gwas <- pred.gwas[pred.gwas$analysisId == idGwas,]
  tableOfTypes <- with(pred.gwas, table(trait, environment))
  tableOfTypes <- data.frame(tableOfTypes)

  field <- unique(pred.gwas$environment)
  modeling <- result$modeling
  modeling <- modeling[which(modeling$analysisId == idGwas  & modeling$environment != "general"),]
  if("(Intercept)" %in% modeling$environment){mod <- "mta"}else{mod <- "sta"}
  pred.gwas <- result$predictions
  pred <- pred.gwas[grep(mod,pred.gwas$module),]
  tableOfType <- with(pred, table(trait, environment))
  tableOfType <- data.frame(tableOfType)
  tableOfType <- tableOfType[which(tableOfType$environment %in% field),]
  tableOfTyp <- merge.data.frame(tableOfType, tableOfTypes, by = c("trait", "environment"))
  colnames(tableOfTyp)[3:4] <- c("TGV_tested", "Number of markers")
  
  if("params" %in% ls()){ # we will download the document
    
    printTable(tableOfTyp)
    
  } else{ # display in website
    numericTraits <- setdiff(c("trait", "environment"), colnames(tableOfTyp))
    DT::renderDT({DT::formatRound(DT::datatable(
      tableOfTyp, extensions = 'Buttons', rownames = FALSE, class = 'cell-border',
      options = list(
        dom = 'Blfrtip', scrollY = "200px", scrollX = "400px",
        buttons = c('copy','csv','excel','pdf','print'),
        paging = F)
    ),numericTraits)
    })
    
  }
  
}
```


### Marker effect per trait

This Manhattan plot allows you easily see the threshold for significant marker and significant markers for trait by environment combinations. The Quantile-quantile (QQ) plot allows you to inspect quality of the model fitted.

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  idGwas <- result$status[which(result$status$module == "gwas"),"analysisId"]
  idGwas <- idGwas[length(idGwas)]
  pred <- result$predictions
  pred <- pred[pred$analysisId == idGwas,]
  traitsManP <- unique(pred$trait)
  fieldManP <- unique(pred$environment)

  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    shiny::inputPanel(
      selectInput(ns("traitGwasManP"),
                  label = "Trait to visualize:",
                  choices = traitsManP),
      selectInput(ns("fieldGwasManP"),
                  label = "Environment to visualize:",
                  choices = fieldManP)
    )
  }
}
```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
idGwas <- result$status[which(result$status$module == "gwas"),"analysisId"]
idGwas <- idGwas[length(idGwas)]
pred.gwas <- result$predictions
pred.gwas <- pred.gwas[pred.gwas$analysisId == idGwas,]
traitsManP <- unique(pred.gwas$trait)
fieldManP <- unique(pred.gwas$environment)
if("(Intercept)" %in% modeling$environment){mod <- "mta"}else{mod <- "sta"}
idMarkers <- result$status[which(result$status$module == "qaGeno"),"analysisId"]
idMarkers <- idMarkers[length(idMarkers)]


  if("params" %in% ls()){ # we will download the document
    cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
    QQ <- list()
Man <- list()
  for(iTrait in traitsManP){
    for(iField in fieldManP){
      pred.gwas.i <- pred.gwas[pred.gwas$trait == iTrait,]
      pred.gwas.i <- pred.gwas.i[pred.gwas.i$environment == iField,]
      o = sort(pred.gwas.i$predictedValue, decreasing = TRUE)
      e = -log10(ppoints(length(pred.gwas.i$predictedValue)))
      pred.gwas.i$o <- o
      pred.gwas.i$e <- e
      QQ[[iField]] <- ggplot2::ggplot(pred.gwas.i, ggplot2::aes(x=e, y=o)) + ggplot2::geom_point() +
        ggplot2::geom_abline(intercept = 0, slope = 1, col = "deepskyblue2") + ggplot2::theme_classic() +
        ggplot2::expand_limits(x=c(0,round(max(o)+1)),y=c(0,round(max(o)+1))) + ggplot2::scale_x_continuous(breaks = seq(0,round(max(o)+1))) + ggplot2::scale_y_continuous(breaks = seq(0,round(max(o)+1))) + ggplot2::theme(axis.text = ggplot2::element_text(colour = "black"), axis.title = ggplot2::element_text(colour = "black")) + 
        ggpubr::border() +
        ggplot2::labs(x = expression(Expected ~ ~-log[10](italic(p))), y = expression(Observed ~ ~-log[10](italic(p))))
      
      pred = pred.gwas.i[order(pred.gwas.i$stdError),]
      "%>%" <- dplyr::`%>%`
      don <- pred %>%
        dplyr::group_by(stdError) %>%
        dplyr::summarise(chr_len = max(reliability)) %>%
        dplyr::mutate(tot=cumsum(chr_len) - chr_len) %>%
        dplyr::left_join(pred, ., by=c("stdError" = "stdError")) %>%
        dplyr::arrange(stdError, reliability) %>%
        dplyr::mutate(BPcum=reliability+tot)
      don$BPcum <- as.numeric(don$BPcum)
      #don$text <- paste("SNP: ", don$SNP, "\nPosition: ", don$BP, "\nChromosome: ", don$CHR, "\n-log10(P) :", -log10(don$P) %>% round(2), sep="")
      
    # don.i <- don[don$trait == iTrait,]
    cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
    Markers <- result$data$geno_imp[[ as.character(idMarkers) ]]
    marker_map <- data.frame(SNP = Markers@loc.names,
                         CHR = Markers@chromosome,
                         BP = Markers@position)
    id <- result$status[which(grepl(mod,result$status$module)),"analysisId"]
    id <- id[length(id)]
    phenoData <- result$predictions[which(result$predictions$analysisId == id),]
    phenoData <- phenoData[which(phenoData$environment == iField),]
    phenoData <- phenoData %>%
      dplyr::select(ID = designation, trait, value = predictedValue) %>%
      tidyr::spread(key = trait, value = value) %>%
      dplyr::arrange(ID) %>%
      dplyr::filter(!is.na(!!dplyr::sym(iTrait))) %>%
      dplyr::select(c("ID", all_of(iTrait)))
    
    common_ids <- intersect(Markers@ind.names, phenoData$ID)
    markers <- as.matrix(Markers[which(Markers@ind.names %in% common_ids),])
    phenoData <- phenoData[phenoData$ID %in% common_ids, ]
    
    ids_to_keep <- colnames(markers)
    marker_map <- marker_map[match(ids_to_keep, marker_map$SNP), , drop = FALSE]
    
    meta <- marker_map %>%
      dplyr::group_by(CHR) %>%
      dplyr::summarise(N = length(CHR))
    
    CORR <- list()
    CORR[[1]] <- cor(markers[,1:meta$N[meta$CHR==1]])
    for(i in 2:nrow(meta)){
      dat = markers[,(meta$N[meta$CHR==(i-1)]+1):(meta$N[meta$CHR==(i-1)]+meta$N[meta$CHR==i])]
      CORR[[i]] = cor(dat)
    }
    z = list()
    for(i in 1:nrow(meta)){
      for(j in 1:nrow(CORR[[i]])){
        xx = as.vector(which(is.na(CORR[[i]][j,])))
      }
      if(length(xx)>0) {z[[i]] = CORR[[i]][-xx,-xx]} else {z[[i]] = CORR[[i]]}
    }
    corr = as.vector(NULL)
    for(i in 1:nrow(meta)){
      corr[i] = poolr::meff(z[[i]], method = "liji")
    }
    Meff = sum(corr)
    p_threshold = (1-(1-0.05))/Meff # p-value to use as threshold for significance
    Man[[iField]] <- cgiarPipeline::manhattanPlot(don, threshold = -log10(p_threshold), field = iField)
  }
  }

# grid::grid.null() #grid.newpage()
# # Create layout : nrow = 3, ncol = 2
# grid::pushViewport(grid::viewport(layout = grid::grid.layout(nrow = length(Man), ncol = 3)))
# # A helper function to define a region on the layout
# define_region <- function(row, col){
#   grid::viewport(layout.pos.row = row, layout.pos.col = col)
# } 
# # Arrange the plots
# for(i in 1:length(Man)){
#   print(Man[[i]], vp = define_region(row = i, col = 1:2))
#   print(QQ[[i]], vp = define_region(row = i, col = 3))
# }
  }else{
    #shiny::renderPlot({
    plotly::renderPlotly({
          QQ <- list()
          Man <- list()
      inputTrait <- input$traitGwasManP
      inputField <- input$fieldGwasManP

      #for(iField in field){
      #pred.gwas.i <- data.frame()
      pred.gwas.i <- pred.gwas[pred.gwas$trait == inputTrait,]
      pred.gwas.i <- pred.gwas.i[pred.gwas.i$environment == inputField,]
      #o <- sort(pred.gwas.i$predictedValue, decreasing = TRUE)
      o <- pred.gwas.i$predictedValue
      e <- -log10(ppoints(length(pred.gwas.i$predictedValue)))
      pred.gwas.i$o <- o
      pred.gwas.i$e <- e
      qq <- ggplot2::ggplot(pred.gwas.i, ggplot2::aes(x=e, y=o)) + ggplot2::geom_point() +
        ggplot2::geom_abline(intercept = 0, slope = 1, col = "deepskyblue2") + ggplot2::theme_classic() +
        # ggplot2::expand_limits(x=c(0,round(max(o)+1)),y=c(0,round(max(o)+1))) + 
        ggplot2::scale_x_continuous(breaks = seq(0,round(max(o)+1))) + ggplot2::scale_y_continuous(breaks = seq(0,round(max(o)+1))) +
        # ggplot2::labs(x = expression(Expected ~ ~-log[10](italic(p))), y = expression(Observed ~ ~-log[10](italic(p)))) +
        ggplot2::theme(axis.text = ggplot2::element_text(colour = "black"), axis.title = ggplot2::element_text(colour = "black")) + 
        ggpubr::border() #+
        #ggplot2::labs(x = expression(Expected ~ ~-log[10](italic(p))), y = expression(Observed ~ ~-log[10](italic(p))))
      # QQ$across <- plotly::plotly_build(QQ$across)
      # QQ$across <- QQ$across %>% plotly::layout(xaxis = list(showline = TRUE, mirror = TRUE, linecolor = 'black'),
      #                             yaxis = list(showline = TRUE, mirror = TRUE, linecolor = 'black'))
      
      qq <- plotly::plotly_build(qq)
      QQ[[inputField]] <- qq %>% plotly::layout(xaxis = list(showline = TRUE, mirror = TRUE, linecolor = 'black'),
                                  yaxis = list(showline = TRUE, mirror = TRUE, linecolor = 'black'))

      pred.gwas.i = pred.gwas.i[!is.na(pred.gwas.i$stdError),]
      pred = pred.gwas.i[order(pred.gwas.i$stdError),]
      "%>%" <- dplyr::`%>%`
      don <- pred %>%
        dplyr::group_by(stdError) %>%
        dplyr::summarise(chr_len = max(reliability)) %>%
        dplyr::mutate(tot=cumsum(chr_len) - chr_len) %>%
        dplyr::left_join(pred, ., by=c("stdError" = "stdError")) %>%
        dplyr::arrange(stdError, reliability) %>%
        dplyr::mutate(BPcum=reliability+tot)
      don$BPcum <- as.numeric(don$BPcum)
      #don$text <- paste("SNP: ", don$SNP, "\nPosition: ", don$BP, "\nChromosome: ", don$CHR, "\n-log10(P) :", -log10(don$P) %>% round(2), sep="")
      
      # don.i <- don[don$trait == inputTrait,]
      Markers <- result$data$geno_imp[[ as.character(idMarkers) ]]
      marker_map <- data.frame(SNP = Markers@loc.names,
                         CHR = Markers@chromosome,
                         BP = Markers@position)

      id <- result$status[which(grepl(mod,result$status$module)),"analysisId"]
      id <- id[length(id)]
      phenoData <- result$predictions[which(result$predictions$analysisId == id),]
      phenoData <- phenoData[which(phenoData$environment == inputField),]
      phenoData <- phenoData %>%
        dplyr::select(ID = designation, trait, value = predictedValue) %>%
        tidyr::spread(key = trait, value = value) %>%
        dplyr::arrange(ID) %>%
        dplyr::filter(!is.na(!!dplyr::sym(inputTrait))) %>%
        dplyr::select(c("ID", all_of(inputTrait)))
      common_ids <- intersect(Markers@ind.names, phenoData$ID)
      markers <- as.matrix(Markers[which(Markers@ind.names %in% common_ids),])
      phenoData <- phenoData[phenoData$ID %in% common_ids, ]

      ids_to_keep <- colnames(markers)
      marker_map <- marker_map[match(ids_to_keep, marker_map$SNP), , drop = FALSE]

      meta <- marker_map %>%
        dplyr::group_by(CHR) %>%
        dplyr::summarise(N = length(CHR))

      # CORR <- list()
      # CORR[[1]] <- cor(markers[,1:meta$N[meta$CHR==1]])
      # for(i in 2:nrow(meta)){
      #   dat = markers[,(meta$N[meta$CHR==(i-1)]+1):(meta$N[meta$CHR==(i-1)]+meta$N[meta$CHR==i])]
      #   CORR[[i]] = cor(dat)
      # }
      # z = list()
      # for(i in 1:nrow(meta)){
      #   for(j in 1:nrow(CORR[[i]])){
      #     xx = as.vector(which(is.na(CORR[[i]][j,])))
      #   }
      #   if(length(xx)>0) {z[[i]] = CORR[[i]][-xx,-xx]} else {z[[i]] = CORR[[i]]}
      # }
      # corr = as.vector(NULL)
      # for(i in 1:nrow(meta)){
      #   corr[i] = poolr::meff(poolr::mvnconv(z[[i]], target = "p", cov2cor = TRUE), method = "liji") # corr[i] = poolr::meff(z[[i]], method = "liji")
      # }
      # Meff = sum(corr)
      p_threshold = 1e-4 # (1-(1-0.05))/Meff # p-value to use as threshold for significance
      Man[[inputField]] <- cgiarPipeline::manhattanPlot(don, threshold = -log10(p_threshold)) # , field = inputField 
      #}
      # plotly::subplot(Man$across, QQ$across, widths = c(0.7, 0.3), nrows = 2) # length(Man)
      # plotly::subplot(M, widths = c(0.7, 0.3), nrows = 2) #

      s <- c(Man, QQ); s
      w <- s[c(seq(1,2*length(Man),2),seq(1,2*length(Man))[-seq(1,2*length(Man),2)])]; w
      plotly::subplot(w, widths = c(0.7, 0.3), nrows = length(Man))

#       grid::grid.null() #grid.newpage()
# # Create layout : nrow = 3, ncol = 2
# grid::pushViewport(grid::viewport(layout = grid::grid.layout(nrow = length(Man), ncol = 3)))
# # A helper function to define a region on the layout
# define_region <- function(row, col){
#   grid::viewport(layout.pos.row = row, layout.pos.col = col)
# } 
# # Arrange the plots
# for(i in 1:length(Man)){
#   print(Man[[i]], vp = define_region(row = i, col = 1:2))
#   print(QQ[[i]], vp = define_region(row = i, col = 3))
# }
    })
  }
}
```

The following boxplots allow you to inspect for each of the top 5 significant markers, the distribution of adjusted means (y-axis) by genotype levels (x-axis). The adjusted means are from the single or multi trial analysis for each trait that were used as input for the analysis.

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
idGwas <- result$status[which(result$status$module == "gwas"),"analysisId"]
idGwas <- idGwas[length(idGwas)]
pred.gwas <- result$predictions
pred.gwas <- pred.gwas[pred.gwas$analysisId == idGwas,]
traitsManP <- unique(pred.gwas$trait)
fieldManP <- unique(pred.gwas$environment)
if("(Intercept)" %in% modeling$environment){mod <- "mta"}else{mod <- "sta"}

  if("params" %in% ls()){ # we will download the document
    cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
  for(iTrait in traitsManP){
    for(iField in fieldManP){
      pred.gwas.i <- pred.gwas[pred.gwas$trait == iTrait,]
      pred.gwas.i <- pred.gwas.i[pred.gwas.i$environment == iField,]

      pred = pred.gwas.i[order(pred.gwas.i$stdError),]
      "%>%" <- dplyr::`%>%`
      don <- pred %>%
        dplyr::group_by(stdError) %>%
        dplyr::summarise(chr_len = max(reliability)) %>%
        dplyr::mutate(tot=cumsum(chr_len) - chr_len) %>%
        dplyr::left_join(pred, ., by=c("stdError" = "stdError")) %>%
        dplyr::arrange(stdError, reliability) %>%
        dplyr::mutate(BPcum=reliability+tot)
      don$BPcum <- as.numeric(don$BPcum)
      #don$text <- paste("SNP: ", don$SNP, "\nPosition: ", don$BP, "\nChromosome: ", don$CHR, "\n-log10(P) :", -log10(don$P) %>% round(2), sep="")
      
    # don.i <- don[don$trait == iTrait,]
    cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
    Markers <- result$data$geno_imp[[ as.character(idMarkers) ]]
    marker_map <- data.frame(SNP = Markers@loc.names,
                         CHR = Markers@chromosome,
                         BP = Markers@position)
    id <- result$status[which(grepl(mod,result$status$module)),"analysisId"]
    id <- id[length(id)]
    phenoData <- result$predictions[which(result$predictions$analysisId == id),]
    phenoData <- phenoData[which(phenoData$environment == iField),]
    phenoData <- phenoData %>%
      dplyr::select(ID = designation, trait, value = predictedValue) %>%
      tidyr::spread(key = trait, value = value) %>%
      dplyr::arrange(ID) %>%
      dplyr::filter(!is.na(!!dplyr::sym(iTrait))) %>%
      dplyr::select(c("ID", all_of(iTrait)))
    
    common_ids <- intersect(Markers@ind.names, phenoData$ID)
    markers <- as.matrix(Markers[which(Markers@ind.names %in% common_ids),])
    phenoData <- phenoData[phenoData$ID %in% common_ids, ]
    
    ids_to_keep <- colnames(markers)

  }
  }

  }else{
    plotly::renderPlotly({
      
      s <- list()
      inputTrait <- input$traitGwasManP
      inputField <- input$fieldGwasManP

      pred.gwas.i <- pred.gwas[pred.gwas$trait == inputTrait,]
      pred.gwas.i <- pred.gwas.i[pred.gwas.i$environment == inputField,]

      pred = pred.gwas.i[order(pred.gwas.i$stdError),]
      "%>%" <- dplyr::`%>%`
      don <- pred %>%
        dplyr::group_by(stdError) %>%
        dplyr::summarise(chr_len = max(reliability)) %>%
        dplyr::mutate(tot=cumsum(chr_len) - chr_len) %>%
        dplyr::left_join(pred, ., by=c("stdError" = "stdError")) %>%
        dplyr::arrange(stdError, reliability) %>%
        dplyr::mutate(BPcum=reliability+tot)
      don$BPcum <- as.numeric(don$BPcum)
      #don$text <- paste("SNP: ", don$SNP, "\nPosition: ", don$BP, "\nChromosome: ", don$CHR, "\n-log10(P) :", -log10(don$P) %>% round(2), sep="")
      
      # don.i <- don[don$trait == inputTrait,]
      Markers <- result$data$geno_imp[[ as.character(idMarkers) ]]
      marker_map <- data.frame(SNP = Markers@loc.names,
                         CHR = Markers@chromosome,
                         BP = Markers@position)
      id <- result$status[which(grepl(mod,result$status$module)),"analysisId"]
      id <- id[length(id)]
      phenoData <- result$predictions[which(result$predictions$analysisId == id),]
      phenoData <- phenoData[which(phenoData$environment == inputField),]
      phenoData <- phenoData %>%
        dplyr::select(ID = designation, trait, value = predictedValue) %>%
        tidyr::spread(key = trait, value = value) %>%
        dplyr::arrange(ID) %>%
        dplyr::filter(!is.na(!!dplyr::sym(inputTrait))) %>%
        dplyr::select(c("ID", all_of(inputTrait)))

      common_ids <- intersect(Markers@ind.names, phenoData$ID)
      markers <- as.matrix(Markers[which(Markers@ind.names %in% common_ids),])
      phenoData <- phenoData[phenoData$ID %in% common_ids, ]

      ids_to_keep <- colnames(markers)
      
      don_sorted <- don[order(don$predictedValue, decreasing = TRUE),]
      top5markers <- don_sorted$designation[which(don_sorted$predictedValue %in% don_sorted$predictedValue[1:5])]

top5matrix <- markers[,which(colnames(markers) %in% top5markers)]
top5matrix <- as.data.frame(top5matrix)
top5matrix$ID <- rownames(top5matrix)

top5matrix <- merge.data.frame(phenoData, top5matrix, by = "ID")
top5matrix <- reshape2::melt(top5matrix, id.var = c("ID", colnames(top5matrix)[2]), 
                           variable.name = 'topmarkers') 

s <- ggplot2::ggplot(data = top5matrix, ggplot2::aes(x = factor(value), y = top5matrix[,2])) + # [[inputField]]
  #labs(x = "Genotype", y = "Phenotype") + 
  ggplot2::theme_bw() + 
  ggplot2::geom_boxplot(fill='#A4A4A4') + ggplot2::geom_jitter(alpha = .1) +
  ggplot2::theme(axis.text = ggplot2::element_text(color="black"),
        axis.title = ggplot2::element_text(color="black", size = 10)) + 
  ggpubr::border() + ggplot2::labs(x = "Genotype", y = "Phenotype") + ggplot2::facet_grid(. ~ topmarkers, scales='free')

#plotly::plotly_build(s) # p[[inputField]] <- 

plotly::ggplotly(s)
      # u <- c(s)
      # # plotly::plotly_build(u)
      # w <- u[seq(1,length(u))]
      # #plotly::ggplotly(w)
      # plotly::subplot(w)
      
    })
  }
}
```

### Predictions

The following table allows you to inspect the marker effect predictions in wide format for each trait by environment combination to get an overall overview of marker effect across chromosomes.

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  #numericalColumns <- c("P","-log10(P)") 
  idGwas <- result$status[which(result$status$module == "gwas"),"analysisId"]
  idGwas <- idGwas[length(idGwas)]
  pred.gwas <- result$predictions
  pred.gwas <- pred.gwas[pred.gwas$analysisId == idGwas,]
  traits <- unique(pred.gwas$trait)
  pred.gwas <- subset(pred.gwas, select = -c(module,analysisId,pipeline,gid,mother,father,entryType,effectType))
  pred.gwas.wide <- reshape(
    data=pred.gwas,
    timevar = c("trait"),
    idvar = c("designation","environment","stdError","reliability"),
    direction="wide"
  )
  names(pred.gwas.wide) <- c("SNP","environment","CHR","BP",traits)

  if("params" %in% ls()){ # we will download the document
    printTable(pred.gwas.wide)
  }else{
    DT::renderDT({DT::formatRound(DT::datatable(
      pred.gwas.wide,
      extensions = 'Buttons',
      rownames = FALSE,
      class = 'cell-border',
      options = list(
        dom = 'Blfrtip',
        scrollY = "400px",
        scrollX = "400px",
        buttons = c('copy','csv','excel','pdf','print'),
        paging = F)
    ), traits, digits = 5)
    })
  }
}
```


### References of methods used

Boer M, van Rossum B (2022). LMMsolver: Linear Mixed Model Solver. R package version 1.0.4.9000.

Covarrubias-Pazaran G. 2016. Genome assisted prediction of quantitative traits using the R package sommer. PLoS ONE 11(6):1-15.

R Core Team (2024). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

<p>&nbsp;</p>
