---
title: "Selection Index Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
 toDownload: FALSE
---

```{r setup, include=FALSE}
# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
options(knitr.kable.NA = '')
# loading necessary R packages ####
## data manipulation
# library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric
## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, 
                       scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,...) {
  oldnames <- colnames(DT)
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  colnames(DT) <- oldnames
  table <- DT::datatable(data = DT, 
                         colnames = colNames,
                         filter = "top", 
                         options = list(autoWidth = autoWidthOpt,
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'csv', 'excel', 'print'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35),
                                        scrollX = scrollXOpt),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% DT::formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r, include=FALSE}
# Init Step to make sure that the dependencies are loaded
htmltools::tagList(printTable(mtcars))
htmltools::tagList(ggplotly(ggplot()))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
```

```{r}
suppressWarnings(tryCatch({
  if(file.exists("./outputs/result.RData")){
    load("./outputs/resultIndex.RData")
  }else{
    load("resultIndex.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

if (is.null(result)){
  # invisible(NULL)
} else {
  mydata <- result$predictions
  idIndex <- result$status[which(result$status$module == "indexD"),"analysisId"];
  idIndex <- idIndex[length(idIndex)]
  idMta <-result$modeling[which(result$modeling$parameter %in% "analysisId" & result$modeling$analysisId == idIndex),"value"]
  mydata <- mydata[which(mydata$analysisId %in% c(idMta,idIndex)),]
  envsUsed <- unique(result$modeling[which(result$modeling$analysisId %in% idIndex),"environment"])
  mydata <- mydata[which(mydata$environment %in% envsUsed),]
  mTrait <- unique(result$modeling[which(result$modeling$analysisId %in% idIndex),"trait"])
  uTrait <- c(mTrait[which(mTrait %in% unique(mydata$trait))],"desireIndex")
  mydata <- mydata[which(mydata$trait %in% uTrait),]
  wide <- stats::reshape(mydata[,c(c("designation"),"trait",c("predictedValue", "entryType"))], direction = "wide", idvar = c("designation"),
                         timevar = "trait", v.names = c("predictedValue"), sep= "_")
  colnames(wide) <- gsub("predictedValue_","",colnames(wide))
  numeric.output <- colnames(wide)[-c(1:2)]
  
}
```

### Objectives of a Selection Index

The objective of this dashboard is to help scientist to understand the following points:

1. Individual across environment predictions for each trait (input) 

2. Individual sensitivity values from the Finlay-Wilkinson model (Finlay & Wilkinson, 1963) (input)

3. Relative distance between the desired values (target product profiles) and the population means (output)

4. Expected response to selection in each trait after using the index (output)

5. Relationship between the index with the input traits (output)

Understanding these data features should allow the scientist to identify which traits are driving the desire selection index (Pesek & Baker, 1969) and understand what is the approximate expected change of the population in the next generation. We would recommend to use the index as the input trait for an optimal contribution selection algorithm.

### Across-environment trait table

The following table allows you to inspect the across-environment predictions for the different entries (rows) and different traits (columns) that were used as input to calculate the selection index.

<p>&nbsp;</p>

```{r, results='asis'}

if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    printTable(wide, autoWidthOpt = TRUE, scrollXOpt = FALSE,
           colNames = colnames(wide),
           numericColNames = numeric.output,
           numRound = 4)
  }else{
    
    DT::renderDT(printTable(wide,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(wide),
                        numericColNames = numeric.output,
                        numRound = 4), server = FALSE)
  }
  
}
```

<p>&nbsp;</p>

### Modeling table

The following table aims to keep record of the desire values selected and corresponding weights for this run.

```{r, results='asis'}

if (is.null(result)){
  # invisible(NULL)
} else {
  
  modelingTable <- result$modeling
  modelingTable <- modelingTable[which(modelingTable$analysisId %in% c(idIndex)),]
  
  if("params" %in% ls()){ # we will download the document
    printTable(modelingTable, autoWidthOpt = TRUE, scrollXOpt = FALSE,
           colNames = colnames(modelingTable))
  }else{
    
    DT::renderDT(printTable(modelingTable,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(modelingTable)), server = FALSE)
  }
  
}
```


### Radar Plot

This spider plot allows you to visualize the distance between the population means and the target values (hopefully coming from your product profile).

<p>&nbsp;</p>

```{r, results='asis'}
library(dplyr)
if (is.null(result)){
  # invisible(NULL)
} else {
  
  dtIdxD <- result$predictions
  idIndex <- result$status[which(result$status$module == "indexD"),"analysisId"];
  idIndex <- idIndex[length(idIndex)]
  idMta <- result$modeling[which(result$modeling$parameter %in% "analysisId" & result$modeling$analysisId == idIndex),"value"]
  mydata <- dtIdxD[which(dtIdxD$analysisId %in% idMta),setdiff(colnames(dtIdxD),c("module","analysisId"))]
  envsUsed <- unique(result$modeling[which(result$modeling$analysisId %in% idIndex),"environment"])
  mydata <- mydata[which(mydata$environment %in% envsUsed),]
  # subset to only traits being part of the index
  desirev <-result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"value"]
  entryTypes <-result$modeling[which(result$modeling$parameter %in% "entryTypeToUse" & result$modeling$analysisId == idIndex),"value"]
  uTraits <- result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"trait"]
  uTraits <- gsub("_scaled","",uTraits)
  dtIdxD <- dtIdxD[which(dtIdxD$trait %in% uTraits),]
  ##
  environmentPredictionsRadar2=envsUsed
  traitFilterPredictionsRadar2=unique(dtIdxD[which(dtIdxD$analysisId %in% idMta),"trait"])
  ## ensure product profile means come sorted
  meanGroupPredictionsRadar= result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),c("trait","value")]
  meanGroupPredictionsRadar <- meanGroupPredictionsRadar[with(meanGroupPredictionsRadar, order(as.numeric(as.factor(trait)))), ]
  meanGroupPredictionsRadar <- meanGroupPredictionsRadar[,"value"]
  
  fontSizeRadar=12; r0Radar=NULL; neRadar=NULL; plotSdRadar=FALSE
  
  mydata = mydata[which( (mydata$environment %in% environmentPredictionsRadar2) &
                         (mydata$trait %in% traitFilterPredictionsRadar2) &
                         (mydata$entryType %in% entryTypes)
),]
  
  if(!is.null(environmentPredictionsRadar2) & !is.null(traitFilterPredictionsRadar2)){
    mm <- stats::aggregate(predictedValue~trait, data=mydata, FUN=mean, na.rm=TRUE)
    mmsd <- stats::aggregate(predictedValue~trait, data=mydata, FUN=stats::sd, na.rm=TRUE)
    mm2 <- stats::aggregate(predictedValue~trait, data=mydata, FUN=min, na.rm=TRUE)
    mm3 <- stats::aggregate(predictedValue~trait, data=mydata, FUN=max, na.rm=TRUE)
    namesA <- mm[,1]
    meanA <- mm[,2]/(mm3[,2])
    
    # meanA <- rep(0.75,nrow(mm))
    fig <-  plotly::plot_ly(
      type = 'scatterpolar',
      fill = 'toself'
    )
    fig <- fig %>%
      plotly::add_trace(
        r = meanA, #c(39, 28, 8, 7, 28, 39),
        theta =paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
        name = 'Breeding population means',
        text=round(mm[,2],3)
      )
    ## add product profile means
    if(!is.null(meanGroupPredictionsRadar)){
      meanB <- as.numeric(unlist(strsplit(meanGroupPredictionsRadar,",")))
      if(length(meanB) == length(mm[,2])){
        meanB <- meanB + mm[,2]
        # r2 = (meanB*0.75)/mm[,2]
        r2 = meanB/(mm3[,2])
        dToSearch <- meanB #- mm[,2]
        fig <- fig %>%
          plotly::add_trace(
            r = r2, # c(1.5, 10, 39, 31, 15, 1.5),
            theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
            name = "Desired population", #'Group B'
            # mode="text",
            text=round(dToSearch,2)
            # textfont = list(color = '#000000', size = fontSizeRadar)
          )
      }
    }
    # add SD polygon
    if(plotSdRadar){
      meanC <- (mm[,2]+mmsd[,2])/(mm3[,2])
      fig <- fig %>%
        plotly::add_trace(
          r = meanC, # c(1.5, 10, 39, 31, 15, 1.5),
          theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
          name = "Mean + Standard Deviation from pop", #'Group B'
          text=round(mm[,2]+mmsd[,2],3)
        )
      meanD <- (mm[,2]-mmsd[,2])/(mm3[,2])
      fig <- fig %>%
        plotly::add_trace(
          r = meanD, # c(1.5, 10, 39, 31, 15, 1.5),
          theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
          name = "Mean - Standard Deviation from pop", #'Group B'
          text=round(mm[,2]-mmsd[,2],3)
        )
    }
    # add selection limits polygon
    ## add product profile means'
    if(!is.null(r0Radar)){
      meanR <- as.numeric(unlist(strsplit(r0Radar,",")))/100
      if(length(meanR) == length(mm[,2])){
        meanR <- mm[,2] + ((meanR*mm[,2])*neRadar*2) # selection limits 2*Ne*R0
        r3 = (meanR*0.75)/mm[,2]
        dToSearchR <- meanR #- mm[,2]
        fig <- fig %>%
          plotly::add_trace(
            r = r3, # c(1.5, 10, 39, 31, 15, 1.5),
            theta = paste(1:length(namesA),namesA,sep="."),# c('A','B','C', 'D', 'E', 'A'),
            name = "Selection limits", #'Group B'
            # mode="text",
            text=round(dToSearchR,2)
            # textfont = list(color = '#000000', size = fontSizeRadar)
          )
      }
    }
  }
  
  if("params" %in% ls()){ # we will download the document
    
    
    print(htmltools::tagList(plotly::ggplotly(fig)))
    
  }else{
    
    plotly::renderPlotly({
      fig
    }) # enf of render plotly
    
  }
  
}

```

### Expected Response Plot

The following plot allows the user to visualize the expected response in the new generation given the individuals selected (different proportions can be investigated) using the selection index.

<p>&nbsp;</p>

```{r}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  dtIdxD <- result$predictions
  idIndex <- result$status[which(result$status$module == "indexD"),"analysisId"];
  idIndex <- idIndex[length(idIndex)]
  idMta <-result$modeling[which(result$modeling$parameter %in% "analysisId" & result$modeling$analysisId == idIndex),"value"]
  mydataMta <- dtIdxD[which(dtIdxD$analysisId == idMta),setdiff(colnames(dtIdxD),c("module","analysisId"))]
  mydataIndex <- dtIdxD[which(dtIdxD$analysisId == idIndex),setdiff(colnames(dtIdxD),c("module","analysisId"))]
  mydata <- rbind(mydataMta,mydataIndex)
  envsUsed <- unique(result$modeling[which(result$modeling$analysisId %in% idIndex),"environment"])
  mydata <- mydata[which(mydata$environment %in% envsUsed),]
  desirev <-result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"value"]
  entryTypes <-result$modeling[which(result$modeling$parameter %in% "entryTypeToUse" & result$modeling$analysisId == idIndex),"value"]
  uTraits <- result$modeling[which(result$modeling$parameter %in% "desire" & result$modeling$analysisId == idIndex),"trait"]#unique(mydata$trait)
  uTraits <- gsub("_scaled","",uTraits)
  uTraits <- c(uTraits, "desireIndex")
  mydata$selected <- "no"
  mydata <- mydata[which(mydata$entryType %in% entryTypes),]
  
  if("params" %in% ls()){ # we will download the document
    proportionTrait <- c(0.05,0.10,0.15,0.20)
  }else{
    inputPanel(
      shiny::numericInput(ns("proportionTrait"), label = "Selected proportion", value = 0.1, min=0.001,max=1, step=0.05),
      selectInput(ns("perTrait"),
                  label = "Trait",
                  choices = uTraits),
      shiny::checkboxInput(ns("perET"),
                             label = "Show per Entry Type",
                             value = TRUE)
    )
    
  }
  
}
```

<div style = "width:auto; height:auto; overflow:auto">

```{r, results='asis'}
library(dplyr)
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    
    # cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
    cat(paste0("### Proportion selected  {.tabset .tabset-pills}    \n\n  "))
    for(iProportionTrait in proportionTrait){
      cat("\n\n####", iProportionTrait, "{.tabset .tabset-pills}       \n\n")
      
      for(iTrait in uTraits){ # iTrait = uTraits[1]
        cat("\n\n#####", iTrait, "{.tabset .tabset-pills}       \n\n")
        
        prov <- mydata[mydata$trait == "desireIndex", ]
        prov <- prov[with(prov, order(-predictedValue)), ]
        selected <- prov[1:round(nrow(prov)*iProportionTrait),"designation"]
        mydata[which( (mydata$trait == iTrait) & (mydata$designation %in% selected) ), "selected"] <- "yes"
        
        for(iEntryType in c("per trait", "per trait and entryType")){
          cat("\n\n######", iEntryType, "{.tabset .tabset-pills}       \n\n")
          
          if(iEntryType == "per trait"){
            p <- ggplot2::ggplot(mydata[which(mydata$trait == iTrait),], ggplot2::aes(x=predictedValue, color=selected, fill = selected)) +
              ggplot2::geom_histogram() +
              ggplot2::ylab("Frequency") +
              ggplot2::geom_rug(sides="t", length = ggplot2::unit(0.3, "cm")) 
            
            print(htmltools::tagList(plotly::ggplotly(p) %>% layout(xaxis = list(tickangle =-45),autosize = T, height = 400)))
          }else{
            p <- ggplot2::ggplot(mydata[which(mydata$trait == iTrait),], ggplot2::aes(x=predictedValue, color=selected, fill = selected)) +
              ggplot2::geom_histogram() +
              ggplot2::ylab("Frequency") +
              ggplot2::geom_rug(sides="t", length = ggplot2::unit(0.3, "cm")) +
              ggplot2::facet_wrap(~entryType, ncol=2, scales = "free") + ggplot2::theme(panel.spacing = unit(1, 'lines'))
            
            print(htmltools::tagList(plotly::ggplotly(p) %>% layout(xaxis = list(tickangle =-45),autosize = T, 
                                                                    height = ceiling(length(unique(mydata[which(mydata$trait == iTrait), 
                                                                                                   "entryType"]))/2)*400)))
          }
        }
        cat("\n")
      }
      cat("\n")
      cat("\n")
    }
    
  }else{
    
    # shiny::renderPlot({ # 
    plotly::renderPlotly({
      ## if user provides threshold values let them know the selection differentials
      prov <- mydata[mydata$trait == "desireIndex", ]
      prov <- prov[with(prov, order(-predictedValue)), ]
      selected <- prov[1:round(nrow(prov)*input$proportionTrait),"designation"]
      mydata[which( (mydata$trait == input$perTrait) & (mydata$designation %in% selected) ), "selected"] <- "yes"
      
      if(input$perET == TRUE){
        p <- ggplot2::ggplot(mydata[which( mydata$trait == input$perTrait),], ggplot2::aes(x=predictedValue, color=selected, fill = selected)) +
        ggplot2::geom_histogram() +
        ggplot2::ylab("Frequency") +
        ggplot2::geom_rug(sides="t", length = ggplot2::unit(0.3, "cm")) +
        ggplot2::facet_wrap(~entryType, ncol=2, scales = "free")
        
        # p
        plotly::ggplotly(p) %>% layout(xaxis = list(tickangle =-45),autosize = T, 
                                       height = ceiling(length(unique(mydata[which(mydata$trait == input$perTrait),"entryType"]))/2)*400)
      } else{
        p <- ggplot2::ggplot(mydata[which( mydata$trait == input$perTrait),], ggplot2::aes(x=predictedValue, color=selected, fill = selected)) +
        ggplot2::geom_histogram() +
        ggplot2::ylab("Frequency") +
        ggplot2::geom_rug(sides="t", length = ggplot2::unit(0.3, "cm"))
        
        # p
        plotly::ggplotly(p) %>% layout(xaxis = list(tickangle =-45),autosize = T, height = 400)
      }

      
    })
    
  }
  
  
  
}

```

</div>

### Index versus Trait Plot

The following plot allows you to visualize the relationship between the selection index with other traits. This can be specially useful for selecting materials with high total merit (index) and stability for yield (for example, for product development).

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  
  
  mtaTraits <- setdiff( colnames(wide), c("designation","desireIndex","entryType"))
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("traitMtaScatter"), 
                       label = "Trait:", 
                       multiple = FALSE,
                       choices = mtaTraits, 
    )
  }
  
}

```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){ # we will download the document
    
    cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
    
    for(iTrait in mtaTraits){ # iTrait="Yield_Mg_ha"
      cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
      
      wide$trait1 <- wide[,iTrait]
      p <- ggplot(wide, aes(x=trait1, y=desireIndex, text=designation)) + 
        geom_point(size=2, alpha = 0.6, aes(shape=entryType)) + xlab(iTrait) + ylab("Selection Index") 
      
      if(nrow(wide) > 1000){ # go fixed format
        print(p)
      }else{ # go dynamic
        print(htmltools::tagList(plotly::ggplotly(p)))
      }
    }
    
  }else{ # render in screen
    
    plotly::renderPlotly({
      wide$trait1 <- wide[,input$traitMtaScatter]
      p <- ggplot(wide, aes(x=trait1, y=desireIndex, text=designation)) + 
        geom_point(size=2, alpha = 0.6, aes(shape=entryType) ) + xlab(input$traitMtaScatter) + ylab("Selection Index") 
      
      plotly::ggplotly(p)
    })
    
  }
  
}
```

`r if( length(na.omit(unique(result$predictions$mother)))> 0 & length(na.omit(unique(result$predictions$father)))>0 ){"### Parental Performance \n
The following plots allow the user to review the performance of crosses, mothers and fathers for the available traits. If there are too many crosses (more than 150), only mother and fathers will be plotted. \n"}`

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  
  modeling <- result$modeling
  idIndex <- result$status[which(result$status$module == "indexD"),"analysisId"];
  idIndex <- idIndex[length(idIndex)]
  idMta <- modeling[modeling$analysisId == idIndex & modeling$trait == "inputObject","value"]
  predTrait <- result$predictions
  predTrait <- predTrait[predTrait$analysisId %in% c(idMta, idIndex),]
  envsUsed <- unique(result$modeling[which(result$modeling$analysisId %in% idIndex),"environment"])
  predTrait <- predTrait[which(predTrait$environment %in% envsUsed),]
  predTrait$cross <- paste(predTrait$mother, predTrait$father, sep="x")
  
  traitsForBoxplot <- na.omit(unique(predTrait$trait))
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    if( length(na.omit(unique(predTrait$mother)))> 0 & length(na.omit(unique(predTrait$father)))>0 ){
      shiny::selectInput(ns("traitIndexBox"), 
                         label = "Trait:", 
                         multiple = FALSE,
                         choices = traitsForBoxplot, 
      )
    }
    
  }
  
}

```

```{r, results='asis'}

if( length(na.omit(unique(predTrait$mother)))> 0 & length(na.omit(unique(predTrait$father)))>0 ){
  if("params" %in% ls()){ # we will download the document
    
    cat("\n\n###  {.tabset .tabset-pills}       \n\n")
    
    for (iTrait in traitsForBoxplot){ # 
      cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
      
      dd <- predTrait[which(predTrait$trait %in% iTrait),]
      
      p1 <- ggplot2::ggplot(dd, ggplot2::aes(x=reorder(father, -predictedValue), y=predictedValue) ) + 
        # ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0), ggplot2::aes(colour=father), alpha=0.9) +
        ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=father)) +
        ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
                       axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
        ggplot2::labs(x="fathers", y="Trait value")  #+ ggplot2::facet_grid(trait~environment, scales = "free")
      p2 <- ggplot2::ggplot(dd, ggplot2::aes(x=reorder(mother, -predictedValue), y=predictedValue) ) + 
        # ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0), ggplot2::aes(colour=father), alpha=0.9) +
        ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=mother)) +
        ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
                       axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
        ggplot2::labs(x="mothers", y="Trait value") #+ ggplot2::facet_grid(trait~environment, scales = "free")
      
      ply1 <- plotly::ggplotly(p1)
      ply2 <- plotly::ggplotly(p2)
      myPlots <- list(ply1, ply2)
      if(length(unique(dd$cross)) < 150){ # if there's more than 300 crosses don't plot
        p3 <- ggplot2::ggplot(dd, ggplot2::aes(x=reorder(cross, -predictedValue), y=predictedValue) ) + 
          # ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0), ggplot2::aes(colour=father), alpha=0.9) +
          ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=cross)) +
          ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
                         axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
          ggplot2::labs(x="crosses", y="Trait value") #+ ggplot2::facet_grid(trait~environment, scales = "free")
        ply3 <- plotly::ggplotly(p3)
        myPlots <- c(myPlots, ply3)
      }
      print(htmltools::tagList( plotly::subplot( myPlots, nrows=1) ) )
      
      cat("\n")
    }
    
    cat("\n")
    
  }else{
    
    # shiny::renderPlot({
    plotly::renderPlotly({
      if(length(input$traitIndexBox) > 0){
        dd <- predTrait[which(predTrait$trait %in% input$traitIndexBox),]
        
        p1 <- ggplot2::ggplot(dd, ggplot2::aes(x=reorder(father, -predictedValue), y=predictedValue) ) + 
          # ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0), ggplot2::aes(colour=father), alpha=0.9) +
          ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=father)) +
          ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
                         axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
          ggplot2::labs(x="fathers", y="Trait value") #+ ggplot2::facet_grid(trait~environment, scales = "free")
        p2 <- ggplot2::ggplot(dd, ggplot2::aes(x=reorder(mother, -predictedValue), y=predictedValue) ) + 
          # ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0), ggplot2::aes(colour=father), alpha=0.9) +
          ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=mother)) +
          ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
                         axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
          ggplot2::labs(x="mothers", y="Trait value") #+ ggplot2::facet_grid(trait~environment, scales = "free")
        
        ply1 <- plotly::ggplotly(p1)
        ply2 <- plotly::ggplotly(p2)
        myPlots <- list(ply1, ply2)
        if(length(unique(dd$cross)) < 150){ # if there's more than 300 crosses don't plot
          p3 <- ggplot2::ggplot(dd, ggplot2::aes(x=reorder(cross, -predictedValue), y=predictedValue) ) + 
            # ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0), ggplot2::aes(colour=father), alpha=0.9) +
            ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=cross)) +
            ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
                           axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
            ggplot2::labs(x="crosses", y="Trait value") #+ ggplot2::facet_grid(trait~environment, scales = "free")
          ply3 <- plotly::ggplotly(p3)
          myPlots <- c(myPlots, ply3)
        }
        plotly::subplot(myPlots, nrows=1)
        
      }
    })
    
  }
}

```

### References of methods used

Pesek, J., & Baker, R. J. (1969). Desired improvement in relation to selection indices. Canadian journal of plant science, 49(6), 803-804.

Ceron-Rojas, J. J., & Crossa, J. (2018). Linear selection indices in modern plant breeding (p. 256). Springer Nature.

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

<p>&nbsp;</p>

