---
title: "Genotype Verification Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
  toDownload: FALSE
---

```{r setup, include=FALSE}

# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
options(knitr.kable.NA = '')

# loading necessary R packages ####
## data manipulation
library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric

## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(grafify)
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                         numericColNames = NULL, numRound = 3, 
                         scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,...) {
    
    DT <- data.frame(lapply(X = DT, 
                            FUN = function(x) {
                              if(is.numeric(x)){
                                round(x, numRound)
                              } else {
                                x
                              }
                            }))
    
    table <- DT::datatable(data = DT, 
                           colnames = colNames,
                           filter = "top", 
                           options = list(autoWidth = autoWidthOpt,
                                          dom = 'l<<t>pB>', 
                                          buttons = c('copy', 'excel'),
                                          pageLength = pageLength,
                                          searchHighlight = TRUE,
                                          lengthMenu = c(7, 14, 21, 28, 35),
                                          scrollX = scrollXOpt),
                           extensions = 'Buttons',
                           rownames = FALSE,
                           ...)
    if (length(numericColNames) > 0){
      table <- table %>% DT::formatRound(columns = numericColNames,
                                     digits = numRound)
    }
    
    table
  }
```

```{r }

suppressWarnings(tryCatch({
  if(file.exists("./outputs/result.RData")){
    load("./outputs/resultQaPheno.RData")
  }else{
    load("resultQaPheno.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

```


### Verification results
<p>&nbsp;</p>

The following table shows the values associated for the individuals on different metrics such as probability of matching the expected genotype, heterozygosity, among others.

```{r }

if (is.null(result)){
  # invisible(NULL)
} else {
  predictions <- result$predictions
  idQa <- result$status[which(result$status$module %in% c("gVerif")),"analysisId"];
  idQa <- idQa[length(idQa)]
  predictions <- predictions[which( predictions$analysisId == idQa),]
  traitsTable <- unique(predictions$trait)
  
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("traitQa"), 
                       label = "Verification measure:", 
                       choices = traitsTable, 
    )
  }
  
}

```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  predictionsX <- result$predictions[which( result$predictions$analysisId == idQa),]
  
  if("params" %in% ls()){ # we will download the document
    cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
    
    
    for (j in 1:length(traitsTable)){
      cat("\n\n####", traitsTable[j], "{.tabset .tabset-pills}       \n\n")
      
      predictions2 <- predictionsX[which(predictionsX$trait == traitsTable[j]),]
      p <- ggplot2::ggplot(predictions2, ggplot2::aes(x=designation, y=predictedValue  ) ) +
        # geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
        ggplot2::theme_classic()+
        ggplot2::geom_point(alpha = 0.6)+
        # geom_jitter(alpha = 0.6, width = 0.3) +
        ggplot2::labs(x="Genotype", y = traitsTable[j] ) +
        ggplot2::theme(axis.text.x = ggplot2::element_blank()) 
      print(htmltools::tagList(plotly::ggplotly(p)))
      
    }
    
  }else{ # regular rendering
    plotly::renderPlotly({
      # myPreds <- myResult %>% filter(trait == input$traitQa)
      predictions2 <- predictionsX[which(predictionsX$trait == input$traitQa),]
      p <- ggplot(predictions2, aes(x=designation, y=predictedValue  ) ) +
        # geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
        theme_classic()+
        geom_point(alpha = 0.6)+
        # geom_jitter(alpha = 0.6, width = 0.3) +
        labs(x="Genotype", y = input$traitQa) +
        theme(axis.text.x = element_blank() ) 
      fig <-  plotly::ggplotly(p)#, height = get_h()*2);
      fig
    })
  }
  
}

```


```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  numericalColumns <- c("predictedValue") 
  
  if("params" %in% ls()){ # we will download the document
    printTable(predictions, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(predictions),
           numericColNames = numericalColumns,
           numRound = 4)
  }else{
    DT::renderDT(printTable(predictions[predictions$trait==input$traitQa,c("trait","gid","designation","mother","father","predictedValue")],
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(predictions[predictions$trait==input$traitQa,c("trait","gid","designation","mother","father","predictedValue")]),
                        numericColNames = numericalColumns,
                        numRound = 4), server = FALSE)
  }
  
}

```



### Verification Metrics

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  numericalColumns2 <- c("value") 
  metricsVeri <- result$metrics[which( result$metrics$analysisId == idQa),]
  
  if("params" %in% ls()){ # we will download the document
    printTable(metricsVeri, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(metricsVeri),
           numericColNames = numericalColumns2,
           numRound = 4)
  }else{
    DT::renderDT(printTable(metricsVeri,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(metricsVeri),
                        numericColNames = numericalColumns2,
                        numRound = 4), server = FALSE)
  }
  
}

```

<p>&nbsp;</p>

### Verification Modeling

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  # numericalColumns2 <- c("value") 
  modelingVeri <- result$modeling[which( result$modeling$analysisId == idQa),]
  
  if("params" %in% ls()){ # we will download the document
    printTable(modelingVeri, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(modelingVeri))
  }else{
    DT::renderDT(printTable(modelingVeri,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(modelingVeri)), server = FALSE)
  }
  
}

```

<p>&nbsp;</p>

### References on methods used

Tukey, J. W. (1977). Exploratory Data Analysis. Section 2C.

Velleman, P. F. and Hoaglin, D. C. (1981). Applications, Basics and Computing of Exploratory Data Analysis. Duxbury Press.


<p>&nbsp;</p>



