---
  title: "Quality Assurance Pheno Report"
author: ""
date: "December 2023"
output: html_document
params:
  toDownload: FALSE
---

```{r setup, include=FALSE}

# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
echo = FALSE,
cache = FALSE,
warning = FALSE,
message = FALSE,
comment = NA,
out.width = "100%",
error = TRUE)
options(knitr.kable.NA = '')

# loading necessary R packages ####
## data manipulation
library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric

## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(grafify)
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, ...) {
  
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  
  table <- DT::datatable(data = DT, 
                         filter = "top", 
                         options = list(autoWidth = TRUE, 
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'excel'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35)),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r }

suppressWarnings(tryCatch({
  if(file.exists("./outputs/result.RData")){
    load("./outputs/resultQaPheno.RData")
  }else{
    load("resultQaPheno.RData")
  }
}, error = function(e) {
  # NULL
}))

```


### Modifications table
<p>&nbsp;</p>

The following table shows the modifications saved to be applied for the raw phenotypic data when is used in one of the analytical modules.

```{r }

if (is.null(result)){
  # invisible(NULL)
} else {
  modifications <- result$modifications$pheno
  idQa <- result$status[which(result$status$module %in% c("qaRaw","qaMb","qaFilter","qaDesign")),"analysisId"];
  idQa <- idQa[length(idQa)]
  idQaType <- result$status[result$status$analysisId == idQa, "module"]
  modifications <- modifications[which( modifications$analysisId == idQa & modifications$module == idQaType),]
  traitsTable <- unique(modifications[which( modifications$analysisId == idQa),"trait"])
  
  
  if("params" %in% ls()){ # we will download the document
    
  }else{
    shiny::selectInput(ns("traitQa"), 
                       label = "Trait:", 
                       choices = traitsTable, 
    )
  }
  
}

```


```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  numericalColumns <- c("value","row") 
  
  if("params" %in% ls()){ # we will download the document
    printTable(modifications)
  }else{
    DT::renderDT({DT::formatRound(DT::datatable(
      modifications[modifications$trait==input$traitQa,],
      rownames = FALSE,
      class = 'cell-border stripe',
      options = list(
        scrollY = "400px",
        scrollX = "400px",
        paging = F)
    ),numericalColumns)
    })
  }
  
}

```



<p>&nbsp;</p>

### Modifications plot

The following plot allows you to visualize which records were tagged as outliers and will be ignored in posterior analyses if this QA time stamp is added to the posterior analyses.

<p>&nbsp;</p>

```{r }

if (is.null(result)){
# invisible(NULL)
} else {

traitsBoxplot <- unique(modifications[which( modifications$analysisId == idQa),"trait"])
if(idQaType == "qaMb"){traitsBoxplot <- paste0(traitsBoxplot,"-residual")}

if("params" %in% ls()){ # we will download the document

}else{
shiny::selectInput(ns("traitQaBox"), 
label = "Trait:", 
choices = traitsBoxplot, 
)
}

}

```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  # object <- result
  mydata <- result$data$pheno
  ### change column names for mapping
  paramsPheno <- result$metadata$pheno
  paramsPheno <- paramsPheno[which(paramsPheno$parameter != "trait"),]
  colnames(mydata) <- cgiarBase::replaceValues(colnames(mydata), 
                                               Search = paramsPheno$value, 
                                               Replace = paramsPheno$parameter )
  ###
  mydata$rowindex <- 1:nrow(mydata)
  mydata[, "environment"] <- as.factor(mydata[, "environment"])
  mydata[, "designation"] <- as.factor(mydata[, "designation"])
  mo <-result$modifications$pheno
  analysisIds <- unique(mo$analysisId)
  analysisIds <- analysisIds[length(analysisIds)]
  mo <- mo[which(mo$analysisId == analysisIds),]
  mydata$color <- "valid"
  if(nrow(mo) > 0){mydata$color[which(mydata$rowindex %in% unique(mo$row))]="tagged"}
  mydata$color <- as.factor(mydata$color)
  
  if("params" %in% ls()){ # we will download the document
    
    cat(paste0("##   {.tabset .tabset-pills}    \n\n  "))
    
    for(iTrait in traitsBoxplot){
      cat("\n\n###", iTrait, "{.tabset .tabset-pills}       \n\n")
      
      mydata$predictedValue <- mydata[,iTrait]
      
      p <- ggplot(mydata, aes(x=as.factor(environment), y=predictedValue)) +
        geom_boxplot(fill='#A4A4A4', color="black", notch = TRUE)+
        theme_classic()+
        geom_jitter(aes(colour = color), alpha = 0.4) +
        xlab("Environment") + ylab("Trait value") +
        theme(axis.text.x = element_text(angle = 45))
      print(p)
    }
    
  }else{
    
    shiny::renderPlot({
      mydata$predictedValue <- mydata[,input$traitQaBox]
      
      ggplot(mydata, aes(x=as.factor(environment), y=predictedValue)) +
        geom_boxplot(fill='#A4A4A4', color="black", notch = TRUE)+
        theme_classic()+
        geom_jitter(aes(colour = color), alpha = 0.4) +
        xlab("Environment") + ylab("Trait value") +
        theme(axis.text.x = element_text(angle = 45))
      
    })
    
  }
  
}
```









