---
title: "STA Report"
author: ""
date: "2023-11-03"
output: html_document
params:
 toDownload: FALSE
---

```{r setup, include=FALSE}
# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
options(knitr.kable.NA = '')
# loading necessary R packages ####
## data manipulation
# library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric
## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, ...) {
  
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  
  table <- DT::datatable(data = DT, 
                         filter = "top", 
                         options = list(autoWidth = TRUE, 
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'excel'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35)),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r, include=FALSE}
# Init Step to make sure that the dependencies are loaded
htmltools::tagList(printTable(mtcars))
htmltools::tagList(ggplotly(ggplot()))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
```

### Entries and traits by environment

The following table allows to see which locations had data for the different traits.

<p>&nbsp;</p>

```{r, results='asis'}
suppressWarnings(tryCatch({
  if(file.exists("./outputs/resultSta.RData")){
    load("./outputs/resultSta.RData")
  }else{
    load("resultSta.RData")
  }
}, error = function(e) {
  # NULL
}))

if (is.null(result)){
  # invisible(NULL)
} else {
  pred <- result$predictions
  traits <- unique(pred$trait)
  traits.env <- aggregate(designation ~ environment + trait, data = pred, FUN = length)
  traits.env <- reshape(data=traits.env, timevar = "trait", idvar = "environment", direction="wide")
  colnames(traits.env) <- c('environment',traits)
  summarySta <- as.data.frame(traits.env[,'environment'])
  for (i in 1:length(traits)){
    temp <- traits.env[,c('environment',traits[i])]
    colnames(temp)[2] <- "temp.col"
    temp <- as.data.frame(temp)
    j=i+1
    tempp=within(temp,{
      temp.col.n=NA
      temp.col.n[!is.na(temp.col)]='✅ '
      temp.col.n[is.na(temp.col)]=''
    })
    temp <- tempp[,3]
    summarySta <- cbind(summarySta,temp)
  }
  colnames(summarySta) <- c('environment',traits)
  
  pheno <- result$data$pheno
  ### change column names for mapping
  paramsPheno <- result$metadata$pheno
  paramsPheno <- paramsPheno[which(paramsPheno$parameter != "trait"),]
  colnames(pheno) <- cgiarBase::replaceValues(colnames(pheno), Search = paramsPheno$value, Replace = paramsPheno$parameter )
  ###
  pheno.env <- pheno[,c('environment','designation')]
  pheno.env<- unique(pheno.env[c('designation','environment')]) 
  pheno.env.designation <-  aggregate(designation ~environment, data = pheno.env, FUN = length)
  colnames(pheno.env.designation)[2] <- "# of entries"
  summarySta <- merge(x = summarySta, y = pheno.env.designation, by = "environment", all = TRUE)
  
  if("params" %in% ls()){ # we will download the document
    printTable(summarySta)
  }else{
    summarySta <- kableExtra::kbl(summarySta)
    summarySta <-kableExtra::kable_styling(
      summarySta, 
      bootstrap_options = c("hover","condensed","responsive"), fixed_thead = T)
    kableExtra::scroll_box(summarySta,width = "100%")
  }
  
}
```

<p>&nbsp;</p>

### Summary statistics

The following table allows you to verify different quality metrics (KPIs) for the different environments. The selector button allows you to filter by trait.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  modeling <- result$modeling
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"];
  idSta <- idSta[length(idSta)]
  modeling <- modeling[which(modeling$analysisId == idSta),"trait"]
  
  if("params" %in% ls()){ # we will download the document
    
  }else{
    shiny::selectInput(ns("traitSta"), 
                       label = "Trait:", 
                       multiple = FALSE,
                       choices = unique(modeling), 
    )
  }
  
}

```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  pred.traits <- subset(pred, select = c(trait,designation,environment,predictedValue))
  pred.traits <- reshape(
    data=pred.traits, 
    timevar = c("trait"), 
    idvar = c("environment","designation"), 
    direction="wide"
  )
  names(pred.traits)[-c(1,2)] <- gsub("predictedValue.", "", names(pred.traits)[-c(1,2)]) 
  pred.traits <- reshape(pred.traits, 
                         direction = "long",
                         varying = list(names(pred.traits)[3:ncol(pred.traits)]),
                         v.names = "Value",
                         idvar = c("designation", "environment"),
                         timevar = "trait",
                         times = traits)
  rownames(pred.traits) <- NULL
  ## Max
  pred.traits.max <- aggregate(Value ~ environment + trait, data = pred.traits, function(x) max(x,na.rm=T))
  colnames(pred.traits.max)[3] <- "max"
  ## Min
  pred.traits.min <- aggregate(Value ~ environment + trait, data = pred.traits, function(x) min(x,na.rm=T))
  colnames(pred.traits.min)[3] <- "min"
  ## Mean
  pred.traits.mean <- aggregate(Value ~ environment + trait, data = pred.traits, function(x) mean(x,na.rm=T))
  colnames(pred.traits.mean)[3] <- "mean"
  ## H2
  metrics <- result$metrics
  metrics.h2 <- metrics[metrics$parameter=="plotH2",]
  metrics.h2 <- subset(metrics.h2, select = c(trait,environment,value))
  colnames(metrics.h2)[3] <- "H2"
  ## CV
  metrics.CV <- metrics[metrics$parameter=="CV",]
  metrics.CV <- subset(metrics.CV, select = c(trait,environment,value))
  colnames(metrics.CV)[3] <- "CV"
  ## r2
  metrics.r2 <- metrics[metrics$parameter=="r2",]
  metrics.r2 <- subset(metrics.r2, select = c(trait,environment,value))
  colnames(metrics.r2)[3] <- "r2"
  ## Vg
  metrics.Vg <- metrics[metrics$parameter=="Vg",]
  metrics.Vg <- subset(metrics.Vg, select = c(trait,environment,value))
  colnames(metrics.Vg)[3] <- "Vg"
  ## Vr
  metrics.Vr <- metrics[metrics$parameter=="Vr",]
  metrics.Vr <- subset(metrics.Vr, select = c(trait,environment,value))
  colnames(metrics.Vr)[3] <- "Vr"
  
  pred.traits.summary <- merge(x = pred.traits.min, y = pred.traits.mean, by = c("environment","trait"), all = TRUE)
  pred.traits.summary <- merge(x = pred.traits.summary, y = pred.traits.max, by = c("environment","trait"), all = TRUE)
  pred.traits.summary <- merge(x = pred.traits.summary, y = metrics.h2, by = c("environment","trait"), all = TRUE)
  pred.traits.summary <- merge(x = pred.traits.summary, y = metrics.CV, by = c("environment","trait"), all = TRUE)
  pred.traits.summary <- merge(x = pred.traits.summary, y = metrics.r2, by = c("environment","trait"), all = TRUE)
  pred.traits.summary <- merge(x = pred.traits.summary, y = metrics.Vg, by = c("environment","trait"), all = TRUE)
  pred.traits.summary <- merge(x = pred.traits.summary, y = metrics.Vr, by = c("environment","trait"), all = TRUE)
  numeric.output <- c("min","mean","max","H2","CV","r2","Vg","Vr")
  
  if("params" %in% ls()){ # we will download the document
    printTable(pred.traits.summary)
  }else{
    DT::renderDT({DT::formatRound(DT::datatable(
      pred.traits.summary[pred.traits.summary$trait==input$traitSta,],
      rownames = FALSE,
      class = 'cell-border stripe',
      options = list(
        scrollY = "400px",
        scrollX = "400px",
        paging = F)
    ),numeric.output)
    })
  }
  
}
```

<p>&nbsp;</p>

### Predictions 

The following table allows you to check the trait predictions in wide format and filter them by environment.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  modeling <- result$modeling
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"];
  idSta <- idSta[length(idSta)]
  envs <- unique(result$predictions[which( result$predictions$analysisId == idSta),"environment"])
  
  if("params" %in% ls()){ # we will download the document
    
  }else{
    shiny::selectInput(ns("envSta"), 
                       label = "Environment:", 
                       choices = envs, 
    )
  }
  
}

```


```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  traits <- unique(pred$trait)
  predictions.designation <- subset(pred, select = c(trait,designation,environment,predictedValue))
  predictions.designation <- reshape(
    data=predictions.designation, 
    timevar = c("trait"), 
    idvar = c("environment","designation"), 
    direction="wide"
  )
  names(predictions.designation) <- c("deignation","environment",traits)
  
  if("params" %in% ls()){ # we will download the document
    printTable(predictions.designation)
  }else{
    DT::renderDT({DT::formatRound(DT::datatable(
      predictions.designation[predictions.designation$environment==input$envSta,],
      rownames = FALSE,
      class = 'cell-border stripe',
      options = list(
        scrollY = "400px",
        scrollX = "400px",
        paging = F)
    ),traits)
    })
  }
  
}

```

<p>&nbsp;</p>

### Boxplot

The following boxplot allows you to see the distribution of predicted values by trait (y-axis) in the different environments.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  modeling <- result$modeling
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"];
  idSta <- idSta[length(idSta)]
  modeling <- modeling[which(modeling$analysisId == idSta),"trait"]
  
  if("params" %in% ls()){ # we will download the document
    
  }else{
    shiny::selectInput(ns("traitStaBox"), 
                       label = "Trait:", 
                       multiple = FALSE,
                       selected = unique(modeling),
                       choices = unique(modeling), 
    )
  }
  
}

```

```{r, results='asis'}
predTrait <- result$predictions

if("params" %in% ls()){ # we will download the document
  traits <- unique(modeling)
  cat("\n\n## Trait view {.tabset .tabset-pills}       \n\n")
  cat("\n\n The following boxplot allows you to see the distribution of predicted values by trait in the different environments.       \n\n")
  
  for (j in 1:length(traits)){
    cat("\n\n###", traits[j], "{.tabset .tabset-pills}       \n\n")
    
    p <- ggplot2::ggplot(predTrait[which(predTrait$trait %in% traits[j]),], ggplot2::aes(x=environment, y=predictedValue, label=designation) ) + 
      ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0.2), ggplot2::aes(colour=environment), alpha=0.9) +
      # ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=factor(environment))) +
      ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
                     axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
      ggplot2::labs(x="", y="Trait value") #+ ggplot2::facet_grid(trait~., scales = "free") 
    print(htmltools::tagList(ggplotly(p)))
    
    cat("\n")
  }
  
  cat("\n")
  
}else{
  plotly::renderPlotly({
    if(length(input$traitStaBox) > 0){
      p <- ggplot2::ggplot(predTrait[which(predTrait$trait %in% input$traitStaBox),], ggplot2::aes(x=environment, y=predictedValue, label=designation) ) +
        ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0.2), ggplot2::aes(colour=environment), alpha=0.9) +
        # ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=factor(environment))) +
        ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"),
                       axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
        ggplot2::labs(x="", y="Trait value") #+ ggplot2::facet_grid(trait~., scales = "free") #+
      # ggplot2::guides(color=ggplot2::guide_legend("Environment"))

      fig2 <-  plotly::ggplotly(p)#, height = get_h()*2); # fig2
      fig2
    }

  })
  
    # if(length(input$traitStaBox) > 0){
    #   p <- ggplot2::ggplot(predTrait[which(predTrait$trait %in% input$traitStaBox),], ggplot2::aes(x=environment, y=predictedValue, label=designation) ) + 
    #     ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0.2), ggplot2::aes(colour=environment), alpha=0.9) +
    #     # ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=factor(environment))) +
    #     ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
    #                    axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
    #     ggplot2::labs(x="", y="Trait value") #+ ggplot2::facet_grid(trait~., scales = "free") #+
    #   # ggplot2::guides(color=ggplot2::guide_legend("Environment"))
    #   
    #   p
    # }
}

```
