---
title: "Quality Assurance Pheno Report"
author: ""
date: "December 2023"
output: html_document
params:
 toDownload: FALSE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
library(dplyr)
```

```{r plot_init, include = FALSE}
# Init Step to make sure that the dependencies are loaded
# was done to enable proper printing of the tab header and the graphs within the for loop below using print(htmltools::tagList(ggplotly(...)))
htmltools::tagList(plotly::ggplotly(ggplot2::ggplot()))

# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}

```

### Modifications table
<p>&nbsp;</p>

The following table shows the modifications saved to be applied for the raw phenotypic data when is used in one of the analytical modules.

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  modifications <- result$modifications$pheno
  idQa <- result$status[which(result$status$module %in% c("qaRaw","qaMb")),"analysisId"];
  idQa <- idQa[length(idQa)]
  traits <- unique(modifications[which( modifications$analysisId == idQa),"trait"])
  
  if("params" %in% ls()){ # we will download the document
    
  }else{
    shiny::selectInput(ns("traitQa"), 
                       label = "Trait:", 
                       choices = traits, 
    )
  }
  
}

```


```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  traits <- c("value","row") 
  
  if("params" %in% ls()){ # we will download the document
    printTable(modifications)
  }else{
    DT::renderDT({DT::formatRound(DT::datatable(
      modifications[modifications$trait==input$traitQa,],
      rownames = FALSE,
      class = 'cell-border stripe',
      options = list(
        scrollY = "400px",
        scrollX = "400px",
        paging = F)
    ),traits)
    })
  }
  
}

```

<p>&nbsp;</p>

### Modifications plot

The following density plot allows you to see the expected change in the next generation compared to the current generation given the current genetic gain parameters (accuracy, intensity, genetic variance) obtained in a particular MET for the selected traits.

<p>&nbsp;</p>

```{r }

suppressWarnings(tryCatch({
  if(file.exists("./outputs/resultPgg.RData")){
    load("./outputs/resultPgg.RData")
  }else{
    load("resultPgg.RData")
  }
}, error = function(e) {
  # NULL
}))

if (is.null(result)){
  # invisible(NULL)
} else {
  modeling <- result$modeling
  idPgg <- result$status[which(result$status$module == "pgg"),"analysisId"];
  idPgg <- idPgg[length(idPgg)]
  modeling <- modeling[which(modeling$analysisId == idPgg & modeling$environment == "across"),"trait"]
  if("params" %in% ls()){
    
  }else{
    shiny::selectInput(ns("traitFilterPredictions2D2"),
                       label = "Trait:",
                       choices = unique(modeling),
    )
  }
}

```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  metrics <- result$metrics
  idPgg <- result$status[which(result$status$module == "pgg"),"analysisId"];
  idPgg <- idPgg[length(idPgg)]
  current.metrics <- metrics[which(metrics$analysisId %in% c(idPgg)),]
  
  if("params" %in% ls()){
    cat(paste0("## Selection differentials and expected response  {.tabset .tabset-pills}    \n\n  "))
    # l <- htmltools::tagList()
    # counter <- 1
    for(iTrait in unique(modeling)){
      cat("\n\n###", iTrait, "{.tabset .tabset-pills}       \n\n")
      mu0 <-current.metrics[which(current.metrics$parameter == "meanG" & current.metrics$trait == iTrait ),"value"]
      sd0 <-current.metrics[which(current.metrics$parameter == "sigmaG" & current.metrics$trait == iTrait ),"value"]
      R <-current.metrics[which(current.metrics$parameter == "R" & current.metrics$trait == iTrait ),"value"]
      vline <- function(x = 0, color = "darkblue") {list( type = "line",y0 = 0,y1 = 1,yref = "paper",x0 = x, x1 = x,line = list(color = color, dash="dot"))}
      vline2 <- function(x = 0, color = "blue") {list( type = "line",y0 = 0,y1 = 1,yref = "paper",x0 = x, x1 = x,line = list(color = color))}
      vline3 <- function(x = 0, color = "red") {list( type = "line",y0 = 0,y1 = 1,yref = "paper",x0 = x, x1 = x,line = list(color = color))}
      i0 <-current.metrics[which(current.metrics$parameter == "i" & current.metrics$trait == iTrait ),"value"]
      p <- plotly::plot_ly(alpha = 0.6) %>% plotly::add_histogram(x = rnorm(5000, mean=mu0, sd = sd0 ), histnorm = "probability", name="Current Generation" ) %>% plotly::add_histogram(x = rnorm(5000, mean=mu0+R, sd = sd0 ), histnorm = "probability", name="Expected New Generation" ) %>% plotly::layout(barmode = "overlay",xaxis = list(title = "Trait value"), yaxis = list(title = "Probability" ),  shapes = list(vline(mu0+(sd0*i0)), vline2(mu0), vline3(mu0+R) ) )
      print(htmltools::tagList(plotly::ggplotly(p)))
      # l[[counter]] <- tagList(HTML(markdown::mark(text=paste0("\n\n#### ", iTrait, "\n"))),
      # 
      #                        plotly::plot_ly(alpha = 0.6) %>% plotly::add_histogram(x = rnorm(5000, mean=mu0, sd = sd0 ), histnorm = "probability", name="Current Generation" ) %>% plotly::add_histogram(x = rnorm(5000, mean=mu0+R, sd = sd0 ), histnorm = "probability", name="Expected New Generation" ) %>% plotly::layout(barmode = "overlay",xaxis = list(title = "Trait value"), yaxis = list(title = "Probability" ),  shapes = list(vline(mu0+(sd0*i0)), vline2(mu0), vline3(mu0+R) ) )
      # 
      # )
      # counter <- counter + 1
    }
  }else{
    plotly::renderPlotly({
      fig <- plotly::plot_ly(alpha = 0.6)
      ## plot for the general population
      mu0 <-current.metrics[which(current.metrics$parameter == "meanG" & current.metrics$trait == input$traitFilterPredictions2D2 ),"value"];
      sd0 <-current.metrics[which(current.metrics$parameter == "sigmaG" & current.metrics$trait == input$traitFilterPredictions2D2 ),"value"];
      fig <- fig %>% plotly::add_histogram(x = rnorm(5000, mean=mu0, sd = sd0 ), histnorm = "probability", name="Current Generation" )
      ## plot for the expected new population
      R <-current.metrics[which(current.metrics$parameter == "R" & current.metrics$trait == input$traitFilterPredictions2D2 ),"value"];
      fig <- fig %>% plotly::add_histogram(x = rnorm(5000, mean=mu0+R, sd = sd0 ), histnorm = "probability", name="Expected New Generation" )
      ## add line for the selected population
      vline <- function(x = 0, color = "darkblue") {list( type = "line",y0 = 0,y1 = 1,yref = "paper",x0 = x, x1 = x,line = list(color = color, dash="dot"))}
      vline2 <- function(x = 0, color = "blue") {list( type = "line",y0 = 0,y1 = 1,yref = "paper",x0 = x, x1 = x,line = list(color = color))}
      vline3 <- function(x = 0, color = "red") {list( type = "line",y0 = 0,y1 = 1,yref = "paper",x0 = x, x1 = x,line = list(color = color))}
      # overlay the histograms
      i0 <-current.metrics[which(current.metrics$parameter == "i" & current.metrics$trait == input$traitFilterPredictions2D2 ),"value"];
      fig <- fig %>% plotly::layout(barmode = "overlay",xaxis = list(title = "Trait value"), yaxis = list(title = "Probability" ),  shapes = list(vline(mu0+(sd0*i0)), vline2(mu0), vline3(mu0+R) ) )
      fig
    })
  }
}
```






