---
title: "Single Trial Analysis Report"
author: ""
date: "2023-11-03"
output: html_document
params:
 toDownload: FALSE
---

```{r setup, include=FALSE}
# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
options(knitr.kable.NA = '')
# loading necessary R packages ####
## data manipulation
# library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric
## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, ...) {
  
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  
  table <- DT::datatable(data = DT, 
                         filter = "top", 
                         options = list(autoWidth = TRUE, 
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'excel'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35)),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r, include=FALSE}
# Init Step to make sure that the dependencies are loaded
htmltools::tagList(printTable(mtcars))
htmltools::tagList(ggplotly(ggplot()))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}
```

### Entries and traits by environment table

The following table allows to see which locations had data for the different traits.

<p>&nbsp;</p>

```{r, results='asis'}
suppressWarnings(tryCatch({
  if(file.exists("./outputs/resultSta.RData")){
    load("./outputs/resultSta.RData")
  }else{
    load("resultSta.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

if (is.null(result)){
  
} else {
  
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"];
  idSta <- idSta[length(idSta)]
  
  pred <- result$predictions
  pred <- pred[pred$analysisId == idSta,]
  traits <- unique(pred$trait)
  traits.env <- aggregate(designation ~ environment + trait, data = pred, FUN = length)
  traits.env <- reshape(data=traits.env, timevar = "trait", idvar = "environment", direction="wide")
  colnames(traits.env) <- c('environment',traits)
  summarySta <- as.data.frame(traits.env[,'environment'])
  for (i in 1:length(traits)){
    temp <- traits.env[,c('environment',traits[i])]
    colnames(temp)[2] <- "temp.col"
    temp <- as.data.frame(temp)
    j=i+1
    tempp=within(temp,{
      temp.col.n=NA
      temp.col.n[!is.na(temp.col)]='✅ '
      temp.col.n[is.na(temp.col)]=''
    })
    temp <- tempp[,3]
    summarySta <- cbind(summarySta,temp)
  }
  colnames(summarySta) <- c('environment',traits)
  
  pheno <- result$data$pheno
  ### change column names for mapping
  paramsPheno <- result$metadata$pheno
  paramsPheno <- paramsPheno[which(paramsPheno$parameter != "trait"),]
  colnames(pheno) <- cgiarBase::replaceValues(colnames(pheno), Search = paramsPheno$value, Replace = paramsPheno$parameter )
  ###
  pheno.env <- pheno[,c('environment','designation')]
  pheno.env<- unique(pheno.env[c('designation','environment')]) 
  pheno.env.designation <-  aggregate(designation ~environment, data = pheno.env, FUN = length)
  colnames(pheno.env.designation)[2] <- "Number of entries"
  summarySta <- merge(x = summarySta, y = pheno.env.designation, by = "environment", all = TRUE)
  
  if("params" %in% ls()){ # we will download the document
    
    printTable(summarySta)
  }else{
    summarySta <- kableExtra::kbl(summarySta)
    summarySta <-kableExtra::kable_styling(
      summarySta, 
      bootstrap_options = c("hover","condensed","responsive"), fixed_thead = T)
    kableExtra::scroll_box(summarySta,width = "100%")
  }
  
}
```

<p>&nbsp;</p>

### Summary statistics

The following table allows you to verify different quality metrics (KPIs) for the different environments. The selector button allows you to filter by trait.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  modeling <- result$modeling
  staTraits <- setdiff( unique(modeling[which(modeling$analysisId == idSta),"trait"]), "inputObject")
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    
    shiny::selectInput(ns("traitSta"), 
                       label = "Trait:", 
                       multiple = FALSE,
                       choices = staTraits, 
    )
  }
  
}

```

Table of statistics:

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  pred.traits <- subset(pred, select = c(trait,designation,environment,predictedValue))
  pred.traits <- reshape(
    data=pred.traits, 
    timevar = c("trait"), 
    idvar = c("environment","designation"), 
    direction="wide"
  )
  names(pred.traits)[-c(1,2)] <- gsub("predictedValue.", "", names(pred.traits)[-c(1,2)]) 
  pred.traits <- reshape(pred.traits, 
                         direction = "long",
                         varying = list(names(pred.traits)[3:ncol(pred.traits)]),
                         v.names = "Value",
                         idvar = c("designation", "environment"),
                         timevar = "trait",
                         times = traits)
  rownames(pred.traits) <- NULL
  ## Max
  pred.traits.max <- aggregate(Value ~ environment + trait, data = pred.traits, function(x) max(x,na.rm=T))
  colnames(pred.traits.max)[3] <- "max"
  ## Min
  pred.traits.min <- aggregate(Value ~ environment + trait, data = pred.traits, function(x) min(x,na.rm=T))
  colnames(pred.traits.min)[3] <- "min"
  ## Mean
  pred.traits.mean <- aggregate(Value ~ environment + trait, data = pred.traits, function(x) mean(x,na.rm=T))
  colnames(pred.traits.mean)[3] <- "mean"
  ## H2
  metrics <- result$metrics
  metrics.h2 <- metrics[metrics$parameter=="plotH2",]
  metrics.h2 <- subset(metrics.h2, select = c(trait,environment,value))
  colnames(metrics.h2)[3] <- "H2"
  ## CV
  metrics.CV <- metrics[metrics$parameter=="CV",]
  metrics.CV <- subset(metrics.CV, select = c(trait,environment,value))
  colnames(metrics.CV)[3] <- "CV"
  ## r2
  metrics.r2 <- metrics[metrics$parameter=="r2",]
  metrics.r2 <- subset(metrics.r2, select = c(trait,environment,value))
  colnames(metrics.r2)[3] <- "r2"
  ## Vg
  metrics.Vg <- metrics[metrics$parameter=="Vg",]
  metrics.Vg <- subset(metrics.Vg, select = c(trait,environment,value))
  colnames(metrics.Vg)[3] <- "Vg"
  ## Vr
  metrics.Vr <- metrics[metrics$parameter=="Vr",]
  metrics.Vr <- subset(metrics.Vr, select = c(trait,environment,value))
  colnames(metrics.Vr)[3] <- "Vr"
  
  pred.traits.summary <- merge(x = pred.traits.min, y = pred.traits.mean, by = c("environment","trait"), all = TRUE)
  pred.traits.summary <- merge(x = pred.traits.summary, y = pred.traits.max, by = c("environment","trait"), all = TRUE)
  pred.traits.summary <- merge(x = pred.traits.summary, y = metrics.h2, by = c("environment","trait"), all = TRUE)
  pred.traits.summary <- merge(x = pred.traits.summary, y = metrics.CV, by = c("environment","trait"), all = TRUE)
  pred.traits.summary <- merge(x = pred.traits.summary, y = metrics.r2, by = c("environment","trait"), all = TRUE)
  pred.traits.summary <- merge(x = pred.traits.summary, y = metrics.Vg, by = c("environment","trait"), all = TRUE)
  pred.traits.summary <- merge(x = pred.traits.summary, y = metrics.Vr, by = c("environment","trait"), all = TRUE)
  numeric.output <- c("min","mean","max","H2","CV","r2","Vg","Vr")
  
  if("params" %in% ls()){ # we will download the document
    printTable(unique(pred.traits.summary))
  }else{
    DT::renderDT({DT::formatRound(DT::datatable(
      unique(pred.traits.summary[pred.traits.summary$trait==input$traitSta,]),
      rownames = FALSE,
      class = 'cell-border stripe',
      options = list(
        scrollY = "300px",
        scrollX = "300px",
        paging = F)
    ),numeric.output)
    })
  }
  
}
```

Variance components ratios per environment 

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  modeling <- result$modeling
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"];
  idSta <- idSta[length(idSta)]
  staTraits2 <- setdiff( unique(modeling[which(modeling$analysisId == idSta),"trait"]), "inputObject")
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("traitSta0"), 
                       label = "Trait:", 
                       multiple = FALSE,
                       choices = staTraits2, 
    )
  }
  
}

```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  df2 <- result$metrics
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"];
  idSta <- idSta[length(idSta)]
  df2 <- df2[df2$analysisId==idSta,]
  df2 <- df2[df2$parameter %in% c("Vg","Vr"), ]
  total <- aggregate(value~environment+trait, FUN = sum, data=df2); 
  colnames(total)[3] <- "totalVar"
  df3 <- merge(df2,total, by=c("environment","trait") )
  df3$valueProp <- df3$value/df3$totalVar
  
  if("params" %in% ls()){ # we will download the document
    
    cat("\n\n###  {.tabset .tabset-pills}       \n\n")
    
    for (j in 1:length(staTraits2)){
      cat("\n\n####", staTraits2[j], "{.tabset .tabset-pills}       \n\n")
      df4 <- df3[df3$trait == staTraits2[j],]
      p <- ggplot2::ggplot(data=df4, ggplot2::aes(x=environment, y=valueProp, fill=parameter)) +
        ggplot2::geom_bar(stat="identity") +
        ggplot2::xlab("Environment") + ggplot2::ylab("Proportion of variance") +
        ggplot2::scale_fill_brewer(palette="Accent") +
        # ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust=1, vjust = 1)) +
        ggplot2::theme(axis.text.x = ggplot2::element_blank())
      ggplot2::geom_hline(yintercept = seq(0.25,0.75,0.25), linetype = 3, colour = 'red')
      # print(p)
      p2 <- ggplot2::ggplot(data=df4, ggplot2::aes(x=environment, y=value, fill=parameter)) +
        ggplot2::geom_bar(stat="identity") +
        ggplot2::xlab("Environment") + ggplot2::ylab("Value of variance") +
        ggplot2::scale_fill_brewer(palette="Accent") +
        ggplot2::theme(axis.text.x = ggplot2::element_blank())
      
      print(htmltools::tagList( plotly::subplot(plotly::ggplotly(p), plotly::ggplotly(p2), nrows=2) ) )
      
      cat("\n")
    }
    
    cat("\n")
    
  }else{
    
    # shiny::renderPlot({
    plotly::renderPlotly({
      df3 <- df3[df3$trait == input$traitSta0,]
      p <- ggplot2::ggplot(data=df3, ggplot2::aes(x=environment, y=valueProp, fill=parameter)) +
        ggplot2::geom_bar(stat="identity") +
        ggplot2::xlab("Environment") + ggplot2::ylab("Proportion of variance") +
        scale_fill_brewer(palette="Accent") +
        # ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust=1, vjust = 1)) +
        ggplot2::theme(axis.text.x = ggplot2::element_blank())
      ggplot2::geom_hline(yintercept = seq(0.25,0.75,0.25), linetype = 3, colour = 'red')
      p2 <- ggplot2::ggplot(data=df3, ggplot2::aes(x=environment, y=value, fill=parameter)) +
        ggplot2::geom_bar(stat="identity") +
        ggplot2::xlab("Environment") + ggplot2::ylab("Value of variance") +
        scale_fill_brewer(palette="Accent") +
        # ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust=1, vjust = 1))
        ggplot2::theme(axis.text.x = ggplot2::element_blank())
      # gridExtra::grid.arrange(p,p2)
      ply1 <- plotly::ggplotly(p)
      ply2 <- plotly::ggplotly(p2)
      plotly::subplot(ply1, ply2, nrows=2)
    })
  }
  
}
```

<p>&nbsp;</p>

```{r}
if (is.null(result)){
  # invisible(NULL)
} else {
  metrics <- result$metrics
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"];
  idSta <- idSta[length(idSta)]
  parameters <- metrics[which(metrics$analysisId == idSta),"parameter"]
  if("params" %in% ls()){  # if user will download the report
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("parameterMetrics"), 
                       label = "Parameter:", 
                       multiple = FALSE,
                       selected = setdiff( unique(parameters), c("Vg","Vr") ),
                       choices = setdiff( unique(parameters), c("Vg","Vr") )
    )
  }
}
```

Barplot for parameter values

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  mydata = result$metrics
  mydata <- mydata[which(mydata$analysisId %in% idSta),]
  
  if("params" %in% ls()){  # if user will download the report
    
    cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
    counter <- 1
    # for(iTrait in staTraits){
    for(iParam in unique(parameters)){
      cat("\n\n####", iParam, "{.tabset .tabset-pills}       \n\n")
      
      mydata2 = mydata[which(mydata$parameter %in% iParam),]
      p <- ggplot(data=mydata2, aes(x=environment, y=value, fill=trait)) +
        ggplot2::geom_bar(stat="identity", position=position_dodge()) +
        ggplot2::theme(axis.text.x = ggplot2::element_blank(), legend.position = "bottom") +
        # ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), axis.text.x = ggplot2::element_text(angle = 45,hjust=1, vjust = 1), legend.position = "bottom") +
        ggplot2::labs(x="Environment", y="Parameter value")
      print(htmltools::tagList(plotly::ggplotly(p)))
      # print(p)
      cat("\n")
    }
    # }
    cat("\n")
  }else{ # display in website
    
    plotly::renderPlotly({
      # shiny::renderPlot({
      mydata2 = mydata[which(mydata$parameter %in% input$parameterMetrics),]
      p <- ggplot2::ggplot(data=mydata2, aes(x=environment, y=value, fill=trait)) +
        ggplot2::geom_bar(stat="identity", position=position_dodge()) +
        ggplot2::theme(axis.text.x = ggplot2::element_blank(), legend.position = "bottom") + 
        # ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"),  axis.text.x = ggplot2::element_text(angle = 45,hjust=1, vjust = 1), legend.position = "bottom") +
        ggplot2::labs(x="Environment", y="Parameter value")
      # p
      plotly::ggplotly(p)
    })
    
  }
  
}
```

<p>&nbsp;</p>

`r if (length(which( result$metadata$pheno$parameter %in% c("row","col") ))==2 ) {
coordNames <- result$metadata$pheno[result$metadata$pheno$parameter %in% c("row","col"),"value" ]
if( any( apply(result$data$pheno[,coordNames],2, sommer::propMissing) == 1) ){shinyjs::hide()}else{"### Field view\n
The following graphs allow the user to inspect the spatial trends in the different fields.\n"}

}`

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  modeling <- result$modeling
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"];
  idSta <- idSta[length(idSta)]
  traitFieldView <- unique(result$predictions[which( result$predictions$analysisId == idSta),"trait"])
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    
    if(length(which( result$metadata$pheno$parameter %in% c("row","col") ))==2 ){ # user mapped the columns
      
      coordNames <- result$metadata$pheno[result$metadata$pheno$parameter %in% c("row","col"),"value" ]
      
      if( any( apply(result$data$pheno[,coordNames],2, sommer::propMissing) == 1) ){ # we can't proceed, there's no information in one of the coordinates
        shinyjs::hide()
      }else{
        shiny::selectInput(ns("traitStaFieldView"), 
                           label = "Trait:", 
                           choices = traitFieldView, 
        )
      }
      
    }
    
  }
  
}
```

```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if(length(which( result$metadata$pheno$parameter %in% c("row","col") ))==2 ){
    
    
    if( any( apply(result$data$pheno[,coordNames],2, sommer::propMissing) == 1) ){ # we can't proceed, there's no information in one of the coordinates
      shinyjs::hide()
    }else{
      pheno <- result$data$pheno
      metaPheno <- result$metadata$pheno; metaPheno <- metaPheno[metaPheno$parameter != "trait", ]
      colnames(pheno) <- cgiarBase::replaceValues( colnames(pheno), Search = metaPheno$value, Replace = metaPheno$parameter)
      
      if("params" %in% ls()){ # we will download the document
        
        cat(paste0("###   {.tabset .tabset-pills}    \n\n  "))
        
        for(iTrait in traitFieldView){
          cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
          
          pheno$value <- pheno[, iTrait]
          maxVal <- max(pheno$value, na.rm = TRUE) # get the maximum value found in the matrix of connectivity
          minVal <- min(pheno$value, na.rm = TRUE) # get the maximum value found in the matrix of connectivity
          meanVal <- mean(pheno$value, na.rm = TRUE) # get the maximum value found in the matrix of connectivity
          
          p <- ggplot2::ggplot(data = pheno, ggplot2::aes(row, col, fill = value))+
            ggplot2::geom_tile(color = "white")+
            ggplot2::scale_fill_gradient2(low = "#E46726", high = "#038542", mid = "gold",
                                          midpoint = meanVal, limit = c(minVal,maxVal), space = "Lab",
                                          name=iTrait) +
            ggplot2::theme_minimal()+
            ggplot2::ylab("") + ggplot2::xlab("") +
            ggplot2::facet_wrap(~environment, scales = "free") +
            ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank() )
          # print(p) # static
          print(htmltools::tagList(plotly::ggplotly(p))) # dynamic
          
        }
        
      }else{ # we will display in the GUI
        
        # shiny::renderPlot({ # static
        plotly::renderPlotly({ # dynamic
          
          pheno$value <- pheno[, input$traitStaFieldView]
          maxVal <- max(pheno$value, na.rm = TRUE) # get the maximum value found in the matrix of connectivity
          minVal <- min(pheno$value, na.rm = TRUE) # get the maximum value found in the matrix of connectivity
          meanVal <- mean(pheno$value, na.rm = TRUE) # get the maximum value found in the matrix of connectivity
          
          p <- ggplot2::ggplot(data = pheno, ggplot2::aes(row, col, fill = value))+
            ggplot2::geom_tile(color = "white")+
            ggplot2::scale_fill_gradient2(low = "#E46726", high = "#038542", mid = "gold",
                                          midpoint = meanVal, limit = c(minVal,maxVal), space = "Lab",
                                          name=input$traitStaFieldView) +
            ggplot2::theme_minimal()+
            ggplot2::ylab("") + ggplot2::xlab("") +
            ggplot2::facet_wrap(~environment, scales = "free") +
            ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank() )
          plotly::ggplotly(p)
          
        })
        
      }
    } # end of if user has coordinate data
    
  } # end of if user mapped the coordinate columns
  
}

```

### Predictions

The following table allows you to check the trait predictions in wide format and filter them by environment.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  modeling <- result$modeling
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"];
  idSta <- idSta[length(idSta)]
  envs <- unique(result$predictions[which( result$predictions$analysisId == idSta),"environment"])
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("envSta"), 
                       label = "Environment:", 
                       choices = envs, 
    )
  }
  
}
```


```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  traits <- unique(pred$trait)
  predictions.designation <- subset(pred, select = c(trait,designation,environment,predictedValue))
  predictions.designation <- reshape(
    data=predictions.designation, 
    timevar = c("trait"), 
    idvar = c("environment","designation"), 
    direction="wide"
  )
  names(predictions.designation) <- c("designation","environment",traits)
  
  if("params" %in% ls()){ # we will download the document
    printTable(predictions.designation)
  }else{
    DT::renderDT({DT::formatRound(DT::datatable(
      predictions.designation[predictions.designation$environment==input$envSta,],
      rownames = FALSE,
      class = 'cell-border stripe',
      options = list(
        scrollY = "400px",
        scrollX = "400px",
        paging = F)
    ),traits)
    })
  }
  
}

```

<p>&nbsp;</p>

The following boxplot allows you to see the distribution of predicted values by trait (y-axis) in the different environments.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  modeling <- result$modeling
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"];
  idSta <- idSta[length(idSta)]
  modeling <- modeling[which(modeling$analysisId == idSta),"trait"]
  modeling <- setdiff(modeling, "inputObject")
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("traitStaBox"), 
                       label = "Trait:", 
                       multiple = FALSE,
                       selected = unique(modeling),
                       choices = unique(modeling), 
    )
  }
  
}

```

```{r, results='asis'}
predTrait <- result$predictions

if("params" %in% ls()){ # we will download the document
  traits <- unique(modeling)
  cat("\n\n###  {.tabset .tabset-pills}       \n\n")
  
  for (j in 1:length(traits)){ # 
    cat("\n\n####", traits[j], "{.tabset .tabset-pills}       \n\n")
    
    p <- ggplot2::ggplot(predTrait[which(predTrait$trait %in% traits[j]),], ggplot2::aes(x=environment, y=predictedValue, label=designation) ) + 
      # ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0.2), ggplot2::aes(colour=environment), alpha=0.9) +
      ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=factor(environment))) +
      # ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
      #                axis.text.x = ggplot2::element_text(angle = 45, hjust=1, vjust = 1), legend.position = "none") +
      ggplot2::theme(axis.text.x = ggplot2::element_blank(), legend.position = "bottom") + 
      ggplot2::labs(x="", y="Trait value") #+ ggplot2::facet_grid(trait~., scales = "free") 
    # print(p) # static
    print(htmltools::tagList(ggplotly(  p  ))) # dynamic
    
    cat("\n")
  }
  
  cat("\n")
  
}else{
  
  # shiny::renderPlot({
  plotly::renderPlotly({
    if(length(input$traitStaBox) > 0){
      dd <- predTrait[which(predTrait$trait %in% input$traitStaBox),]
      p <- ggplot2::ggplot(dd, ggplot2::aes(x=environment, y=predictedValue, label=designation) ) +
        # ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0.2), ggplot2::aes(colour=environment), alpha=0.9) +
        ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=factor(environment))) +
        # ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), axis.text.x = ggplot2::element_text(angle = 45, hjust=1, vjust = 1), legend.position = "none") +
        ggplot2::theme(axis.text.x = ggplot2::element_blank(), legend.position = "bottom") + 
        ggplot2::labs(x="", y="Trait value")
      # p
      plotly::ggplotly(p)
    }
  })
}

```

<p>&nbsp;</p>

### Correlation between environments

The following plot aims to show the correlation between BLUEs or BLUPs (depending on the parameter settings) among the different environments.

<p>&nbsp;</p>

```{r }
if (is.null(result)){
  # invisible(NULL)
} else {
  modeling <- result$modeling
  idSta <- result$status[which(result$status$module == "sta"),"analysisId"];
  idSta <- idSta[length(idSta)]
  modeling <- modeling[which(modeling$analysisId == idSta),"trait"]
  modeling <- setdiff(modeling, "inputObject")
  
  if("params" %in% ls()){ # we will download the document
    shinyjs::hide()
  }else{
    shiny::selectInput(ns("traitStaCor"), 
                       label = "Trait:", 
                       multiple = FALSE,
                       selected = unique(modeling),
                       choices = unique(modeling), 
    )
  }
  
}

```


```{r, results='asis'}
if (is.null(result)){
  # invisible(NULL)
} else {
  
  predictions.gcorrE <- subset(pred, select = c(trait,designation,environment,predictedValue))
  
  # addTextInPlots <- ifelse(length(envs) > 30, FALSE, TRUE)
  if("params" %in% ls()){ # we will download the document
    
    cat("\n\n###  {.tabset .tabset-pills}       \n\n")
    
    for(iTrait in traits){
      cat("\n\n####", iTrait, "{.tabset .tabset-pills}       \n\n")
      
      predictions.gcorrE2 <- predictions.gcorrE[predictions.gcorrE$trait == iTrait, ]
      wide <- stats::reshape(predictions.gcorrE2,
                             direction = "wide", idvar = "designation",
                             timevar = "environment", v.names = "predictedValue", sep= "")
      colnames(wide) <- gsub("predictedValue","",colnames(wide))
      wide2 <- as.data.frame(wide[,-c(1:2)]); colnames(wide2) <- colnames(wide)[-c(1:2)]
      corr <- round(stats::cor(wide2, use="pairwise.complete.obs"),2)
      vv <- cgiarBase::matToTab(corr)
      p <- ggplot2::ggplot(data = vv, ggplot2::aes(Var2, Var1, fill = Freq))+
        ggplot2::geom_tile(color = "white")+
        ggplot2::scale_fill_gradient2(low = "#E46726", high = "#038542", mid = "white", 
                                      midpoint = 0, limit = c(-1,1), space = "Lab", 
                                      name="Pearson\nCorrelation") +
        ggplot2::theme_minimal()+ 
        # ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, size = 8, hjust = 1))+
        # ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0, vjust = 1, size = 8, hjust = 1))+
        ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank()) + 
        ggplot2::ylab("") + ggplot2::xlab("") +
        ggplot2::coord_fixed()
      # print(p)
      print(htmltools::tagList(plotly::ggplotly(p)))
      cat("\n")
    }
    cat("\n")
    cat("\n")
  }else{
    
    # shiny::renderPlot({
    plotly::renderPlotly({
      predictions.gcorrE2 <- predictions.gcorrE[predictions.gcorrE$trait == input$traitStaCor, ]
      
      wide <- stats::reshape(predictions.gcorrE2,
                             direction = "wide", idvar = "designation",
                             timevar = "environment", v.names = "predictedValue", sep= "")
      colnames(wide) <- gsub("predictedValue","",colnames(wide))
      wide2 <- as.data.frame(wide[,-c(1:2)]); colnames(wide2) <- colnames(wide)[-c(1:2)]
      corr <- round(stats::cor(wide2, use="pairwise.complete.obs"),2)
      vv <- cgiarBase::matToTab(corr)
      p <- ggplot2::ggplot(data = vv, ggplot2::aes(Var2, Var1, fill = Freq))+
        ggplot2::geom_tile(color = "white")+
        ggplot2::scale_fill_gradient2(low = "#E46726", high = "#038542", mid = "white", 
                                      midpoint = 0, limit = c(-1,1), space = "Lab", 
                                      name="Pearson\nCorrelation") +
        ggplot2::theme_minimal()+ 
        # ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, size = 8, hjust = 1))+
        # ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0, vjust = 1, size = 8, hjust = 1))+
        ggplot2::theme(axis.text.x = ggplot2::element_blank(), axis.text.y = ggplot2::element_blank()) + 
        ggplot2::ylab("") + ggplot2::xlab("") +
        ggplot2::coord_fixed()
      # p
      plotly::ggplotly(p)
    })
    
    
  }
  
}
```


### References of methods used

Velazco, J. G., Rodriguez-Alvarez, M. X., Boer, M. P., Jordan, D. R., Eilers, P. H., Malosetti, M., & Van Eeuwijk, F. A. (2017). Modelling spatial trends in sorghum breeding field trials using a two-dimensional P-spline mixed model. Theoretical and Applied Genetics, 130, 1375-1392.

Rodriguez-Alvarez, M. X., Boer, M. P., van Eeuwijk, F. A., & Eilers, P. H. (2018). Correcting for spatial heterogeneity in plant breeding experiments with P-splines. Spatial Statistics, 23, 52-71.

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

Boer M, van Rossum B (2022). LMMsolver: Linear Mixed Model Solver. R package version 1.0.4.9000.

<p>&nbsp;</p>

