---
title: "Advancement Meeting Report"
output: html_document
params:
  toDownload: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = TRUE
)

library(DT)
library(magrittr)
library(shiny)
library(ggplot2)

load("resultAdvMeeting.RData")
```

```{r, include=FALSE}
printTable <- function(DT, pageLength = 25,
                       numericColNames = NULL, numRound = 3,
                       scrollXOpt = TRUE, filt = "top", ...) {

  oldnames <- colnames(DT)

  DT <- data.frame(lapply(DT, function(x) {
    if (is.numeric(x)) round(x, numRound) else x
  }))

  colnames(DT) <- oldnames

  table <- DT::datatable(
    data = DT,
    filter = filt,
    options = list(
      dom = "l<<t>pB>",
      buttons = c("copy", "csv", "excel", "print"),
      pageLength = pageLength,
      searchHighlight = TRUE,
      lengthMenu = c(10, 25, 50, 100),
      scrollX = scrollXOpt
    ),
    extensions = "Buttons",
    rownames = FALSE,
    ...
  )

  if (length(numericColNames) > 0) {
    table <- table %>%
      DT::formatRound(columns = numericColNames, digits = numRound)
  }

  table
}
```

### About this analysis

**Advancement Meeting Dashboard** enables multi-stakeholder consensus on product advancement decisions. Each breeder independently completes the Pre-advancement workflow and saves their output. A meeting facilitator merges all stakeholder files and uses this module to compare decisions, discuss disagreements, and record consensus outcomes.

**Inputs used:**

- **Multiple Final Selection Stamps:** One per stakeholder, each containing per-designation selection decisions (SELECTED, NOT SELECTED, REVISE, CHECK).
- **Shared MTA predictions (BLUPs):** The underlying multi-trial analysis providing trait performance data for all candidates.

**Outputs produced:**

- A **Meeting Stamp** recording the consensus decision for every designation, combining auto-assigned unanimous decisions with manually resolved controversial candidates.

---

```{r summary-section, results='asis'}
# Expected variables from .RData:
# comparison_matrix, meeting_decisions_df, mta_stamp, participants, meeting_name,
# trait_directions, predictions_data

if (!exists("meeting_name") || is.null(meeting_name)) meeting_name <- "Unnamed meeting"
if (!exists("participants") || is.null(participants)) participants <- character(0)
if (!exists("comparison_matrix") || is.null(comparison_matrix)) comparison_matrix <- data.frame()
if (!exists("meeting_decisions_df") || is.null(meeting_decisions_df)) meeting_decisions_df <- data.frame()

n_participants <- length(participants)
participant_names <- paste(participants, collapse = ", ")
n_total <- nrow(meeting_decisions_df)

n_selected <- if (n_total > 0 && "decision" %in% colnames(meeting_decisions_df)) {
  sum(meeting_decisions_df$decision == "SELECTED", na.rm = TRUE)
} else { 0 }

n_not_selected <- if (n_total > 0 && "decision" %in% colnames(meeting_decisions_df)) {
  sum(meeting_decisions_df$decision == "NOT SELECTED", na.rm = TRUE)
} else { 0 }

# Discussed and resolved = controversial candidates that received a joint decision
# These are candidates whose controversy_score is strictly between 0 and 1
# (or all REVISE) that were manually resolved
if (n_total > 0 && "controversy_score" %in% colnames(meeting_decisions_df)) {
  n_discussed <- sum(
    meeting_decisions_df$controversy_score > 0 &
    meeting_decisions_df$controversy_score < 1,
    na.rm = TRUE
  )
} else {
  n_discussed <- 0
}

cat(paste0(
  "### Summary\n\n",
  "**Meeting name:** ", meeting_name, "\n\n",
  "- **Number of participants:** ", n_participants, "\n",
  "- **Participants:** ", participant_names, "\n",
  "- **Total candidates evaluated:** ", n_total, "\n",
  "- **Consensus SELECTED:** ", n_selected, "\n",
  "- **Consensus NOT SELECTED:** ", n_not_selected, "\n",
  "- **Discussed and resolved:** ", n_discussed, "\n\n"
))
```

---

### Final Meeting Decisions Table

The table below shows all candidates with their meeting consensus decision, controversy score, and each stakeholder's individual decision. Cells are color-coded by decision status: blue = SELECTED, orange = NOT SELECTED, yellow = REVISE, pink = CHECK.

```{r, include=FALSE}
htmltools::tagList(printTable(mtcars))
```

```{r decisions-table}
# STATUS_COLORS: SELECTED = "#0072B2", NOT SELECTED = "#D55E00",
#                REVISE = "#F9A825", CHECK = "#C2185B"
STATUS_COLORS <- c(
  "SELECTED"     = "#0072B2",
  "NOT SELECTED" = "#D55E00",
  "REVISE"       = "#F9A825",
  "CHECK"        = "#C2185B"
)

# Light background versions for table cells
STATUS_BG_COLORS <- c(
  "SELECTED"     = "#D6EAF8",
  "NOT SELECTED" = "#F5C4A5",
  "REVISE"       = "#FFF3CD",
  "CHECK"        = "#F5DCE9"
)

if (nrow(comparison_matrix) > 0 && nrow(meeting_decisions_df) > 0) {

  # Build display table: designation, meeting_decision, controversy_score, stakeholder columns
  stakeholder_cols <- setdiff(
    colnames(comparison_matrix),
    c("designation", "controversy_score")
  )

  display_df <- data.frame(
    designation = comparison_matrix$designation,
    stringsAsFactors = FALSE
  )

  # Add meeting decision
  display_df$meeting_decision <- meeting_decisions_df$decision[
    match(display_df$designation, meeting_decisions_df$designation)
  ]

  # Add controversy score
  if ("controversy_score" %in% colnames(comparison_matrix)) {
    display_df$controversy_score <- comparison_matrix$controversy_score
  }

  # Add stakeholder columns
  for (sc in stakeholder_cols) {
    display_df[[sc]] <- comparison_matrix[[sc]]
  }

  # Sort by controversy (most controversial first: closest to 0.5)
  if ("controversy_score" %in% colnames(display_df)) {
    display_df <- display_df[
      order(abs(display_df$controversy_score - 0.5)), , drop = FALSE
    ]
  }

  # Render DT with color coding
  dt <- printTable(
    display_df,
    pageLength = 25,
    scrollXOpt = TRUE,
    numericColNames = "controversy_score"
  )

  # Apply color coding to decision columns
  decision_cols <- c("meeting_decision", stakeholder_cols)
  decision_cols <- intersect(decision_cols, colnames(display_df))

  for (col in decision_cols) {
    dt <- DT::formatStyle(
      dt,
      columns = col,
      backgroundColor = DT::styleEqual(
        c("SELECTED", "NOT SELECTED", "CHECK", "REVISE"),
        c(STATUS_BG_COLORS[["SELECTED"]], STATUS_BG_COLORS[["NOT SELECTED"]],
          STATUS_BG_COLORS[["CHECK"]], STATUS_BG_COLORS[["REVISE"]])
      ),
      fontWeight = "bold"
    )
  }

  htmltools::tagList(dt)
} else {
  cat("No meeting decisions data available.\n")
}
```

---

### Agreement Heatmap

This heatmap shows the comparison matrix of stakeholder decisions. Rows represent designations sorted by controversy level (most disagreement at top), columns represent stakeholders. Cell colors indicate each stakeholder's decision.

```{r agreement-heatmap, fig.height=10, fig.width=10}
if (nrow(comparison_matrix) > 0) {

  stakeholder_cols_hm <- setdiff(
    colnames(comparison_matrix),
    c("designation", "controversy_score")
  )

  if (length(stakeholder_cols_hm) > 0) {

    # Sort by controversy (most controversial at top = closest to 0.5)
    hm_df <- comparison_matrix
    if ("controversy_score" %in% colnames(hm_df)) {
      hm_df <- hm_df[order(abs(hm_df$controversy_score - 0.5)), , drop = FALSE]
    }

    # Convert to long format for ggplot
    hm_long <- do.call(rbind, lapply(stakeholder_cols_hm, function(st) {
      data.frame(
        designation = hm_df$designation,
        stakeholder = st,
        decision = hm_df[[st]],
        stringsAsFactors = FALSE
      )
    }))

    # Create ordered factor for designation (preserve controversy sort order)
    hm_long$designation <- factor(
      hm_long$designation,
      levels = rev(hm_df$designation)
    )

    # Ensure decision is a factor with known levels
    hm_long$decision <- factor(
      hm_long$decision,
      levels = c("SELECTED", "NOT SELECTED", "REVISE", "CHECK")
    )

    p_hm <- ggplot(hm_long, aes(x = stakeholder, y = designation, fill = decision)) +
      geom_tile(color = "white", linewidth = 0.5) +
      scale_fill_manual(
        values = STATUS_COLORS,
        name = "Decision",
        drop = FALSE
      ) +
      theme_minimal(base_size = 12) +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = 11),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size = 11),
        legend.title = element_text(size = 12, face = "bold"),
        panel.grid = element_blank()
      ) +
      labs(
        title = "Agreement Heatmap (sorted by controversy)",
        subtitle = "Most controversial candidates at top"
      )

    print(p_hm)
  } else {
    cat("No stakeholder columns found in comparison matrix.\n")
  }
} else {
  cat("No comparison matrix data available for heatmap.\n")
}
```

---

### Population Statistics

For every metric - the consensus selection index and each trait used in the meeting - the table below contrasts the **selected** set against **all candidates** (checks excluded) and against the **checks** on their own. Rows are colour-coded to match the distribution plot: blue = selected, orange = all candidates, purple = checks.

```{r meeting-population-statistics, results='asis'}
have_stats <- exists("review_long") && !is.null(review_long) &&
  is.data.frame(review_long) && nrow(review_long) > 0

if (!have_stats) {
  cat("_No trait distribution data available for this meeting._\n\n")
}
```

```{r meeting-population-statistics-tables, results='asis'}
# Rendered as plain styled HTML rather than DT: these are small fixed 3-row
# summaries, and DT's JavaScript width calculation detaches the header from the
# body when several tables are emitted into a static document.
if (have_stats) {
  group_colors <- if (exists("REVIEW_GROUP_COLORS")) {
    REVIEW_GROUP_COLORS
  } else {
    c("Selected" = "#85C1E9", "All candidates" = "#E8A87C", "Checks" = "#C39BD3")
  }

  # Wash the group colours toward white so the text stays readable
  fade <- function(hex, opacity = 0.35) {
    r <- strtoi(substr(hex, 2, 3), 16L)
    g <- strtoi(substr(hex, 4, 5), 16L)
    b <- strtoi(substr(hex, 6, 7), 16L)
    sprintf("#%02X%02X%02X",
            as.integer(r * opacity + 255 * (1 - opacity)),
            as.integer(g * opacity + 255 * (1 - opacity)),
            as.integer(b * opacity + 255 * (1 - opacity)))
  }
  faded <- vapply(group_colors, fade, character(1))
  group_levels <- names(group_colors)
  stat_cols <- c("Min", "Max", "Mean", "Median", "SD")

  fmt <- function(x) if (is.na(x)) "&ndash;" else formatC(round(x, 3), format = "f", digits = 3)

  stats_for <- function(sub) {
    do.call(rbind, lapply(group_levels, function(g) {
      v <- sub$value[sub$group == g]
      v <- v[is.finite(v)]
      if (length(v) == 0) {
        data.frame(Group = g, Min = NA_real_, Max = NA_real_, Mean = NA_real_,
                   Median = NA_real_, SD = NA_real_, stringsAsFactors = FALSE)
      } else {
        data.frame(Group = g, Min = min(v), Max = max(v), Mean = mean(v),
                   Median = stats::median(v), SD = stats::sd(v),
                   stringsAsFactors = FALSE)
      }
    }))
  }

  for (m in unique(review_long$metric)) {
    sub <- review_long[review_long$metric == m, , drop = FALSE]
    label <- if (identical(m, "index_value")) "Index Value" else m
    sdf <- stats_for(sub)

    header <- paste0(
      "<tr>",
      "<th style='text-align:left; padding:7px 10px; border-bottom:2px solid #2C3E50;'>Group</th>",
      paste0("<th style='text-align:right; padding:7px 10px; border-bottom:2px solid #2C3E50;'>",
             stat_cols, "</th>", collapse = ""),
      "</tr>"
    )

    body <- paste0(vapply(seq_len(nrow(sdf)), function(i) {
      bg <- faded[[sdf$Group[i]]]
      paste0(
        "<tr style='background-color:", bg, ";'>",
        "<td style='text-align:left; padding:6px 10px; border-bottom:1px solid #fff;'>",
        sdf$Group[i], "</td>",
        paste0("<td style='text-align:right; padding:6px 10px; border-bottom:1px solid #fff;'>",
               vapply(stat_cols, function(cl) fmt(sdf[[cl]][i]), character(1)),
               "</td>", collapse = ""),
        "</tr>"
      )
    }, character(1)), collapse = "")

    cat(paste0(
      "<div style='margin: 0 0 22px 0;'>",
      "<div style='font-weight:700; font-size:14px; color:#2C3E50;",
      " padding:0 0 5px 2px; border-bottom:2px solid #2C3E50; margin-bottom:0;'>",
      label, "</div>",
      "<table style='width:100%; border-collapse:collapse; font-size:13px;",
      " table-layout:fixed; margin-top:0;'>",
      "<thead>", header, "</thead><tbody>", body, "</tbody></table>",
      "</div>\n"
    ))
  }
}
```

The plot below shows, for each metric, how the consensus-selected set is distributed against the full candidate set. The dashed vertical line marks the mean of the checks.

```{r meeting-population-statistics-height, include=FALSE}
# Scale the figure with the number of facet rows (2 panels per row)
stats_fig_height <- 4
if (have_stats) {
  n_panels <- length(unique(review_long$metric))
  stats_fig_height <- max(3.5, ceiling(n_panels / 2) * 2.6)
}
```

```{r meeting-population-statistics-plot, fig.width=11, fig.height=stats_fig_height}
if (have_stats) {
  plot_df <- review_long[review_long$group %in% c("All candidates", "Selected"), , drop = FALSE]

  if (nrow(plot_df) > 0) {
    metric_order <- unique(review_long$metric)
    label_order <- vapply(metric_order, function(m) {
      if (identical(m, "index_value")) "Index Value" else m
    }, character(1))

    plot_df$label <- factor(plot_df$label, levels = unname(label_order))
    plot_df$group <- factor(plot_df$group, levels = c("All candidates", "Selected"))

    check_df <- review_long[review_long$group == "Checks", , drop = FALSE]
    check_means <- NULL
    if (nrow(check_df) > 0) {
      check_means <- stats::aggregate(value ~ label, data = check_df,
                                      FUN = mean, na.rm = TRUE)
      check_means$label <- factor(check_means$label, levels = unname(label_order))
    }

    group_colors <- if (exists("REVIEW_GROUP_COLORS")) {
      REVIEW_GROUP_COLORS
    } else {
      c("Selected" = "#85C1E9", "All candidates" = "#E8A87C", "Checks" = "#C39BD3")
    }

    p <- ggplot2::ggplot(plot_df, ggplot2::aes(x = value, fill = group)) +
      ggplot2::geom_histogram(position = "identity", alpha = 0.55,
                              bins = 30, colour = NA) +
      ggplot2::facet_wrap(~ label, scales = "free", ncol = 2) +
      ggplot2::scale_fill_manual(
        values = c(
          "All candidates" = unname(group_colors[["All candidates"]]),
          "Selected"       = unname(group_colors[["Selected"]])
        ),
        name = NULL
      ) +
      ggplot2::labs(x = "Predicted value", y = "Count") +
      ggplot2::theme_minimal(base_size = 12) +
      ggplot2::theme(legend.position = "top",
                     strip.text = ggplot2::element_text(face = "bold"))

    if (!is.null(check_means) && nrow(check_means) > 0) {
      p <- p + ggplot2::geom_vline(
        data = check_means,
        ggplot2::aes(xintercept = value),
        colour = unname(group_colors[["Checks"]]),
        linetype = "dashed", linewidth = 0.9
      )
    }

    # Mean lines for Selected and All candidates
    sel_means <- stats::aggregate(value ~ label, data = plot_df[plot_df$group == "Selected", , drop = FALSE], FUN = mean, na.rm = TRUE)
    all_means <- stats::aggregate(value ~ label, data = plot_df[plot_df$group == "All candidates", , drop = FALSE], FUN = mean, na.rm = TRUE)
    if (nrow(sel_means) > 0) {
      sel_means$label <- factor(sel_means$label, levels = levels(plot_df$label))
      p <- p + ggplot2::geom_vline(
        data = sel_means,
        ggplot2::aes(xintercept = value),
        colour = unname(group_colors[["Selected"]]),
        linetype = "dashed", linewidth = 0.7
      )
    }
    if (nrow(all_means) > 0) {
      all_means$label <- factor(all_means$label, levels = levels(plot_df$label))
      p <- p + ggplot2::geom_vline(
        data = all_means,
        ggplot2::aes(xintercept = value),
        colour = unname(group_colors[["All candidates"]]),
        linetype = "dashed", linewidth = 0.7
      )
    }

    print(p)
  }
}
```

---

### TPP Breakdown

```{r meeting-tpp-breakdown, results='asis'}
have_bd <- exists("breakdown_detail") && !is.null(breakdown_detail) &&
  !is.null(breakdown_detail$table) && is.data.frame(breakdown_detail$table) &&
  nrow(breakdown_detail$table) > 0

if (!have_bd) {
  cat("_No TPP information available for this meeting._\n\n")
} else {
  cat(paste0(
    "This table compares what the Target Product Profile asks for against what was ",
    "actually evaluated on the phenotypic data. The `_tpp` columns reproduce the TPP ",
    "verbatim; the `_data` columns show what was resolved against the data - relative ",
    "specifications are converted to the equivalent absolute threshold using the mean of ",
    "the mapped checks. `Sel_pct_meeting_criteria` is the share of the **consensus-selected** ",
    "individuals meeting the resolved threshold and `All_pct_meeting_criteria` the share ",
    "across all candidates, so comparing the two shows how much the meeting decisions ",
    "enriched each criterion. Traits with no phenotypic mapping read \"not evaluated\" and ",
    "traits the stakeholders dropped read \"trait not used\".\n\n"
  ))
}
```

```{r meeting-tpp-breakdown-table}
if (have_bd) {
  # Colour-code the two Target_proximity columns. Defined inline because the
  # knit environment has no access to the package's internal helpers.
  .prox_bg <- c(
    "Comfortably meeting"   = "#D5F5E3",
    "Moderately meeting"    = "#E8F8F5",
    "Barely meeting"        = "#FCF3CF",
    "Barely failing"        = "#FDEBD0",
    "Moderately failing"    = "#FADBD8",
    "Substantially failing" = "#F5B7B1"
  )
  .prox_fg <- c(
    "Comfortably meeting"   = "#186A3B",
    "Moderately meeting"    = "#148F77",
    "Barely meeting"        = "#7D6608",
    "Barely failing"        = "#935116",
    "Moderately failing"    = "#943126",
    "Substantially failing" = "#78281F"
  )

  .bd <- printTable(breakdown_detail$table, pageLength = 20,
                    scrollXOpt = TRUE, filt = "none")
  for (.cl in intersect(c("Target_proximity_sel", "Target_proximity_all"),
                        colnames(breakdown_detail$table))) {
    .bd <- DT::formatStyle(
      .bd,
      columns = .cl,
      backgroundColor = DT::styleEqual(names(.prox_bg), unname(.prox_bg)),
      color = DT::styleEqual(names(.prox_fg), unname(.prox_fg)),
      fontWeight = DT::styleEqual(names(.prox_bg), rep("600", length(.prox_bg)))
    )
  }

  htmltools::tagList(.bd)
}
```

```{r meeting-tpp-breakdown-notes, results='asis'}
if (have_bd && length(breakdown_detail$env_notes) > 0) {
  cat("**Traits evaluated on a subset of environments**\n\n")
  for (n in breakdown_detail$env_notes) cat(paste0("- ", n, "\n"))
  cat("\n")
}
```

---

### Global Options Summary

```{r meeting-global-options, results='asis'}
have_opts <- exists("stakeholder_tables") && !is.null(stakeholder_tables) &&
  length(stakeholder_tables) > 0 &&
  !(length(stakeholder_tables) == 1 && "No Data" %in% names(stakeholder_tables))

if (!have_opts) {
  cat("_No stakeholder configuration data available for this meeting._\n\n")
} else {
  cat(paste0(
    "This section records the **actual selection criteria each stakeholder chose** for ",
    "their own initial selection - not what the TPP asks for. For each trait it shows the ",
    "index weight it carried, the threshold value and direction from that stakeholder's ",
    "trait rule card, and the reference check(s) used where the rule was defined relative ",
    "to a check. Comparing the tables shows where the stakeholders' criteria diverged, ",
    "which is often the source of the disagreements resolved in the meeting.\n\n"
  ))
}
```

```{r meeting-global-options-tables}
if (have_opts) {
  opt_tables <- lapply(names(stakeholder_tables), function(sname) {
    tbl_df <- stakeholder_tables[[sname]]
    if (is.null(tbl_df) || nrow(tbl_df) == 0) return(NULL)
    DT::datatable(
      tbl_df,
      rownames = FALSE,
      options = list(dom = "t", paging = FALSE, searching = FALSE,
                     ordering = FALSE, scrollX = TRUE),
      caption = htmltools::tags$caption(
        style = "caption-side: top; text-align: left; font-weight: 700; color: darkcyan;",
        sname
      )
    )
  })
  opt_tables <- opt_tables[!vapply(opt_tables, is.null, logical(1))]
  htmltools::tagList(opt_tables)
}
```
