---
title: "Population structure report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
toDownload: FALSE
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, out.width = "100%")
```

```{r printfxn, include=FALSE}
# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, ...) {
  
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  
  table <- DT::datatable(data = DT, 
                         filter = "top", 
                         options = list(autoWidth = TRUE, 
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'excel'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35)),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r plot_init, include = FALSE}
# Init Step to make sure that the dependencies are loaded
# was done to enable proper printing of the tab header and the graphs within the for loop below using print(htmltools::tagList(ggplotly(...)))
htmltools::tagList(printTable(mtcars))
#htmltools::tagList(plotly::ggplotly(ggplot2::ggplot()))

# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}

```

### Summary diversity

The following table allows you to inspect the principal statistics for the diversity analysis.

<p>&nbsp;</p>

```{r, results='asis'}
suppressWarnings(tryCatch({
  if(file.exists(normalizePath("R/outputs/resultPopStr.RData"))){
    load(normalizePath("R/outputs/resultPopStr.RData"))   
  }else{
    load("resultPopStr.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

    seedatosum<-as.data.frame(result[["PopStr"]][["SummaryDiversityAnalysis"]])
    colnames(seedatosum)=c("Parameter","Value")
    seedatosum[,2]=as.numeric(seedatosum[,2])
    #if("params" %in% ls()){ 
      printTable(seedatosum)
    #}else{
    #  DT::renderDT(DT::datatable(seedatosum, extensions = 'Buttons',
    #                   options = list(dom = 'Blfrtip',scrollX = TRUE,buttons = c('copy', 'csv','excel', 'pdf', 'print'),
    #                   lengthMenu = list(c(5,20,50,-1), c(5,20,50,'All'))))
    #              )
    #}

```

<p>&nbsp;</p>

### Percent of missing values
  
  This plot allows the user to visualize the distribution of percent of missing values for genotypes and for markers.

<p>&nbsp;</p>
```{r, results='asis'}
suppressWarnings(tryCatch({
  if(file.exists(normalizePath("R/outputs/resultPopStr.RData"))){
    load(normalizePath("R/outputs/resultPopStr.RData"))
  }else{
    load("resultPopStr.RData")	
  }
}, error = function(e) {
  shinyjs::hide()
}))

      NAmark=result[["PopStr"]][["CalculusPerMarker"]][["%NA"]]
      NAgen=result[["PopStr"]][["CalculusPerGenotype"]][["%NA"]]

      fig <- plotly::plot_ly(y = ~NAmark, type = "box", name="missing values in markers")
      fig <- fig %>% plotly::add_trace(y = ~NAgen, name="missing values in genotypes")
      fig <- fig %>% plotly::layout(yaxis = list(title = 'Percent'))
      fig 

```

<p>&nbsp;</p>

### MDS of population structure
  
  This plot allows the user to visualize the closeness or distance between genotypes to check if the population structure was as expected.

<p>&nbsp;</p>
```{r, results='asis'}
suppressWarnings(tryCatch({
  if(file.exists(normalizePath("R/outputs/resultPopStr.RData"))){
    load(normalizePath("R/outputs/resultPopStr.RData"))
	load(normalizePath("R/outputs/parmDendMDS.RData"))
  }else{
    load("resultPopStr.RData")
	load("parmDendMDS.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

      uno=as.data.frame(result[["PopStr"]][["Plots"]][[2]])
	if(length(names(uno)>5)){catv<-names(uno)[6]}else{catv <-"GroupClust"}
	eti="Gen"

	txlab <- paste0('Factor 1 (',result[["PopStr"]][["Plots"]][[1]][1],'%)')
	tylab <- paste0('Factor 2 (',result[["PopStr"]][["Plots"]][[1]][2],'%)')
	tzlab <- paste0('Factor 3 (',result[["PopStr"]][["Plots"]][[1]][3],'%)')

	txlab2=txlab
	if(plist[[1]][[1]]=="Factor2") txlab2=tylab
	if(plist[[1]][[1]]=="Factor3") txlab2=tzlab
	tylab2=txlab
	if(plist[[1]][[2]]=="Factor2") tylab2=tylab
	if(plist[[1]][[2]]=="Factor3") tylab2=tzlab

	p<-plotly::plot_ly(data=uno,x=uno[,plist[[1]][[1]]],y=uno[,plist[[1]][[2]]],color=uno[,catv],
                   type="scatter",mode="markers",colors = plist[[1]][[3]],xaxis=F, yaxis=F,
                   text=uno[,eti],marker=list(size=plist[[1]][[4]]))
	#color de fondo del grafico
	p<-p %>% plotly::layout(plot_bgcolor=plist[[1]][[5]],autosize=F, height = "350px",width = "550px")
	#titulo y etiquetas ejes
	p<-p %>% plotly::layout(title=plist[[1]][[6]],titlefont=list(size=plist[[1]][[7]],color=plist[[1]][[8]]), xaxis = list(title = txlab2, titlefont=list(size=plist[[1]][[9]],color=plist[[1]][[10]])),
                        yaxis = list(title = tylab2,titlefont=list(size=plist[[1]][[9]],color=plist[[1]][[10]])))
	p
```

<p>&nbsp;</p>

### Dendogram
  
  This plot allows the user to visualize in one way the groupings of genotypes.

<p>&nbsp;</p>
```{r, results='asis',fig.width=14, fig.height=14}
suppressWarnings(tryCatch({
  if(file.exists(normalizePath("R/outputs/resultPopStr.RData"))){
    load(normalizePath("R/outputs/resultPopStr.RData"))
	load(normalizePath("R/outputs/parmDendMDS.RData"))
  }else{
    load("resultPopStr.RData")
	load("parmDendMDS.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

      data=as.data.frame(result[["PopStr"]][["Plots"]][[2]])
      if(length(names(data)>5)){catv<-names(data)[6]}else{catv <-"GroupClust"}
	  txlab <- paste0('Factor 1 (',result[["PopStr"]][["Plots"]][[1]][1],'%)')
	  tylab <- paste0('Factor 2 (',result[["PopStr"]][["Plots"]][[1]][2],'%)')
	  tzlab <- paste0('Factor 3 (',result[["PopStr"]][["Plots"]][[1]][3],'%)')
	  
      info<- data[,c("Gen",catv)]
      info<- cbind(ID=info$Gen,info)
      names(info)=c("ID","Gen","Group")
      tree=result[["PopStr"]][["Plots"]][[3]]

	if (plist[[2]][[1]]=="rectangular"){
		plot(tree, type = "phylogram", cex = plist[[2]][[2]], label.offset = plist[[2]][[3]], show.tip.label = TRUE, edge.color = "black", edge.width =plist[[2]][[4]], edge.lty = 1,tip.color = plist[[2]][[5]][info$Group])
		legend(plist[[2]][[6]], legend=levels(info$Group), fill=plist[[2]][[5]],box.lty=0)
	}else{
		plot(tree, type = "fan", cex = plist[[2]][[2]], label.offset = plist[[2]][[3]], show.tip.label = TRUE, edge.color = "black", edge.width =plist[[2]][[4]], edge.lty = 1,tip.color = plist[[2]][[5]][info$Group])
		legend(plist[[2]][[6]], legend=levels(info$Group), fill=plist[[2]][[5]],box.lty=0)
	}

```

