---
title: "Number of Founders Report"
author: "Contact:<a href = 'https://github.com/Breeding-Analytics/bioflow' target = '_blank'>Breeding Analytics Team, OneCGIAR</a> breedinganalytics@cgiar.org"
date: "`r format(Sys.time(), '%B %d, %Y')`"  
output: html_document
params:
  toDownload: FALSE
---

```{r setup, include=FALSE}

# knitr R markdown chunk options
knitr::opts_chunk$set(dependson = knitr::all_labels(),
                      echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      out.width = "100%",
                      error = TRUE)
options(knitr.kable.NA = '')

# loading necessary R packages ####
## data manipulation
library(dplyr)    # %>%, data cleaning functions
library(magrittr) # coerce col to factors or numeric

## outputs - graphs, tables
library(ggplot2)  # ggplot(), etc.
library(plotly)  # ggplot(), etc.
library(DT)       # datatable()
library(knitr)    # kable
library(grafify)
library(data.table)
library(shiny)
```

```{r printfxn, include=FALSE}

# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                         numericColNames = NULL, numRound = 3, 
                         scrollXOpt = FALSE, colNames = NULL, autoWidthOpt = FALSE,...) {
    
    DT <- data.frame(lapply(X = DT, 
                            FUN = function(x) {
                              if(is.numeric(x)){
                                round(x, numRound)
                              } else {
                                x
                              }
                            }))
    
    table <- DT::datatable(data = DT, 
                           colnames = colNames,
                           filter = "top", 
                           options = list(autoWidth = autoWidthOpt,
                                          dom = 'l<<t>pB>', 
                                          buttons = c('copy', 'excel'),
                                          pageLength = pageLength,
                                          searchHighlight = TRUE,
                                          lengthMenu = c(7, 14, 21, 28, 35),
                                          scrollX = scrollXOpt),
                           extensions = 'Buttons',
                           rownames = FALSE,
                           ...)
    if (length(numericColNames) > 0){
      table <- table %>% DT::formatRound(columns = numericColNames,
                                     digits = numRound)
    }
    
    table
  }
```

```{r }

suppressWarnings(tryCatch({
  if(file.exists("./outputs/result.RData")){
    load("./outputs/resultQaPheno.RData")
  }else{
    load("resultQaPheno.RData")
  }
}, error = function(e) {
  shinyjs::hide()
}))

```


### Table of metrics
<p>&nbsp;</p>

The following table shows the summary of the allele coverage and the standard errors for the different number of founders sampled.

```{r }

if (is.null(result)){
  # invisible(NULL)
} else {
  metrics <- result$metrics
  tail(metrics)
  idNe <- result$status[which(result$status$module %in% c("neMarker")),"analysisId"];
  idNe <- idNe[length(idNe)]
  metrics <- metrics[which( metrics$analysisId == idNe),]
  nes <- metrics[which(metrics$parameter == "Ne"),"value"]
  metrics <- metrics[which(metrics$parameter == "allelesCovered"),]
  metrics$Ne <- nes
  metrics <- metrics[,c(setdiff(colnames(metrics), c("Ne","value","stdError","trait","environment") ), c("Ne","value","stdError"))]
  numericalColumns <- c("Ne","value","stdError")
  
  if("params" %in% ls()){ # we will download the document
    printTable(metrics, autoWidthOpt = TRUE, scrollXOpt = FALSE,
             colNames = colnames(metrics),
           numericColNames = numericalColumns,
           numRound = 4)
  }else{
    DT::renderDT(printTable(metrics,
                        autoWidthOpt = FALSE, scrollXOpt = TRUE,
                        colNames = colnames(metrics),
                        numericColNames = numericalColumns,
                        numRound = 4), server = FALSE)
  }
}

```

### Graph of allele coverage
<p>&nbsp;</p>

The following graph shows the allele coverage (y axis) at different number of founders sampled (x-axis). It should be understood that once we hit the golden line (0.95) and the green line (0.975) we have probably hit the number of founders of our population. Still, if you notice that the line follows a consistent steep you may want to declare the number of founders as the number where the line stabilizes.

```{r, results='asis'}

if (is.null(result)){
  # invisible(NULL)
} else {
  
  p <- ggplot2::ggplot(metrics,ggplot2::aes(x=Ne,y=value))+
    ggplot2::geom_ribbon(ggplot2::aes(x=Ne,ymin=value-stdError,
                                      ymax=value+stdError),
                         alpha=0.2,linetype=0)+
    ggplot2::geom_line(linewidth=1)+
    ggplot2::guides()+
    ggplot2::theme_bw()+
    ggplot2::scale_x_continuous("Number of Founders")+
    ggplot2::scale_y_continuous("Allele coverage")  +
    ggplot2::geom_hline(yintercept = 0.95, color="gold", linetype="dashed") +
    ggplot2::geom_hline(yintercept = 0.975, color="forestgreen", linetype="dashed")
  
  
  if("params" %in% ls()){ # we will download the document
    print(p)
  }else{
    shiny::renderPlot({
      p
    })
  }
  
  
}

```

<p>&nbsp;</p>

### References on methods used

Tukey, J. W. (1977). Exploratory Data Analysis. Section 2C.

Velleman, P. F. and Hoaglin, D. C. (1981). Applications, Basics and Computing of Exploratory Data Analysis. Duxbury Press.


<p>&nbsp;</p>



