---
title: "Optimal Cross Selection Report"
author: ""
date: "December 2023"
output: html_document
params:
 toDownload: FALSE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
```

```{r printfxn, include=FALSE}
# functions ####
# for printing tables (data.frames) - DT::datatable()
printTable <- function(DT, pageLength = 7, 
                       numericColNames = NULL, numRound = 3, ...) {
  
  DT <- data.frame(lapply(X = DT, 
                          FUN = function(x) {
                            if(is.numeric(x)){
                              round(x, numRound)
                            } else {
                              x
                            }
                          }))
  
  table <- DT::datatable(data = DT, 
                         filter = "top", 
                         options = list(autoWidth = TRUE, 
                                        dom = 'l<<t>pB>', 
                                        buttons = c('copy', 'excel'),
                                        pageLength = pageLength,
                                        searchHighlight = TRUE,
                                        lengthMenu = c(7, 14, 21, 28, 35)),
                         extensions = 'Buttons',
                         rownames = FALSE,
                         ...)
  if (length(numericColNames) > 0){
    table <- table %>% formatRound(columns = numericColNames,
                                   digits = numRound)
  }
  
  table
}
```

```{r plot_init, include = FALSE}
# Init Step to make sure that the dependencies are loaded
# was done to enable proper printing of the tab header and the graphs within the for loop below using print(htmltools::tagList(ggplotly(...)))
htmltools::tagList(plotly::ggplotly(ggplot2::ggplot()))
htmltools::tagList(printTable(mtcars))
# Get the current figure size in pixels:
get_w <- function() {
  with(knitr::opts_current$get(c("fig.width", "dpi", "fig.retina")),
       fig.width*dpi/fig.retina)
}

get_h <- function() {
  with(knitr::opts_current$get(c("fig.height", "dpi", "fig.retina")),
       fig.height*dpi/fig.retina)
}

```

### Cross performance table

The following table provides an overview of the performance of the predicted crosses for the different traits. There should be as many crosses as requested in the interface. If you requested different number of crosses or degrees, the treatment column allows you to visualize the results from different runs.

<p>&nbsp;</p>

```{r, results='asis'}
suppressWarnings(tryCatch({
  if(file.exists("./outputs/resultOcs.RData")){
    load("./outputs/resultOcs.RData")
  }else{
    load("resultOcs.RData")
  }
}, error = function(e) {
  NULL
}))

if (is.null(result)){
  invisible(NULL)
} else {
  mydata = result$predictions
  idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
  idOcs <- idOcs[length(idOcs)]
  ocsTrait <- unique(result$modeling[which(result$modeling$analysisId == idOcs & result$modeling$parameter == "ocsFormula"),"trait"]);
  mydata <- mydata[which(mydata$analysisId %in% c(idOcs)),]
  
  treatments <- unique(mydata$environment)
  wideList <- list()
  for(iTreatment in treatments){
    wideList[[iTreatment]] <- stats::reshape(mydata[which(mydata$environment == iTreatment),c(c("designation"),"trait",c("predictedValue"))], 
                                             direction = "wide", idvar = c("designation"),
                                             timevar = "trait", v.names = c("predictedValue"), sep= "_")
    wideList[[iTreatment]]$treatment <- iTreatment
    colnames(wideList[[iTreatment]]) <- gsub("predictedValue_","",colnames(wideList[[iTreatment]]))
    numeric.output <- colnames(wideList[[iTreatment]])[-c(1)]
  }
  wide <- do.call(rbind, wideList)
  rownames(wide) <- NULL
  
  cat("The trait used for OCS was: ",ocsTrait)
  
  if("params" %in% ls()){ # we will download the document
    printTable(wide)
  }else{
    DT::renderDT(DT::formatRound(DT::datatable(wide, extensions = 'Buttons',
                                               options = list(dom = 'Blfrtip',scrollX = TRUE,buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                                                              lengthMenu = list(c(5,20,50,-1), c(5,20,50,'All')))
    ), numeric.output)
    )
  }
  
}
```

<p>&nbsp;</p>

### Number of crosses per parent

The following barplot allows you identify the number of times (y-axis) that a parent (x-axis) is suggested to be used in the crossing block for a given treatment (number of crosses by degrees).

<p>&nbsp;</p>

```{r}
if (is.null(result)){
  invisible(NULL)
} else {
  predictions <- result$predictions
  idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
  idOcs <- idOcs[length(idOcs)]
  myEnvs <- predictions[which(predictions$analysisId == idOcs),"environment"]
  if("params" %in% ls()){  # if user will download the report
  }else{
    shiny::selectInput(ns("environ"), 
                       label = "Treatment:", 
                       multiple = TRUE,
                       selected = unique(myEnvs),
                       choices = unique(myEnvs)
    )
  }
}
```

```{r, results='asis'}
if (is.null(result)){
  invisible(NULL)
} else {
  
  if("params" %in% ls()){  # if user will download the report
    
    cat(paste0("### A  {.tabset .tabset-pills}    \n\n  "))
    counter <- 1
    for(iEnv in unique(myEnvs)){
      cat("\n\n####", iEnv, "{.tabset .tabset-pills}       \n\n")
      mydata = result$predictions
      idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
      idOcs <- idOcs[length(idOcs)]
      ocsTrait <- result$modeling[which(result$modeling$analysisId == idOcs & result$modeling$parameter == "ocsFormula"),"trait"];
      mydata <- mydata[which(mydata$analysisId %in% c(idOcs[length(idOcs)])),]
      mydata <- mydata[which(mydata$trait == ocsTrait),]
      mydata <- mydata[which(mydata$environment %in% iEnv),]
      
      treatments <- unique(mydata$environment)
      wideList <- list()
      for(iTreatment in treatments){
        overlayCounts <- with(mydata[which(mydata$environment == iTreatment),],sommer::overlay(mother,father))
        countsCol <- data.frame(times=apply(overlayCounts,2,sum))
        countsCol$parent <- as.factor(rownames(countsCol))
        countsCol$treatment <- iTreatment
        wideList[[iTreatment]] <- countsCol
      }
      countsCol <- do.call(rbind, wideList)
      
      p2 <- ggplot2::ggplot(data=countsCol, ggplot2::aes(x=parent, y=times)) +
        ggplot2::geom_bar(stat="identity", fill="steelblue")+
        ggplot2::geom_text(ggplot2::aes(label=times), vjust=6, size=3.5)+
        ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45)) + ggplot2::facet_grid(~treatment) + 
        ggplot2::labs(x="Parent", y="Times used in a cross") 
      print(htmltools::tagList(plotly::ggplotly(p2)))
      cat("\n")
    }
    cat("\n")
  }else{
    
    plotly::renderPlotly({
      
      mydata = result$predictions
      idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
      idOcs <- idOcs[length(idOcs)]
      ocsTrait <- result$modeling[which(result$modeling$analysisId == idOcs & result$modeling$parameter == "ocsFormula"),"trait"];
      mydata <- mydata[which(mydata$analysisId %in% c(idOcs[length(idOcs)])),]
      mydata <- mydata[which(mydata$trait == ocsTrait),]
      mydata <- mydata[which(mydata$environment %in% input$environ),]
      
      treatments <- unique(mydata$environment)
      wideList <- list()
      for(iTreatment in treatments){
        overlayCounts <- with(mydata[which(mydata$environment == iTreatment),],sommer::overlay(mother,father))
        countsCol <- data.frame(times=apply(overlayCounts,2,sum))
        countsCol$parent <- as.factor(rownames(countsCol))
        countsCol$treatment <- iTreatment
        wideList[[iTreatment]] <- countsCol
      }
      countsCol <- do.call(rbind, wideList)
      
      p2 <- ggplot2::ggplot(data=countsCol, ggplot2::aes(x=parent, y=times)) +
        ggplot2::geom_bar(stat="identity", fill="steelblue")+
        ggplot2::geom_text(ggplot2::aes(label=times), vjust=6, size=3.5)+
        ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45)) + ggplot2::facet_grid(~treatment) + 
        ggplot2::labs(x="Parent", y="Times used in a cross") 
      fig2 <-  plotly::ggplotly(p2)#, height = get_h()); 
      fig2
      
    })
    
  }
  
}
```

<p>&nbsp;</p>

### Parent performance table

This table shows the performance of the parents selected by the OCS algorithm. The parental means are extracted from the multi-trial analysis results.

<p>&nbsp;</p>

```{r}

if (is.null(result)){
  invisible(NULL)
} else {
  mydata = result$predictions
  idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
  idOcs <- idOcs[length(idOcs)]
  # get parents selected in crosses
  mydataOcs <- mydata[which(mydata$analysisId %in% c(idOcs)),]
  parentsSelected <- unique(c(mydataOcs$mother, mydataOcs$father))
  # what was the input of ocs? index or mta?
  idInputOcs <- result$modeling[which(result$modeling$analysisId == idOcs & result$modeling$trait == "inputObject"), "value"]
  moduleInputOcs <- result$status[which(result$status$analysisId == idInputOcs),"module"];
  if(moduleInputOcs != "mta"){ # we keep looking for mta since we had an index
    idInputOcs <- result$modeling[which(result$modeling$analysisId == idInputOcs & result$modeling$trait == "inputObject"), "value"]
    moduleInputOcs <- result$status[which(result$status$analysisId == idInputOcs),"module"];
  }else{} # we stop
  
  mydata <- mydata[which(mydata$analysisId %in% c(idInputOcs)),]
  wideParents <- stats::reshape(mydata[,c(c("designation"),"trait",c("predictedValue"))], direction = "wide", idvar = c("designation"),
                                timevar = "trait", v.names = c("predictedValue"), sep= "_")
  colnames(wideParents) <- gsub("predictedValue_","",colnames(wideParents))
  numeric.output <- colnames(wideParents)[-c(1)]
  wideParents <- wideParents[which(wideParents$designation %in% parentsSelected),]
  rownames(wideParents) <- NULL
  
  if("params" %in% ls()){  # if user will download the report
    printTable(wideParents)
  }else{ # if user will just examinate the report on the website
    DT::renderDT(DT::formatRound(DT::datatable(wideParents, extensions = 'Buttons',
                                               options = list(dom = 'Blfrtip',scrollX = TRUE,buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                                                              lengthMenu = list(c(5,20,50,-1), c(5,20,50,'All')))
    ), numeric.output)
    )
  }
  
}
```

<p>&nbsp;</p>

### Treatment comparison

The following table summarizes the trait means for the different OCS runs (number of crosses and degrees treatments). 

<p>&nbsp;</p>

```{r}

if (is.null(result)){
  invisible(NULL)
} else {
  
  idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
  idOcs <- idOcs[length(idOcs)]
  metrics <- result$metrics[which(result$metrics$analysisId == idOcs),]
  metrics <- metrics[which(metrics$parameter == "meanValue"),]
  metrics <- metrics[with(metrics, order(trait)), c("trait","environment","parameter","value","stdError")]
  colnames(metrics) <- cgiarBase::replaceValues(colnames(metrics), Search = "environment", Replace = "treatment")
  rownames(metrics) <- NULL
  numeric.output2 <- c("value","stdError")
  
  if("params" %in% ls()){  # if user will download the report
    printTable(metrics)
  }else{ # if user will just examinate the report on the website
    DT::renderDT(DT::formatRound(DT::datatable(metrics, extensions = 'Buttons',
                                               options = list(dom = 'Blfrtip',scrollX = TRUE,buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                                                              lengthMenu = list(c(5,20,50,-1), c(5,20,50,'All')))
    ), numeric.output2)
    )
  }
  
}
```

<p>&nbsp;</p>

### Cross performance per parent

The following boxplot allows you to see the performance of the different crosses predicted (dots in the boxplot) grouped by the different parents (x-axis).

<p>&nbsp;</p>
```{r}
if (is.null(result)){
  invisible(NULL)
} else {
  predictions <- result$predictions
  idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
  idOcs <- idOcs[length(idOcs)]
  myTraits <- predictions[which(predictions$analysisId == idOcs),"trait"]
  
  if("params" %in% ls()){  # if user will download the report
    
  }else{ # if user will just visualize the report in the website
    
    shiny::selectInput(ns("traitFilterPredictions2D2"), 
                       label = "Trait:", 
                       multiple = TRUE,
                       selected = unique(myTraits),
                       choices = unique(myTraits)
    )
    
    
  }
  
}
```

```{r, results='asis'}
library(dplyr)
if (is.null(result)){
  # invisible(NULL)
} else {
  
  if("params" %in% ls()){  # if user will download the report
    
    cat(paste0("## B  {.tabset .tabset-pills}    \n\n  "))
    for(iTrait in unique(myTraits)){
      cat("\n\n###", iTrait, "{.tabset .tabset-pills}       \n\n")
      mydata = result$predictions
      idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
      idOcs <- idOcs[length(idOcs)]
      mydata <- mydata[which(mydata$analysisId %in% c(idOcs)),]
      ## get the required traits
      mydata <- mydata[which(mydata$trait %in% c(iTrait)),]
      ##
      mydataF <- mydataM <- mydata
      mydataF <- mydataF[,-which(colnames(mydataF)=="mother")]
      colnames(mydataF)[which(colnames(mydataF)=="father")] <- "parent"
      mydataM <- mydataM[,-which(colnames(mydataM)=="father")]
      colnames(mydataM)[which(colnames(mydataM)=="mother")] <- "parent"
      mydata2 <- rbind(mydataM,mydataF[,colnames(mydataM)])
      
      p <- ggplot2::ggplot(mydata2, ggplot2::aes(x=parent, y=predictedValue, label=designation) ) + 
        ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0), ggplot2::aes(colour=parent), alpha=0.9) +
        ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=parent)) +
        ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
                       axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
        ggplot2::labs(x="", y="Trait value") #+ ggplot2::facet_grid(trait~environment, scales = "free")
      print(htmltools::tagList(plotly::ggplotly(p)))
      
    }
    
  }else{ # if user will just visualize the report in the website
    plotly::renderPlotly({
      mydata = result$predictions
      idOcs <- result$status[which(result$status$module == "ocs"),"analysisId"];
      idOcs <- idOcs[length(idOcs)]
      mydata <- mydata[which(mydata$analysisId %in% c(idOcs)),]
      ## get the required traits
      mydata <- mydata[which(mydata$trait %in% c(input$traitFilterPredictions2D2)),]
      ##
      mydataF <- mydataM <- mydata
      mydataF <- mydataF[,-which(colnames(mydataF)=="mother")]
      colnames(mydataF)[which(colnames(mydataF)=="father")] <- "parent"
      mydataM <- mydataM[,-which(colnames(mydataM)=="father")]
      colnames(mydataM)[which(colnames(mydataM)=="mother")] <- "parent"
      mydata2 <- rbind(mydataM,mydataF[,colnames(mydataM)])
      
      
      p <- ggplot2::ggplot(mydata2, ggplot2::aes(x=parent, y=predictedValue, label=designation) ) + 
        ggplot2::geom_jitter(position=ggplot2::position_jitter(width=0.3, height=0), ggplot2::aes(colour=parent), alpha=0.9) +
        ggplot2::geom_boxplot(alpha = 0.5, show.legend = FALSE, ggplot2::aes(fill=parent)) +
        ggplot2::theme(strip.text.x = ggplot2::element_text(size=9, color="black", face="bold"), 
                       axis.text.x = ggplot2::element_text(angle = 45), legend.position = "none") +
        ggplot2::labs(x="", y="Trait value") + ggplot2::facet_grid(trait~environment, scales = "free")# +
      # ggplot2::guides(color=ggplot2::guide_legend(""))
      
      fig <-  plotly::ggplotly(p)#, height = get_h()*2);
      fig
    })
  }
  
}
```






