---
title: "MTA Report"
author: ""
date: "2023-11-03"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
font.plot_ly <- list(family="Courier New, monospace", size=9)
```

```{r }
library(tidyverse)
suppressWarnings(tryCatch({
        load("./outputs/resultMta.RData")
      }, error = function(e) {
        NULL
      }))
if (is.null(resultMta)){
  invisible(NULL)
} else {
  idMta <- resultMta$status[which(resultMta$status$module == "mta"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred <- resultMta$predictions
  pred <- pred[pred$analysisId == idMta,]
  traits <- unique(pred$trait)
}
```


### Distribution by trait

The following boxplot allows to see the distribution of predicted values (y-axis) in the different environments for each **trait**.

<p>&nbsp;</p>

```{r }
if (is.null(resultMta)){
  invisible(NULL)
} else {
  selectInput(ns("traitMta"), 
            label = "", 
            choices = traits
            )
}

```

```{r }
if (is.null(resultMta)){
  invisible(NULL)
} else {
  idSta <- resultMta$status[which(resultMta$status$module == "sta"),"analysisId"]
  idSta <- idSta[length(idSta)]
  pred.sta <- resultMta$predictions
  pred.sta <- pred.sta[pred.sta$analysisId == idSta,]
  pred.sta.boxplot <- subset(pred.sta, select = c(trait,designation,environment,predictedValue))
  
  plotly::renderPlotly({
  temp.pred.box <- pred.sta.boxplot[pred.sta.boxplot$trait==input$traitMta,]
  p <- plotly::plot_ly(
  data = temp.pred.box,
  y = ~predictedValue,
  x = ~environment,
  type = "box",
  color = ~environment
  )
  plotly::layout(
    p,
    font=font.plot_ly,
    xaxis=list( title=""), 
    yaxis=list( title="Trait value")
    )
})
}
```

<p>&nbsp;</p>

### Environmental correlations

The following graph allows to see the genetic correlations of the different environment for each **trait**.

<p>&nbsp;</p>

```{r }
if (is.null(resultMta)){
  invisible(NULL)
} else {
  idMta <- resultMta$status[which(resultMta$status$module == "mta"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred <- resultMta$predictions
  pred <- pred[pred$analysisId == idMta,]
  traits <- unique(pred$trait)
  selectInput(ns("traitPredictionsCorrelation"),
              label = "",
              choices = traits
              )
}

```


```{r }
if (is.null(resultMta)){
  invisible(NULL)
} else {
  idSta <- resultMta$status[which(resultMta$status$module == "sta"),"analysisId"]
  idSta <- idSta[length(idSta)]
  pred.sta <- resultMta$predictions
  pred.sta <- pred.sta[pred.sta$analysisId == idSta,]
  traits <- unique(pred.sta$trait)
  predictions.gcorrE <- subset(pred.sta, select = c(trait,designation,environment,predictedValue))
  
  plotly::renderPlotly({
    p.cor.pred.env <- corPlotPredictions(predictions.gcorrE, input$traitPredictionsCorrelation, unitOfCorrelation="designation", correlatedAcross="environment",
                             valueForCorrelation="predictedValue",FALSE, FALSE, TRUE)
    plotly::layout(p.cor.pred.env,font=font.plot_ly)
    })
}
```

<p>&nbsp;</p>

## Trait correlations

The following graph allows to see the genetic correlations of the different traits for across environments.

<p>&nbsp;</p>

```{r }
if (is.null(resultMta)){
  invisible(NULL)
} else {
  idMta <- resultMta$status[which(resultMta$status$module == "mta"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred.mta <- resultMta$predictions
  pred.mta <- pred.mta[pred.mta$analysisId == idMta,]
  predictions.gcorrT <- subset(pred.mta, select = c(trait,designation,environment,predictedValue))
  plotly::renderPlotly({
    p.cor.pred.traits <- corPlotPredictions(predictions.gcorrE, traitPredictionsCorrelation=NULL, unitOfCorrelation="designation", correlatedAcross="trait",
                             valueForCorrelation="predictedValue",FALSE, FALSE, TRUE)
    plotly::layout(p.cor.pred.traits,font=font.plot_ly)
    })
}
```

<p>&nbsp;</p>

## Biplot by trait

The following graph allows to see the performance of the genotypes over the enevironments for each **trait**.

<p>&nbsp;</p>

```{r }
if (is.null(resultMta)){
  invisible(NULL)
} else {
  idMta <- resultMta$status[which(resultMta$status$module == "mta"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred.mta <- resultMta$predictions
  pred.mta <- pred.mta[pred.mta$analysisId == idMta,]
  traits <- unique(pred.mta$trait)
  # tags$style(type='text/css', 
  #            ".selectize-input { font-size: 10px; line-height: 6px;}
  #            .selectize-dropdown { font-size: 10px; line-height: 6px;}"
  #            )
  selectInput(ns("traitBiplot"),
            label = "",
            choices = unique(pred.mta$trait)
            )
}

```

```{r }
library(tidyverse)
if (is.null(resultMta)){
  invisible(NULL)
} else {
  idSta <- resultMta$status[which(resultMta$status$module == "sta"),"analysisId"]
  idSta <- idSta[length(idSta)]
  pred.sta <- resultMta$predictions
  pred.sta <- pred.sta[pred.sta$analysisId == idSta,]
  
plotly::renderPlotly({
  pred.mta.biplot <- subset(pred.sta, select = c(trait,designation,environment,predictedValue)) %>%
  filter(trait==input$traitBiplot) %>% #
  droplevels() %>%
  pivot_wider(names_from = environment, values_from = predictedValue) %>%
  select(-trait) %>%
  column_to_rownames("designation")

pred.mta.biplot <- na.omit(pred.mta.biplot[,2:4])

prin_comp <- prcomp(pred.mta.biplot, rank = 2)
components <- prin_comp[["x"]]
components <- data.frame(components)
components$PC2 <- -components$PC2

explained_variance <- summary(prin_comp)[["sdev"]]
explained_variance <- explained_variance[1:2]
comp <- prin_comp[["rotation"]]
comp[,'PC2'] <- - comp[,'PC2']
loadings <- comp
for (i in seq(explained_variance) ){
  loadings[,i] <- comp[,i] * explained_variance[i]
}

p.mta.biplot <- plotly::plot_ly (type = 'scatter', mode = 'markers')
p.mta.biplot <- plotly::add_trace(p.mta.biplot, 
                  x=components$PC1,
                  y=components$PC2,
                  text=rownames(components),
                  marker=list(color="blue"),
                  hoverinfo='text',
                  showlegend=FALSE
                  )


for (i in 1:ncol(pred.mta.biplot)){
  p.mta.biplot <- p.mta.biplot %>%
    plotly::add_segments(x = 0, xend = loadings[i, 1], y = 0, yend = loadings[i, 2], line = list(color = 'red'),inherit = FALSE, showlegend = FALSE) %>%
    plotly::add_annotations(x=loadings[i, 1], y=loadings[i, 2], ax = 0, ay = 0,text = rownames(loadings)[i], xanchor = 'center', yanchor= 'bottom')
}
    p.mta.biplot
    plotly::layout(p.mta.biplot,font=font.plot_ly)
    })


}

```


<p>&nbsp;</p>

### Predictions 

The following table allows to check the trait predictions in wide format.

<p>&nbsp;</p>

```{r }
if (is.null(resultMta)){
  invisible(NULL)
} else {
  idMta <- resultMta$status[which(resultMta$status$module == "mta"),"analysisId"]
  idMta <- idMta[length(idMta)]
  pred.mta <- resultMta$predictions
  pred.mta <- pred.mta[pred.mta$analysisId == idMta,]
  traits <- unique(pred$trait)
  pred.mta <- subset(pred.mta, select = c(trait,designation,environment,predictedValue))
  pred.mta.wide <- reshape(
    data=pred.mta, 
    timevar = c("trait"), 
    idvar = c("environment","designation"), 
    direction="wide"
  )
  names(pred.mta.wide) <- c("designation","environment",traits)
  DT::renderDT({DT::formatRound(DT::datatable(
    pred.mta.wide,
    extensions = 'Buttons',
    rownames = FALSE,
    class = 'cell-border',
    options = list(
      dom = 'Blfrtip',
      scrollY = "400px",
      scrollX = "400px",
      buttons = c('copy','csv','excel','pdf','print'),
      paging = F)
  ),traits)
  })
}

```




